### 2024-03-18 report.qmd
add 1st draft of basic report (to move later)

---
title: "The Temporal Network of Hospital Patient Transfers Paper"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: OTHER AUTHORS
date: today
format:
    html: default
    pdf: default
---

# Introduction

Antimicrobial resistance is bad. In Victoria, Australia, we have observed CPE, and want to understand how the underlying movement of patients could affect the spread of CPE and other related AMR in the future.
The focus of this paper is on the patient movement network: ways of representing, simplifying, and approximating those patient movements; and how those representations of patient movement affect the spread of a hypothetical infection between hospitals in the Victorian healthcare system.

# Methods

## Generating the temporal network
We have line-listed patient movement data between {start_date} and {end_date}. This allows us to generate exact patient movement networks up to minute precision.
By arranging patient journeys within this dataset, we can construct a network of patient transfers when two consecutive (but not necessarily temporally adjacent) admissions of the same patient occur at different healthcare facilities.
These then define a temporal network where a (single) patient being discharged from hospital $A$ at time $t_1$ and admitted to hospital $B$ at a later time $t_2$, where there are no other event in between, contributes to a weight of 1 on the edge $(A, t_1) \to (B, t_2)$.
To simplify the computational expense of such a network, and to protect patient anonymity, we choose to discretise events in the system. This is done by choosing some window size $W$, and treating events that occur between $t$ and $t+W$ as all occuring at $t$. This causes some transfers to "disappear" depending on the choice of $W$.
The patient transfers can be "indirect", where the patient is not present in the hospital system for some amount of time between consecutive admissions, i.e. when $t_2 \neq t_1$. From inspecting the dataset, we see that there is a significant number of transfers that are indirect, and the time between consecutive admissions is distributed sub-exponentially, witha  heavy tail.
These types of transfers imply causal edges between hospitals that are neither synchronous (occur at the same time), nor persistent (where the effect of a cause at time $t_1$ is not felt over some interval $(t_1, t_2)$), which make it challenging to fit a temporal network generated by this dataset into classical temporal network frameworks of representation.

## Projecting the temporal network
We can project these causal edges by making their effects persistent in a hospital. 
So an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:

1. $(A, t_1) \to (B, t_1)$ and $(B, t_1) \to \dots \to (B, t_2)$
2. $(A, t_1) \to \dots \to (A, t_2)$ and $(A, t_2) \to (B, t_2)$

This causes an implied increased infection pressure at certain hospital nodes. This sort of projection also loses specificity of temporal causality. By inspecting the temporal network that results from this projection, you cannot tell if a patient is moving from $t_1$ to $t_2$ or to $t_3$.

There is also a mechanistically distinct projection, where we introduce "home" nodes for each facility. For some facility $A$, the home node is denoted $A'$.
Thus we could also project the edge as

- $(A, t_1) \to (A', t_1)$ and $(A', t_1) \to \dots \to (A', t_2)$ and $(A', t_2) \to (B, t_2)$.

This is an ideal projection, but is difficult to achieve. Specifically, it is difficult to control the duration that an individual spends in the home node.
However, this projection still loses temporal causal specificty, as above.

## Simulating on Networks

We use a discrete time Markov chain model to simulate the spread of infection. Simulations step forward in timesteps of $\delta t$, and events within the window $\delta t$ are assumed to be independent.
For simplicity, we model the infection as SI or SIS, depending on the choice of removal rate; we assume that all facilities are at capacity all the time, which allows us to remove the explciit tracking of susceptibles.
When individuals are removed, they are randomly chosen to either never return or eventually return (to the hospital system); if they do, they are randomly placed into a transfer bucket which has two properties:

- where they will readmit
- when they will readmit

For a "full" temporal simulation, both where and when are specified; for a "reduced" simulation, only "where" is specified, and "when" falls into one of two buckets: immediately, or not immediately.
We use a "full" temporal simulation with the unprojected temporal network, and a "reduced" simulation for any projection of that temporal network that has lost temporal information.

<Specification of the model here>

| Parameter | Description |
|-----------|---------------------------------------|
| $\beta$   | Rate of infection per unit infected per day |
| $\eta$    | Rate of removal/recovery |
| $\gamma$  | Rate of discharge from hospital |
| $M$       | Transition matrix inferred from network |

# Results

We run a range of experiments.

## Leading results

## Linelist simulations

We can also run this model structure on top of the actual line list, instead of aggregating patient movement into a network.

## No infection simulations

To isolate the effect of the projections on the movement of patients, we can simulate the model with parmeters $\beta = 0, \eta=0$.

## Home node projection simulations

## Markovian Approximation of movement

### 2024-06-06 network_conversion.py
Add util for conforming other mappings to ordering

@@ -36,6 +36,9 @@ class Ordering():
         size_mapping = esz.quick_read(file_name)
         return cls(size_mapping)
 
+    def conform(self, other: Mapping):
+        return [other.get(x) for x in self]
+
 
 class TemporalNetworkConverter():
     def __init__(self, network: ig.Graph, ordering: Ordering, weight: str|None=None):

### 2024-06-06 network_simulation.py
Add skeleton export

@@ -1,10 +1,9 @@
 """this is a base refactor of on_network_simulation"""
 
-
 from typing import Sequence, Mapping
 import numpy as np
 from util import BlackHole
-from numba_sample import multinomial_sparse_full
+from numba_sample import multinomial_sparse_full, truncated_poisson
 
 class Simulation():
     def __init__(self, full_size: Sequence[int], parameters: Mapping, dt=1.0):
@@ -14,13 +13,20 @@ class Simulation():
         self.parameters = parameters
         self.dt = dt
 
-        self.ts = []
-        self._history = []
+        self.ts = [0.0]
+        self._history = [self.state] # softref here should update as initial self.state updates in seed
 
     @property
     def history(self):
         return np.hstack(self._history)
-    
+
+    def export_history(self, to):
+        np.savez_compressed(
+            file = to,
+            ts = self.ts,
+            history = self.history,
+        )
+
     def reset(self, soft=True):
         if soft:
             self.state = self._history[0]
@@ -108,6 +114,13 @@ class SnapshotWithHome(Simulation):
 
         self.shadow_state = np.zeros((self.NHOSP, self.NHOSP), dtype=np.int64)
         self.transient_shadow = np.zeros_like(self.shadow_state, dtype=np.int64)
+
+    def export_history(self, to):
+        np.savez_compressed(
+            file = to,
+            ts = self.ts,
+            history = self.history,
+        )
 
     def reset(self, soft=True):
         super().reset(soft=soft)

### 2024-06-06 network_simulation.py
Update truncated poisson draw (~10x)

@@ -161,7 +161,7 @@ class SnapshotWithHome(Simulation):
         indirect_move_to = move_to[:, self.NHOSP :].reshape(WIDE)
 
         indirect_return_rate = TX_in[self.current_index] * X
-        indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, 0, X)
+        indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, X)
         indirect_returns = indirect_returns_raw.sum(axis=1).reshape(NARROW)
         self.mover_in.append(indirect_returns)
 

### 2024-06-06 numba_sample.py

@@ -85,16 +85,49 @@ def multinomial_sparse_full(trials, prob_matrix: sparse.csr_matrix):
     )
 
 
-@guvectorize([(float64[:], int64, int64[:], int64[:])], "(n),(),(n)->(n)")
-def _trunc_pois_u(lam, lb, ub, res):
-    for i, (v, u) in enumerate(zip(lam, ub)):
-        res[i] = max(min(np.random.poisson(v), u), lb)
+@njit()
+def truncated_poisson(rates, upper_bound):
+    """Performs an upper-truncated poisson sampling over a 2D array of rates
+    
+    Arguments
+    ---
+    rates: <np.ndarray[float]> 2D array of rates to sample a poisson-distributed random variable at
+    upper_bound: <np.ndarray[int]> 2D array of upper bounds to truncate the random draw at
 
+    Notes
+    ---
+    Structure taken from https://stackoverflow.com/a/72465167
+    Things to note are 
+        - explicit bound checking avoidance via assert
+        - iteration via indexing
+        - reduction of branching using min
+            - branch for the zero is justified, as we expect rates to be relatively sparse
+                - avoids expensive random draw
+    """
 
-def truncated_poisson(rates, lower_bound, upper_bound):
-    x = _trunc_pois_u(
-        rates.flatten(),
-        lower_bound,
-        upper_bound.flatten(),
-    )
-    return x.reshape(rates.shape)
+    ny, nx = rates.shape
+    assert rates.shape == upper_bound.shape
+
+    draws = np.zeros((nx, ny), dtype=np.int64)
+
+    for i in range(nx):
+        rates_row = rates[i, :]
+        ub_row = upper_bound[i, :]
+
+        for j in range(ny):
+            assert j >= 0
+
+            lmbd = rates_row[j]
+            ub = ub_row[j]
+
+            if lmbd == 0 or ub == 0:
+                # draws[i,j] = 0 
+                continue
+
+            draw = np.random.poisson(lmbd)
+
+            draw = min(draw, ub)
+
+            draws[i,j] = draw
+
+    return draws

### 2024-06-06 on_network_simulation.py

@@ -538,7 +538,7 @@ class SnapshotNoveauSimulation(Simulation):
         indirect_move_to = move_to[:, self.NHOSP :].reshape(WIDE)
 
         indirect_return_rate = self.transition_matrices["in"][self.current_index] * X
-        indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, 0, X)
+        indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, X)
         indirect_returns = indirect_returns_raw.sum(axis=1).reshape(NARROW)
         self.mover_in.append(indirect_returns)
 

### 2024-06-07 report.qmd
add notes on how the snapshot proj works

@@
     - name: OTHER AUTHORS
-date: today
+date: last-modified
 format:
@@
 ## Projecting the temporal network
-We can project these causal edges by making their effects persistent in a hospital. 
-So an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:
+We can project these causal edges by making their effects persistent in a hospital. This reduces the amount of information stored, so in theory should improve performance of simulations.
+
+With a naive approach, an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:
 
@@
 However, this projection still loses temporal causal specificty, as above.
+
+### Our solution
+
+We project the temporal network into a series of multilayer networks.
+Instantaneous movements (consisting of direct hospital-hospital transfers) make up one layer; movements from hospital to home make up another layer; and movements from home to hospital make up the third layer.
+By doing this, we can more accurately capture the time that a patient spends at home between hospital visits, but without explicitly tracking individuals (just the number of patients along each path).
 
@@
 ## Markovian Approximation of movement
+
+# Appendix
+
+## Estimation of parameters for the projected network simulations
+
+The rate / transition matrix that defines the rate of return from home to hospital is characterised by the matrix $M_{in}$, such that
+
+$$M_{in}(i, j, t) = \frac{n_r(i, j, t)}{n_h(i, j, t)},$$
+
+where $i$ is the index for the hospital an individual was at prior to going home, $j$ the index for the hospital an individual is going to from home, $t$ the current time, $n_r$ the number of individuals that are returning from home, and $n_h$ the number of individuals that are currently at home (and will return to a hospital at some point in the future).
+
+We observe $n_r$ in the data, but we do not have any direct observations of $n_h$.
+We could compute this by iterating through the data, counting the number of individuals going home and returning to hospital, but this is expensive, and not necessarily available in all cases.
+Instead, we utilise the available values of $n_g$, the number of individuals that go home from hospital.
+We have a estimated return time distribution for individuals characterised by the pdf $r(\tau)$, so we apply this by estimating $n_h$ as
+
+$$n_h(i, j, t) \approx \int_{t-T}^t n_g(i, j, \tau) r(t - \tau) d\tau.$$
+
+We truncate this approximation at some bandwith time $T$.
+
+The retrun time distribution is presented in @figure-??. We fit a power law model $r(\tau) \sim \tau ^k$ with coefficient $k \approx -0.6$.

### 2024-08-19 network_conversion.py
Add in static network simulation

@@ -1,15 +1,18 @@
+import abc
+from os import PathLike
 import numpy as np
 import numba as nb
 from scipy import sparse
 import polars as pl
 import igraph as ig
-from typing import Mapping, Hashable, Sequence
+from typing import Mapping, Hashable, Sequence, Any, SupportsFloat as Numeric, Iterable
 from functools import lru_cache
 from pathlib import Path
-from util import Iden, nparr_find, NotFound
+from util import Iden, nparr_find, NotFound, SupportsGet
 import examine_transfers_for_sizes as esz
 
-class Ordering():
+
+class Ordering:
     def __init__(self, order_mapping: Mapping):
         """Constructs an ordering of objectes, based on a mapping of values that can be ordered"""
         self.order = np.asanyarray(sorted(order_mapping.keys()))
@@ -22,6 +25,9 @@ class Ordering():
     def __iter__(self):
         for item in self.order:
             yield item
+
+    def __len__(self):
+        return len(self.order)
 
     @lru_cache()
     def __getitem__(self, key):
@@ -40,15 +46,22 @@ class Ordering():
     def conform(self, other: Mapping):
         return [other.get(x) for x in self]
 
-class ColumnDict(dict):
+    def __call__(self, key):
+        """returns the size of the object"""
+        idx = self[key]
+        return self.sizes[idx]
 
+
+class ColumnDict(dict):
     @classmethod
     def from_file(cls, file, key_name, value_name):
         df = pl.read_csv(file)
-        return cls(zip(
+        return cls(
+            zip(
                 df.select(key_name).to_series().to_list(),
                 df.select(value_name).to_series().to_list(),
-            ))
+            )
+        )
 
     @classmethod
     def from_prob_final_file(cls, file):
@@ -57,17 +70,29 @@ class ColumnDict(dict):
     def organise_by(self, ordering: Ordering):
         return ordering.conform(self)
 
-class TemporalNetworkConverter():
-    def __init__(self, network: ig.Graph, ordering: Ordering, weight: str|None=None):
+
+class Converter(abc.ABC):
+    @abc.abstractmethod
+    def __init__(self, network: ig.Graph, ordering: Ordering):
+        pass
+
+    @abc.abstractmethod
+    def map_parameters(self, parameters: Mapping):
+        pass
+
+
+class TemporalNetworkConverter(Converter):
+    def __init__(
+        self, network: ig.Graph, ordering: Ordering, weight: str | None = None
+    ):
[…]

### 2024-08-19 network_simulation.py

@@ -7,7 +7,8 @@ from util import BlackHole
 from numba_sample import multinomial_sparse_full, truncated_poisson
 import h5py
 
-class Simulation():
+
+class Simulation:
     def __init__(self, full_size: Sequence[int], parameters: Mapping, dt=1.0):
         self.N = np.asanyarray(full_size).reshape((-1, 1))
         self.NHOSP, _ = self.N.shape
@@ -16,17 +17,19 @@ class Simulation():
         self.dt = dt
 
         self.ts = [0.0]
-        self._history = [self.state] # softref here should update as initial self.state updates in seed
+        self._history = [
+            self.state
+        ]  # softref here should update as initial self.state updates in seed
 
     @property
     def history(self):
         return np.hstack(self._history)
 
     def export_history(self, to, identity, **kwargs):
-        with h5py.File(to, mode='a') as h5f:
+        with h5py.File(to, mode="a") as h5f:
             g = h5f.create_group(identity)
-            g.create_dataset('ts', data=self.ts, compression='gzip')
-            g.create_dataset('history', data=self.history, compression='gzip')
+            g.create_dataset("ts", data=self.ts, compression="gzip")
+            g.create_dataset("history", data=self.history, compression="gzip")
             g.attrs.update(kwargs)
 
     def reset(self, soft=True):
@@ -55,10 +58,10 @@ class Simulation():
         """Performs a sinble time step of the simulation
         Returns the state at the end of the time step"""
 
-        beta = self.parameters['beta']
-        gamma = self.parameters['gamma']
-        pstay = self.parameters['prob_final_stay']
-        WW = self.parameters['weighting_matrix']
+        beta = self.parameters["beta"]
+        gamma = self.parameters["gamma"]
+        pstay = self.parameters["prob_final_stay"]
+        WW = self.parameters["weighting_matrix"]
 
         I = self.state
 
@@ -75,7 +78,7 @@ class Simulation():
 
         # compute movements
         # first, draw how many stay
-        n_mov = self.rng.binomial(n_rec, 1-pstay)
+        n_mov = self.rng.binomial(n_rec, 1 - pstay)
         # second, draw where they go
         M_mov_I = self.rng.multinomial(n_mov, WW)
         n_mov_I = M_mov_I.sum(axis=0)
@@ -96,16 +99,9 @@ class Simulation():
                 break
 
 
-class SnapshotWithHome(Simulation):
[…]

### 2024-08-19 staticify_network.py

@@ -9,11 +9,21 @@ def load_temporal_network(path):
     return gim.make_graph(path)
 
 
+def temporal_to_static(temponet, direct_threshold=4):
+    edge_list = convert_temporal_network_to_static_edge_list(
+        temponet=temponet, direct_threshold=direct_threshold
+    )
+
+    static_graph = generate_graph_from_edge_list(edge_df=edge_list)
+    static_graph["time_span"] = max(temponet.vs["time"])
+    return static_graph
+
+
 def convert_temporal_network_to_static_edge_list(
     temponet: ig.Graph, direct_threshold=4
 ) -> pl.DataFrame:
     """Convert a full temporal network into a dataframe of edges of the corresponding static network
-    
+
     Differentiates between direct transfers and indirect transfers"""
 
     # extract network information
@@ -33,7 +43,7 @@ def convert_temporal_network_to_static_edge_list(
     # add in column for direct transfer bool flag
     edge_df = edge_df.with_columns(
         direct=((pl.col("dest_time") - pl.col("source_time") < direct_threshold)),
-        link_time=(pl.col('dest_time') - pl.col('source_time'))*pl.col('weight')
+        link_time=(pl.col("dest_time") - pl.col("source_time")) * pl.col("weight"),
     )
 
     # combine edges across time
@@ -53,7 +63,7 @@ def convert_temporal_network_to_static_edge_list(
             pl.col("link_time").list.sum(),
         )
         .with_columns(
-            (pl.col("link_time") / pl.col("indirect_weight")).alias('link_time')
+            (pl.col("link_time") / pl.col("indirect_weight")).alias("link_time")
         )
     )
 
@@ -74,23 +84,26 @@ def generate_graph_from_edge_list(edge_df: pl.DataFrame) -> ig.Graph:
         .to_series()
         .to_list()
     )
-    node_index = {nd: i for i,nd in enumerate(nodes)}
+    node_index = {nd: i for i, nd in enumerate(nodes)}
 
-    G.add_vertices(len(nodes), attributes={'node': nodes})
+    G.add_vertices(len(nodes), attributes={"node": nodes})
 
+    # drop direct self loops and map node indices
     mapped_edge_df = edge_df.with_columns(
-        pl.col('source_loc').map_dict(node_index),
-        pl.col('dest_loc').map_dict(node_index),
[…]

### 2024-08-19 util.py

@@ -1,4 +1,4 @@
-from typing import Any
+from typing import Any, Protocol, TypeVar
 import numpy as np
 import numba as nb
 
@@ -11,7 +11,7 @@ class Iden():
     
     def __setitem__(self, x):
         raise TypeError("Identity objects do not support assignment")
-    
+
 class BlackHole():
     def __init__(self):
         pass
@@ -30,3 +30,9 @@ def nparr_find(array: np.ndarray, item: Any):
         
 class NotFound(Exception):
     pass
+
+T = TypeVar('T', covariant=True)
+K = TypeVar('K', contravariant=True)
+class SupportsGet(Protocol[K, T]):
+    def __getitem__(self, key: K) -> T:
+        ...

### 2024-08-19 report.qmd
Fixup: snapshot sims now use accurate shadow state values to compute in rate

@@
 The retrun time distribution is presented in @figure-??. We fit a power law model $r(\tau) \sim \tau ^k$ with coefficient $k \approx -0.6$.
+
+## Simulation Details and Model
+
+We keep records of two sets of states:
+
+1. $n(t, x)$, the current number of infected at location $x$. This represents individuals in hospital.
+2. $u(t, x, y)$, the current number of individuals that came from location $x$ and will depart for location $y$ at some point in the future. This represents the number of people at home.
+
+Each snapshot can be considered a graph composed of vertices $V$ that correspond to locations, and have edges $E$.
+
+Edges of the temporal network, $e(x, s, y, t)$ from location $x$ at time $s$ to location $y$ at time $t$ with some weight $w$ can be decomposed:
+
+1. if $s = t$, construct an edge $e(x, y)$ with weight $w$
+2. if $s \neq t$ construct two edges:
+   1. $e_o(x, y, s)$ if it doesn't exist, and add weight $w$
+   2. $e_i(x, y, t)$ if it doesn't exist, and add weight $w$
+
+We can alternatively conceptualise this as decomposing the temporal edges into movements to a "home" node for each pair of hospitals.
+The edges $e_o(x, y, s)$ can be conceptualised as an edge between $x$ and a new "home" node denoted by the ordered pair $(x, y)$ at time $s$.
+
+```{python}
+#| echo: false
+
+# we're going to use networkx and its layout engine to draw the diagram cause the other engines are dogwater
+import networkx as nx
+from matplotlib import pyplot as plt
+
+fig = plt.figure()
+grid = fig.add_gridspec(2, 1, hspace=0)
+ax1, ax2 = grid.subplots(sharex=True, squeeze=True)
+for ax in (ax1, ax2):
+    ax.set_box_aspect(1)
+
+ABC = nx.DiGraph(['AB', 'BC', 'BA', 'CB', 'AC', 'CA'])
+nx.draw_networkx(ABC, ax=ax1, with_labels=True)
+
+```
+
+## Searching snapshots for information on unlikely paths
+
+We observed some empirical simulated hitting time survival curves that suggested that paths that did exist and were likely on the temporal network did not exist or were very unlikely on the equivalent snapshots.
+This seems confusing, since the snapshot representation _should_ represent a superset of possible movements.

### 2024-08-19 netsim_summariser.py

@@ -1,47 +1,32 @@
 import h5py
 import polars as pl
 import numpy as np
-
+import multiprocessing
+from itertools import repeat
 
 class Summariser():
 
     def __init__(self, file):
         self.target_file = file
-        self._open = False
-
-    def __enter__(self):
-        self._file = h5py.File(self.target_file)
-        self._open = True
-        return self
-    
-    def __exit__(self, exc_type, exc_val, exc_tb):
-        self._open = False
-        self._file.close()
-
-    def ensure_open(self):
-        if self._open:
-            return
-        else:
-            self._file = h5py.File(self.target_file)
-            self._open = True
 
     @staticmethod
     def hitting_time(target : h5py.Group, null=np.nan, **kwargs):
         mask = (target['history'][:] != 0)
-        return np.where(mask.any(axis=1), mask.argmax(axis=1), null)
+        return list(np.where(mask.any(axis=1), mask.argmax(axis=1), null))
 
     @staticmethod
     def movement_out_total(target: h5py.Group, **kwargs):
-        return target['mover_out'][:].sum(axis=1)
+        return list(target['mover_out'][:].sum(axis=1))
     
     @staticmethod
     def movement_in_total(target: h5py.Group, **kwargs):
-        return target['mover_in'][:].sum(axis=1)
+        return list(target['mover_in'][:].sum(axis=1))
 
     @staticmethod
     def steady_state_infected(target: h5py.Group, **kwargs):
         init = target['history'][:,0].sum()
-        return target['history'][:].sum(axis=0) / init
+        normed = target['history'][:].sum(axis=0) / init
+        return [np.mean(normed), np.std(normed)]
     
     @staticmethod
     def emptying_time(target: h5py.Group, null=np.nan, **kwargs):
@@ -49,29 +34,65 @@ class Summariser():
         mask = (xs == 0)
         return mask.argmax() if mask.any() else null
 
-    def collect(self, over=None, verbose=False):
-        self.ensure_open()
-        if over is None:
-            over = list(self._file)
+    @property
+    def groups(self):
+        with h5py.File(self.target_file, 'r') as fp:
+            grps = list(fp)
+        return grps
 
-        results = []
-
-        recall = lambda *x: None
+    def compute_metrics(self, target_str, no_move=False, verbose=False):
         if verbose:
-            recall = print
+            print(target_str)
 
-        for sim in over:
-            recall(sim)
-            target = self._file[sim]
[…]

### 2024-08-19 new_network_sim.py

@@ -3,11 +3,11 @@ import numpy as np
 import network_conversion as netcon
 import network_simulation as netsim
 
-def create_snaphome_parameter_converter(ordering_file, snapshots_directory, kerpow=-0.6, trunc=26):
+def create_snaphome_parameter_converter(ordering_file, snapshots_directory):
 
     ordering = netcon.Ordering.from_file(ordering_file)
 
-    converter = netcon.SnapshotWithHomeConverter.from_directory(snapshots_directory, ordering=ordering, kernel_expo=kerpow, trunc=trunc)
+    converter = netcon.SnapshotWithHomeConverter.from_directory(snapshots_directory, ordering=ordering)
 
     return converter
 
@@ -65,15 +65,15 @@ def simulate_sim_and_record(simulation: netsim.SnapshotWithHome, simid, until=10
 if __name__ == "__main__":
     import datetime
 
-    # conv = create_snaphome_parameter_converter(
-    #     ordering_file="./concordant_networks/size_14.csv",
-    #     snapshots_directory="./conc_tempo_14_detailed/",
-    # )
+    conv = create_snaphome_parameter_converter(
+        ordering_file="./concordant_networks/size_14.csv",
+        snapshots_directory="./conc_tempo_14_detailed/",
+    )
 
-    import pickle
+    # import pickle
 
-    with open("snapconverter.pkl", 'rb') as ifp:
-        conv = pickle.load(ifp)
+    # with open("snapconverter.pkl", 'rb') as ifp:
+    #     conv = pickle.load(ifp)
 
     params = init_zero_params(
         conv, "./probability_of_final_stay_by_shuffled_campus.csv"
@@ -96,5 +96,5 @@ if __name__ == "__main__":
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_ss_new_sims/sim_all.h5"
+            outname = f"zero_ss_exact_home_sims/sim_all.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)

### 2024-08-19 summary_stats_zero.py

@@ -2,9 +2,11 @@
 
 import re
 from pathlib import Path
+from typing import Iterable
 from matplotlib import pyplot as plt
 import numpy as np
 import polars as pl
+import h5py
 
 """
 - hitting time
@@ -15,7 +17,7 @@ import polars as pl
 class Realisation:
     __slots__ = [
         "path",
-        "t",
+        "ts",
         "history",
         "mover_out",
         "mover_in",
@@ -95,7 +97,14 @@ class RealisationArray():
         for rlz in self.realisations[filter]:
             collation.append(getattr(Realisation(rlz), metric)(*args, **kwargs))
         return collation
-    
+
+    def multicollect(self, metrics: Iterable[str], filter=slice(0, None), *args, **kwargs):
+        collation = {metric:[] for metric in metrics}
+        for rlz in self.realisations[filter]:
+            for metric in metrics:
+                collation[metric].append(getattr(Realisation(rlz), metric)(*args, **kwargs))
+        return collation
+
     @property
     def metadata(self):
         if self._metadata is None:
@@ -106,3 +115,19 @@ class RealisationArray():
                 schema=Realisation.base_name_schema[self.name_pattern]
             ).with_row_count()
         return self._metadata
+    
+
+    def to_hdf5(self, out):
+        with h5py.File(out, 'w') as h5f:
+            for rlz in self.metadata.to_dicts():
+                g = h5f.create_group(f"sim_{rlz['row_nr']}")
+                g.attrs['seed'] = rlz['seed']
+                g.attrs['simdate'] = rlz['simdate']
+                g.attrs['simtime'] = rlz['simtime']
+                rfl = self.realisations[rlz['row_nr']]
+                with np.load(rfl) as ifp:
+                    tarr = ifp.get('ts') or ifp['t']
+                    g.create_dataset('ts', data=tarr.flatten(), compression='gzip')
+                    g.create_dataset('history', data=ifp['history'], compression='gzip')
+                    g.create_dataset('mover_in', data=np.squeeze(ifp['mover_in']), compression='gzip')
+                    g.create_dataset('mover_out', data=np.squeeze(ifp['mover_out']), compression='gzip')

### 2024-08-19 new_network_sim.py
remove pkl cache of converter

@@ -70,11 +70,6 @@ if __name__ == "__main__":
         snapshots_directory="./conc_tempo_14_detailed/",
     )
 
-    # import pickle
-
-    # with open("snapconverter.pkl", 'rb') as ifp:
-    #     conv = pickle.load(ifp)
-
     params = init_zero_params(
         conv, "./probability_of_final_stay_by_shuffled_campus.csv"
     )

### 2024-09-18 reports/report.qmd
add and modify analysis stuff this is pre-dist upgrade panic

--- 2024-08-19 report.qmd
@@
 ---
-title: "The Temporal Network of Hospital Patient Transfers Paper"
+title: "Temporality is Important for Networks of Hospital Patient Transfers"
 authors:
@@
     - name: Nic Geard
-    - name: OTHER AUTHORS
+    - name: "******"
 date: last-modified
@@
 ---
+
+# Key Points
+
+- There is a significant amount of patient movement in the Victorian hospital system where the patient returns home for a substative amount of time between discharge from one hospital and subsequent admission at another hospital.
+- This "indirect" movement has implications on the rate of spread of a hypothetical contagion through the hospital system. It causes dispersal of patients to be slower than expected.
+- We can recover some of this behaviour by introducing an additional layer in the network to represent indirect movement.
 
@@
 
-# Methods
+# Network modelling of hospital patient movement
 
@@
 By arranging patient journeys within this dataset, we can construct a network of patient transfers when two consecutive (but not necessarily temporally adjacent) admissions of the same patient occur at different healthcare facilities.
-These then define a temporal network where a (single) patient being discharged from hospital $A$ at time $t_1$ and admitted to hospital $B$ at a later time $t_2$, where there are no other event in between, contributes to a weight of 1 on the edge $(A, t_1) \to (B, t_2)$.
-To simplify the computational expense of such a network, and to protect patient anonymity, we choose to discretise events in the system.
-This is done by choosing some window size $W$, and treating events that occur between $t$ and $t+W$ as all occuring at $t$.
-This causes some transfers to "disappear" depending on the choice of $W$.
-The patient transfers can be "indirect", where the patient is not present in the hospital system for some amount of time between consecutive admissions, i.e.
+We can model this data as a temporal network where a (single) patient being discharged from hospital $A$ at time $t_1$ and admitted to hospital $B$ at a later time $t_2$, given the patient is not admitted to any other hospital in the interim, contributes to a weight of 1 on the edge $(A, t_1) \to (B, t_2)$.
+To reduce the computational expense of such a network, and to protect patient anonymity, we aggregate these patient transfer events to a discrete set of times.
+By choosing some window size $W$, we treat events that occur in the time interval $[t_i, t_i+W)$ as all occuring at $t_i$.
+This aggregation classfies patient transfers into two classes. 
+Transfers can be "direct", where the discharge and subsequent admission both occur within the same time window $[t_i, t_i+W)$.
+Alternatively they can be "indirect", where the patient is not present in the hospital system for some amount of time between discharge and admission, i.e.
 when $t_2 \neq t_1$.
-From inspecting the dataset, we see that there is a significant number of transfers that are indirect, and the time between consecutive admissions is distributed sub-exponentially, witha  heavy tail.
+From inspecting the dataset, we see that there is a significant number of transfers that are indirect, and the time between consecutive admissions is distributed sub-exponentially, with a heavy tail.
 These types of transfers imply causal edges between hospitals that are neither synchronous (occur at the same time), nor persistent (where the effect of a cause at time $t_1$ is not felt over some interval $(t_1, t_2)$), which make it challenging to fit a temporal network generated by this dataset into classical temporal network frameworks of representation.
-
-## Projecting the temporal network
-We can project these causal edges by making their effects persistent in a hospital. This reduces the amount of information stored, so in theory should improve performance of simulations.
-
-With a naive approach, an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:
-
-1. $(A, t_1) \to (B, t_1)$ and $(B, t_1) \to \dots \to (B, t_2)$
-2. $(A, t_1) \to \dots \to (A, t_2)$ and $(A, t_2) \to (B, t_2)$
-
-This causes an implied increased infection pressure at certain hospital nodes.
-This sort of projection also loses specificity of temporal causality.
-By inspecting the temporal network that results from this projection, you cannot tell if a patient is moving from $t_1$ to $t_2$ or to $t_3$.
-
-There is also a mechanistically distinct projection, where we introduce "home" nodes for each facility. For some facility $A$, the home node is denoted $A'$.
-Thus we could also project the edge as
-
-- $(A, t_1) \to (A', t_1)$ and $(A', t_1) \to \dots \to (A', t_2)$ and $(A', t_2) \to (B, t_2)$.
-
-This is an ideal projection, but is difficult to achieve. Specifically, it is difficult to control the duration that an individual spends in the home node.
-However, this projection still loses temporal causal specificty, as above.
-
-### Our solution
-
-We project the temporal network into a series of multilayer networks.
-Instantaneous movements (consisting of direct hospital-hospital transfers) make up one layer; movements from hospital to home make up another layer; and movements from home to hospital make up the third layer.
-By doing this, we can more accurately capture the time that a patient spends at home between hospital visits, but without explicitly tracking individuals (just the number of patients along each path).
 
@@
 
-## Linelist simulations
-
-We can also run this model structure on top of the actual line list, instead of aggregating patient movement into a network.
-
 ## No infection simulations
@@
 
-## Home node projection simulations
-
-## Markovian Approximation of movement
-
 # Appendix
 
-## Estimation of parameters for the projected network simulations
+## Projecting the temporal network
+We can project these causal edges by making their effects persistent in a hospital. This reduces the amount of information stored, so in theory should improve performance of simulations.
+
+With a naive approach, an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:
+
+1. $(A, t_1) \to (B, t_1)$ and $(B, t_1) \to \dots \to (B, t_2)$
+2. $(A, t_1) \to \dots \to (A, t_2)$ and $(A, t_2) \to (B, t_2)$
+
+This causes an implied increased infection pressure at certain hospital nodes.
+This sort of projection also loses specificity of temporal causality.
+By inspecting the temporal network that results from this projection, you cannot tell if a patient is moving from $t_1$ to $t_2$ or to $t_3$.
+
+There is also a mechanistically distinct projection, where we introduce "home" nodes for each facility. For some facility $A$, the home node is denoted $A'$.
+Thus we could also project the edge as
+
+- $(A, t_1) \to (A', t_1)$ and $(A', t_1) \to \dots \to (A', t_2)$ and $(A', t_2) \to (B, t_2)$.
+
+This is an ideal projection, but is difficult to achieve. Specifically, it is difficult to control the duration that an individual spends in the home node.
+However, this projection still loses temporal causal specificty, as above.
+
+We project the temporal network into a series of multilayer networks.
+Instantaneous movements (consisting of direct hospital-hospital transfers) make up one layer; movements from hospital to home make up another layer; and movements from home to hospital make up the third layer.
+By doing this, we can more accurately capture the time that a patient spends at home between hospital visits, but without explicitly tracking individuals (just the number of patients along each path).
+
+<!-- ## Estimation of parameters for the projected network simulations
 
@@
 
-The retrun time distribution is presented in @figure-??. We fit a power law model $r(\tau) \sim \tau ^k$ with coefficient $k \approx -0.6$.
+The retrun time distribution is presented in @figure-??. We fit a power law model $r(\tau) \sim \tau ^k$ with coefficient $k \approx -0.6$. -->
 
@@
 
-```{python}
+<!-- ```{python}
 #| echo: false
@@
 
-```
+``` -->
 
@@
 This seems confusing, since the snapshot representation _should_ represent a superset of possible movements.
+
+## Approximations made with the Static Network
+
+A standard static network consists of a single layer of nodes and edges.
+In order to capture patients leaving and then subsequently returning to hospital, we model the network with two layers: a "standard" layer represeting 'direct' transfers, and a secondary layer representing indirect transfers.
+This secondary layer is bipartite in nature, where each (potential) edge in the primary layer is represented by replacing it with a "home" node, and adding two additional edges to it: one from the original source node to the home node, and another from the home node to the original target node.
+This is implicitly captured in the network data structure by setting two additional edge attributes that represent the amount of flow on each of the two new secondary layer edges respectively.
+
+We remove self-loops from the static network; we attempt one model without self-loops in only the direct layer (since this does not make physical sense), and another without self-loops in both the direct layer, and the indirect layer (we uspected that these self-loops slowed the movement of individuals down, but the resulting effect is the explosion of movement in the first simulation step)
+
+## Quantifying the effect of including indirect transfers
+
+A simplistic model of the effect of indirect transfer can be modelled with the following figures:
+
+:::{#fig-model-going-home layout-ncol="2"}
+
+```{dot}
+//| fig-width: 50%
+digraph G {
+    S₀ -> S₁ [ label="λ₀" ]
+}
+```
+
+```{dot}
+//| fig-width: 50%
+digraph G {
+    S₀ -> S₁ [ label="λ₀p" ]
+    S₀ -> S₂ [ label="λ₀(1-p)" ]
+    S₂ -> S₁ [ label="λ₂" ]
+}
+```
+
+Toy model of the effect of indirect transfer on hitting time
+:::
+
+In the model without indirect transfer, we can compute the expected hitting time $k$ from $S_0$ to $S_1$
+
+$$k = {\lambda_0}^{-1}$$
+
+In a model with indirect transfer, we introduce an additional state $S_2$ which represents the indirect pathway.
+If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k$ of $S_1$ given we start at $S_0$:
+
+$$k = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
+
+i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$.
+
+In the full network, we can roughly

### 2024-09-18 hitting_markov.py

@@ -2,8 +2,9 @@
 
 from typing import Iterable
 import numpy as np
+from numpy.typing import NDArray
 import igraph as ig
-from scipy import optimize
+from scipy import optimize, sparse
 
 
 def Q_mat_naive(G: ig.Graph):
@@ -37,8 +38,15 @@ def _Q_mat(Z: np.ndarray, p_null: np.ndarray):
 
     return Q
 
+def _Q_mat_sparse(A: sparse.sparray):
 
-def min_nonnegative_soln(A: np.ndarray, b: np.ndarray):
+    rowsums = A.sum(axis=1)
+    Q = A.copy()
+    Q.setdiag(-rowsums)
+
+    return Q
+
+def min_nonnegative_soln(A: np.ndarray, b: np.ndarray, full=False):
     """Generic wrapper for solving an ill-conditioned Ax=b linear equation for minimal non-neg x
     
     We wrap the use of scipy.optimize.linprog to solve the linear program:
@@ -69,14 +77,14 @@ def min_nonnegative_soln(A: np.ndarray, b: np.ndarray):
         bounds=bounds,
     )
 
-    if not soln.success:
+    if not soln.success or full:
         return soln
 
     return soln.x
 
-def build_hitting_time_problem(Q: np.ndarray, target_set: Iterable):
+def build_hitting_time_problem(Q: NDArray, target_set: Iterable):
 
-    A = -np.array(Q)
+    A = -Q.copy()
     b = np.ones(Q.shape[0])
     
     for i in target_set:
@@ -86,14 +94,15 @@ def build_hitting_time_problem(Q: np.ndarray, target_set: Iterable):
 
     return A, b
 
-def solve_hitting_time(Q: np.ndarray, target_set: Iterable):
+
+def solve_hitting_time(Q: NDArray, target_set: Iterable):
 
     A, b = build_hitting_time_problem(Q, target_set)
     return min_nonnegative_soln(A, b)
 
 def build_hitting_prob_problem(Q: np.ndarray, target_set: Iterable):
 
-    A = np.array(Q)
+    A = Q.copy()
     b = np.zeros(Q.shape[0])
 
     for i in target_set:

### 2024-09-18 hitting_time_analysis.py

@@ -1,9 +1,13 @@
 import numpy as np
 import igraph as ig
 from matplotlib import pyplot as plt, colors
+from scipy import sparse
 
 import network_conversion as ntc
 import hitting_markov as hit
+
+
+### STANDARD STATIC MODEL
 
 G = ig.Graph.Read("./concordant_networks/shuf_static_network.graphml")
 
@@ -33,6 +37,91 @@ hitting_time_arr = np.hstack([h.reshape((-1, 1)) for h in hitting_times])
 
 plt.matshow(hitting_time_arr, norm=colors.LogNorm())
 plt.colorbar(label='hitting time (days)')
-plt.xlabel('from')
-plt.ylabel('to')
-plt.savefig("zstatic_no_home_time.png")
+plt.xlabel('to')
+plt.ylabel('from')
+plt.savefig("zstatic_no_home_time.png", dpi=360)
+
+### STATIC MODEL WITH DELAYED RETURN
+
+Z = ig.Graph.Read("./concordant_networks/trimmed_static_base_1_threshold_4.graphml")
+
+zgraph_map = {int(k):v for k,v in zip(Z.vs['node'], Z.vs['id'])}
+zsorted_graph_map = {k: zgraph_map[k] for k in sorted(zgraph_map.keys())}
+zgraph_ordering = ntc.Ordering(zsorted_graph_map)
+zN = len(zgraph_ordering.sizes)
+
+T_direct = ntc.transition_matrix_from_graph(Z, 
+                                            ordering=zgraph_ordering, 
+                                            global_scaling=(10*365), 
+                                            ordering_key='node', 
+                                            adjacency_attribute='direct_weight'
+                                           )
+T_indirect = ntc.transition_matrix_from_graph(Z, 
+                                            ordering=zgraph_ordering, 
+                                            global_scaling=(10*365), 
+                                            ordering_key='node', 
+                                            adjacency_attribute='indirect_weight'
+                                           )
+T_return = ntc.transition_matrix_from_graph(Z, 
+                                            ordering=zgraph_ordering, 
+                                            ordering_key='node', 
+                                            adjacency_attribute='link_time'
+                                           )
+
[…]

### 2024-09-18 network_conversion.py

@@ -1,5 +1,6 @@
 import abc
 from os import PathLike
+import random
 import numpy as np
 import numba as nb
 from scipy import sparse
@@ -39,7 +40,7 @@ class Ordering:
         return value[0]
 
     @classmethod
-    def from_file(cls, file_name):
+    def from_file(cls, file_name: PathLike):
         size_mapping = esz.quick_read(file_name)
         return cls(size_mapping)
 
@@ -54,7 +55,7 @@ class Ordering:
 
 class ColumnDict(dict):
     @classmethod
-    def from_file(cls, file, key_name, value_name):
+    def from_file(cls, file: PathLike, key_name, value_name):
         df = pl.read_csv(file)
         return cls(
             zip(
@@ -201,13 +202,27 @@ class SnapshotWithHomeConverter(Converter):
         self,
         network_snapshots: Mapping[int, ig.Graph],
         ordering: Ordering,
+        shuffle_snapshots=False,
+        infer_durations=False,
     ):
         self.ordering = ordering
 
         self.snapshot_times = sorted(network_snapshots.keys())
-        self.snapshot_durations = np.diff(self.snapshot_times)
+        if infer_durations:
+            base_snapshot_durations = np.diff(self.snapshot_times)
+            base_duration_map = {k:d for k,d in zip(self.snapshot_times, base_snapshot_durations)}
 
-        # construct raw transition matrices that are to be parsed into weighting matrices
+        if shuffle_snapshots:
+            random.shuffle(self.snapshot_times)
+
+        # infer the duration scaling for each snapshot
+        if infer_durations:
+            self.snapshot_durations = [base_duration_map[k] for k in self.snapshot_times]
+        else:
+            self.snapshot_durations = [network_snapshots[k]['duration'] for k in self.snapshot_times]
+        
+
+        # construct raw transition matrices that are to be parsed into weighti ng matrices
         self.raw_transition_matrices = {
             adjtype: [
                 self.parse_transition_matrix_from_graph(
@@ -300,6 +315,10 @@ class SnapshotWithHomeConverter(Converter):
         for direct, departs in zip(
             self.raw_transition_matrices["direct"], self.raw_transition_matrices["out"]
         ):
+            # purge direct self-loops as they are non-physical
+            diagidx = np.diag_indices_from(direct)
+            direct[diagidx] = 0.0
+
             leaves = (direct + departs).sum(axis=1).reshape((-1, 1))
[…]

### 2024-09-18 new_network_sim.py

@@ -71,7 +71,7 @@ if __name__ == "__main__":
     )
 
     params = init_zero_params(
-        conv, "./probability_of_final_stay_by_shuffled_campus.csv"
+        conv, "./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv"
     )
 
     sim = create_sim(
@@ -91,5 +91,5 @@ if __name__ == "__main__":
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_ss_exact_home_sims/sim_all.h5"
+            outname = f"zero_sims/snapshot/sim_all_nodirectselfloop.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)

### 2024-09-18 numba_sample.py

@@ -73,7 +73,7 @@ def _mn_sparse_full(
     return result
 
 
-def multinomial_sparse_full(trials, prob_matrix: sparse.csr_matrix):
+def multinomial_sparse_full(trials, prob_matrix: sparse.csr_array):
     """Performs numba-acceled multinomial smapling for a large 2D probability matrix
 
     Arguments

### 2024-09-18 projection_metrics.py

@@ -1,3 +1,4 @@
+import numpy as np
 import graph_importer as gim
 import polars as pl
 import igraph as ig
@@ -118,11 +119,15 @@ _snapshot_mode = {
 
 def make_snapshots(graph: ig.Graph, output_dir, mode="out", aggmode="sum"):
     times = sorted(set(graph.vs["time"]))
+    durations = np.diff(times)
+    if len(durations) < 1:
+        raise ValueError("Not enough times to construct sane snapshots")
 
     snapper = _snapshot_mode[mode]
 
-    for t in times:
+    for t, dt in zip(times, durations):
         snapshot = snapper(graph, t, aggmode=aggmode)
+        snapshot['duration'] = dt
         snapshot.write_graphml(f"{output_dir}/{t:04}.graphml")
 
 

### 2024-09-26 reports/report.qmd
report minor edits (+refs)

@@
     pdf: default
+bibliography: refs.bib
+csl: plos.csl
 ---
@@
 # Introduction
+
+Antimicrobial resistance poses a great threat to human health and development. [@murray_global_2022]
+Globally, it poses a large burden, operationally and economically, on hospital systems.
 
@@
 
+<!-- I don't know why, but the first invocation of the dot engine craps out -->
+```{dot}
+```
+
 :::{#fig-model-going-home layout-ncol="2"}
@@
 
-In the full network, we can roughly
+For a larger network, the number of possible paths causes complications for explicit computation of the effect, but we can always expect it to follow a similar principle, where splitting transfers into a direct path and a slower indirect path will add to the expected hitting time.

### 2024-09-26 hitting_time_analysis.py
relative change in expected hitting time

@@ -7,7 +7,7 @@ import network_conversion as ntc
 import hitting_markov as hit
 
 
-### STANDARD STATIC MODEL
+## STANDARD STATIC MODEL
 
 G = ig.Graph.Read("./concordant_networks/shuf_static_network.graphml")
 
@@ -117,7 +117,7 @@ plt.matshow(zhitting_time_arr, norm=colors.LogNorm())
 plt.colorbar(label='hitting time (days)')
 plt.xlabel('to')
 plt.ylabel('from')
-plt.savefig("z_z_static_no_home_time.png", dpi=360)
+plt.savefig("z_z_s.png", dpi=360)
 
 zresidence_times = -1/R.diagonal()
 plt.figure()
@@ -125,3 +125,48 @@ plt.hist(np.log10(zresidence_times), bins=31)
 plt.yscale('log')
 plt.xlabel('$log_{10}$ residence time (log-days)')
 plt.savefig('zstatic_no_home_residence.png', dpi=360)
+
+# we know relative change is usually quite large
+relative_change = zhitting_time_arr / (hitting_time_arr+1) 
+plt.matshow(relative_change)
+plt.colorbar(label='relative change in hitting time')
+plt.xlabel('to')
+plt.ylabel('from')
+plt.savefig("zrelative_change_hitting_time_home.png", dpi=360)
+
+plt.figure()
+plt.hist(relative_change.flatten(), bins=31)
+plt.xlabel("relative change")
+plt.ylabel("frequency")
+plt.yscale('log')
+
+
+
+direct_movement_total_out = T_direct.sum(axis=1)
+direct_movement_total_in = T_direct.sum(axis=0)
+
+indirect_movement_total_out = T_indirect.sum(axis=1)
+indirect_movement_total_in = T_indirect.sum(axis=0)
+
+
+large_indices = np.unravel_index(np.argsort(relative_change, axis=None)[::-1][:5], relative_change.shape)
+large_index_list = np.array(large_indices).T
+
+
+plt.figure()
+plt.scatter(direct_movement_total_out, direct_movement_total_in)
+plt.xscale('log')
+plt.yscale('log')
+
+
+for indices, colour in zip(large_index_list, (f"C{i+1}" for i in range(5))):
+    xs = direct_movement_total_out[list(indices)]
+    ys = direct_movement_total_in[list(indices)]
+    plt.annotate(
+        "",
+        xy = (xs[1], ys[1]),
[…]

### 2024-10-01 netsim_summariser.py
add extent metric

@@ -5,42 +5,66 @@ import multiprocessing
 from itertools import repeat
 
 class Summariser():
+    """Utility class to sumamrise metrics from an h5 database of simulation outputs"""
 
     def __init__(self, file):
         self.target_file = file
 
     @staticmethod
     def hitting_time(target : h5py.Group, null=np.nan, **kwargs):
+        """Yields list of hitting times to all locations"""
         mask = (target['history'][:] != 0)
         return list(np.where(mask.any(axis=1), mask.argmax(axis=1), null))
 
     @staticmethod
+    def extent(target: h5py.Group, time_limit=30, **kwargs):
+        """Yields the number of locations hit by the time limit"""
+        mask = (target['history'][:] != 0)
+        return (
+            sum(np.where(mask.any(axis=1), mask.argmax(axis=1), np.nan) < time_limit, -1),
+            time_limit,
+        )
+
+    @staticmethod
     def movement_out_total(target: h5py.Group, **kwargs):
+        """Yields the list of number of movements out of all locations"""
         return list(target['mover_out'][:].sum(axis=1))
-    
+
     @staticmethod
     def movement_in_total(target: h5py.Group, **kwargs):
+        """Yields the list of number of movements into all locations"""
         return list(target['mover_in'][:].sum(axis=1))
 
     @staticmethod
     def steady_state_infected(target: h5py.Group, **kwargs):
+        """Yields the mean and standard deviation of the number of infected individuals (over locs)"""
         init = target['history'][:,0].sum()
         normed = target['history'][:].sum(axis=0) / init
         return [np.mean(normed), np.std(normed)]
-    
+
     @staticmethod
     def emptying_time(target: h5py.Group, null=np.nan, **kwargs):
+        """Yields the first time that no location has any infected individuals"""
         xs = target['history'][:].sum(axis=0)
         mask = (xs == 0)
         return mask.argmax() if mask.any() else null
 
     @property
     def groups(self):
+        """The list of simulation keys (h5 groups)"""
         with h5py.File(self.target_file, 'r') as fp:
             grps = list(fp)
         return grps
 
-    def compute_metrics(self, target_str, no_move=False, verbose=False):
+    def compute_metrics(self, target_str, no_move=False, verbose=False, **kwargs):
+        """Computes (all) metrics on a target simulation given by the key target_str
+        
+        Args:
+            no_move: if True, does not return the movement metrics
+            verbose: if True, also prints the target string to stdout
+            **kwargs: keywords to pass into the metrics
[…]

### 2024-10-03 netsim_summariser.py
Add extent metric computations

@@ -3,6 +3,7 @@ import polars as pl
 import numpy as np
 import multiprocessing
 from itertools import repeat
+from functools import wraps
 
 class Summariser():
     """Utility class to sumamrise metrics from an h5 database of simulation outputs"""
@@ -11,26 +12,39 @@ class Summariser():
         self.target_file = file
 
     @staticmethod
+    def _sanitise(fn):
+        @wraps(fn)
+        def sanitised_fn(*args, **kwargs):
+            return list(fn(*args, **kwargs))
+        return sanitised_fn
+
+    @staticmethod
+    @_sanitise
     def hitting_time(target : h5py.Group, null=np.nan, **kwargs):
         """Yields list of hitting times to all locations"""
         mask = (target['history'][:] != 0)
-        return list(np.where(mask.any(axis=1), mask.argmax(axis=1), null))
+        indices = np.where(mask.any(axis=1), mask.argmax(axis=1), null)
+        times = np.nan * np.empty_like(indices)
+        times[np.isfinite(indices)] = target['ts'][:][indices[np.isfinite(indices)].astype(int)]
+        return times
+
+    # @staticmethod
+    # def extent(target: h5py.Group, time_limit=30, **kwargs):
+    #     """Yields the number of locations hit by the time limit"""
+    #     hitting_times = np.array(Summariser.hitting_time)
+    #     return (
+    #         sum(np.where(mask.any(axis=1), mask.argmax(axis=1), np.nan) < time_limit, -1),
+    #         time_limit,
+    #     )
 
     @staticmethod
-    def extent(target: h5py.Group, time_limit=30, **kwargs):
-        """Yields the number of locations hit by the time limit"""
-        mask = (target['history'][:] != 0)
-        return (
-            sum(np.where(mask.any(axis=1), mask.argmax(axis=1), np.nan) < time_limit, -1),
-            time_limit,
-        )
-
-    @staticmethod
+    @_sanitise
     def movement_out_total(target: h5py.Group, **kwargs):
         """Yields the list of number of movements out of all locations"""
         return list(target['mover_out'][:].sum(axis=1))
 
     @staticmethod
+    @_sanitise
     def movement_in_total(target: h5py.Group, **kwargs):
         """Yields the list of number of movements into all locations"""
         return list(target['mover_in'][:].sum(axis=1))
@@ -78,7 +92,7 @@ class Summariser():
                 self.hitting_time,
                 self.steady_state_infected,
                 self.emptying_time,
-                self.extent,
+                # self.extent,
[…]

### 2024-10-03 reports/report.qmd
Report updates (mostly commenting out)

@@
 
-- There is a significant amount of patient movement in the Victorian hospital system where the patient returns home for a substative amount of time between discharge from one hospital and subsequent admission at another hospital.
+- There is a significant amount of patient movement in the Victorian hospital system where the patient returns home for a substantive amount of time between discharge from one hospital and subsequent admission at another hospital.
 - This "indirect" movement has implications on the rate of spread of a hypothetical contagion through the hospital system. It causes dispersal of patients to be slower than expected.
@@
 Globally, it poses a large burden, operationally and economically, on hospital systems.
+In Australia, we see [significant] economic burden from AMR.
+this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR.
+One X of interest is carbapenemase-producing enterobacteriacea (CPE).
+Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level.
+This already causes some level of burden, and introduced strains would further multiply this.
 
-Antimicrobial resistance is bad.
-In Victoria, Australia, we have observed CPE, and want to understand how the underlying movement of patients could affect the spread of CPE and other related AMR in the future.
-The focus of this paper is on the patient movement network: ways of representing, simplifying, and approximating those patient movements; and how those representations of patient movement affect the spread of a hypothetical infection between hospitals in the Victorian healthcare system.
+Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon.
+The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes.
+However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network.
+Typically, this is due to the lack of granularity in the base dataset.
+It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents.
+In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison.
 
-# Network modelling of hospital patient movement
+# Methods
 
-## Generating the temporal network 
+## Setting and data sources
+
+<- yank from [@lydeamore_carbapenemase-producing_2024]
+
+Admissions data is from the Victorian Admitted Episdoes Dataset (VAED).
+
+Some exploratory analysis was done in by [@lydeamore_carbapenemase-producing_2024]
+
+## Network Modelling
+
+A network or graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ is made up of nodes or vertices $\mathcal{V} = \{v_1, v_2, \dots\}$ and edges $\mathcal{E} = \{e_{ij} \coloneqq (v_i, v_j), \dots\}$ that join those nodes.
+Each edge $e_{ij}$ can be assigned attributes, including an edge weight $w_{ij}$.
+For our application, this edge weight will typically represent the number of observed movements in a given time period.
+Thus the network can encode a representation of the movement between hospitals over a given time period.
+
+We can use the network to construct a Markov chain, where the expected rate of movement is governed by the observed movement in the data.
+For some edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, the rate of movement along that edge $\lambda_{ij} = w_{ij} / \tau$, such that the time between movement events along that edge are exponentially distributed $\text{Exp}(\lambda_{ij})$.
+
+[Stuff about previous implementations of this idea]
+
+## Explicitly Temporal Models
+Previous analyses of movement and spread of contagions through these networks have made the assumption that the weight $w_{ij}$ for a given edge does not change over time, i.e. the network is static.
+One advantage that the VAED provide us is the minute-scale granularity of patient admissions, which allows us to construct temporal networks and models of patient movement.
+
+<- Description of temporal networks and models>
+
+<- Description of temporally reductive networks and models>
+
+<!-- 
 We have line-listed patient movement data between {start_date} and {end_date}. This allows us to generate exact patient movement networks up to minute precision.
@@
 To reduce the computational expense of such a network, and to protect patient anonymity, we aggregate these patient transfer events to a discrete set of times.
-By choosing some window size $W$, we treat events that occur in the time interval $[t_i, t_i+W)$ as all occuring at $t_i$.
-This aggregation classfies patient transfers into two classes. 
+By choosing some window size $W$, we treat events that occur in the time interval $[t_i, t_i+W)$ as all occurring at $t_i$.
+This aggregation classifies patient transfers into two classes. 
 Transfers can be "direct", where the discharge and subsequent admission both occur within the same time window $[t_i, t_i+W)$.
@@
 These types of transfers imply causal edges between hospitals that are neither synchronous (occur at the same time), nor persistent (where the effect of a cause at time $t_1$ is not felt over some interval $(t_1, t_2)$), which make it challenging to fit a temporal network generated by this dataset into classical temporal network frameworks of representation.
+-->
 
-## Simulating on Networks
-
+<!-- 
 We use a discrete time Markov chain model to simulate the spread of infection.
@@
 | $\gamma$  | Rate of discharge from hospital |
-| $M$       | Transition matrix inferred from network |
+| $M$       | Transition matrix inferred from network | -->
 
@@
 
-We run a range of experiments.
+We simulate the movement of individuals
 
-## Leading results
-
-## No infection simulations
-
-To isolate the effect of the projections on the movement of patients, we can simulate the model with parmeters $\beta = 0, \eta=0$.
+<!-- 
+To isolate the effect of the projections on the movement of patients, we can simulate the model with parmeters $\beta = 0, \eta=0$. -->
 

### 2024-10-30 reports/report.qmd
Add report content and diagram filter

@@
 csl: plos.csl
+jupyter: python3
+filters:
+    - diagram
+execute:
+    echo: false
+    warning: false
+    error: false
 ---
+
+```{python}
+import numpy as np
+import polars as pl
+from matplotlib import pyplot as plt
+import seaborn as sns
+import itertools
+import re
+
+from pathlib import Path
+
+# setup printing formats for tables
+pl.Config.set_tbl_hide_column_data_types(True)
+pl.Config.set_tbl_hide_dataframe_shape(True)
+
+sns.set_theme(style='ticks')
+sns.set_context('paper')
+# plt.style.use('ggplot')
+```
 
@@
 However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network.
-Typically, this is due to the lack of granularity in the base dataset.
 It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents.
 In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison.
+In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granularity of the temporality]]]]].
 
@@
 
-## Network Modelling
+## Network Modelling {#sec-network-modelling}
 
@@
 <- Description of temporally reductive networks and models>
+
+```{python}
+#| label: fig-readmission-window
+#| fig-cap: Distribution of readmission time (time between discharge and successive admission of a single patient)
+
+# sourced from readmission_time.py
+readmit_window = pl.read_parquet("../concordant_networks/readmission_time.parquet")
+
+fig, ax = plt.subplots()
+ax.scatter(*(readmit_window.to_numpy().T), marker='.')
+ax.set_xscale('asinh') # todo fix xticks
+ax.set_yscale('log')
+ax.set_xlabel('Readmission Window (days)')
+ax.set_ylabel('Count')
+ax.set_aspect(0.5)
+ax.set_xticks([0, 1, 1e1, 1e2])
+ax.set_xticks([2, 5, 2e1, 5e1, 2e2, 5e2], minor=True)
+sns.despine()
+```
+
+{skip}
+
+### Networks and Projections
+
+Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_m$ to a target hospital $x_j$ at some time $t_n$.
+We could work with the exact times, but this will lead to an overly sparse representation that is difficult to work with.
+Instead, we discretise time into bins by treating movements involving some arbitrary hospital $x_i$ between times $t_n$ and $t_n + W$ as instead involving a single node at $v_{i,n} \coloneqq (x_i, t_n)$.
+Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$.
+
+The network analysis of the temporal network as described above is computationally expensive since it is relatively large ($N_x = 388$ hospitals and $N_t = 259$ fortnightly bins, for $\approx 10000$ nodes).
+We can instead analyse slices of the temporal network, where a slice at $t_n$ consists of nodes at that time discretisation bin, and the related nodes.
+
+### Dynamics
+
+<!-- TODO need to work on indexing consistency -->
+
+For the temporal network, we can model the number of patients discharged from some given location ($x_i$) at a given time ($t_n$) as a Poisson random variable with rate $\gamma(x_i, t_n)$.
+Some proportion $q(x_i, \emptyset)$ of these individuals will not re-admit at any other hospital, and can be removed from the system; the remaining discharged patients will have some probability $p(x_i, t_n, x_j, t_m)$ to re-admit at location $x_j$ at time $t_m$.
+These probabilites $p$ wil depend on the edges...
+Patients that have $t_i \neq t_j$ will be moved into an auxiliary state $z_{i,j}$...
+Movemnts $z_{ij}$ are distriuted as Uniform(0, W)
 
@@
 
-We simulate the movement of individuals
-
 <!-- 
 To isolate the effect of the projections on the movement of patients, we can simulate the model with parmeters $\beta = 0, \eta=0$. -->
+
+```{python}
+# simulation archives from run_zero*.py
+simulation_archives = {
+    'temporal': '../zero_sims/temporal/sim_all_30s.h5',
+    'static': '../zero_sims/static/sim_all_30s.h5',
+    'snapshot': '../zero_sims/snapshot/sims_with_30seeds_pseudocap.h5',
+    'naive_static': '../zero_sims/naive_static/sim_all_30s.h5',
+}
+
+processed_results = {}
+
+# metrics precomputed in compute_summary_stats.ipynb
+for label, archive in simulation_archives.items():
+        processed_results[label] = pl.read_parquet(Path(archive).with_name('metrics_30s.parquet'))
+```
+
+We define the _extent_ of a simulation by the number of facilities that some patient has moved to or through within a given number of days.
+We see in @fig-extent-30 that we get starkly different distriubtions of extent for the naïve static model, commpared to the other models.
+
+```{python}
+#| label: fig-extent-30
+#| fig-cap: "Distribution of 30-day extent in different simulation models on different network types"
+
+combined_all_df = pl.concat(
+    [(df.select('seed', 'extent')
+        .with_columns(
+            pl.lit(label).alias('simulation'), 
+            pl.col('extent').cast(pl.Int64)
+        )
+     ) for label, df in processed_results.items()
+    ]
+)
+
+sim_type_order = ['temporal', 'snapshot', 'static', 'naive_static']
+extent_bins = np.arange(0, combined_all_df.select('extent').max().item()+1)
+
+sns.histplot(
+    combined_all_df, 
+    x = 'extent', 
+    hue = 'simulation', 
+    multiple = 'dodge', 
+    bins = extent_bins, 
+    shrink = 0.8, 
+    hue_order = sim_type_order,
+)
+sns.despine()
+```
+
+```{python}
+# comparisons from compute_summary_stats.ipynb
+comparative_results = dict()
+name_pattern = re.compile(r"(\w+)_x_([\w_]+)\.30s")
+for comp_file in Path("../zero_sims/comparisons/").glob("*.30s.parquet"):
+    pair = name_pattern.match(comp_file.stem).groups()
+    if pair is None:
+        raise RuntimeError(f"Unable to parse file name: {comp_file.name}")
+    paired_comparison = pl.read_parquet(comp_file)
+    comparative_results[pair] = paired_comparison
+    comparative_results[tuple(reversed(pair))] = paired_comparison.with_columns(-pl.col('statistic'))
+```
+
+```{python}
+#| label: tbl-example
+#| tbl-cap: An example table
+
+comparative_results['static', 'snapshot'].head(5)
+```
 

### 2024-10-30 graph_importer.py
Update computational backend code

@@ -1,7 +1,8 @@
 import igraph as ig
+from os import PathLike
 
 
-def make_graph(filename: str) -> ig.Graph:
+def make_graph(filename: str | PathLike) -> ig.Graph:
     G = ig.Graph.Read_Ncol(filename)
     G.vs["loc"] = [int(x.split(",")[0].strip("()")) for x in G.vs["name"]]
     G.vs["time"] = [int(x.split(",")[1].strip("()")) for x in G.vs["name"]]

### 2024-10-30 network_conversion.py

@@ -2,11 +2,12 @@ import abc
 from os import PathLike
 import random
 import numpy as np
+from numpy import typing as npt
 import numba as nb
 from scipy import sparse
 import polars as pl
 import igraph as ig
-from typing import Mapping, Hashable, Sequence, Any, SupportsFloat as Numeric, Iterable
+from typing import Mapping, Hashable, Sequence, Any, SupportsFloat
 from functools import lru_cache
 from pathlib import Path
 
@@ -80,8 +81,9 @@ class ColumnDict(dict):
     def from_prob_final_file(cls, file):
         return cls.from_file(file, "loc", "final_stay")
 
-    def organise_by(self, ordering: Ordering):
-        return ordering.conform(self)
+    def organise_by(self, ordering: Ordering) -> npt.NDArray:
+        output = ordering.conform(self)
+        return np.array(output)
 
 
 def transition_matrix_from_graph(
@@ -178,7 +180,7 @@ class TemporalNetworkConverter(Converter):
         return {"NLOC": self.NLOC, "NT": self.NT}
     
     @classmethod
-    def from_file(cls, network_filepath: str|PathLike, **kwargs):
+    def from_file(cls, network_filepath: str | PathLike, **kwargs):
         graph = gim.make_graph(network_filepath)
         return cls(network=graph, **kwargs)
 
@@ -203,6 +205,14 @@ class TemporalNetworkConverter(Converter):
         ).tocsr()
 
         return mapped_parameters
+    
+    def delay(self, n: int):
+        """Reformats internal attributes so that the starting time is delayed by n time steps"""
+
+        self.NT -= n
+
+        nidx = int(n * self.NLOC)
+        self.A = self.A[nidx:, nidx:] # CSR sparse -> CSR sparse
 
 
 class SnapshotWithHomeConverter(Converter):
@@ -370,7 +380,7 @@ class StaticConverter(Converter):
     }
 
     def __init__(
-        self, network: ig.Graph, ordering: Ordering, time_span: Numeric | None = None, purge_selfloops=False
+        self, network: ig.Graph, ordering: Ordering, time_span: SupportsFloat | None = None, purge_selfloops=False
     ):
         self.ordering = ordering
 
@@ -464,7 +474,7 @@ class StaticConverter(Converter):
 class NaiveStaticConverter(Converter):
 
     def __init__(
-        self, network: ig.Graph, ordering: Ordering, time_span: Numeric | None=None
[…]

### 2024-10-30 network_simulation.py

@@ -369,7 +369,3 @@ class TemporalSim(SimulationWithMovers):
         super().reset(soft=soft)
 
         self.time_travellers = np.zeros_like(self.time_travellers, dtype=np.int64)
-
-    def delay(self, n: int):
-
-        raise NotImplementedError("old code uses a different state setup, which makes this tricky to port")

### 2024-10-30 readmission_time.py

@@ -1,0 +1,22 @@
+import polars as pl
+import graph_importer as gim # imports igraph
+
+def construct_readmission_counts(file="./concordant_networks/temponet_1_365.lgl") -> pl.DataFrame:
+
+    G = gim.make_graph(file)
+
+    counts = [
+        {
+            'readmission_time': edge.target_vertex['time'] - edge.source_vertex['time'],
+            'count': edge['weight'],
+        }
+        for edge in G.es
+    ]
+
+    return pl.from_dicts(counts).with_columns(pl.col('count').cast(pl.Int64))
+
+
+if __name__ == "__main__":
+    
+    df = construct_readmission_counts(file="./concordant_networks/temponet_1_365.lgl")
+    df.write_parquet("./concordant_networks/readmission_time.parquet")

### 2024-10-30 reports/report.qmd
more dumping of images

@@
 
+.....
+
+For the naïve and improved static network models, we can compute the hitting times directly and compare them.
+We can model the dyanmics on these networks as a (continuous time) Markov Chain and extract the mean hitting time to any target hospital from all other hospitals by solving a linear system.
+For the naïve static network, this is relatively straightforward, as we can simply model the rate of movement from hospital $i$ to hospital $j$ as the mean number of observed movements over the observation period (XXX days).
+For the improved static network, we expand the state by including an additional state for each pair of hospitals, so that the residence time at home (between hospital admissions) for indirect transfers can be modelled.
+Then as with the naïve model, the movement from $i$ to $j$ directly is the mean observed number of transfers over the observation epriod.
+The indirect movements from $i$ to the intermediate $i,j$ state are derived from the observed number of movements, and the rate from $i,j$ to $j$ is derived from the mean readmission period (as seed in @fig-readmission-window).
+
+Comparing the relative difference in the mean hitting times gives us @fig-hitting-time-difference
+
+```{python}
+#| label: fig-hitting-time-difference
+#| fig-cap: Relative difference between the expected hitting times for the naïve static and improved static models
+
+# temp hack
+# todo port across computation code
+from PIL import Image
+image = Image.open("./precompiled/zrelative_change_hitting_time_home.png")
+image
+```
+
 <!-- 
@@
     hue_order = sim_type_order,
+    discrete = True,
 )
@@
 
+We can frame the empirical cumulative distribution (ECDF) of hitting times from a given seed hospital to a target hospital in a survival analysis view, and compare different experimental setups by comparing the ECDFs of the combinations of hitting times with log-rank tests.
+Each pair of hospitals yields a separate log-rank test statistic.
+Using a crude Bonferroni correction to counteract the multiple comparisons (which occur due to the $N=338^2$ combinations of seed and target hospitals), we find that there are significant differences between each type of model.
+We note spikes at very low p-values when comparing the naïve static model to other models.
+
 ```{python}
-#| label: tbl-example
-#| tbl-cap: An example table
+#| label: fig-logrank-comparison
+#| fig-cap: Log-rank test comparisons between experiments
 
-comparative_results['static', 'snapshot'].head(5)
+comparable_pairs = map(lambda x: sorted(x, key=sim_type_order.index), 
+                       itertools.combinations(sim_type_order, 2))
+
+NN = len(sim_type_order) - 1
+
+fig, axs = plt.subplots(nrows=NN, ncols=NN, squeeze=True, sharex=True, sharey=True)
+
+confidence_threshold = 0.01 / (338**2)
+
+for pair in comparable_pairs:
+    index = (sim_type_order.index(pair[1]) - 1, sim_type_order.index(pair[0]))
+    comparison = comparative_results[tuple(pair)]
+
+    sns.histplot(comparison, x='pvalue', log_scale=True, ax=axs[index], kde=True)
+
+    axs[index].axvline(confidence_threshold, color='r')
+    axs[index].set_yscale('log')
+    sns.despine(ax=axs[index])
+
+for label, ax in zip(sim_type_order[1:], axs[:,0]):
+    ax.set_ylabel(label)
+
+for label, ax in zip(sim_type_order, axs[-1,:]):
+    ax.set_xlabel(label)
+
+fig.supxlabel('pvalue')
+fig.supylabel('Count')
+
+for index in zip(*np.triu_indices_from(axs, k=1)):
+    fig.delaxes(axs[index])
 ```

### 2024-10-30 reports/report.qmd
Move reduced model from apdx

@@
 
-.....
-
-For the naïve and improved static network models, we can compute the hitting times directly and compare them.
-We can model the dyanmics on these networks as a (continuous time) Markov Chain and extract the mean hitting time to any target hospital from all other hospitals by solving a linear system.
-For the naïve static network, this is relatively straightforward, as we can simply model the rate of movement from hospital $i$ to hospital $j$ as the mean number of observed movements over the observation period (XXX days).
-For the improved static network, we expand the state by including an additional state for each pair of hospitals, so that the residence time at home (between hospital admissions) for indirect transfers can be modelled.
-Then as with the naïve model, the movement from $i$ to $j$ directly is the mean observed number of transfers over the observation epriod.
-The indirect movements from $i$ to the intermediate $i,j$ state are derived from the observed number of movements, and the rate from $i,j$ to $j$ is derived from the mean readmission period (as seed in @fig-readmission-window).
-
-Comparing the relative difference in the mean hitting times gives us @fig-hitting-time-difference
-
-```{python}
-#| label: fig-hitting-time-difference
-#| fig-cap: Relative difference between the expected hitting times for the naïve static and improved static models
-
-# temp hack
-# todo port across computation code
-from PIL import Image
-image = Image.open("./precompiled/zrelative_change_hitting_time_home.png")
-image
-```
-
 <!-- 
@@
 # Results
+
+## A Reduced Model
+
+For the naïve and improved static network models, we can compute the hitting times directly and compare them.
+We can model the dyanmics on these networks as a (continuous time) Markov Chain and extract the mean hitting time to any target hospital from all other hospitals.
+
+A simplistic model of the effect of indirect transfer can be modelled with the following figures:
+
+:::{#fig-model-going-home layout-ncol="2"}
+
+```{dot}
+//| fig-width: 50%
+digraph G {
+    S₀ -> S₁ [ label="λ₀" ]
+}
+```
+
+```{dot}
+//| fig-width: 50%
+digraph G {
+    S₀ -> S₁ [ label="λ₀p" ]
+    S₀ -> S₂ [ label="λ₀(1-p)" ]
+    S₂ -> S₁ [ label="λ₂" ]
+}
+```
+
+Toy model of the effect of indirect transfer on hitting time
+:::
+
+In the model without indirect transfer, we can compute the expected hitting time $k$ from $S_0$ to $S_1$
+
+$$k = {\lambda_0}^{-1}$$
+
+In a model with indirect transfer, we introduce an additional state $S_2$ which represents the indirect pathway.
+If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k$ of $S_1$ given we start at $S_0$:
+
+$$k = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
+
+i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$.
+
+For a larger network, the number of possible paths causes complications for explicit computation of the effect, but we can always expect it to follow a similar principle, where splitting transfers into a direct path and a slower indirect path will add to the expected hitting time.
+These can be obtained by solving a linear system @Norris_Markov_Chains_1997.
+For the naïve static network, we can model the rate of movement from hospital $i$ to hospital $j$ as the mean number of observed movements over the observation period (XXX days).
+For the improved static network, we expand the state by including an additional state for each pair of hospitals, so that the residence time at home (between hospital admissions) for indirect transfers can be modelled.
+Then as with the naïve model, the movement from $i$ to $j$ directly is the mean observed number of transfers over the observation epriod.
+The indirect movements from $i$ to the intermediate $i,j$ state are derived from the observed number of movements, and the rate from $i,j$ to $j$ is derived from the mean readmission period (as seed in @fig-readmission-window).
+
+Comparing the relative difference in the mean hitting times gives us @fig-hitting-time-difference
+
+```{python}
+#| label: fig-hitting-time-difference
+#| fig-cap: Relative difference between the expected hitting times for the naïve static and improved static models
+
+# temp hack
+# todo port across computation code
+from PIL import Image
+image = Image.open("./precompiled/zrelative_change_hitting_time_home.png")
+image
+```
+
+## Movement Simulation Studies
 
@@
 We remove self-loops from the static network; we attempt one model without self-loops in only the direct layer (since this does not make physical sense), and another without self-loops in both the direct layer, and the indirect layer (we uspected that these self-loops slowed the movement of individuals down, but the resulting effect is the explosion of movement in the first simulation step)
-
-## Quantifying the effect of including indirect transfers
-
-A simplistic model of the effect of indirect transfer can be modelled with the following figures:
-
-<!-- I don't know why, but the first invocation of the dot engine craps out -->
-```{dot}
-```
-
-:::{#fig-model-going-home layout-ncol="2"}
-
-```{dot}
-//| fig-width: 50%
-digraph G {
-    S₀ -> S₁ [ label="λ₀" ]
-}
-```
-
-```{dot}
-//| fig-width: 50%
-digraph G {
-    S₀ -> S₁ [ label="λ₀p" ]
-    S₀ -> S₂ [ label="λ₀(1-p)" ]
-    S₂ -> S₁ [ label="λ₂" ]
-}
-```
-
-Toy model of the effect of indirect transfer on hitting time
-:::
-
-In the model without indirect transfer, we can compute the expected hitting time $k$ from $S_0$ to $S_1$
-
-$$k = {\lambda_0}^{-1}$$
-
-In a model with indirect transfer, we introduce an additional state $S_2$ which represents the indirect pathway.
-If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k$ of $S_1$ given we start at $S_0$:
-
-$$k = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
-
-i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$.
-
-For a larger network, the number of possible paths causes complications for explicit computation of the effect, but we can always expect it to follow a similar principle, where splitting transfers into a direct path and a slower indirect path will add to the expected hitting time.

### 2024-10-30 reports/report.qmd
Add in exponential approximation curves to readmit fig

@@
 
+# naive static distribution
+mean_window = (
+    readmit_window
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+num_readmits = readmit_window.select(pl.col('count').sum()).item()
+
+xs = np.arange(0, 366, dtype=np.float64)
+naive_static_dist = num_readmits / mean_window * np.exp(-xs / mean_window) # scaling to N
+
+# piecewise distribution (improved static)
+threshold = 4
+shorts = readmit_window.filter(pl.col('readmission_time') <= threshold)
+longs = readmit_window.filter(pl.col('readmission_time') > threshold)
+
+short_window = (
+    shorts
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+n_shorts = shorts.select(pl.col('count').sum()).item()
+
+long_window = (
+    longs
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+n_longs = longs.select(pl.col('count').sum()).item()
+
+sxs = np.arange(0, threshold+1, dtype=np.float64)
+short_dist = n_shorts / short_window * np.exp(-sxs / short_window)
+
+lxs = np.arange(threshold, 366, dtype=np.float64)
+long_dist = n_longs / long_window * np.exp(-lxs / long_window)
+
 fig, ax = plt.subplots()
-ax.scatter(*(readmit_window.to_numpy().T), marker='.')
-ax.set_xscale('asinh') # todo fix xticks
+ax.scatter(*(readmit_window.to_numpy().T), marker='.', color='k')
+h_0 = ax.plot(xs, naive_static_dist, color='C0', label='Naïve Static')
+h_1 = ax.plot(sxs, short_dist, color='C1', label='Improved Static')
+ax.plot(lxs, long_dist, color='C1')
+ax.axvline(threshold, color='C1', linestyle='dashed')
+ax.set_xscale('asinh')
 ax.set_yscale('log')
@@
 ax.set_xticks([2, 5, 2e1, 5e1, 2e2, 5e2], minor=True)
+ax.set_xlim(-1e-1, 1e3) # !! this is manual
+ax.set_ylim(5e3, 1e8) # !! this is manual
+ax.legend(handles=[*h_0, *h_1])
 sns.despine()
@@
 #| fig-cap: Log-rank test comparisons between experiments
+#| cache: true
 

### 2024-10-30 reports/report.qmd
support coloneqq

@@
     html: default
-    pdf: default
+    pdf:
+        include-in-header:
+            - text: |
+                \usepackage{mathtools}
+        papersize: a4
 bibliography: refs.bib
@@
 
-A network or graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ is made up of nodes or vertices $\mathcal{V} = \{v_1, v_2, \dots\}$ and edges $\mathcal{E} = \{e_{ij} \coloneqq (v_i, v_j), \dots\}$ that join those nodes.
+A network or graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ is made up of nodes or vertices $\mathcal{V} = \{v_1, v_2, \dots\}$ and edges $\mathcal{E} = \{e_{ij} \coloneqq (v_i, v_j)\}$ that join those nodes.
 Each edge $e_{ij}$ can be assigned attributes, including an edge weight $w_{ij}$.
@@
     bins = extent_bins, 
-    shrink = 0.8, 
     hue_order = sim_type_order,

### 2024-11-04 reports/report.qmd
yep report updates

@@
 jupyter: python3
-filters:
-    - diagram
+# filters:
+#     - diagram
 execute:
@@
 - We can recover some of this behaviour by introducing an additional layer in the network to represent indirect movement.
+
+# Glossary
+
+| symbol | name | description |
+|--------|------|-------------|
+| v | vertex | |
+| e | edge | |
+| i | index of vertex | |
+| j | index of other vertex | |
+| t | time | |
+| x | location | hospital |
+| W | bin width | discretisation of the temporal network |
+| $\gamma$ | discharge | rate of discharge |
+| N | num | size of something |
+| k | hitting time | (expected) |
+| $\lambda$ | rate | of an event / process |
+| T | time period | of the _entire_ simulation |
+
+define a vertex $v_i = (x_i, t_i)$ made up of a location and time resp.
 
@@
 Previous analyses of movement and spread of contagions through these networks have made the assumption that the weight $w_{ij}$ for a given edge does not change over time, i.e. the network is static.
-One advantage that the VAED provide us is the minute-scale granularity of patient admissions, which allows us to construct temporal networks and models of patient movement.
+This implies that the rate movement of patients between hospitals is Poisson, or that the time between successive admissions for a given patient is exponentially distributed.
+Empirically, this is poor modelling choice, since the readmission window, as shown in @fig-readmission-window, is definitely not exponentially distributed, and has a significant heavy tail.
+A naïve static approximation that matches the mean readmission window will be inaccurate; in the following, we will attempt to resolve this problem in three increasingly granular ways:
+
+1. using a simple piecewise exponential approximation, by choosing a single threshold
+2. using a series of static networks with ...
+3. using a temporal network that exactly captures the readmission window distribution
+
+<!-- One advantage that the VAED provide us is the minute-scale granularity of patient admissions, which allows us to construct temporal networks and models of patient movement. -->
 
@@
 
-Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_m$ to a target hospital $x_j$ at some time $t_n$.
+Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$.
 We could work with the exact times, but this will lead to an overly sparse representation that is difficult to work with.
-Instead, we discretise time into bins by treating movements involving some arbitrary hospital $x_i$ between times $t_n$ and $t_n + W$ as instead involving a single node at $v_{i,n} \coloneqq (x_i, t_n)$.
+Instead, we discretise time into bins by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$. 
 Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$.
@@
 We define the _extent_ of a simulation by the number of facilities that some patient has moved to or through within a given number of days.
-We see in @fig-extent-30 that we get starkly different distriubtions of extent for the naïve static model, commpared to the other models.
+We see in @fig-extent-30 that we get starkly different distributions of extent for the naïve static model, compared to the other models.
 
@@
 Each pair of hospitals yields a separate log-rank test statistic.
-Using a crude Bonferroni correction to counteract the multiple comparisons (which occur due to the $N=338^2$ combinations of seed and target hospitals), we find that there are significant differences between each type of model.
+Using a crude Bonferroni correction to counteract the multiple comparisons (which occur due to the ${N_x}^2=338^2$ combinations of seed and target hospitals), we find that there are significant differences between each type of model.
 We note spikes at very low p-values when comparing the naïve static model to other models.
@@
 
-fig, axs = plt.subplots(nrows=NN, ncols=NN, squeeze=True, sharex=True, sharey=True)
+fig, axs = plt.subplots(
+    nrows=NN, ncols=NN, 
+    squeeze=True, 
+    sharex=True, sharey=True, 
+    layout="constrained"
+)
 
@@
     axs[index].axvline(confidence_threshold, color='r')
-    axs[index].set_yscale('log')
+    axs[index].set_yscale('asinh')
     sns.despine(ax=axs[index])
@@
 We remove self-loops from the static network; we attempt one model without self-loops in only the direct layer (since this does not make physical sense), and another without self-loops in both the direct layer, and the indirect layer (we uspected that these self-loops slowed the movement of individuals down, but the resulting effect is the explosion of movement in the first simulation step)
+
+## Snapshots
+
+Network analysis of the explicitly-temporal network is difficult, since the number of nodes is relatively large ($N \sim N_x \times N_t$), and inversely scales with the temporal discretisation chosen ($N_t \sim 1/W$).
+
+We can instead look at the series of static network induced by projection onto the locations at each time point. This allows

### 2024-11-04 projection_metrics.py
actually pass args

@@ -184,7 +184,7 @@ def snapshot_metric_ranks(
 ):
     metric_df = metric_over_snapshots(snapshots, metric)
 
-    return rank_metrics(metric_df, ordered=True, descending=True, fill=fill)
+    return rank_metrics(metric_df, ordered=ordered, descending=descending, fill=fill)
 
 
 def plot_rankings(

### 2024-11-12 reports/report.qmd
Update report

@@
 jupyter: python3
-# filters:
-#     - diagram
 execute:
@@
 import polars as pl
-from matplotlib import pyplot as plt
+from matplotlib import pyplot as plt, colors
 import seaborn as sns
+import h5py
 import itertools
@@
 
-# Glossary
+<!-- # Glossary
 
@@
 
-define a vertex $v_i = (x_i, t_i)$ made up of a location and time resp.
+define a vertex $v_i = (x_i, t_i)$ made up of a location and time resp. -->
 
@@
 
-<- yank from [@lydeamore_carbapenemase-producing_2024]
+Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of more than 300 hospital campuses, which vary in size, case mix and services.
 
-Admissions data is from the Victorian Admitted Episdoes Dataset (VAED).
+Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
+In particular, admissions data from [] to [] ([] days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admissions is []
+This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.
 
-Some exploratory analysis was done in by [@lydeamore_carbapenemase-producing_2024]
+Some exploratory analysis of this data was done in by [@lydeamore_carbapenemase-producing_2024].
 
@@
 Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$. 
-We could work with the exact times, but this will lead to an overly sparse representation that is difficult to work with.
-Instead, we discretise time into bins by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$. 
-Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e. we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$.
+Because we cannot export exact times, we instead discretise time.
+This is done by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$. 
+Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e.
+we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$, but with some explicit temporal representation.
 
@@
 
-### Dynamics
-
-<!-- TODO need to work on indexing consistency -->
+## Dynamics
 
@@
 
-# Results
-
-## A Reduced Model
+### A Reduced Model
 
@@
 ```{dot}
-//| fig-width: 50%
+//| fig-width: 30%
 digraph G {
@@
 
-In the model without indirect transfer, we can compute the expected hitting time $k$ from $S_0$ to $S_1$
+In the model without indirect transfer, we can compute the expected hitting time $k_{1}$ from $S_0$ to $S_1$ as
 
-$$k = {\lambda_0}^{-1}$$
+$$k_{1} = {\lambda_0}^{-1}$$
 
 In a model with indirect transfer, we introduce an additional state $S_2$ which represents the indirect pathway.
-If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k$ of $S_1$ given we start at $S_0$:
+If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k_{1^*}$ of $S_1$ given we start at $S_0$:
 
-$$k = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
+$$k_{1^*} = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
 
@@
 For a larger network, the number of possible paths causes complications for explicit computation of the effect, but we can always expect it to follow a similar principle, where splitting transfers into a direct path and a slower indirect path will add to the expected hitting time.
-These can be obtained by solving a linear system @Norris_Markov_Chains_1997.
+These can be obtained by solving a linear system [@Norris_Markov_Chains_1997].
+
+# Results
+
+## Temporal Network Analysis
+
+## Effect of Home-return Mechanic
+
 For the naïve static network, we can model the rate of movement from hospital $i$ to hospital $j$ as the mean number of observed movements over the observation period (XXX days).
@@
 
-Comparing the relative difference in the mean hitting times gives us @fig-hitting-time-difference
+Comparing the (uniform) relative difference in the mean hitting times ($= k_{1^*} / (k_1 + 1)$) gives us @fig-hitting-time-difference.
+We see that there is significant banding in the structure of the matrix.
 
@@
 
-# temp hack
-# todo port across computation code
-from PIL import Image
-image = Image.open("./precompiled/zrelative_change_hitting_time_home.png")
-image
+with h5py.File('../hitting_time_analysis/hitting_time_arrays.h5', 'r') as fp:
+    naive_static_times = fp['naive_static'][:]
+    delayed_static_times = fp['delayed_static'][:]
+
+relative_change = delayed_static_times / (naive_static_times + 1)
+plt.matshow(relative_change, cmap='viridis', fignum=99)
+plt.colorbar(label='Relative difference in hitting time')
+plt.xlabel('Hospital to hit')
+plt.ylabel('Starting hospital');
 ```
@@
 
+# Discussion
+
+The modelling of patient transfers between hospitals has historically been done by using static networks.
+
 # Appendix

### 2024-11-12 hitting_time_analysis.py
refactor hitting time analysis script for reproduc

@@ -5,168 +5,212 @@ from scipy import sparse
 
 import network_conversion as ntc
 import hitting_markov as hit
+from realisation_summary_stats import hitting_time
 
+## NAIVE STATIC MODEL
+def compute_static_hitting_times(
+    graph_file = "./concordant_networks/shuf_static_network.graphml",
+    scaling = (10*365),
+    plot = False,
+):
 
-## STANDARD STATIC MODEL
+    G = ig.Graph.Read(graph_file)
 
-G = ig.Graph.Read("./concordant_networks/shuf_static_network.graphml")
+    graph_map = {int(k):v for k,v in zip(G.vs['name'], G.vs['id'])}
+    sorted_graph_map = {k: graph_map[k] for k in sorted(graph_map.keys())}
+    graph_ordering = ntc.Ordering(sorted_graph_map)
+    N = len(graph_ordering.sizes)
 
-graph_map = {int(k):v for k,v in zip(G.vs['name'], G.vs['id'])}
-sorted_graph_map = {k: graph_map[k] for k in sorted(graph_map.keys())}
-graph_ordering = ntc.Ordering(sorted_graph_map)
-N = len(graph_ordering.sizes)
+    # we use the naive construction, since the null set (never return) version creates 
+    # results that misrepresent the amount of time it takes to move through the system
+    # We also use the internal function in order to directly pass the transition matrix 
+    # constructed, as opposed to the non-scaled weighted adjacency matrix
+    T = ntc.transition_matrix_from_graph(G, 
+                                        ordering=graph_ordering, 
+                                        global_scaling=scaling, 
+                                        ordering_key='name', 
+                                        adjacency_attribute='weight'
+                                        )
+    Q = hit._Q_mat_naive(T)
 
-## we use the naive construction, since the null set (never return) version creates 
-## results that misrepresent the amount of time it takes to move through the system
-## We also use the internal function in order to directly pass the transition matrix 
-## constructed, as opposed to the non-scaled weighted adjacency matrix
-T = ntc.transition_matrix_from_graph(G, 
-                                     ordering=graph_ordering, 
-                                     global_scaling=(10*365), 
-                                     ordering_key='name', 
[…]

### 2024-11-12 analyse_temporal_snapshots.py
Basic time series analysis of cluster belonging

@@ -1,0 +1,46 @@
+# this will rely on clustering labelling
+
+import polars as pl
+import itertools
+from sklearn import metrics as sklm
+from matplotlib import pyplot as plt
+
+def import_clusters(cluster_file = "conc_snapshots_deprc/conc_tempo_14_snapshots/results/infomap_clusters.csv"):
+    return pl.read_csv(cluster_file)
+
+
+
+
+
+def main():
+
+    clusters = import_clusters()
+
+    adjusted_clusters = clusters.fill_null(-1)
+
+    persistence_functions = {
+        'rand': sklm.adjusted_rand_score,
+        'mutual_info': sklm.adjusted_mutual_info_score,
+    }
+
+    cl_persist_metric = {
+        score_name: [
+            score_function(
+                adjusted_clusters.select(cl_left).to_series().to_numpy(),
+                adjusted_clusters.select(cl_right).to_series().to_numpy(),
+            )
+            for cl_left, cl_right in itertools.pairwise(adjusted_clusters.columns[1:])
+        ]
+        for score_name, score_function in persistence_functions.items()
+    }
+
+    for score_name, score in cl_persist_metric.items():
+        plt.plot(score, marker='.', label=score_name)
+    plt.legend()
+
+
+
+
+
+if __name__ == "__main__":
+    main()

### 2024-11-14 reports/report.qmd
Add diagram

@@
 csl: plos.csl
+filters:
+    - diagram
 jupyter: python3
@@
 
-Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of more than 300 hospital campuses, which vary in size, case mix and services.
+Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.
 
 Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
-In particular, admissions data from [] to [] ([] days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admissions is []
+In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study.
+All Victorian hospital admissions during this study period are included.
+The total number of admission records is 26876787.
 This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.
@@
 
-## Network Modelling {#sec-network-modelling}
+## Existing Network Modelling {#sec-network-modelling}
 
@@
 ## Explicitly Temporal Models
-Previous analyses of movement and spread of contagions through these networks have made the assumption that the weight $w_{ij}$ for a given edge does not change over time, i.e. the network is static.
-This implies that the rate movement of patients between hospitals is Poisson, or that the time between successive admissions for a given patient is exponentially distributed.
-Empirically, this is poor modelling choice, since the readmission window, as shown in @fig-readmission-window, is definitely not exponentially distributed, and has a significant heavy tail.
+Previous analyses of movement and spread of contagions through these networks have made assumptions that result in the weight $w_{ij}$ of any given edge being constant over time, i.e.
+the network is static.
+[.....Typically, this is a result of performing an aggregation, and approximating the patient movement between hospitals as being Poisson, or equivalently that the residence time  hospitals being exponentially distributed......]
+As shown in @fig-readmission-window, these approximations lead to misleading dynamics on the hospital system of Victoria: the empirical distribution of inter-transfer times [....].
 A naïve static approximation that matches the mean readmission window will be inaccurate; in the following, we will attempt to resolve this problem in three increasingly granular ways:
@@
 <!-- One advantage that the VAED provide us is the minute-scale granularity of patient admissions, which allows us to construct temporal networks and models of patient movement. -->
-
-<- Description of temporal networks and models>
-
-<- Description of temporally reductive networks and models>
 
@@
 
-### Networks and Projections
-
 Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$. 
@@
 This is done by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$.
+
+<!-- ::: {#fig-}
+```{.tikz}
+\begin{tikzpicture}
+\tikzstyle{every node}=[font=\LARGE]
+\draw [short] (2.75,14.75) -- (16.5,14.75);
+\draw [short] (2.75,11) -- (16.5,11);
+\draw [dashed] (3.5,16.25) -- (3.5,9.25);
+\draw [dashed] (8.25,16.5) -- (8.25,9.25);
+\draw [dashed] (12.25,16.5) -- (12.25,9.25);
+\draw [dashed] (16,16.5) -- (16,9.25);
+\draw [->, >=Stealth] (7.25,14.75) -- (9.25,11);
+\draw [->, >=Stealth] (8.75,14.75) -- (13.25,11);
+\draw [->, >=Stealth] (9.5,11) -- (10.75,14.75);
+\draw [->, >=Stealth] (4.25,11) -- (5,14.75);
+\draw [->, >=Stealth] (5.5,14.75) -- (5.5,11);
+\draw [->, >=Stealth] (6.25,11) -- (7,14.75);
+\draw [->, >=Stealth] (10,14.75) .. controls (11.75,15.75) and (12.5,15.75) .. (14.25,14.75) ;
+\draw [->, >=Stealth] (15,11) -- (15,14.75);
+\end{tikzpicture}
+``` -->
+
+<!-- ::: {#fig-time-discretisation}
+
+```{.tikz}
+%%| caption="This is an image, created by **TikZ i.e. LaTeX**."
+%%| additionalPackages="\usepackage{adjustbox}"
+\usetikzlibrary{arrows}
+
+\resizebox{16cm}{!}{%
+\begin{tikzpicture}[auto]
+    \tikzstyle{every node}=[font=\LARGE]
+    \draw [short] (0.275,1.475) -- (1.65,1.475);
+    \draw [short] (0.275,1.1) -- (1.65,1.1);
+    \draw [dashed] (0.35,1.65) -- (0.35,0.925);
+    \draw [dashed] (0.825,1.65) -- (0.825,0.925);
+    \draw [dashed] (1.225,1.65) -- (1.225,0.925);
+    \draw [dashed] (1.6,1.65) -- (1.6,0.925);
+    \draw [O-Latex] (0.725,1.475) -- (0.925,1.1);
+    \draw [->, >=Stealth] (0.875,1.475) -- (1.325,1.1);
+    \draw [->, >=Stealth] (0.95,1.1) -- (1.075,1.475);
+    \draw [->, >=Stealth] (0.425,1.1) -- (0.5,1.475);
+    \draw [->, >=Stealth] (0.55,1.475) -- (0.55,1.1);
+    \draw [->, >=Stealth] (0.625,1.1) -- (0.7,1.475);
+    \draw [->, >=Stealth] (1.0,1.475) .. controls (1.175,1.575) and (1.25,1.575) .. (1.425,1.475) ;
+    \draw [->, >=Stealth] (1.5,1.1) -- (1.5,1.475);
+\end{tikzpicture}
+}
+```
+
+Example schematic for time discretisation
+::: -->
+
 Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e.
@@
 These can be obtained by solving a linear system [@Norris_Markov_Chains_1997].
+
+![An example of a small 2-layer static network](diagrams/precompiled/BiLayer_ManimCE_v0.18.1.png){#fig-bilayer}
 
@@
 
+{{< pagebreak >}}
 # Appendix

### 2024-11-14 reports/diagrams/multilayer_network.py

@@ -1,0 +1,106 @@
+#! /usr/bin/env -S manim -qh
+
+from manim import *
+
+config.background_color = WHITE
+config.background_opacity = 0.0
+
+class BiLayer(ThreeDScene):
+
+    def construct(self):
+
+        self.set_camera_orientation(
+            phi=45 * DEGREES,
+        )
+
+        separation = 2.5
+        levels = [separation, -separation]
+        plane_height = 3
+        plane_width = 4
+        plane_shear_angle = 60 * DEGREES
+        plane_shear = plane_height / np.tan(plane_shear_angle)
+        colours = [YELLOW, GREEN]
+        names = ['Direct', 'Indirect']
+        for level, colour, name in zip(levels, colours, names):
+            plane = Polygon(
+                *[
+                    (-plane_width - plane_shear, -plane_height, level),
+                    (-plane_width + plane_shear, plane_height, level),
+                    (plane_width + plane_shear, plane_height, level),
+                    (plane_width - plane_shear, -plane_height, level),
+                ],
+                color = colour, 
+                fill_opacity=0.4,
+                stroke_width=0.0,
+            )
+            plane.set_z_index(0)
+            self.add(plane)
+            label_text = Text(name, color=BLACK, font_size=18)
+            label_text.align_to(label_text.get_corner(DOWN + LEFT), DOWN + LEFT)
+            label_text.move_to(plane.get_vertices()[0] + UP/2 + RIGHT, aligned_edge=DOWN + LEFT)
+            self.add(label_text)
+
+        vertices = ['A', 'B', 'C']
+        
+        G0 = DiGraph(
+            vertices = vertices,
+            edges = [
+                ('A', 'B'),
+                ('B', 'C'),
+                ('A', 'C'),
+            ],
+            labels = True,
+            layout = "circular",
+            vertex_config = {
+                'fill_opacity': 1.0,
+            },
+            edge_config = {
+                "path_arc" : 1.2,
+                "color": BLACK,
+            },
+        )
+
+        G0.set_z(levels[0])
+        for v in G0.vertices.values():
+            v.set_z_index(5)
+        for e in G0.edges.values():
+            e.set_z_index(4)
+
+        self.add(G0)
+
+        G1 = DiGraph(
+            vertices = vertices,
+            edges = [
+                ('A', 'B'),
+                ('B', 'A'),
+                ('B', 'C'),
+                ('C', 'A'),
+                ('A', 'A'),
+                ('B', 'B'),
[…]

### 2024-11-25 staticify_network.py
fix deprc pl.Expr.map_dict -> pl.Expr.replace

@@ -95,15 +95,13 @@ def generate_graph_from_edge_list(edge_df: pl.DataFrame) -> ig.Graph:
         .otherwise(pl.col("direct_weight"))
         .alias("direct_weight")
     ).with_columns(
-        pl.col("source_loc").map_dict(node_index).alias("mapped_source"),
-        pl.col("dest_loc").map_dict(node_index).alias("mapped_dest"),
+        pl.col("source_loc").replace_strict(node_index, return_dtype=pl.Int32).alias("mapped_source"),
+        pl.col("dest_loc").replace_strict(node_index, return_dtype=pl.Int32).alias("mapped_dest"),
     )
 
-    edge_list = mapped_edge_df.select("mapped_source", "mapped_dest").to_dict()
-    edges = zip(edge_list["mapped_source"], edge_list["mapped_dest"])
+    edges = mapped_edge_df.select(pl.concat_list("mapped_source", "mapped_dest")).to_series().to_list()
 
     edge_info = mapped_edge_df.drop("mapped_source", "mapped_dest").to_dict()
 
-    G.add_edges(es=list(edges), attributes=edge_info)
-
+    G.add_edges(es=edges, attributes=edge_info)
     return G

### 2024-11-25 network_conversion.py
replace bad self-built where with inbuilt np.where

@@ -36,11 +36,11 @@ class Ordering:
     @lru_cache()
     def __getitem__(self, key):
         """A reverse search of the order list"""
-        value = nparr_find(self.order, int(key))
-        if value is None:
-            raise NotFound
-        # value is a 1D tuple; we should return the interior item
-        return value[0]
+        value = np.nonzero(self.order == key)[0]
+        # check for not found
+        if len(value) == 0:
+            raise NotFound(key)
+        return value.item()
 
     @classmethod
     def from_file(cls, file_name: PathLike):
@@ -106,7 +106,7 @@ def transition_matrix_from_graph(
     else:
         graph_order_base = graph.vs[ordering_key]
 
-    graph_paste_order = [ordering[int(k)] for k in graph_order_base]
+    graph_paste_order = [ordering[k] for k in graph_order_base]
     A = graph.get_adjacency(attribute=adjacency_attribute)
     if matrix_size is None:
         MAXN = len(graph_paste_order)

### 2024-11-25 probability_over_networks.py
start examining the simulations and network code for small case

@@ -1,0 +1,175 @@
+# test the static probability theory
+
+import numpy as np
+import polars as pl
+import igraph as ig
+import warnings
+
+from collections import Counter
+
+from pytz import BaseTzInfo
+
+import staticify_network as stcfy
+import projection_metrics as projm
+
+import network_conversion as nconv
+import network_simulation as nsiml
+
+def make_graph_from_name_edges(name_edges, vertex_names=None):
+
+    if vertex_names is None:
+        vertex_names = sorted({name for vertices in name_edges.keys() for name in vertices})
+    if isinstance(vertex_names, list | set):
+        vertex_map = [{"name": name, "id": i} for i, name in enumerate(vertex_names)]
+    elif isinstance(vertex_names, dict):
+        vertex_map = [
+            {"name": name, "id": i, **values}
+            for i, (name, values) in enumerate(vertex_names.items())
+        ]
+    else:
+        raise TypeError("vertex_names provided is not a list, set or dict")
+    vertex_map_df = pl.from_dicts(vertex_map)
+
+    edge_df = pl.from_dicts(
+        [
+            {"source_vertex": v0, "target_vertex": v1, **values}
+            for (v0, v1), values in name_edges.items()
+        ]
+    )
+
+    mapped_edge_df = (
+        edge_df.join(
+            vertex_map_df.select("name", "id"), left_on="source_vertex", right_on="name"
+        )
+        .rename({"id": "source_id"})
+        .join(
+            vertex_map_df.select("name", "id"), left_on="target_vertex", right_on="name"
+        )
+        .rename({"id": "target_id"})
+        .with_columns(
+            edge_name=pl.concat_list("source_vertex", "target_vertex"),
+            edge_id=pl.concat_list("source_id", "target_id"),
+        )
+    )
+
+    edge_list = mapped_edge_df.select("edge_id").to_series().to_list()
+
+    vertex_attrs = vertex_map_df.sort("id").drop("id").to_dict(as_series=False)
+    edge_attrs = mapped_edge_df.to_dict(as_series=False)
+
+    return ig.Graph(
+        n=len(vertex_map),
+        edges=edge_list,
+        directed=True,
+        edge_attrs=edge_attrs,
+        vertex_attrs=vertex_attrs,
+    )
+
+
+def basic_example():
+    locs = ["A", "B", "C"]
+    times = list(range(20))
+    vertex_info = {
[…]

### 2024-11-25 reports/report.qmd
Add newer report ideas

@@
 sns.set_context('paper')
-# plt.style.use('ggplot')
+plt.style.use('tableau-colorblind10')
 ```
@@
 digraph G {
-    S₀ -> S₁ [ label="λ₀" ]
+    A -> B [ label="λ₀" ]
 }
@@
 digraph G {
-    S₀ -> S₁ [ label="λ₀p" ]
-    S₀ -> S₂ [ label="λ₀(1-p)" ]
-    S₂ -> S₁ [ label="λ₂" ]
+    A -> B [ label="λ†p" ]
+    A -> H [ label="λ†(1-p)" ]
+    H -> B [ label="λₕ" ]
 }
@@
 
-In the model without indirect transfer, we can compute the expected hitting time $k_{1}$ from $S_0$ to $S_1$ as
+::: {#new-eq}
+
+Let us denote the (empirical) distribution of return times as $f(T)$, where $T$ is the return time.
+Then for a naive model, we have that the hazard of moving between two hospitals is approximately
+$\lambda_0 = (\mathbb{E}(A \to B))^{-1} = (\int_0^{\infty} f(T) T dT)^{-1}$
+
+For the refined model, we have instead two processes
+
+$A \to B$ and $A \to H \to B$,
+
+where
+
+$\mathbb{E}(T_{A\to B}) = \int_0^{\Omega} f(T) T dT$
+
+and
+
+$\mathbb{E}(T_{A\to H\to B}) = \int_\Omega^{\infty} f(T) T dT$
+
+We could simulate with some rate $\lambda_\dagger$ so that
+
+$\lambda_d = \mathbb{E}(T_{A\to B})^{-1} = \lambda_\dagger \times p$
+
+and
+
+$\lambda_x = \lambda_\dagger \times (1-p)$ (indirect)
+
+AND
+
+$p = \int_0^{\Omega} f(T) dT$ is the proportion of individuals that take the direct path.
+
+Thus we can recover that $\lambda_\dagger = \left(\int_0^{\Omega} f(T) T dT\right)^{-1} \int_0^{\infty} f(T) dT$.
+
+And if we require that
+
+$(\lambda_x)^{-1} + (\lambda_h)^{-1} = \int_\Omega^{\infty} f(T) T dT$
+
+and so
+
+$$
+\begin{aligned}
+(\lambda_h)^{-1} &= \left[\int_\Omega^{\infty} f(T) T dT \right] - (\lambda_\dagger (1-p))^{-1}\\
+&= \left[\int_\Omega^{\infty} f(T) T dT \right] - \left[\left[\int_0^{\Omega} f(T) T dT \right]^{-1} {\int_\Omega^{\infty} f(T) dT} \right]^{-1}\\
+&= \left[\int_\Omega^{\infty} f(T) T dT \right] - \left[\int_0^{\Omega} f(T) T dT \frac{1}{\int_\Omega^\infty f(T) dT}\right]
+\end{aligned}
+$$
+
+We also need that $(\lambda_\dagger (1-p))^{-1} < \int_\Omega^{\infty} f(T) T dT$ (is this always possible?)
+
+If $T \sim Exp(\lambda)$, then $f(T) = \lambda \exp(-\lambda T)$ and $F(T) = 1 - \exp(-\lambda T)$.
+We can also find that $\int_a^b f(T) T dT = \left[ -(T + \lambda^{-1}) \exp(-\lambda T) \right]^b_a$, so that
+
+$\int_0^\Omega f(T) T dT = -(\Omega + \lambda^{-1}) \exp(-\lambda \Omega) + (0 + \lambda^{-1}) \exp(-\lambda 0) = \lambda^{-1} (1- \exp(-\lambda \Omega)) - \Omega \exp(-\lambda \Omega)$
+
+and
+
+$\int_\Omega^\infty f(T) T dT = (\Omega \lambda^{-1}) \exp(-\lambda \Omega)$
+
+so that
+
+$(\lambda_\dagger (1-p))^{-1} = [\lambda^{-1} (1- \exp(-\lambda \Omega)) - \Omega \exp(-\lambda \Omega)]  \frac{1}{\exp(-\lambda \Omega)}$
+
+We can then say that 
+:::
+
+<!-- In the model without indirect transfer, we can compute the expected hitting time $k_{1}$ from $S_0$ to $S_1$ as
 
@@
 
-i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$.
+i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$. -->
 

### 2024-11-25 reports/diagrams/multilayer_network.py

@@ -42,6 +42,23 @@ class BiLayer(ThreeDScene):
 
         vertices = ['A', 'B', 'C']
         
+        graph_config = dict(
+            labels = True,
+            layout = 'circular',
+            vertex_config = dict(
+                fill_opacity = 1.0,
+                stroke_color = BLACK,
+            ),
+            edge_config = dict(
+                path_arc = 1.2,
+                loop_config = {
+                    "angle_between_points": PI, 
+                    "path_arc": 3 * PI / 2,
+                },
+                color = BLACK,
+            )
+        )
+
         G0 = DiGraph(
             vertices = vertices,
             edges = [
@@ -49,15 +66,7 @@ class BiLayer(ThreeDScene):
                 ('B', 'C'),
                 ('A', 'C'),
             ],
-            labels = True,
-            layout = "circular",
-            vertex_config = {
-                'fill_opacity': 1.0,
-            },
-            edge_config = {
-                "path_arc" : 1.2,
-                "color": BLACK,
-            },
+            **graph_config
         )
 
         G0.set_z(levels[0])
@@ -79,16 +88,7 @@ class BiLayer(ThreeDScene):
                 ('B', 'B'),
                 ('C', 'C'),
             ],
-            labels = True,
-            layout = "circular",
-            vertex_config={
-                'fill_opacity': 1.0,
-            },
-            edge_config = {
-                "path_arc" : 1.2,
-                "loop_config": {"angle_between_points": PI, "path_arc": 3 * PI / 2},
-                "color": BLACK,
-            },
+            **graph_config
         )
 
         G1.set_z(levels[1])

### 2024-11-26 probability_over_networks.py
Add in snapshot sim explorations

@@ -1,5 +1,6 @@
 # test the static probability theory
 
+from typing import Hashable
 import numpy as np
 import polars as pl
 import igraph as ig
@@ -104,9 +105,9 @@ def get_snapshots(G: ig.Graph):
         snapshot['time'] = t
         snapshot['duration'] = dt
 
-    return snapshots
+    return {snapshot['time']: snapshot for snapshot in snapshots}
 
-def simulate_on(G: ig.Graph | list[ig.Graph], ordering, type='temporal', repeats=30):
+def simulate_on(G: ig.Graph | dict[Hashable, ig.Graph], ordering, type='temporal', repeats=30):
     ordering = nconv.Ordering(ordering)
     prob_zero = np.array([0.0 for _ in ordering])
     base_params = {
@@ -115,6 +116,8 @@ def simulate_on(G: ig.Graph | list[ig.Graph], ordering, type='temporal', repeats
     }
 
     sim_hists = []
+    move_in = []
+    move_out = []
     match type:
         case "temporal":
             converter = nconv.TemporalNetworkConverter(
@@ -127,10 +130,18 @@ def simulate_on(G: ig.Graph | list[ig.Graph], ordering, type='temporal', repeats
                 dt = 1.0,
                 num_times=converter.NT,
                 discretisation_size=converter.DT,
-                track_movement=False,
+                track_movement=True,
             )
         case "snapshot":
-            pass
+            converter = nconv.SnapshotWithHomeConverter(G, ordering=ordering)
+            parameters = converter.map_parameters(base_params)
+            sim = nsiml.SnapshotWithHome(
+                full_size=converter.ordering.sizes,
+                parameters=parameters,
+                timings=converter.snapshot_times,
+                track_movement=True,
+                dt=1.0
+            )
         case "static":
             converter = nconv.StaticConverter(G, ordering=ordering, time_span=20)
             parameters = converter.map_parameters(base_params)
@@ -138,15 +149,25 @@ def simulate_on(G: ig.Graph | list[ig.Graph], ordering, type='temporal', repeats
                 full_size = converter.ordering.sizes,
                 parameters=parameters,
                 dt =1.0,
-                track_movement=False
+                track_movement=True
             )
+        case _:
+            raise ValueError(f"invalid type: {type}")
     sim.seed(n_seed_events=0, n_seed_number=0)
     for repeat in range(repeats):
[…]

### 2024-11-26 network_simulation.py
bugfix: axis of summation for returning

@@ -190,7 +190,7 @@ class SnapshotWithHome(SimulationWithMovers):
 
         indirect_return_rate = TX_in[self.current_index] * X
         indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, X)
-        indirect_returns = indirect_returns_raw.sum(axis=1).reshape(NARROW)
+        indirect_returns = indirect_returns_raw.sum(axis=0).reshape(NARROW)
         self.mover_in.append(indirect_returns)
 
         I_new = np.clip(
@@ -259,7 +259,7 @@ class StaticWithHome(SimulationWithMovers):
 
         indirect_return_rate = TX_in * X
         indirect_returns_raw = truncated_poisson(indirect_return_rate * self.dt, X)
-        indirect_returns = indirect_returns_raw.sum(axis=1).reshape(NARROW)
+        indirect_returns = indirect_returns_raw.sum(axis=0).reshape(NARROW)
         self.mover_in.append(indirect_returns)
 
         I_new = np.clip(

### 2025-01-15 reports/report.qmd
Add multichain hitting among other things

@@
 sns.set_theme(style='ticks')
+sns.set_palette('colorblind')
 sns.set_context('paper')
-plt.style.use('tableau-colorblind10')
 ```
@@
     'temporal': '../zero_sims/temporal/sim_all_30s.h5',
-    'static': '../zero_sims/static/sim_all_30s.h5',
-    'snapshot': '../zero_sims/snapshot/sims_with_30seeds_pseudocap.h5',
+    'static': '../zero_sims/static/sim_all_30s_fixed_axis.h5',
+    'snapshot': '../zero_sims/snapshot/sims_with_30seeds_pseudocap_fixed_axis.h5',
     'naive_static': '../zero_sims/naive_static/sim_all_30s.h5',
+}
+
+metrics_archives = {
+    'temporal': '../zero_sims/temporal/metrics_30s.parquet',
+    'static': '../zero_sims/static/metrics_30s_fa.parquet',
+    'snapshot': '../zero_sims/snapshot/metrics_30s_fa.parquet',
+    'naive_static': '../zero_sims/naive_static/metrics_30s.parquet'
 }
@@
 # metrics precomputed in compute_summary_stats.ipynb
-for label, archive in simulation_archives.items():
-        processed_results[label] = pl.read_parquet(Path(archive).with_name('metrics_30s.parquet'))
+for label, archive in metrics_archives.items():
+        processed_results[label] = pl.read_parquet(archive)
 ```

### 2025-01-15 clusters.py

@@ -78,7 +78,8 @@ def multi_infomap(G_base: str | Path, conv=True):
 
 
 def heat_wavelet(G: ig.Graph, tau=1.2, normalised=True):
-    assert not G.is_directed(), "Laplacian not defined for directed networks"
+    if G.is_directed():
+        raise ValueError("Laplacian not defined for directed networks")
     A = np.array(G.get_adjacency().data)
     D = np.diag(np.array(G.degree()))
     W = 1 / np.sqrt(D)

### 2025-01-15 deprecated_simulation_code/summary_stats_zero.py

@@ -124,6 +124,8 @@ class RealisationArray():
                 g.attrs['seed'] = rlz['seed']
                 g.attrs['simdate'] = rlz['simdate']
                 g.attrs['simtime'] = rlz['simtime']
+                if 'delay' in rlz:
+                    g.attrs['delay'] = rlz['delay']
                 rfl = self.realisations[rlz['row_nr']]
                 with np.load(rfl) as ifp:
                     tarr = ifp.get('ts') or ifp['t']

### 2025-01-15 hitting_time_analysis.py

@@ -8,10 +8,10 @@ import hitting_markov as hit
 from realisation_summary_stats import hitting_time
 
 ## NAIVE STATIC MODEL
-def compute_static_hitting_times(
+def get_naive_static_Q(
     graph_file = "./concordant_networks/shuf_static_network.graphml",
-    scaling = (10*365),
-    plot = False,
+    size_file="./concordant_networks/size_14.csv",
+    time_span = (10*365),
 ):
 
     G = ig.Graph.Read(graph_file)
@@ -19,19 +19,32 @@ def compute_static_hitting_times(
     graph_map = {int(k):v for k,v in zip(G.vs['name'], G.vs['id'])}
     sorted_graph_map = {k: graph_map[k] for k in sorted(graph_map.keys())}
     graph_ordering = ntc.Ordering(sorted_graph_map)
-    N = len(graph_ordering.sizes)
+
+    size_order = ntc.Ordering.from_file(size_file)
+    sizes = np.array([size_order(i) for i in graph_ordering.order])
 
     # we use the naive construction, since the null set (never return) version creates 
     # results that misrepresent the amount of time it takes to move through the system
     # We also use the internal function in order to directly pass the transition matrix 
     # constructed, as opposed to the non-scaled weighted adjacency matrix
     T = ntc.transition_matrix_from_graph(G, 
-                                        ordering=graph_ordering, 
-                                        global_scaling=scaling, 
+                                        ordering=graph_ordering,
+                                        scaling_per_node=sizes,
+                                        global_scaling=time_span,
                                         ordering_key='name', 
-                                        adjacency_attribute='weight'
+                                        adjacency_attribute='weight',
                                         )
     Q = hit._Q_mat_naive(T)
+
+    return Q
+
+
+def compute_static_hitting_times(
+    Q,
+    plot = False,
+):
+
+    N, _ = Q.shape
 
     ## I wish I could speed this up, but I think we need the absorption
     hitting_times = [
@@ -47,10 +60,10 @@ def compute_static_hitting_times(
         plt.xlabel('to')
         plt.ylabel('from')
 
-        ax = axim.axes
+        ax = [axim.axes]
 
     # plt.savefig("zstatic_no_home_time.png", dpi=360)
-    return hitting_time_arr, [ax]
+    return hitting_time_arr, ax
 
 
 ### STATIC MODEL WITH DELAYED RETURN
[…]

### 2025-01-15 hitting_time_multichain.py

@@ -1,0 +1,51 @@
+"""
+This is an implementation of the computation of expected minimal hitting time over N identically parameterised continuous-time Markov chains {M_i(λ,Q)} 
+"""
+import hitting_markov as mkv  # single chain computations
+import numpy as np
+from scipy import linalg
+from scipy import integrate
+
+def construct_surival_ode(Q, index=-1):
+    """Implement the problem u'(t) = Q_{A^cA^c} u(t)"""
+    # remove row and column
+    Qcc = np.delete(np.delete(Q, index, axis=0), index, axis=1)
+
+    def backward_equation(t, y):
+        return (Qcc @ y[:,np.newaxis]).flatten()
+    
+    u_0 = np.ones(Qcc.shape[0])
+
+    return backward_equation, u_0
+
+
+"""
+STRATEGY
+
+-> Set up the integration scheme using Gauss-Laguerre quadrature.
+    - requires the setup of Laguerre polynomials to solve for roots.
+        - can use np.polynomial.laguerre
+-> Use scipy.integrate.solve_ivp to get u(t) at the given roots
+-> Use w_i * (u(t) * Exp(t)) to solve the integral and return hitting times.
+"""
+
+def determine_laguerre_roots_and_weights(precision=30):
+    return np.polynomial.laguerre.laggauss(precision)
+
+def solve_survival_ode(Nchains, Q, index=-1, precision=30):
+    roots, weights = determine_laguerre_roots_and_weights(precision)
+    gl_weights = weights * np.exp(roots)
+
+    survival_ode_fun, survival_ode_y0 = construct_surival_ode(Q, index=index)
+
+    tspan = [0, max(roots)*1.1]
+
+    survival_sol = integrate.solve_ivp(
+        fun=survival_ode_fun,
+        t_span=tspan,
+        y0=survival_ode_y0,
+        t_eval=roots
+    )
+
+    return np.dot(gl_weights, survival_sol.y.T ** Nchains)
+

### 2025-01-15 network_conversion.py

@@ -7,7 +7,7 @@ import numba as nb
 from scipy import sparse
 import polars as pl
 import igraph as ig
-from typing import Mapping, Hashable, Sequence, Any, SupportsFloat
+from typing import Mapping, Hashable, Optional, Sequence, Any, SupportsFloat
 from functools import lru_cache
 from pathlib import Path
 
@@ -88,7 +88,7 @@ class ColumnDict(dict):
 
 def transition_matrix_from_graph(
     graph: ig.Graph,
-    ordering: SupportsGet | None = None,
+    ordering: Optional[SupportsGet] = None,
     scaling_per_node: Sequence | None = None,
     global_scaling: float = 1,
     ordering_key: Hashable | None = None,

### 2025-01-15 numba_sample.py
New numba accelled samples

@@ -1,6 +1,6 @@
 import numpy as np
 from numba import njit
-from numba import guvectorize, float64, int64
+from numba import guvectorize, float64, int64, prange
 from scipy import sparse
 
 
@@ -72,6 +72,36 @@ def _mn_sparse_full(
 
     return result
 
+@njit()
+def _mn_sparse_full_partitioned(
+    input_vector, prob_matrix_data, prob_matrix_indices, prob_matrix_indptr, left_partition_size, right_partition_size
+):
+    n_in = len(input_vector)
+    x_left, y_left = left_partition_size
+    x_right, y_right = right_partition_size
+    assert n_in == x_left
+    assert n_in == x_right
+
+    collated_results = np.zeros((n_in, y_left+y_right), dtype=np.int64)
+    left_part = np.zeros(left_partition_size, dtype=np.int64)
+    right_part = np.zeros(right_partition_size, dtype=np.int64)
+
+    for i in range(n_in):
+        v = input_vector[i]
+        if v == 0:
+            continue
+
+        row_start = prob_matrix_indptr[i]
+        row_end = prob_matrix_indptr[i + 1]
+        probs = prob_matrix_data[row_start:row_end]
+
+        samples = np.random.multinomial(v, probs)
+        collated_results[i, prob_matrix_indices[row_start:row_end]] += samples
+
+    left_part += collated_results[:, :y_left]
+    right_part += collated_results[:, y_left:] 
+
+    return left_part, right_part
 
 def multinomial_sparse_full(trials, prob_matrix: sparse.csr_array):
     """Performs numba-acceled multinomial smapling for a large 2D probability matrix
@@ -87,6 +117,26 @@ def multinomial_sparse_full(trials, prob_matrix: sparse.csr_array):
         prob_matrix.indices,
         prob_matrix.indptr,
         prob_matrix.shape,
+    )
+
+def multinomial_sparse_full_partitioned(trials, prob_matrix: sparse.csr_array, partition_index: int):
+    """Performs numba-acceled multinomial smapling for a large 2D probability matrix
+
+    Arguments
+    ---
+    trials: <np.ndarray[int]> vector of number of trials, corresponds to rows of prob_matrix
+    prob_matrix: <np.ndarray[float]> 2D matrix of probabilities. Normalised so that each row sums to 1.
+    """
+    n_x, n_y = prob_matrix.shape
+    n_l = partition_index
+    n_r = n_y - partition_index
[…]

### 2025-02-19 small_case.qmd
some hitting_time analysis

---
title: "Small Case: Checking Hitting Times in the Year of Our Lord"
format:
  html:
    code-fold: true
---

```{python Imports} 
import numpy as np
import networkx as nx
import igraph as ig

import hitting_markov as hmk
import hitting_time_analysis as ht
import hitting_time_multichain as mch
import network_conversion as ntc

from matplotlib import pyplot as plt
```

Here we want to check that our simulations and analytical results have parity.

The idea is to setup a small network that has at least one very rare link, and compute the hitting times analytically for N=[1, 5, 10, 30] individuals, and also simualte for those cases.

We generate a small 5-node directed, weighted network.
```{python}
G = ig.Graph(
  edges = [
    (0, 1),
    (0, 3),
    (1, 2),
    (1, 3),
    (2, 0),
    (3, 2),
    (3, 4),
    (4, 3),
  ],
  directed = True,
  vertex_attrs = {
    'name': list('ABCDE'),
    'capacity': [10, 12, 20, 10, 3],
  },
  edge_attrs = {
    'weight': [
      30, # (0, 1),
      5, # (0, 3),
      5, # (1, 2),
      10, # (1, 3),
      15, # (2, 0),
      12, # (3, 2),
      6, # (3, 4),
      1, # (4, 3),
    ]
  }
)

G.write_graphml("small_case/graph.graphml")

fig, ax = plt.subplots()
ig.plot(
  G,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=G.es['weight'],
  edge_background='white',
  edge_label_dist=0.7,
)
```

We compute Q, the transition matrix.
We extract T, the base transition matrix, by allowing defaults of using the vertex indices as the ordering, since we have the sizes attached to the graph.
```{python}
T = ntc.transition_matrix_from_graph(
  graph=G,
  scaling_per_node=G.vs['capacity'],
  global_scaling=28,
  adjacency_attribute='weight'
)
Q = hmk._Q_mat_naive(T)
```

This is our ground truth, the hitting time for N=1, as extracted from the linear system that results.
```{python}
def draw_ecdf(values, ax=None, **kwargs):
  value_arr = np.sort(values.flatten())
  ecdf_quantiles = np.linspace(0, 1, len(value_arr), endpoint=False)

  if ax is None:
    fig, ax = plt.subplots()
  ax.step(value_arr, ecdf_quantiles, **kwargs)
  return ax
```

```{python}
hitting_time_one,_ = ht.compute_static_hitting_times(Q)
draw_ecdf(hitting_time_one)
```

Here we use a scaling=100 to get proper convergence of the integrals. This value is relatively arbitrary, but it must be >1 since there are slow species that ruin our time.
```{python}
hitting_time_numerical = mch.compute_all_multichain_hittings(1, Q=Q, scaling=100)
ax = draw_ecdf(hitting_time_numerical)
```

These are our numerical base results. We "hope" they are correct (insofar as numerical error)
```{python}
ht_numers = {
  n_chains: mch.compute_all_multichain_hittings(n_chains, Q=Q, scaling=100)
  for n_chains in [1, 5, 10, 30]
}

fig, ax = plt.subplots()
for k, v in ht_numers.items():
  draw_ecdf(v, ax=ax, label=f"N={k}")
plt.legend()
```

We can then simulate on this naive static network.
We will offload this work to another script, mostly because the simulations take a significant amount of time.

![Script]()

### 2025-02-19 hitting_time_multichain.py

@@ -46,13 +46,13 @@ def solve_survival_ode(Q, index=-1, locations=None, endpoint=200):
     )
 
 
-def compute_multichain_hitting_time(Nchains, Q, index=-1, precision=30):
+def compute_multichain_hitting_time(Nchains, Q, index=-1, precision=30, scaling=1.0):
     roots, weights = determine_laguerre_roots_and_weights(precision)
     gl_weights = weights * np.exp(roots)
 
-    survival_sol = solve_survival_ode(Q, index=index, locations=roots)
+    survival_sol = solve_survival_ode(Q, index=index, locations=(roots * scaling))
 
-    return np.dot(gl_weights, survival_sol.y.T**Nchains)
+    return scaling * np.dot(gl_weights, survival_sol.y.T**Nchains)
 
 
 def conform_compacted_array(hitting_arr, idx):
@@ -68,10 +68,15 @@ def stitch_hitting_arrays(hitting_arrs: Iterable):
         [conform_compacted_array(arr, idx) for idx, arr in enumerate(hitting_arrs)]
     ).T
 
-def compute_all_multichain_hittings(Nchains, Q, precision=30):
+def compute_all_multichain_hittings(Nchains, Q, precision=30, scaling=1.0):
     return stitch_hitting_arrays(
         [
-            compute_multichain_hitting_time(Nchains, Q, index=idx, precision=precision)
+            compute_multichain_hitting_time(
+                Nchains, 
+                Q, index=idx, 
+                precision=precision, 
+                scaling=scaling
+            )
             for idx in range(Q.shape[0])
         ]
     )

### 2025-02-24 network_conversion.py
Fix transition probability bug in naive static

@@ -513,7 +513,7 @@ class NaiveStaticConverter(Converter):
         observed_departure_rate = self.transition_matrix.sum(axis=1).reshape((-1, 1))
         mapped_parameters["gamma"] = observed_departure_rate / (1 - p)
 
-        mapped_parameters["transition_matrix"] = self.transition_matrix
+        mapped_parameters["transition_matrix"] = self.transition_matrix / mapped_parameters['gamma']
 
         return mapped_parameters
 

### 2025-02-24 small_case.qmd
Small case: add sim results

@@
 import igraph as ig
+import polars as pl
 
@@
 
-from matplotlib import pyplot as plt
+from matplotlib import pyplot as plt, colormaps
+import seaborn as sns
 ```
@@
     'name': list('ABCDE'),
-    'capacity': [10, 12, 20, 10, 3],
+    'capacity': [100, 120, 200, 100, 30],
   },
@@
     'weight': [
-      30, # (0, 1),
-      5, # (0, 3),
-      5, # (1, 2),
-      10, # (1, 3),
-      15, # (2, 0),
-      12, # (3, 2),
-      6, # (3, 4),
-      1, # (4, 3),
+      300, # (0, 1),
+      50, # (0, 3),
+      50, # (1, 2),
+      100, # (1, 3),
+      150, # (2, 0),
+      120, # (3, 2),
+      60, # (3, 4),
+      10, # (4, 3),
     ]
@@
   edge_background='white',
-  edge_label_dist=0.7,
+  edge_label_dist=1.7,
 )
@@
 ```{python}
+TIMESPAN = 28
 T = ntc.transition_matrix_from_graph(
@@
   scaling_per_node=G.vs['capacity'],
-  global_scaling=28,
+  global_scaling=TIMESPAN,
   adjacency_attribute='weight'
@@
 We can then simulate on this naive static network.
-We will offload this work to another script, mostly because the simulations take a significant amount of time.
+We will offload this work to another script, mostly because the simulations take a significant amount of time (and we only want to run it once)
 
-![Script]()
+[script](small_case_sims.py)
+
+Similarly, we offload the construction of the metrics.
+
+[script](small_case_sim_analysis.py)
+
+Here, we construct the empirical CDF
+```{python}
+metrics = pl.read_parquet("small_case/metrics.parquet")
+hitting_time_cols = pl.selectors.starts_with('hitting_time_')
+aggregated_metrics = metrics.unpivot(
+  on=hitting_time_cols,
+  index=['nchains','seed',],
+  variable_name='target',
+  value_name='hitting_time',
+)
+max_value_plus_one = aggregated_metrics.select(pl.col('hitting_time').max()).item() + 1
+cleaned_metrics = aggregated_metrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)
+
+ax = sns.ecdfplot(cleaned_metrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))
+
+for k, v in ht_numers.items():
+  draw_ecdf(v, ax=ax, label=f"N={k}")
+
+ax.set_xscale('log')
+ax.set_xlim(left=0.3, right=3e3)
+```
+
+We see relatively good agreement here.
+<!-- We detected a bug in this sequence, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node.
+-->

### 2025-02-24 small_case_sim_analysis.py

@@ -1,0 +1,7 @@
+import netsim_summariser as summ
+
+sim_archive = "small_case/simulations.h5"
+summariser = summ.Summariser(sim_archive)
+result_df = summariser.metrics(ncpus=6, no_move=True, drop=[])
+export_file = "small_case/metrics.parquet"
+result_df.write_parquet(export_file)

### 2025-02-24 small_case_sims.py

@@ -1,0 +1,43 @@
+import datetime
+import numpy as np
+import igraph as ig
+import network_conversion as ntc
+import network_simulation as nts
+
+graph = ig.Graph.Read_GraphML("small_case/graph.graphml")
+ordering = ntc.Ordering(
+    {node['name']: node['capacity'] for node in graph.vs}
+)
+converter = ntc.NaiveStaticConverter(graph, ordering=ordering, time_span=28)
+
+prob_no_escape = np.zeros(len(ordering))
+parameters = converter.map_parameters(
+    {
+        'beta': 0.0,
+        'prob_final_stay': prob_no_escape,
+    }
+)
+
+full_sizes = 100_000 * np.ones_like(converter.ordering.sizes, dtype=int)
+
+simulation = nts.Simulation(full_size=full_sizes, parameters=parameters, dt=0.5)
+simulation.seed(n_seed_events=0, n_seed_number=0)
+
+for nchains in [1, 5, 10, 30]:
+    for rep in range(N_REPS:= 100):
+        for i,v in enumerate(converter.ordering.sizes):
+            
+            simulation.reset(soft=False)
+            simulation.state[i] = nchains
+
+            now = datetime.datetime.now()
+            simdate = now.strftime('%y%m%d')
+            simtime = now.strftime('%H%M%S')
+            sim_index = i*N_REPS + rep
+            simid = f"{simdate}_{simtime}_{sim_index}"
+            outname = "small_case/simulations.h5"
+
+            print(simid, end=' ', flush=True)
+
+            simulation.simulate(until = 600, nostop=True)
+            simulation.export_history(outname, identity=simid, seed=i, simdate=simdate, simtime=simtime, nchains=nchains)

### 2025-02-24 network_conversion.py
add a bad hack to fix isolated nodes for naive static

@@ -54,16 +54,18 @@ class Ordering:
         """returns the size of the object"""
         idx = self[key]
         return self.sizes[idx]
-    
+
     def todf(self):
         """Return the dataframe that corresponds to the ordering
         Has columns 'hospital', 'index', and 'size'
         """
-        return pl.from_dict({
-            'hospital': self.order,
-            'index': np.arange(len(self.order)),
-            'size': self.sizes
-        })
+        return pl.from_dict(
+            {
+                "hospital": self.order,
+                "index": np.arange(len(self.order)),
+                "size": self.sizes,
+            }
+        )
 
 
 class ColumnDict(dict):
@@ -136,7 +138,10 @@ class Converter(abc.ABC):
 
 class TemporalNetworkConverter(Converter):
     def __init__(
-        self, network: ig.Graph, ordering: Ordering, weight: str='weight',
+        self,
+        network: ig.Graph,
+        ordering: Ordering,
+        weight: str = "weight",
     ):
         """Extracts transition matrix from a network"""
         self.ordering = ordering
@@ -178,7 +183,7 @@ class TemporalNetworkConverter(Converter):
     @property
     def DIMENSIONS(self):
         return {"NLOC": self.NLOC, "NT": self.NT}
-    
+
     @classmethod
     def from_file(cls, network_filepath: str | PathLike, **kwargs):
         graph = gim.make_graph(network_filepath)
@@ -193,7 +198,9 @@ class TemporalNetworkConverter(Converter):
         mapped_parameters = dict()
 
         mapped_parameters["beta"] = parameters["beta"]
-        mapped_parameters["prob_final_stay"] = parameters["prob_final_stay"].reshape((-1, 1))
+        mapped_parameters["prob_final_stay"] = parameters["prob_final_stay"].reshape(
+            (-1, 1)
+        )
 
         movements_out = self.A.sum(axis=1)
         M_mat = movements_out.reshape((self.NLOC, self.NT), order="F")
@@ -205,14 +212,14 @@ class TemporalNetworkConverter(Converter):
         ).tocsr()
 
         return mapped_parameters
-    
+
     def delay(self, n: int):
         """Reformats internal attributes so that the starting time is delayed by n time steps"""
 
         self.NT -= n
 
         nidx = int(n * self.NLOC)
-        self.A = self.A[nidx:, nidx:] # CSR sparse -> CSR sparse
+        self.A = self.A[nidx:, nidx:]  # CSR sparse -> CSR sparse
 
 
[…]

### 2025-02-25 small_case.qmd
Update some progress

@@
 
-ax = sns.ecdfplot(cleaned_metrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))
+fig, axz = plt.subplots(nrows=2)
+axs = axz.flatten()
+
+for ax in axs:
+  sns.ecdfplot(
+    cleaned_metrics, 
+    x='hitting_time', 
+    hue='nchains', 
+    palette=colormaps.get('jet'),
+    ax=ax
+  )
 
@@
 
-ax.set_xscale('log')
-ax.set_xlim(left=0.3, right=3e3)
+axs[1].set_xlim(-10, 100)
+
 ```
@@
 We see relatively good agreement here.
-<!-- We detected a bug in this sequence, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node.
--->
+
+### Aside: the Bug
+
+We detected a bug in this work, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node.
+
+This causes the ECDFs to clump, and be differentiated by their hitting probabilities.
+
+```{python}
+bug_metrics = pl.read_parquet("small_case/metrics_prebug.parquet")
+hitting_time_cols = pl.selectors.starts_with('hitting_time_')
+aggregated_bugmetrics = bug_metrics.unpivot(
+  on=hitting_time_cols,
+  index=['nchains','seed',],
+  variable_name='target',
+  value_name='hitting_time',
+)
+cleaned_bugmetrics = aggregated_bugmetrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)
+
+ax = sns.ecdfplot(cleaned_bugmetrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))
+```

### 2025-03-07 snapshot_return_policy.qmd
thinking about return time for snapshots

---
title: Examining whether the snapshot return mechanism is correct
---

```{python}
import numpy as np
from matplotlib import pyplot as plt
from scipy import stats
from collections import Counter

rng = np.random.default_rng()
```

We want to investigate whether the snapshot return mechanism produces the correct return time distribution.

Currently, because we do not label individuals, all individuals in the same state are treated the same.
This means that individuals in the "at home" bucket are all treated as if they have been there for some time.

We model return (and the overall return rate) in the snapshot model by balancing the number of observed departures (to home) and arrivals (from home).

We can check the concordance of this model with the temporal distribution by writing an individual based-simulation (or adapt our current simulation, where we now also  track marked individuals pretty explicitly)

We will do this by modelling a single link (single we don't care about network effects), and sending individuals through this link indirectly.

```{python}
SNAPSHOT_DURATION = 14
times = np.arange(0, 100)
readmission_duration = lambda size=1: rng.lognormal(np.log(14), 0.5, size=size)

base_population = 1000
base_move_out = 200
```

```{python}
movements_out = np.random.poisson(base_move_out, size=times.shape)

movements_in = np.zeros_like(times)
movement_dist = []
for _t in times: # implictly this is also the index
    move_times = np.round(readmission_duration(movements_out[_t]), 0).astype(int) + _t
    valid_moves = move_times[move_times < np.max(times)]
    movement_dist.extend(valid_moves)
    if len(valid_moves) < 1:
        continue
    move_counter = Counter(valid_moves)
    move_tos = np.array(list(move_counter.keys()))
    move_counts = np.array(list(move_counter.values()))
    movements_in[move_tos] += move_counts
```

Computing the transition matrices
```{python}
at_home = 0
home_numbers = np.zeros_like(times)
for _t in times:
    at_home += movements_out[_t]
    home_numbers[_t] = at_home
    at_home -= movements_in[_t]

inwards_transition_prob = movements_in / home_numbers

outwards_transition_prob = movements_out / base_population
```

```{python}
# create agents
class Agent():
    _place_map = ['hospital', 'home']
    def __init__(self):
        self.place = 0
        self.go_home_time = np.nan
        self.return_hospital_time = np.nan

    def __repr__(self) -> str:
        return f"Agent()"

    def __str__(self):
        return f"Agent [{self.place} {self.go_home_time:.0f} {self.return_hospital_time:.0f}]"

    @property
    def duration(self):
        return self.return_hospital_time - self.go_home_time

    def go_home(self, time):
        self.go_home_time = time
        self.place = 1

    def return_hospital(self, time):
        self.return_hospital_time = time
        self.place = 0

agent_pool = set()
home_pool = set()

### 2025-03-13 link_time_regr.qmd
update some stuff that's been done to analyse the prlbems

---
title: link_time regression
---

```{python}
import igraph as ig
from matplotlib import pyplot as plt
from scipy import stats, optimize
import numpy as np
```

```{python}
B = ig.Graph.Read("concordant_networks/trimmed_static_base_1_threshold_4.graphml")

direct = np.array(B.es['direct_weight'])
indirect = np.array(B.es['indirect_weight'])
link_time = np.array(B.es['link_time'])

plt.loglog(indirect, link_time, '.')
plt.xlabel('Indirect weight')
plt.ylabel('Link Time')
```

```{python}
valid = np.isfinite(link_time)

# transfrom
log_link_time = np.log(link_time[valid])
log_indirect = np.log(indirect[valid])

# estim mean
log_link_mean = np.mean(log_link_time)

residuals = log_link_time - log_link_mean
composed_variable = log_indirect * residuals

est_mean = np.mean(composed_variable)
est_std = np.std(composed_variable)

plt.plot(log_indirect, composed_variable, '.')
```

```{python}
rng = np.random.default_rng()

indirect_values = np.array([250, 30, 40, 70, 110, 100, 40, 5])
composed_draws = rng.normal(est_mean, est_std, size=indirect_values.shape)
drawn_link_times = np.exp(
    (composed_draws / np.log(indirect_values)) + log_link_mean
) / 12 # for time unit consistency

print(drawn_link_times)
```

### 2025-03-13 sim_movement_over_time.qmd
update some stuff that's been done to analyse the prlbems

---
title: Movement in Simulations over Time
---

We want to see how much movement there is between home and hospital in each simulation.

```{python}
import h5py
import polars as pl
import numpy as np
from matplotlib import pyplot as plt

rng = np.random.default_rng(2025)
```

```{python}
fh = h5py.File("zero_sims/snapshot/sims_with_30seeds_pseudocap_fixedaxis.h5")
fh_grps = list(fh)

test_index = rng.choice(fh_grps)
test_grp = fh[test_index]

data = {
    k: np.array(v[:], dtype=float)
    for k,v in test_grp.items()
}
```

```{python}
def group_window(arr: np.ndarray, W: int):
    return arr.reshape((-1, W)).sum(axis=1)

WINDOW = 10

plt.plot(data['ts'][:-1:WINDOW], group_window(data['mover_out'].sum(axis=1), WINDOW))
plt.plot(data['ts'][:-1:WINDOW], group_window(data['mover_in'].sum(axis=1), WINDOW))
```

```{python}
fh_static = h5py.File("zero_sims/static/sim_all_30s_fixed_axis.h5")
xi_static = rng.choice(list(fh_static))
data_static = {
    k: np.array(v[:], dtype=float) 
    for k,v in fh_static[xi_static].items()
}

WINDOW = 10

plt.plot(
    data_static['ts'][:-1:WINDOW], 
    group_window(data_static['mover_out'].sum(axis=1), WINDOW)
)

plt.plot(
    data_static['ts'][:-1:WINDOW], 
    group_window(data_static['mover_in'].sum(axis=1), WINDOW)
)
```

... want to look at the departure rate of the first snapshot, and see the distrigubtion. Looks very sparse for snapshot 0. Of 4299 edges, we get 4239 edges with weight < 19, 469 with 0 departure (and thus only direct movement). Note that there are 338^2 possible edges.

```{python}
import igraph as ig

G = ig.Graph.Read("conc_tempo_14_detailed/0000.graphml")
S = ig.Graph.Read("concordant_networks/trimmed_static_base_1_threshold_4.graphml")
```

```{python}
snapshot_df = pl.from_dict(
    {
        'departures': G.es['departures'],
        'arrivals': G.es['arrivals'],
        'source': G.es['source'],
        'target': G.es['target'],
    }
).cast(pl.Int64)

static_df = pl.from_dict(
    {
        'indirect': S.es['indirect_weight'],
        # 'link_time': S.es['link_time'],
        'source': S.es['source_loc'],
        'target': S.es['dest_loc'],
    }
).cast(pl.Int64)

size_df = pl.read_csv("concordant_networks/size_14.csv")

snapshot_df = snapshot_df.join(size_df, left_on='source', right_on='hospital')
static_df = static_df.join(size_df, left_on='source', right_on='hospital')

snapshot_df = snapshot_df.with_columns(dep_rate = pl.col('departures').cast(pl.Float64) / pl.col('estimated_size') / 14)
static_df = static_df.with_columns(dep_rate = pl.col('indirect').cast(pl.Float64) / pl.col('estimated_size') / S['time_span'])

dep_rates = pl.concat([
    snapshot_df
        .with_columns(model=pl.lit('snapshot'))
        .select('dep_rate', 'source', 'target', 'model'), 
    static_df
        .with_columns(model=pl.lit('static'))
        .select('dep_rate', 'source', 'target', 'model')
])

```
```{python}
hist_opts = {
    'histtype': 'step',
    'bins': 31,
    'density': True
}
for name, arr in zip(['snapshot', 'static'], [snapshot_moves, static_moves]):
    plt.hist(arr, **hist_opts, label=name)

plt.yscale('log')
plt.legend()

### 2025-03-13 network_conversion.py

@@ -397,6 +397,7 @@ class StaticConverter(Converter):
         ordering: Ordering,
         time_span: SupportsFloat | None = None,
         purge_selfloops=False,
+        ordering_key="node",
     ):
         self.ordering = ordering
 
@@ -413,7 +414,7 @@ class StaticConverter(Converter):
                 ordering=self.ordering,
                 scaling_per_node=self.ordering.sizes,
                 global_scaling=self.time_span,
-                ordering_key="node",
+                ordering_key=ordering_key,
                 adjacency_attribute=adjkey,
                 matrix_size=len(self.ordering.order),
             )
@@ -422,7 +423,7 @@ class StaticConverter(Converter):
         self.raw_transition_matrices["link_time"] = transition_matrix_from_graph(
             graph=network,
             ordering=self.ordering,
-            ordering_key="node",
+            ordering_key=ordering_key,
             adjacency_attribute="link_time",
         )
 

### 2025-03-20 hitting_time_analysis.py
Some hitting time separation of concerns

@@ -1,3 +1,4 @@
+from typing import Mapping
 import numpy as np
 import igraph as ig
 from matplotlib import pyplot as plt, colors
@@ -7,16 +8,25 @@ import network_conversion as ntc
 import hitting_markov as hit
 from realisation_summary_stats import hitting_time
 
+def graph_read(filename: str):
+    return ig.Graph.Read(filename)
+
 ## NAIVE STATIC MODEL
 def get_naive_static_Q(
-    graph_file = "./concordant_networks/shuf_static_network.graphml",
+    graph: str | ig.Graph = "./concordant_networks/shuf_static_network.graphml",
     size_file="./concordant_networks/size_14.csv",
     time_span = (10*365),
+    ordering_key = 'name',
+    value_key = 'id',
 ):
+    """input graph is either the path name or the grpah object"""
 
-    G = ig.Graph.Read(graph_file)
+    if isinstance(graph, ig.GraphBase):
+        G = graph
+    else:
+        G = ig.Graph.Read(graph)
 
-    graph_map = {int(k):v for k,v in zip(G.vs['name'], G.vs['id'])}
+    graph_map = {int(k):v for k,v in zip(G.vs[ordering_key], G.vs[value_key])}
     sorted_graph_map = {k: graph_map[k] for k in sorted(graph_map.keys())}
     graph_ordering = ntc.Ordering(sorted_graph_map)
 
@@ -31,7 +41,7 @@ def get_naive_static_Q(
                                         ordering=graph_ordering,
                                         scaling_per_node=sizes,
                                         global_scaling=time_span,
-                                        ordering_key='name', 
+                                        ordering_key=ordering_key, 
                                         adjacency_attribute='weight',
                                         )
     Q = hit._Q_mat_naive(T)
@@ -67,35 +77,55 @@ def compute_static_hitting_times(
 
 
 ### STATIC MODEL WITH DELAYED RETURN
-def compute_delayed_static_hitting_times(
-    graph_file = "./concordant_networks/trimmed_static_base_1_threshold_4.graphml",
+
+
+def extract_delayed_static_Qmat(
+    graph: str|ig.Graph = "./concordant_networks/trimmed_static_base_1_threshold_4.graphml",
+    size_mapping: str|Mapping = "./concordant_networks/size_14.csv",
     scaling = (10*365),
     eps = 1e-7,
-    plot = False,
+    ordering_key = 'node', 
[…]

### 2025-03-20 hitting_time_multichain.py

@@ -68,6 +68,23 @@ def stitch_hitting_arrays(hitting_arrs: Iterable):
         [conform_compacted_array(arr, idx) for idx, arr in enumerate(hitting_arrs)]
     ).T
 
+def compute_subset_multichain_hittings(Nchains, Q, index_subset=slice(None), precision=30, scaling=1.0):
+    if isinstance(index_subset, slice):
+        index_range = range(*index_subset.indices(Q.shape[0]))
+    else:
+        index_range = index_subset
+    return stitch_hitting_arrays(
+        [
+            compute_multichain_hitting_time(
+                Nchains,
+                Q, index=idx,
+                precision=precision,
+                scaling=scaling,
+            )[index_subset]
+            for idx in index_range 
+        ]
+    )
+
 def compute_all_multichain_hittings(Nchains, Q, precision=30, scaling=1.0):
     return stitch_hitting_arrays(
         [

### 2025-03-20 network_simulation.py
Move locaiton of rgn creation; add naive  snapshot (no home)

@@ -20,6 +20,8 @@ class Simulation:
             self.state
         ]  # softref here should update as initial self.state updates in seed
 
+        self.rng = np.random.default_rng(None)
+
     @property
     def history(self):
         return np.hstack(self._history)
@@ -45,11 +47,21 @@ class Simulation:
         Does not reset the initial state.
         The rng seed can be specified.
         """
-        self.rng = np.random.default_rng(rng_seed)
+        if rng_seed is not None:
+            self.rng = np.random.default_rng(rng_seed)
 
         for _ in range(n_seed_events):
             location = int(self.rng.uniform(0, len(self.state)))
             self.state[location, 0] += n_seed_number
+
+        self.state = np.clip(self.state, 0, self.N)
+
+    def fixedseed(self, location, number_to_seed=1, rng_seed=None):
+        
+        if rng_seed is not None:
+            self.rng = np.random.default_rng(rng_seed)
+
+        self.state[location, 0] += number_to_seed
 
         self.state = np.clip(self.state, 0, self.N)
 
@@ -376,3 +388,41 @@ class TemporalSim(SimulationWithMovers):
         super().reset(soft=soft)
 
         self.time_travellers = np.zeros_like(self.time_travellers, dtype=np.int64)
+
+class SnapshotNaive(Simulation):
+    def __init__(self, full_size: Sequence[int], parameters: Mapping, timings: Sequence, dt=1.0):
+        super().__init__(
+            full_size=full_size,
+            parameters=parameters,
+            dt=dt,
+        )
+
+        self.current_index = 0
+        self.timings = timings
+
+    def swapover(self):
+        # assume current index is correctly set
+        self.parameters['transition_matrix'] = self.parameters['transition_matrices'][self.current_index]
+
+        self.parameters['gamma'] = self.parameters['gammas'][self.current_index]
+
+    def reset(self, soft=True):
+        super().reset(soft=soft)
+        self.current_index = 0
+
+    def simulate(self, until=100, nostop=False):
+        self.swapover()
+        
+        for ti in range(int(until / self.dt)):
+            
+            self.state = self.step()
+            self._history.append(self.state)
+            self.ts.append(self.ts[-1] + self.dt)
+
+            if self.ts[-1] >= self.timings[self.current_index+1]:
+                self.current_index += 1
+                self.swapover()
+
+            if not nostop and np.sum(self.state) < 1:
+                print(f"Early termination: {self.ts[-1] =}")
+                break

### 2025-03-20 sim_movement_over_time.qmd
more investiagtions to network structure

@@
 from matplotlib import pyplot as plt
+import seaborn as sns
 
@@
 import igraph as ig
+from collections import Counter
 
@@
         'indirect': S.es['indirect_weight'],
-        # 'link_time': S.es['link_time'],
+        'direct': S.es['direct_weight'],
         'source': S.es['source_loc'],
@@
 ])
+```
 
+```{python}
+sns.histplot(
+    dep_rates.group_by('source', 'model').agg(pl.col('dep_rate').sum()),
+    x='dep_rate',
+    hue='model',
+)
 ```
+
+I have a hunch about this. Let's construct the "combined" graph of the snapshots over time, and see when it becomes conencted.
+This will be the minimum time before we can hit equilibirum.
 ```{python}
-hist_opts = {
-    'histtype': 'step',
-    'bins': 31,
-    'density': True
-}
-for name, arr in zip(['snapshot', 'static'], [snapshot_moves, static_moves]):
-    plt.hist(arr, **hist_opts, label=name)
+NO_CACHE = False
+connectedness_file = "conc_tempo_14_detailed/analysis/overlay_connectedness.parquet"
 
+if NO_CACHE:
+    Gs = [
+        ig.Graph.Read(f"conc_tempo_14_detailed/{i:04d}.graphml")
+        for i in range(0, 3626+14, 14)
+    ]
+
+    # remove problem vertex (isolated hospital)
+    problem_vertex = 4454
+    for graph in Gs:
+        graph.delete_vertices(graph.vs.select(name_eq=problem_vertex).indices)
+
+    connectedness = []
+    cluster_sizes = []
+    for i in range(len(Gs)):
+        U = ig.union(Gs[:i+1])
+        connectedness.append(U.is_connected())
+        cluster_sizes.append(Counter(map(len, U.connected_components())))
+
+    df = pl.from_dicts([
+        {'snapshot': i, 'fully_connected': full_conn, 'cluster_size': k, 'count': v} 
+        for i, (full_conn, clus_counter) in enumerate(zip(connectedness, cluster_sizes)) 
+        for k,v in clus_counter.items()
+    ])
+
+    df.write_parquet(connectedness_file)
+else:
+    df = pl.read_parquet(connectedness_file)
+```
+
+```{python}
+fig, axs = plt.subplot_mosaic([['c'], ['c'], ['b']], sharex=True, layout='constrained')
+
+cluster_stats = df.group_by('snapshot').agg(
+    pl.col('cluster_size').max().alias('dominant_cluster_size'),
+    pl.col('count').sum().alias('number_of_clusters')
+).with_columns(time = pl.col('snapshot') * 14).sort('snapshot')
+
+sns.lineplot(cluster_stats, x='time', y='dominant_cluster_size', ax=axs['c'])
+sns.lineplot(cluster_stats, x='time', y='number_of_clusters', ax=axs['b'])
+```
+
+## Rare path exploration
+
+We want to see if we can come up with a way of detecting and analysing "rare" paths.
+Specifically rare 1-paths (potentially of the indirect flavour).
+
+To do this, we will want to tabulate edges of the snapshot against the static edge weights
+
+```{python}
+# static ref table already extracted in earlier cell
+
+NO_CACHE = False
+
+snapshot_table_file = "conc_tempo_14_detailed/analysis/combined_outwards.parquet"
+
+if NO_CACHE:
+    snapshot_graph_tables = []
+    for i in range(0, 3626+14, 14):
+        graph = ig.Graph.Read(f"conc_tempo_14_detailed/{i:04d}.graphml")
+        problem_vertex = 4454
+        graph.delete_vertices(graph.vs.select(name_eq=problem_vertex).indices)
+        snapshot_graph_tables.append(
+            (pl.from_dict({
+                'source': graph.es['source'],
+                'target': graph.es['target'],
+                'snapshot_direct': graph.es['weight'],
+                'snapshot_indirect': graph.es['departures'],
+            })
+            .cast({
+                'source': pl.Int64,
+                'target': pl.Int64,
+                'snapshot_direct': pl.Int64,
+                'snapshot_indirect': pl.Int64,
+            })
+            .with_columns(time=pl.lit(i))
+            )
+        )
+    snapshot_table = pl.concat(snapshot_graph_tables)
+    snapshot_table.write_parquet(snapshot_table_file)
+else:
+    snapshot_table = pl.read_parquet(snapshot_table_file)
+
+joined_table = snapshot_table.join(static_df, on=['source', 'target'])
+
+SS_WIDTH = 14
+
+joined_table = (
+    joined_table
+    .with_columns(
+        ss_direct_rate = pl.col('snapshot_direct').cast(pl.Float64) / pl.col('estimated_size') / SS_WIDTH,
+        ss_indirect_rate = pl.col('snapshot_indirect').cast(pl.Float64) / pl.col('estimated_size') / SS_WIDTH,
+        st_direct_rate = pl.col('direct').cast(pl.Float64) / pl.col('estimated_size') / S['time_span'],
+    )
+)
+```
+
+```{python}
+devs = joined_table.filter(pl.col('dep_rate') > 0.0).select(pl.col('ss_indirect_rate') - pl.col('dep_rate')).to_series().to_numpy()
+plt.hist(devs, bins=301)
 plt.yscale('log')
-plt.legend()
+plt.axvline(3 * np.sqrt(np.mean(devs**2)), color='red')
+plt.axvline(-3*np.sqrt(np.mean(devs**2)), color='red')
+```

### 2025-03-20 small_case_sims.py
A small case to detect anomaly in naive static

@@ -4,40 +4,93 @@ import igraph as ig
 import network_conversion as ntc
 import network_simulation as nts
 
-graph = ig.Graph.Read_GraphML("small_case/graph.graphml")
-ordering = ntc.Ordering(
-    {node['name']: node['capacity'] for node in graph.vs}
-)
-converter = ntc.NaiveStaticConverter(graph, ordering=ordering, time_span=28)
 
-prob_no_escape = np.zeros(len(ordering))
-parameters = converter.map_parameters(
-    {
-        'beta': 0.0,
-        'prob_final_stay': prob_no_escape,
-    }
-)
+def run_sims(
+    simulation,
+    n_size,
+    n_reps = 30,
+    outname = "temp_sims.h5"
+):
+    # assume simulation is already initialised via a .seed
 
-full_sizes = 100_000 * np.ones_like(converter.ordering.sizes, dtype=int)
+    for nchains in [1, 5, 10, 30]:
+        for rep in range(n_reps):
+            for i in range(n_size):
+                simulation.reset(soft=False)
+                simulation.state[i] = nchains
 
-simulation = nts.Simulation(full_size=full_sizes, parameters=parameters, dt=0.5)
-simulation.seed(n_seed_events=0, n_seed_number=0)
+                now = datetime.datetime.now()
+                simdate = now.strftime('%y%m%d')
+                simtime = now.strftime('%H%M%S')
+                sim_index = i*n_reps + rep
+                simid = f"{simdate}_{simtime}_{sim_index}"
 
-for nchains in [1, 5, 10, 30]:
-    for rep in range(N_REPS:= 100):
-        for i,v in enumerate(converter.ordering.sizes):
-            
-            simulation.reset(soft=False)
-            simulation.state[i] = nchains
+                print(simid, end=' ', flush=True)
 
-            now = datetime.datetime.now()
-            simdate = now.strftime('%y%m%d')
-            simtime = now.strftime('%H%M%S')
-            sim_index = i*N_REPS + rep
-            simid = f"{simdate}_{simtime}_{sim_index}"
-            outname = "small_case/simulations.h5"
+                simulation.simulate(until = 600, nostop=True)
+                simulation.export_history(outname, identity=simid, seed=i, simdate=simdate, simtime=simtime, nchains=nchains)
 
-            print(simid, end=' ', flush=True)
 
-            simulation.simulate(until = 600, nostop=True)
[…]

### 2025-03-25 rare_path.qmd
Add direct rare pathing experiment

---
title: Rare Pathing
---

## Direct rare pathing
```{python}
import igraph as ig
import numpy as np
import network_simulation as netsim

from matplotlib import pyplot as plt
import seaborn as sns
import polars as pl
```

```{python}

G = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        # (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        1000,
        1000,
        1000,
        # 0,
    ]}
)

R = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        0,
        0,
        0,
        100,
    ]}
)
```

```{python}
N_size = np.array([3000, 3000, 3000, 100]).reshape((-1, 1))

T_G = np.array(G.get_adjacency('weight').data) / N_size

T_R = np.array(R.get_adjacency('weight').data) / N_size

T_stack = np.stack([T_G[:,:,np.newaxis], T_R[:,:,np.newaxis], *[T_G[:,:,np.newaxis] for _ in range(998)]], axis=2)

T_ave = np.squeeze(np.mean(T_stack, axis=2))
```

```{python}
## STATIC
step_size = 0.2
gamma_static = T_ave.sum(axis=1, keepdims=True) # (-1,1)
W_ave = T_ave / gamma_static
W_ave[np.isnan(W_ave)] = 0.0

static_sim = netsim.Simulation(
    full_size = N_size,
    parameters = {
        'beta': 0.0,
        'gamma': gamma_static,
        'prob_final_stay': np.zeros_like(gamma_static),
        'transition_matrix': W_ave,
    },
    dt = step_size,
)

static_sim.fixedseed(0, 1000)
```

```{python}
## SNAPSHOT
gamma_normal = T_G.sum(axis=1, keepdims=True)
gamma_rare = T_R.sum(axis=1, keepdims=True)

W_normal = T_G / gamma_normal
W_rare = T_R / gamma_rare

W_normal[np.isnan(W_normal)] = 0.0
W_rare[np.isnan(W_rare)] = 0.0

snapshot_sim = netsim.SnapshotNaive(
    N_size, 
    parameters={
        'beta': 0.0,
        'transition_matrix': W_normal,
        'gamma': gamma_normal,
        'prob_final_stay': np.zeros_like(gamma_normal),
        'transition_matrices': [W_normal, W_rare, W_normal],
        'gammas': [gamma_normal, gamma_rare, gamma_normal],
    },
    timings=[0, 5, 8],
    dt=step_size,
)

snapshot_sim.fixedseed(0, 1000)
```

```{python}
N_SIMS = 1000

histories = []
for it_ in range(N_SIMS):
    snapshot_sim.reset(soft=True)
    snapshot_sim.simulate(30, nostop=True)
    histories.append(np.array(snapshot_sim.history))
```

## Indirect rare pathing

```{python}
S = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        # (0, 3),
    ], 
    directed=True, 
    edge_attrs={
        'weight': [
            1000,
            1000,
            1000,
            # 0,
        ]
    }
)
```

### 2025-03-27 experiments/small_case/small_case.qmd
Big reshuffle into a more nested directory structure
This is to improve the navgiability and allow for
a more cohesive sharing of code between different parts
of the analysis

---
title: "Small Case: Checking Hitting Times in the Year of Our Lord"
format:
  html:
    code-fold: true
---

# Checking hitting Times: Naive Static Network
```{python Imports} 
import numpy as np
import networkx as nx
import igraph as ig
import polars as pl

import hitting_markov as hmk
import hitting_time_analysis as ht
import hitting_time_multichain as mch
import network_conversion as ntc
import netsim_summariser as summ

from matplotlib import pyplot as plt, colormaps
import seaborn as sns
```

Here we want to check that our simulations and analytical results have parity.

The idea is to setup a small network that has at least one very rare link, and compute the hitting times analytically for N=[1, 5, 10, 30] individuals, and also simulate for those cases.

We generate a small 5-node directed, weighted network.
```{python}
graph_parameters = dict( 
  edges = [
    (0, 1),
    (0, 3),
    (1, 2),
    (1, 3),
    (2, 0),
    (3, 2),
    (3, 4),
    (4, 3),
  ],
  directed = True,
  vertex_attrs = {
    'name': list('ABCDE'),
    'capacity': [100, 120, 200, 100, 30],
  },
  edge_attrs = {
    'weight': [
      300, # (0, 1),
      50, # (0, 3),
      50, # (1, 2),
      100, # (1, 3),
      150, # (2, 0),
      120, # (3, 2),
      60, # (3, 4),
      10, # (4, 3),
    ]
  }
)
G = ig.Graph(**graph_parameters)

G.write_graphml("small_case/graph.graphml")

fig, ax = plt.subplots()
ig.plot(
  G,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=G.es['weight'],
  edge_background='white',
  edge_label_dist=1.7,
)
```

We compute Q, the transition matrix.
We extract T, the base transition matrix, by allowing defaults of using the vertex indices as the ordering, since we have the sizes attached to the graph.
```{python}
TIMESPAN = 28
T = ntc.transition_matrix_from_graph(
  graph=G,
  scaling_per_node=G.vs['capacity'],
  global_scaling=TIMESPAN,
  adjacency_attribute='weight'
)
Q = hmk._Q_mat_naive(T)
```

This is our ground truth, the hitting time for N=1, as extracted from the linear system that results.
```{python}
def draw_ecdf(values, ax=None, **kwargs):
  value_arr = np.sort(values.flatten())
  ecdf_quantiles = np.linspace(0, 1, len(value_arr), endpoint=False)

  if ax is None:
    fig, ax = plt.subplots()
  ax.step(value_arr, ecdf_quantiles, **kwargs)
  return ax
```

```{python}
hitting_time_one,_ = ht.compute_static_hitting_times(Q)
draw_ecdf(hitting_time_one)
```

Here we use a scaling=100 to get proper convergence of the integrals. This value is relatively arbitrary, but it must be >1 since there are slow species that ruin our time.
```{python}
hitting_time_numerical = mch.compute_all_multichain_hittings(1, Q=Q, scaling=100)
ax = draw_ecdf(hitting_time_numerical)
```

These are our numerical base results. We "hope" they are correct (insofar as numerical error)
```{python}
chain_range = [1, 5, 10, 30]
ht_numers = {
  n_chains: mch.compute_all_multichain_hittings(n_chains, Q=Q, scaling=100)
  for n_chains in chain_range
}

fig, ax = plt.subplots()
for k, v in ht_numers.items():
  draw_ecdf(v, ax=ax, label=f"N={k}")
plt.legend()
```

We can then simulate on this naive static network.
We will offload this work to another script, mostly because the simulations take a significant amount of time (and we only want to run it once)

[script](small_case_sims.py)

Similarly, we offload the construction of the metrics.

```{python}
sim_archive = "small_case/simulations.h5"
summariser = summ.Summariser(sim_archive)
metrics = summariser.metrics(ncpus=6, no_move=True, drop=[])
export_file = "small_case/metrics.parquet"
metrics.write_parquet(export_file)
```

Here, we construct the empirical CDF
```{python}
# metrics = pl.read_parquet("small_case/metrics.parquet")
hitting_time_cols = pl.selectors.starts_with('hitting_time_')
aggregated_metrics = metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
max_value_plus_one = aggregated_metrics.select(pl.col('hitting_time').max()).item() + 1
cleaned_metrics = aggregated_metrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)

fig, axz = plt.subplots(nrows=2)
axs = axz.flatten()

for ax in axs:
  sns.ecdfplot(
    cleaned_metrics, 
    x='hitting_time', 
    hue='nchains', 
    palette=colormaps.get('jet'),
    ax=ax
  )

  for k, v in ht_numers.items():
    draw_ecdf(v, ax=ax, label=f"N={k}")

axs[1].set_xlim(-10, 100)

```

We see relatively good agreement here.

### Aside: the Bug

We detected a bug in this work, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node.

This causes the ECDFs to clump, and be differentiated by their hitting probabilities.

```{python}
bug_metrics = pl.read_parquet("small_case/metrics_prebug.parquet")
hitting_time_cols = pl.selectors.starts_with('hitting_time_')
aggregated_bugmetrics = bug_metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
cleaned_bugmetrics = aggregated_bugmetrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)

ax = sns.ecdfplot(cleaned_bugmetrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))
```

# Checking Hitting Times: Improved Static Network

Here, we will want to extend the construction above where we split the graph edge weights between two projections.

```{python}
bigraph_edge_attrs = {
   'direct_weight': [
      50, # (0, 1),
      20, # (0, 3),
      10, # (1, 2),
      30, # (1, 3),
      40, # (2, 0),
      20, # (3, 2),
      20, # (3, 4),
      5, # (4, 3),
    ],
    'indirect_weight': [
      250, # (0, 1),
      30, # (0, 3),
      40, # (1, 2),
      70, # (1, 3),
      110, # (2, 0),
      100, # (3, 2),
      40, # (3, 4),
      5, # (4, 3),
    ],
    # see link_time_regr.qmd
    'link_time': [
      7.52, # (0, 1),
      8.82, # (0, 3),
      20.35, # (1, 2),
      7.10, # (1, 3),
      11.51, # (2, 0),
      10.49, # (3, 2),
      10.99, # (3, 4),
      5.46, # (4, 3),
    ]
}

bigraph_parameters = dict(graph_parameters)
bigraph_parameters['edge_attrs'] = bigraph_edge_attrs

B = ig.Graph(**bigraph_parameters)
B.write_graphml("small_case/bigraph.graphml")

fig, ax = plt.subplots()
ig.plot(
  B,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=list(zip(B.es['direct_weight'], B.es['indirect_weight'], B.es['link_time'])),
  edge_background='white',
  edge_label_dist=1.7,
)
```

So we have split the graph into two layers, a direct alyer, and an indirect layer. I haven't checked the direct/indirect split in the real graph, but we will assume for now that this is an appropriate split.
We have derived link times by empirical resampling of the real graph link times, dividing by 12 to get smaller link times appropriate for a shorter global time scale, and then assigning the sampled link times by matching order of graph edge weight -> link time (higher weight gets shorter time)

We can compute the transition matrix Q for this by performing an expansion of the home states.
```{python}
B_size_map = {nd['name']: nd['capacity'] for nd in G.vs}
Qb = ht.extract_delayed_static_Qmat(
  B, 
  scaling=TIMESPAN,
  size_mapping=B_size_map,
  ordering_key='name', 
  value_key='capacity',
)
```

We can, again, compute the $N=1$ chain case.

```{python}
hitting_time_imprvst_numer = mch.compute_subset_multichain_hittings(
  Nchains = 1,
  Q = Qb.todense(),
  index_subset = slice(Q.shape[0]),
  scaling = 100,
)
ax = draw_ecdf(hitting_time_numerical, label='naive static')
ax = draw_ecdf(hitting_time_imprvst_numer, ax=ax, label='improved static')
ax.legend()
```

And the $N \in \{1, 5, 10, 30\}$ cases
```{python}
ht_imprv_numers = {
  n_chains: mch.compute_subset_multichain_hittings(
    n_chains, 
    Q=Qb.todense(), 
    index_subset=slice(Q.shape[0]), 
    scaling=100
  )
  for n_chains in chain_range
}

fig, ax = plt.subplots()
for i, (k, v) in enumerate(ht_imprv_numers.items()):
  if k == 1: continue
  draw_ecdf(v, ax=ax, label=f"N={k}", color=f"C{i}")
  draw_ecdf(ht_numers[k], ax=ax, label=f"N={k} [naive]", linestyle='dashed', color=f"C{i}")
plt.legend()
```

Let us also look at the scripts that run the simulations.

```{python}
home_sim_archive = "small_case/home_sims.h5"
home_summariser = summ.Summariser(home_sim_archive)
home_metrics = home_summariser.metrics(ncpus=6, no_move=True, drop=[])
home_export_file = "small_case/home_metrics.parquet"
home_metrics.write_parquet(home_export_file)
```

Plotting ...

```{python}
aggregated_home_metrics = home_metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
max_home_value_plus_one = aggregated_home_metrics.select(pl.col('hitting_time').max()).item() + 1
cleaned_home_metrics = aggregated_home_metrics.fill_nan(max_home_value_plus_one).fill_null(max_home_value_plus_one)

fig, axz = plt.subplots(nrows=2)
axs = axz.flatten()

for ax in axs:
  sns.ecdfplot(
    cleaned_home_metrics, 
    x='hitting_time', 
    hue='nchains', 
    palette=colormaps.get('jet'),
    ax=ax
  )

  for k, v in ht_imprv_numers.items():
    draw_ecdf(v, ax=ax, label=f"N={k}")

axs[1].set_xlim(-10, 100)
```

Comparing observed hitting time distributions between the naive and improved static network models

```{python}
fig, axs = plt.subplots(
  nrows=2, ncols=2,
  sharex=True, sharey=True,
  layout='constrained')
axz = axs.flatten()

for ax, nchain in zip(axz, chain_range):
  sns.ecdfplot(cleaned_metrics.filter(pl.col('nchains') == nchain), x='hitting_time', ax=ax, label='Naive')
  sns.ecdfplot(cleaned_home_metrics.filter(pl.col('nchains') == nchain), x='hitting_time', ax=ax, label='Improved')
  ax.set_title(f"Nchains={nchain}")
ax.legend()
```

### 2025-03-27 notebooks/compare_hitting_times.qmd
Big reshuffle into a more nested directory structure
This is to improve the navgiability and allow for
a more cohesive sharing of code between different parts
of the analysis

---
title: Comparing hitting times filtering by seed/target
---

The main aim is to construct a function that defines a quantifiable metric of whether a hitting time 
eCDF is below or above another.

This is to determine whether or not the hitting time distribution for X model is always "above" Y model; and also the extended question of whether X > Y > Z always. (or almost always)

```{python}
import numpy as np
import polars as pl
from scipy import stats
from matplotlib import pyplot as plt
import seaborn as sns
import itertools

from typing import Mapping
```

```{python}
# going to steal from aggregated_compare.ipynb

metrics_archives = {
    'temporal': './zero_sims/temporal/metrics_30s.parquet',
    'static': './zero_sims/static/metrics_30s_fa.parquet',
    'snapshot': './zero_sims/snapshot/metrics_30s_fa.parquet',
    'naive_static': './zero_sims/naive_static/metrics_30s.parquet',
}

raw_metrics = {
  label: pl.read_parquet(archive) for label, archive in metrics_archives.items()
}

hitting_time_prefix = "hitting_time_"
hitting_time_columns = pl.selectors.starts_with(hitting_time_prefix)
```

```{python}
def add_ecdf(df: pl.DataFrame, name='eCDF'):
  augmented_df = (
    df
    .sort('hitting_time', nulls_last=True)
    .with_row_index(name=name)
    .with_columns(pl.col(name) / pl.col(name).max())
  )
  return augmented_df

aggregated_metrics = dict()
for model, metrics in raw_metrics.items():
  agg_df = (
    metrics
    .unpivot(
      on=hitting_time_columns,
      index='seed',
      variable_name='target_name',
      value_name='hitting_time',
    )
    .with_columns(
      target=(
        pl.col('target_name')
        .str.strip_prefix('hitting_time_')
        .str.to_integer()
      )
    )
    .drop('target_name')
    .sort('hitting_time', nulls_last=True)
    .with_row_index(name='CDF')
    .with_columns(
      pl.col('CDF') / pl.col('CDF').max()
    )
  )
  aggregated_metrics[model] = agg_df
```

```{python}
sns.scatterplot(aggregated_metrics['temporal'], x='hitting_time', y='CDF', marker='.')
```

```{python}
def hitting_time_curve(df, x='hitting_time', y='CDF', ax=None, **kwargs):
  if ax is None:
    _, ax = plt.subplots()

  hitting_time = df[x].to_numpy()
  ecdf = df[y].to_numpy()

  ax.plot(hitting_time, ecdf, **kwargs)

  return ax
```

```{python}
hitting_time_curve(aggregated_metrics['temporal'])
```

```{python}
def partial_ecdf_plot(dfs: Mapping[str, pl.DataFrame], filter_fun: pl.Expr):
  ax = None
  for key, df in aggregated_metrics.items():
    aug_df = add_ecdf(df.filter(filter_fun), name="partial eCDF")
    ax = hitting_time_curve(aug_df, y='partial eCDF', ax=ax, label=key)
  ax.legend()
  return ax
```

```{python}
partial_ecdf_plot(aggregated_metrics, pl.col('seed').eq(0))
```

I think the idea here will be to compare N-quantiles of the distribution curve

```{python}
base_quantiles = pl.from_dict(
  {'quantile': np.linspace(0, 2900, 30)}
)

def quantile_metric(df: pl.DataFrame, quantiles=base_quantiles, cdf_col='CDF'):
  searched = quantiles.join_asof(
    df, 
    left_on='quantile',
    right_on='hitting_time', 
  ).select('quantile', cdf_col)

  return searched

def get_partial_quantiles(df: pl.DataFrame):
  # compute partial eCDF
  cdf_col = 'eCDF'
  aug_df = add_ecdf(df, name=cdf_col)
  # search quantiles
  return quantile_metric(aug_df, cdf_col=cdf_col)
```

## Single seed test

Testing for seed = 42
```{python}
test_seed = 42

qmets = {model: get_partial_quantiles(df.filter(pl.col('seed').eq(test_seed))) for model, df in aggregated_metrics.items()}

qmet_agg = pl.concat([
  df.with_columns(model=pl.lit(model))
  for model, df in qmets.items()
])

qmet_matrix = qmet_agg.pivot(on='model', index='quantile', values='eCDF')
```

```{python}
sns.scatterplot(qmet_agg, x='quantile', y='eCDF', hue='model')
```

First attempt: Cramer-von Mises . The problem here is that these goodness of fit criteria are two-sided, and we want a one-sided criterion.
```{python}
comb_symbol = "__x__"
qmet_cramer = qmet_matrix.with_columns(
  *(
    ((pl.col(col1) - pl.col(col2))).alias(f"{col1}{comb_symbol}{col2}")
    for col1, col2 in itertools.combinations(aggregated_metrics.keys(), 2)
  )
).select(pl.selectors.contains(comb_symbol)).sum().to_dicts()[0]
```

Second attempt was to compute P( X > Y ), but we have censored distributions, so this is hopeless.

```{python}
qmet_simple = qmet_matrix.with_columns(
  *(
    (pl.col('naive_static') - pl.col(col)).alias(f'naive_minus_{col}')
    for col in aggregated_metrics.keys()
  )
).select(
  pl.selectors.starts_with('naive_minus_').name.map(
    lambda s: s.removeprefix('naive_minus_')
  )
).sum().to_dicts()[0]

qmet_order = sorted(qmet_simple, key=qmet_simple.get)
```

### Whole of seed

```{python}
def filter_seed(df, seed):
  return df.filter(pl.col('seed').eq(seed))

def filter_target(df, target):
  return df.filter(pl.col('target').eq(target))

def q_simple_order(seed=None, target=None):

  if seed is None and target is None:
    raise ValueError("Must specify at least one of seed or target")
  elif seed is None:
    filterf = lambda df: filter_target(df, target)
  elif target is None:
    filterf = lambda df: filter_seed(df, seed)

  qmets = {
    model: get_partial_quantiles(filterf(df)) 
    for model, df in aggregated_metrics.items()
  }

  qmet_agg = pl.concat([
    df.with_columns(model=pl.lit(model))
    for model, df in qmets.items()
  ])

  qmet_matrix = qmet_agg.pivot(on='model', index='quantile', values='eCDF')

  qmet_simple = qmet_matrix.with_columns(
    *(
      (pl.col('naive_static') - pl.col(col)).alias(f'naive_minus_{col}')
      for col in aggregated_metrics.keys()
    )
  ).select(
    pl.selectors.starts_with('naive_minus_').name.map(
      lambda s: s.removeprefix('naive_minus_')
    )
  ).sum().to_dicts()[0]

  qmet_order = tuple(sorted(qmet_simple, key=qmet_simple.get))

  return qmet_order
```

```{python}
all_q_seed = {i: q_simple_order(seed=i) for i in aggregated_metrics['naive_static'].select(pl.col('seed').unique()).to_series().to_list()}

all_q_targ = {i: q_simple_order(target=i) for i in aggregated_metrics['naive_static'].select(pl.col('target').unique()).to_series().to_list()}
```

```{python}
levels = list(itertools.permutations(aggregated_metrics.keys()))
levels_map = {s: i for i,s in enumerate(levels)}

level_q_seed = [levels_map[all_q_seed[k]] for k in sorted(all_q_seed.keys())]
# not using targ keys, since it includes the shitty 149
level_q_targ = [levels_map[all_q_targ[k]] for k in sorted(all_q_seed.keys())] 
```

```{python}
for lnm, llq in zip(('seed', 'target'), [level_q_seed, level_q_targ]):
  plt.figure()
  plt.title(lnm)
  plt.plot(llq, '.')
  yt = plt.gca().get_yticks()
  plt.gca().set_yticklabels(
    [levels[int(t)] if t < len(levels) else '' for t in yt]
  );
```

Seeing if there is correlation between the ordering and the (estimated) size of the hospital. We know that this estimated size has an artificial lower bound of 5.
```{python}
sizes = pl.read_csv("concordant_networks/size_14.csv").sort('hospital').with_row_index(name='index')

level_df = pl.from_dict({
  'index': sorted(all_q_seed.keys()),
  'seed_order': level_q_seed,
  'targ_order': level_q_targ,
  'seed_str': [all_q_seed[k] for k in sorted(all_q_seed.keys())],
  'targ_str': [all_q_targ[k] for k in sorted(all_q_seed.keys())],
})

level_df = level_df.join(sizes, on='index')
```

```{python}
sns.scatterplot(
  level_df.select('index', 'seed_order', 'estimated_size'),
  x='estimated_size',
  y='seed_order',
)
```

```{python}
jj_level_df = level_df.select('seed_order', 'estimated_size').group_by(pl.all()).len().sort('len', descending=True).with_columns(pl.col('len').log(10).alias('loglen'))
```

```{python}
nonstandard_order_df = level_df.filter(pl.col('seed_order').eq(21))
xhosp = nonstandard_order_df.filter(pl.col('estimated_size').eq(pl.col('estimated_size').max())).select('index').item()

partial_ecdf_plot(aggregated_metrics, pl.col('seed').eq(xhosp))
```

```{python}
nonstandard_hosps = nonstandard_order_df.select('index').to_series().to_list()
for hosp in nonstandard_hosps:

  naive_hosp = add_ecdf(aggregated_metrics['snapshot'].filter(pl.col('seed').eq(hosp)), name='partialCDF')

  maxv = naive_hosp.drop_nulls(pl.col('hitting_time')).select(pl.col('partialCDF').max()).item()

  print(hosp, maxv)
```

### 2025-03-27 notebooks/explore_tau_over_time.qmd
Big reshuffle into a more nested directory structure
This is to improve the navgiability and allow for
a more cohesive sharing of code between different parts
of the analysis

---
title: Exploring the distribution of $\tau$ over time
execute:
  echo: false 
---

We want to explore the distribution of $\tau$, the interadmission time, as a function of the time that an individual leaves hospital.

```{python}
import polars as pl
from matplotlib import pyplot as plt
import seaborn as sns
from pathlib import Path

import graph_importer as gim
import tempograph_extractor as tge

basepath = Path(".")

sns.set_context('talk')
sns.set_style("ticks")
sns.set_palette("colorblind")
```

We load in the 1-day discretisation for temporal fidelity

```{python}
#| echo: false
CACHE = True
if not CACHE:
  G = gim.make_graph((basepath / "concordant_networks/temponet_1_365.lgl").absolute().as_posix())
  edge_df = tge.extract_edge_df(G)
  tau_series = tge.pull_series_tau_dist(edge_df)
  tau_dists = tge.compute_tau_dists(tau_series)
  clean_tau_dists = tau_dists.drop_nans()
else:
  clean_tau_dists = pl.read_csv(basepath / "concordant_networks/analyses/tau_dists.csv").drop_nans()
```

```{python}
fig, axs = plt.subplots(nrows=2, sharex=True)
axz = axs.flatten()
sns.scatterplot(clean_tau_dists, x='time', y='slope', ax=axz[0])
sns.scatterplot(clean_tau_dists, x='time', y='intercept', ax=axz[1])
```

### 2025-03-27 notebooks/snapshot_return_policy.qmd
Big reshuffle into a more nested directory structure
This is to improve the navgiability and allow for
a more cohesive sharing of code between different parts
of the analysis

---
title: Examining whether the snapshot return mechanism is correct
---

```{python}
#| code-fold: true
#| code-summary: imports
import random
import numpy as np
import polars as pl
from matplotlib import pyplot as plt
import seaborn as sns
from scipy import stats
from collections import Counter

rng = np.random.default_rng()
```

We want to investigate whether the snapshot return mechanism produces the correct return time distribution.

Currently, because we do not label individuals, all individuals in the same state are treated the same.
This means that individuals in the "at home" bucket are all treated as if they have been there for some time.

We model return (and the overall return rate) in the snapshot model by balancing the number of observed departures (to home) and arrivals (from home).

We can check the concordance of this model with the temporal distribution by writing an individual based-simulation (or adapt our current simulation, where we now also  track marked individuals pretty explicitly)

We will do this by modelling a single link (single we don't care about network effects), and sending individuals through this link indirectly.

```{python}
#| code-fold: true
SNAPSHOT_DURATION = 14
times = np.arange(0, 1000)
LOGNORMAL = {'mean': np.log(120), 'sigma': 0.5}
readmission_duration = lambda size=1: rng.lognormal(LOGNORMAL['mean'], LOGNORMAL['sigma'], size=size)

log_normal_mean = np.exp(LOGNORMAL['mean'] + LOGNORMAL['sigma']**2 / 2)
log_normal_mode = np.exp(LOGNORMAL['mean'] - LOGNORMAL['sigma']**2)

base_population = 1000
base_move_out = 0.2
```

Forward simulation, generating random movement and random readmission delay times.
```{python}
#| code-fold: true
hospital_state = np.zeros_like(times)
hosp_n = base_population

movements_out = np.zeros_like(times)
movements_in = np.zeros_like(times)
movement_dist = []
for _t in times: # implictly this is also the index
    hosp_n += movements_in[_t]
    hospital_state[_t] = hosp_n
    movements_out[_t] = rng.poisson(hosp_n * base_move_out)
    hosp_n -= movements_out[_t]
    move_times = np.round(readmission_duration(movements_out[_t]), 0).astype(int) + _t
    valid_moves = move_times[move_times < np.max(times)]
    movement_dist.extend(valid_moves - _t)
    if len(valid_moves) < 1:
        continue
    move_counter = Counter(valid_moves)
    move_tos = np.array(list(move_counter.keys()))
    move_counts = np.array(list(move_counter.values()))
    movements_in[move_tos] += move_counts
```

Computing the transition matrices
```{python}
#| code-fold: true
at_home = 0
home_numbers = np.zeros_like(times)
for _t in times:
    at_home += movements_out[_t]
    home_numbers[_t] = at_home
    at_home -= movements_in[_t]

inwards_transition_prob = movements_in / home_numbers

outwards_transition_prob = movements_out / hospital_state 
```

Definitions of data structures of the "agent-based" simulation.
```{python}
#| code-fold: true
# create agents
class Agent():
    _place_map = ['hospital', 'home']
    def __init__(self):
        self.place = 0
        self.go_home_time = np.nan
        self.return_hospital_time = 0

    def __repr__(self) -> str:
        return f"Agent()"

    def __str__(self):
        return f"Agent [{self.place} {self.go_home_time:.0f} {self.return_hospital_time:.0f}]"

    @property
    def duration(self):
        return self.return_hospital_time - self.go_home_time

    def go_home(self, time):
        self.go_home_time = time
        self.place = 1

    def return_hospital(self, time):
        self.return_hospital_time = time
        self.place = 0

class Pool():
    # implementation == ListDict 
    # see stackoverflow 15993447
    def __init__(self):
        self.positions = {}
        self.items = []

    @classmethod
    def of_agents(cls, N: int):
        pool = cls()
        for _ in range(N):
            agent = Agent()
            pool.add(agent)
        return pool

    def add(self, item):
        if item in self.positions:
            return
        self.items.append(item)
        self.positions[item] = len(self.items) - 1

    def pop(self, item):
        position = self.positions.pop(item)
        last_item = self.items.pop()
        if position != len(self.items):
            self.items[position] = last_item
            self.positions[last_item] = position
        return item

    def extend(self, items):
        init_len = len(self.items)
        self.items.extend(items)
        new_len = len(self.items)
        for i, item in enumerate(items):
            self.positions[item] = init_len + i

    def random_choice(self):
        return random.choice(self.items)

    def random_pop(self):
        item = self.random_choice()
        return self.pop(item)

    def queue_pop(self):
        item = self.items[0]
        return self.pop(item)

    def __len__(self):
        return len(self.items)

    def __contains__(self, item):
        return item in self.positions

    def __iter__(self):
        return iter(self.items)
```

Running agent-based sim. 
```{python}
#| code-fold: true

N_AGENTS = 1000
hospital_pool = Pool.of_agents(N_AGENTS)
home_pool = Pool()

return_delays = []
return_timings = []
pool_history = {'hospital': [], 'home': []}
movement_numbers = {'out': [], 'ret': []}
for _t in times:
    outp = outwards_transition_prob[_t]
    inwp = inwards_transition_prob[_t]

    # use implied dt=1
    n_out = np.clip(rng.poisson(outp * len(hospital_pool)), 0, len(hospital_pool))
    n_ret = rng.binomial(n=len(home_pool), p=inwp) 
    movement_numbers['out'].append(n_out)
    movement_numbers['ret'].append(n_ret)

    a_out = [hospital_pool.random_pop() for _ in range(n_out)]
    a_ret = [home_pool.random_pop() for _ in range(n_ret)]

    home_pool.extend(a_out)
    hospital_pool.extend(a_ret)

    for agent in a_out:
        agent.go_home(_t)

    for agent in a_ret:
        agent.return_hospital(_t)
        return_delays.append(agent.duration)
        return_timings.append(_t)

    pool_history['hospital'].append(len(hospital_pool))
    pool_history['home'].append(len(home_pool))
```

Plot of the observed distributions of readmission time
- simulation: from agent-based simulation
- source: from the source movement realisation
- dist: theoretical source distribution for the readmission delay
- static approximation: we approximate using the mean of the source realisation, and passing through to exponential rv dist.
```{python}
bins = np.linspace(min(times), max(times), 61)
plot_args = {'bins': bins, 'density': True, 'histtype': 'step'}
for name, arr in zip(['simulation', 'source', 'dist'], [return_delays, movement_dist, readmission_duration(100_000)]):
    plt.hist(arr, **plot_args, label=name)

tmesh = np.linspace(min(bins), max(bins), 3001)
mean_move_time = np.mean(movement_dist)
lamb = 1. /mean_move_time
plt.plot(tmesh, lamb * np.exp(-lamb * tmesh), label='static approx')

plt.legend()
```

Looking at the return time distribution over time
```{python}
#| code-fold: true
return_df = pl.from_dict({'timing': return_timings, 'delay': return_delays})
```
```{python}
sns.histplot(return_df, x='timing', y='delay', bins=61, discrete=False)
```

```{python}
sns.lineplot(return_df.group_by('timing').len(), x='timing', y='len')

plt.axvline(log_normal_mean, color='red')
plt.axvline(log_normal_mode, color='purple')

ax = plt.gca().twinx()
ax.hist(movement_dist, density=True, bins=61, histtype='step', color='red')
print(log_normal_mean, log_normal_mode)
```

We see transient behaviour as the readmission distribution kicks in, and the system trends toward the pseudo-stable state.

### 2025-03-27 echidna/__init__.py

@@ -1,0 +1,9 @@
+"""Library tools for on-network epidemic simulation
+
+Used for the ECHIDNA project to analyse inter-hospital patient and CPE movement
+in Victoria Australia
+"""
+
+__version__ = "0.0.0a1"
+
+from . import *

### 2025-03-27 echidna/network_simulation.py

@@ -398,7 +398,8 @@ class SnapshotNaive(Simulation):
         )
 
         self.current_index = 0
-        self.timings = timings
+        self.timings = np.array(timings).flatten()
+        self.timings = np.concatenate([self.timings, [np.inf]])
 
     def swapover(self):
         # assume current index is correctly set

### 2025-03-27 env.py

@@ -1,0 +1,15 @@
+import configparser
+import pathlib
+
+parser = configparser.ConfigParser()
+parser.read("env.ini")
+env = {k: pathlib.Path(v) for k,v in parser['filetree'].items()}
+
+subdirs = [
+    "notebooks",
+    "outputs",
+    "report",
+    "scripts",
+]
+for subdir in subdirs:
+    env[subdir] = env['root'] / subdir

### 2025-03-27 scripts/batched_plots.sh

@@ -1,0 +1,7 @@
+#!/usr/bin/env bash
+
+metrics=("pagerank" "harmonic" "closeness" "betweenness" "eigenvector")
+
+for metric in ${metrics[@]}; do
+    python projection_metrics.py plot $1 -g -m $metric
+done

### 2025-03-27 scripts/binomial_sample.py

@@ -1,0 +1,45 @@
+import numpy as np
+from numba import njit, prange
+
+
+@njit(parallel=True)
+def bino_1(N, P):
+    # Initialize an empty matrix to store the samples
+    X = np.empty_like(N)
+
+    # Iterate over each element of N and P in parallel
+    for i in prange(N.shape[0]):
+        for j in prange(N.shape[1]):
+            X[i, j] = np.random.binomial(N[i, j], P[i, j])
+
+    return X
+
+@njit(parallel=True)
+def _bino_2(N, P):
+    X = np.empty_like(N)
+
+    # here, assume N, P flat
+    for i in prange(N.shape[0]):
+        if N[i] == 0:
+            continue
+        X[i] = np.random.binomial(N[i], P[i])
+    
+    return X
+
+def bino_2(N, P):
+    return np.reshape(_bino_2(N.flatten(), P.flatten()), N.shape)
+
+
+@njit(parallel=True)
+def bino_3(N, P):
+    # Initialize an empty matrix to store the samples
+    X = np.empty_like(N)
+
+    # Iterate over each element of N and P in parallel
+    for i in prange(N.shape[0]):
+        for j in prange(N.shape[1]):
+            if N[i, j] == 0 or P[i,j] == 0:
+                continue
+            X[i, j] = np.random.binomial(N[i, j], P[i, j])
+
+    return X

### 2025-03-27 scripts/censored_dist_tests.py

@@ -1,0 +1,37 @@
+"""
+Here, we implement some naive and rudimentary tests for testing if two sets of samples come from the same distribution, when the samples are both right-censored.
+
+We follow the general structure of Chernobai et al. (2015), but adapt for a simple Type I right-censoring.
+
+A. Chernobai, S. T. Rachev, F.J. Fabozzi (2015) "Composite Goodness-of-Fit Tests for Left-Truncated Loss Samples" in "Handbook of Financial Econometrics and Statistics" Springer Science+Business Media, New York.
+"""
+
+import numpy as np
+from scipy import stats
+
+
+def ks_2samp_rc(data1, data2, threshold1, threshold2, mode='auto'):
+    """implements the two-sided kolmogorov-smirnov test for two right-censored distributions"""
+
+    if mode not in ('atuo', 'exact', 'asymp'):
+        raise ValueError(f"Invalid mode: {mode}")
+    MAX_AUTO_N = 10_000
+
+    data1 = np.sort(data1)
+    data2 = np.sort(data2)
+
+    n1 = data1.shape[0]
+    n2 = data2.shape[0]
+
+    if min(n1, n2) == 0:
+        raise ValueError("Sample data must not be empty")
+
+    v1 = np.isfinite(data1)
+    v2 = np.isfinite(data2)
+    vcat = np.concatenate([v1, v2])
+
+    data_all = np.concatenate([data1, data2])
+    cdf1 = np.searchsorted(data1, data_all, side='right') / n1
+    cdf2 = np.searchsorted(data2, data_all, side='right') / n2
+    cdfdiffs = cdf1[vcat] - cdf2[vcat]
+    

### 2025-03-27 scripts/general_plotting_utils.py

@@ -1,0 +1,10 @@
+from matplotlib import colors, cm
+
+class CMapp():
+    def __init__(self, cmap, vmin, vmax, norm_base=colors.Normalize):
+        self.norm = norm_base(vmin, vmax)
+        self.cmap = cmap
+        self.mappable = cm.ScalarMappable(norm=self.norm, cmap=self.cmap)
+
+    def __call__(self, value):
+        return self.mappable.to_rgba(value)

### 2025-03-27 scripts/plot_simulation.py

@@ -1,0 +1,48 @@
+import numpy as np
+from matplotlib import pyplot as plt, colors, cm
+import seaborn as sns
+import itertools
+
+import do_simulation_on_network as dosim
+
+def import_sim(sim_path, size_path):
+    sim_data = dosim.import_batched_realisations(sim_path)
+    sizes = dosim.find_sizes(size_path, None)
+    size_arr = [sz for h,sz in sorted(sizes.items())]
+    sim_data['sizes'] = sizes
+    sim_data['size_arr'] = size_arr
+
+    return sim_data
+
+def plot_sim_timeseries(sim, ax=None, cap=365, **plot_args):
+    
+    if ax is None:
+        _, ax = plt.subplots(layout='constrained')
+    ax.plot(sim.ts[:cap], sim.history.T[:cap,:], **plot_args)
+    return ax
+
+def default_plot_batch(ssims, grid=None, cap=365, lw=0.7, alpha=0.5):
+
+    with sns.plotting_context('talk'):
+        if grid is not None:
+            fig, axs = plt.subplots(
+                *grid, 
+                figsize=(14, 9), 
+                sharex=True, sharey=True,
+                squeeze=True, layout='constrained'
+            )
+            axs = axs.flatten()
+            fig.supxlabel("Time (days)")
+            fig.supylabel("Prevalence (per hospital)")
+        else:
+            axs = itertools.cycle([None])
+        for sim, ax_ in zip(ssims['records'], axs):
+            ax = plot_sim_timeseries(sim, cap=cap, ax=ax_, lw=lw, alpha=alpha)
+            if ax_ is None:
+                ax.set_xlabel('Time (days)')
+                ax.set_ylabel("Prevalence (per hospital)") 
+        
+
+
+def done():
+    plt.close('all')

### 2025-03-27 scripts/run_single_zero_static.py

@@ -1,0 +1,91 @@
+import numpy as np
+import datetime
+import network_conversion as conv
+import network_simulation as simu
+
+def preload():
+
+    ordering = conv.Ordering.from_file("./concordant_networks/size_14.csv")
+    converter = conv.StaticConverter.from_file(
+        "./concordant_networks/static_base_1_threshold_4.graphml", 
+        ordering,
+        purge_selfloops=True,
+    )
+
+
+    return ordering, converter
+
+def gen_params(converter):
+
+    prob_final = conv.ColumnDict.from_prob_final_file("./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
+    prob_final_arr = prob_final.organise_by(converter.ordering)
+    prob_final_zero =  np.zeros_like(prob_final_arr)
+
+    return converter.map_parameters({
+        'beta': 0.0,
+        'prob_final_stay': prob_final_zero,
+    })
+
+def create_sim(converter, parameters, pseudo_capacity=False, seeds=0, numinfperseed=0, dt=1.0, track_movement=True):
+
+    if pseudo_capacity:
+        # fake the size of the hospitals
+        full_sizes = 100_000 * np.ones_like(converter.ordering.sizes, dtype=int)
+    else:
+        full_sizes = converter.ordering.sizes
+
+
+    sim = simu.StaticWithHome(
+        full_size=full_sizes,
+        parameters=parameters,
+        dt=dt,
+        track_movement=track_movement,
+    )
+
+    sim.seed(
+        n_seed_events=seeds,
+        n_seed_number=numinfperseed,
+    )
+
+    return sim
+
+def simulate_sim_and_record(simulation: simu.SimulationWithMovers, simid, until=100, nostop=False, outfile=None, with_movers=False, **kwargs):
+
+    simulation.simulate(until=until, nostop=nostop)
+
+    simulation.export_history(outfile, identity=simid, with_movers=with_movers, **kwargs)
+
+
+
+
+def run_sims(n_reps=20):
+
+    ordering, converter = preload()
+    parameters = gen_params(converter)
+    sim = create_sim(converter, parameters=parameters, pseudo_capacity=True)
+
+    # skips = set([i for i,row in enumerate(converter.outwards_weighting_matrix) if row.nnz == 0])
+    skips = {149}
+
+    for _ in range(n_reps):
+        for i,v in enumerate(converter.ordering.sizes):
+            if i > 3:
+                break
+            print(i, v, sep=',', end=' ', flush=True)
[…]

### 2025-03-27 scripts/run_zero_naive_static.py

@@ -77,7 +77,7 @@ def run_sims(n_reps=20):
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/naive_static/sim_all_30s.h5"
+            outname = f"zero_sims/naive_static/sim_all_30s_fixTbug.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, simdate=simdate, simtime=simtime, seed=i)
 
 if __name__ == "__main__":

### 2025-03-27 scripts/run_zero_snapshot_sim.py

@@ -86,6 +86,7 @@ if __name__ == "__main__":
 
     REPEATS = 20
     for _ in range(REPEATS):
+        print(end='\n', flush=True)
         for i, v in enumerate(conv.ordering.sizes):
             print(i, v, sep=',', end=' ', flush=True)
 
@@ -97,5 +98,5 @@ if __name__ == "__main__":
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/snapshot/sims_with_30seeds_pseudocap.h5"
+            outname = f"zero_sims/snapshot/sims_30s_pc_fa_r2.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)

### 2025-03-27 scripts/run_zero_static_sim.py

@@ -81,7 +81,7 @@ def run_sims(n_reps=20):
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/static/sim_all_30s.h5"
+            outname = f"zero_sims/static/sim_all_30s_fixed_axis.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)
 
 if __name__ == "__main__":

### 2025-03-27 scripts/run_zero_temporal_delays.py

@@ -1,0 +1,96 @@
+import numpy as np
+import datetime
+import network_conversion as conv
+import network_simulation as simu
+
+def preload():
+
+    ordering = conv.Ordering.from_file("./concordant_networks/size_14.csv")
+    converter = conv.TemporalNetworkConverter.from_file(
+        "concordant_networks/temponet_14_365.lgl", 
+        ordering=ordering,
+        weight='weight',
+    )
+
+    return ordering, converter
+
+def gen_params(converter):
+
+    prob_final = conv.ColumnDict.from_prob_final_file("./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
+    prob_final_arr = prob_final.organise_by(converter.ordering)
+    prob_final_zero =  np.zeros_like(prob_final_arr)
+
+    return converter.map_parameters({
+        'beta': 0.0,
+        'prob_final_stay': prob_final_zero,
+    })
+
+def create_sim(
+        converter: conv.TemporalNetworkConverter, 
+        parameters, 
+        pseudo_capacity=False,
+        seeds=0, 
+        numinfperseed=0, 
+        dt=1.0, 
+        track_movement=True,
+) -> simu.TemporalSim:
+
+    if pseudo_capacity:
+        # fake the size of the hospitals
+        full_sizes = 100_000 * np.ones_like(converter.ordering.sizes, dtype=int)
+    else:
+        full_sizes = converter.ordering.sizes
+
+
+    sim = simu.TemporalSim(
+        full_size=full_sizes,
+        parameters=parameters,
+        dt=dt,
+        num_times=converter.NT,
+        discretisation_size=converter.DT,
+        track_movement=track_movement,
+    )
+
+    sim.seed(
+        n_seed_events=seeds,
+        n_seed_number=numinfperseed,
+    )
+
+    return sim
+
+def simulate_sim_and_record(simulation: simu.TemporalSim, simid, until=100, nostop=False, outfile=None, with_movers=False, **kwargs):
+
+    simulation.simulate(until=until, nostop=nostop)
+
+    simulation.export_history(outfile, identity=simid, with_movers=with_movers, **kwargs)
+
+
+
+
+def run_sims(n_reps=20, ndelays=12, size_delay=4):
+
+    ordering, converter = preload()
+    parameters = gen_params(converter)
+
+
+    for delay in range(ndelays):
+        converter.delay(size_delay)
+        delay_id = (delay+1)*size_delay
[…]

### 2025-03-27 scripts/simple_split_model.py

@@ -1,0 +1,137 @@
+"""here we test the simple model construction with an arbitrary truncated power law dist"""
+
+import random
+import numpy as np
+import polars as pl
+import seaborn as sns
+from matplotlib import pyplot as plt
+
+rng = np.random.default_rng()
+
+# generating the empirical cdf
+
+residence_df = pl.read_parquet("./concordant_networks/readmission_time.parquet").sort('readmission_time')
+
+mean_window = (
+    residence_df
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+num_readmits = residence_df.select(pl.col('count').sum()).item()
+
+
+# Naive static model #
+
+BIGN = 25_000
+
+# draw residence times as lambda ~ 1/mean_time
+
+naive_static_draws = [random.expovariate(1.0 / mean_window) for _ in range(BIGN)]
+
+# Improved static model #
+
+# compute components of the empirical cdf
+
+threshold = 4
+shorts = residence_df.filter(pl.col('readmission_time') <= threshold)
+longs = residence_df.filter(pl.col('readmission_time') > threshold)
+
+short_window = (
+    shorts
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+n_shorts = shorts.select(pl.col('count').sum()).item()
+
+long_window = (
+    longs
+    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
+    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
+    .item()
+)
+n_longs = longs.select(pl.col('count').sum()).item()
+
+fast_prop = n_shorts / (n_shorts + n_longs)
+lambd_fast = 1 / short_window
+lambd_star = lambd_fast / fast_prop
+lambd_slow_a = lambd_star * (1 - fast_prop)
+lambd_slow_b = 1 / (long_window - 1 / lambd_slow_a)
+
+lambd_slow_c = 1 / (long_window - 1/ lambd_fast)
+
+improved_static_draws = [None] * BIGN
+for i in range(BIGN):
+    base_time = random.expovariate(lambd_fast)
+    if random.random() < fast_prop:
+        improved_static_draws[i] = base_time
+    else:
[…]

### 2025-03-27 scripts/simple_split_model_2.py

@@ -1,0 +1,41 @@
+import numpy as np
+
+rng = np.random.default_rng()
+
+N = 10_000
+samples = rng.pareto(0.7, size=(N))
+T = np.sum(samples)
+
+Omega = 10  # threshold
+
+small = samples[samples < Omega]
+big = samples[samples >= Omega]
+
+L = np.mean(big)
+mu = T/N
+
+
+def generate_until(generator, T):
+    t = 0
+    ts = []
+    while t < T:
+        dt = generator()
+        t += dt
+        ts.append(dt)
+    return ts
+
+naive_ts = generate_until(lambda : rng.exponential(mu), T)
+
+S = (mu * N + L * len(big)) / len(small)
+def improved_sample():
+    base = rng.exponential(S)
+    if rng.random() < (len(small) / N):
+        return base
+    else:
+        return base + rng.exponential(L - S)
+
+improved_ts = generate_until(improved_sample, T)
+
+print(N, len(naive_ts), len(improved_ts))
+
+

### 2025-03-27 scripts/simulate_nonstat_markov.py

@@ -1,0 +1,27 @@
+# simulate the prob dist of a (non-stationary) markov chain to determine hitting time
+
+# class NonStationaryMarkovSimulation():
+#     def __init__(self, *args):
+
+#         self.transition_matrices = []
+        
+
+
+class StationaryMarkov():
+    def __init__(self, generator):
+        self.Q = generator
+
+    def step(self, state, dt):
+        dP = self.Q @ state
+        return state + dt*dP
+    
+    def solve(self, initial_state, tspan, dt):
+        t, t_end = tspan
+        Pt = [initial_state]
+        ts = [t]
+        while t < t_end:
+            Pt.append(self.step(Pt[-1], dt))
+            t = t + dt
+            ts.append(t)
+
+        return ts, Pt

### 2025-03-27 scripts/tempograph_extractor.py

@@ -1,0 +1,59 @@
+from typing import Iterable
+import graph_importer as gim
+import polars as pl
+from scipy import stats
+
+def extract_edge_df(G: gim.ig.Graph):
+    """Extract a dataframe of edge weights w/ nodal attributes"""
+    graph_data = [
+        {
+            'edge_weight': edge['weight'],
+            'source_loc': edge.source_vertex['loc'],
+            'sink_loc': edge.target_vertex['loc'],
+            'source_time': edge.source_vertex['time'],
+            'sink_time': edge.target_vertex['time'],
+        }
+        for edge in G.es
+    ]
+
+    df = (
+        pl.from_records(graph_data)
+        .with_columns(
+            tau=pl.col('sink_time') - pl.col('source_time'),
+        )
+    )
+
+    return df
+
+
+def pull_series_tau_dist(df: pl.DataFrame):
+
+    grouped_edges = df.group_by('source_time', 'tau').agg(pl.col('edge_weight').sum())
+    return grouped_edges.partition_by('source_time')
+
+def prep_taudist(df: pl.DataFrame):
+
+    sorted_df = df.sort('tau').select(pl.col('tau').log(), pl.col('edge_weight').log())
+    accept_df = sorted_df.filter(pl.col('tau').is_finite())
+
+    tau = accept_df.select(pl.col('tau')).to_series().to_numpy()
+    ews = accept_df.select(pl.col('edge_weight')).to_series().to_numpy()
+
+    return tau, ews
+
+def compute_tau_dists(df_series: Iterable[pl.DataFrame]):
+
+    regression_results = []
+    for df in df_series:
+        key = df.select(pl.col('source_time').unique()).item()
+        res = stats.linregress(*prep_taudist(df))
+        regression_results.append(
+            {
+                'time': key,
+                'slope': res.slope,
+                'intercept': res.intercept,
+                'rvalue': res.rvalue,
+            }
+        )
+
+    return pl.from_dicts(regression_results)

### 2025-03-27 reports/sketch.qmd
Update report

---
title: "DRAFT: Towards a paper on patient movement in the Victorian hospital system"
subtitle: "an exercise in trying to avoid reality"
author: "X-- xXx THE AUTHORS xXx --X"
filters:
  - tikz 
format: html
---

- Introduction
  - CPE is a concern in Victoria, and we want to manage it and prevent it from spreading
    - A good way to understand potential patterns of spread is to understand how patients move between facilities in the system
  - Inter-facility patient movement models for studying CPE
    - mostly static networks, or projected to static networks
    - but temporal fluctuations can potentially affect the dynamics
  - Temporal network methodology is relatively well established
    - but little on systems where information/contact leaves the system for periods of time
  - Our aim is to see whether or not this temporal fidelity is needed for this context
    - Can we get similar movement dynamics with simpler approximations
- Setting and Data Sources
  - Standard spiel
- Models and Methods
  - "Standard" model of

:::{#fig-mineandonlymine}
```{.tikz}
%%| format: svg
%%| filename: transfer_diagram
%%| img-width: 900
\usetikzlibrary{arrows.meta, positioning, shapes.geometric}

\begin{tikzpicture}[
    node distance=2cm, 
    hospital/.style={draw, circle},
    >=latex
  ]

  \begin{scope}[every node/.style = {hospital}]
  \node (A1) {A};
  \node (B1) [below=of A1] {B};
  \node (A2) [right=of A1] {A};
  \node (A3) [right=of A2] {A};
  \node (B2) [below=of A2] {B};
  \node (B3) [below=of A3] {B};
  \end{scope}

  \draw[->, dashed] (A1) edge[bend left] node [above] {indirect self-transfer} (A3);
  \draw[->, dashed] (A1) edge (B2);
  \draw[->, dashed] (A1) edge node [right] {indirect transfer} (B3);
  \draw[->] (A1) edge  node[above] {continuity} (A2);
  \draw[->] (A2) edge  (A3);
  \draw[->] (A1) edge node[left] {direct transfer}  (B1);
\end{tikzpicture}
```

Patient movements from hospital A at a given time: (a) a direct transfer to another facility B at the same time, (b) an indirect transfer to another facility B at a different time, going home in between, (c) an indirect self-transfer, where the patient is discharged home, but readmits at the same facility at a later time, (d) a continuity step where the patient stay at a given facility
:::

Here, we are interested in two processes:

1. the process that discharges patients from hospital
2. the process that readmits patients to hospital

We can characterise each by a random variable that represents the time they spend in the preceding state.
These random variables -- $s$ for the discharge process and $\tau$ for the readmission process -- are modelled differently in each of the different network representations.

| Network Representation | Model of $s$ (per edge) | Model of $\tau$ (per edge) |
|:------:|:------------------:|:------------------:|
| Naïve Static | Exponential | 0 |
| Static with Return Home | Exponential | Exponential |
| Static Snapshots | Exponential, varies over time | Difference of Exponentials, varies over time |
| Temporal | Exponential, varies with time | Empirical |

{{< placeholder >}}
... need an image here for the empirical distribution of $s$.

1. Results
  - Temporal network model yields slower and less diverse spread than other network models.
  - A naïve static network model vastly overestimates the spread 
  - Some improvement can be made by using networks and models that incorporate more temporal fidelity, but they do not capture the magnitude of change in the temporal network model

```{python}
#| echo: false
import netsim_summariser as summ
import polars as pl
import seaborn as sns
import numpy as np
from matplotlib import pyplot as plt
from scipy import stats

sns.set()
sns.set_palette('colorblind')
```

```{python}
metrics_archives = {
    'temporal': './zero_sims/temporal/metrics_30s.parquet',
    'static': './zero_sims/static/metrics_30s_fa.parquet',
    'snapshot': './zero_sims/snapshot/metrics_30s_fa.parquet',
    'naive_static': './zero_sims/naive_static/metrics_30s.parquet',
}

processed_results = {}

for label, archive in metrics_archives.items():
        processed_results[label] = pl.read_parquet(archive)

hitting_time_columns = pl.selectors.starts_with('hitting_time_')

agg_metrics = dict()
hitting_time_dists = dict()
for model, metrics in processed_results.items():
    agg_met = (metrics
                   .unpivot(
                     on=hitting_time_columns, 
                     index='seed', 
                     variable_name='target', 
                     value_name='hitting_time'
                   )
                   .with_columns(target_seed=(
                       pl.col('target')
                       .str.strip_prefix('hitting_time_')
                       .str.to_integer()
                   ))
                   .drop('target')
    )
    agg_metrics[model] = agg_met
    hitting_time_dists[model] = np.sort(agg_met.select('hitting_time').to_series().to_numpy())

analytical_expected_hitting_naive = pl.read_csv("./hitting_time_analysis/numer_hitting_time_30indv_chains.csv")
ana_E_hitT_naive = np.sort(analytical_expected_hitting_naive.select('hitting_time').to_series().to_numpy().flatten())

plt.figure()
for model, dist in hitting_time_dists.items():
    plt.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model)

xlim = plt.gca().get_xlim()
ylim = plt.gca().get_ylim()

plt.plot(ana_E_hitT_naive, np.linspace(0, 1, len(ana_E_hitT_naive)), 
         label='naive\nanalytical', linestyle='dashed')

plt.xlim(xlim)
plt.ylim(None, 0.9)

# plt.ylim(1e-3, None)
plt.xlabel('hitting time')
plt.ylabel('ECDF')

plt.legend(loc='lower right')
```

### 2025-03-27 echidna/__init__.py
Fix echidna lib imports, misc struct

@@ -6,4 +6,19 @@ in Victoria Australia
 
 __version__ = "0.0.0a1"
 
+__all__ = [
+    'clusters',
+    'graph_importer',
+    'hitting_markov',
+    'hitting_time_multichain',
+    'netsim_summariser',
+    'network_conversion',
+    'network_simulation',
+    'numba_sample',
+    'projection_metrics',
+    'sankey',
+    'staticify_network',
+    'util',
+]
+
 from . import *

### 2025-03-27 echidna/hitting_time_multichain.py

@@ -2,7 +2,7 @@
 This is an implementation of the computation of expected minimal hitting time over N identically parameterised continuous-time Markov chains {M_i(λ,Q)} 
 """
 
-import hitting_markov as mkv  # single chain computations
+from . import hitting_markov as mkv  # single chain computations
 import numpy as np
 from scipy import linalg
 from scipy import integrate

### 2025-03-27 echidna/network_conversion.py

@@ -11,9 +11,9 @@ from typing import Mapping, Hashable, Optional, Sequence, Any, SupportsFloat
 from functools import lru_cache
 from pathlib import Path
 
-import graph_importer as gim
-from util import Iden, nparr_find, NotFound, SupportsGet
-import examine_transfers_for_sizes as esz
+from . import graph_importer as gim
+from .util import Iden, NotFound, SupportsGet
+from . import examine_transfers_for_sizes as esz
 
 
 class Ordering:

### 2025-03-27 echidna/network_simulation.py

@@ -2,8 +2,8 @@
 
 from typing import Sequence, Mapping
 import numpy as np
-from util import BlackHole
-from numba_sample import multinomial_sparse_full, truncated_poisson, multinomial_sample_sparse_collapsed
+from .util import BlackHole
+from .numba_sample import multinomial_sparse_full, truncated_poisson, multinomial_sample_sparse_collapsed
 import h5py
 
 

### 2025-03-27 echidna/projection_metrics.py

@@ -1,5 +1,5 @@
 import numpy as np
-import graph_importer as gim
+from . import graph_importer as gim
 import polars as pl
 import igraph as ig
 import glob

### 2025-03-27 echidna/staticify_network.py

@@ -1,8 +1,7 @@
 import igraph as ig
-from networkx import attracting_components
 import polars as pl
 
-import graph_importer as gim
+from . import graph_importer as gim
 
 
 def load_temporal_network(path):

### 2025-03-27 reports/sketch.qmd
attempt path normalisation

@@
 from scipy import stats
+from pathlib import Path
 
@@
 sns.set_palette('colorblind')
+sns.set_context('paper')
+
+root = Path("../")
 ```
@@
 metrics_archives = {
-    'temporal': './zero_sims/temporal/metrics_30s.parquet',
-    'static': './zero_sims/static/metrics_30s_fa.parquet',
-    'snapshot': './zero_sims/snapshot/metrics_30s_fa.parquet',
-    'naive_static': './zero_sims/naive_static/metrics_30s.parquet',
+    'temporal': root / 'simulations/zero_sims/temporal/metrics_30s.parquet',
+    'static': root / 'simulations/zero_sims/static/metrics_30s_fa.parquet',
+    'snapshot': root / 'simulations/zero_sims/snapshot/metrics_30s_fa.parquet',
+    'naive_static': root / 'simulations/zero_sims/naive_static/metrics_30s.parquet',
 }
@@
 
-analytical_expected_hitting_naive = pl.read_csv("./hitting_time_analysis/numer_hitting_time_30indv_chains.csv")
+analytical_expected_hitting_naive = pl.read_csv(root / "outputs/hitting_time_analysis/numer_hitting_time_30indv_chains.csv")
 ana_E_hitT_naive = np.sort(analytical_expected_hitting_naive.select('hitting_time').to_series().to_numpy().flatten())

### 2025-03-27 env.py
rm env util
it was not helpful, better to define the root path at the top of script

@@ -1,15 +1,0 @@
-import configparser
-import pathlib
-
-parser = configparser.ConfigParser()
-parser.read("env.ini")
-env = {k: pathlib.Path(v) for k,v in parser['filetree'].items()}
-
-subdirs = [
-    "notebooks",
-    "outputs",
-    "report",
-    "scripts",
-]
-for subdir in subdirs:
-    env[subdir] = env['root'] / subdir

### 2025-03-31 reports/sketch.qmd
start porting movement

@@
 format: html
+execute:
+  echo: false
 ---
+
+```{python}
+#| echo: false
+from echidna import netsim_summariser as summ
+import polars as pl
+import seaborn as sns
+import numpy as np
+import igraph as ig
+from matplotlib import pyplot as plt
+from scipy import stats
+from pathlib import Path
+from collections import Counter
+
+sns.set()
+sns.set_style('whitegrid')
+sns.set_palette('colorblind')
+sns.set_context('paper')
+
+root = Path("../")
+```
 
@@
 
-:::{#fig-mineandonlymine}
+<!-- :::{#fig-mineandonlymine}
 ```{.tikz}
@@
 Patient movements from hospital A at a given time: (a) a direct transfer to another facility B at the same time, (b) an indirect transfer to another facility B at a different time, going home in between, (c) an indirect self-transfer, where the patient is discharged home, but readmits at the same facility at a later time, (d) a continuity step where the patient stay at a given facility
+::: -->
+
+:::{#fig-tripleburstdragon layout="[[10,-1,10,-1,10]]"}
+
+```{.tikz}
+%%| format: svg
+%%| filename: movement_case_one
+%%| img-width: 800
+%%| fig-cap: direct transfer
+\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}
+
+\begin{tikzpicture}[
+  node distance=2cm,
+  >=latex
+]
+  \tikzset{
+    squarx/.style= {
+      regular polygon,
+      regular polygon sides=4
+    }
+  }
+  \tikzset{
+    house/.style= {
+      single arrow,
+      shape border rotate=90,
+      single arrow head extend=0pt,
+      single arrow tip angle=120,
+      minimum height=0.75cm,
+      minimum width=0.75cm
+    }
+  }
+
+  \node[draw, squarx] (A) {A};
+  \node (X) [right=of A] {};
+  \node[draw, squarx] (B) [right=of X] {B};
+  \node[draw, house] (H) [below=of X] {};
+
+  \draw[->] (A) edge  (B);
+\end{tikzpicture}
+```
+
+```{.tikz}
+%%| format: svg
+%%| filename: movement_case_two
+%%| img-width: 800
+%%| fig-cap: indirect transfer
+\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}
+
+\begin{tikzpicture}[
+  node distance=2cm,
+  >=latex
+]
+  \tikzset{
+    squarx/.style= {
+      regular polygon,
+      regular polygon sides=4
+    }
+  }
+  \tikzset{
+    house/.style= {
+      single arrow,
+      shape border rotate=90,
+      single arrow head extend=0pt,
+      single arrow tip angle=120,
+      minimum height=0.75cm,
+      minimum width=0.75cm
+    }
+  }
+
+  \node[draw, squarx] (A) {A};
+  \node (X) [right=of A] {};
+  \node[draw, squarx] (B) [right=of X] {B};
+  \node[draw, house] (H) [below=of X] {};
+
+  \draw[->] (A) edge[bend right] (H);
+  \draw[->, dashed] (H) edge[bend right]  (B);
+\end{tikzpicture}
+```
+
+```{.tikz}
+%%| format: svg
+%%| filename: movement_case_three
+%%| img-width: 800
+%%| fig-cap: delayed readmission
+\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}
+
+\begin{tikzpicture}[
+  node distance=2cm,
+  >=latex
+]
+  \tikzset{
+    squarx/.style= {
+      regular polygon,
+      regular polygon sides=4
+    }
+  }
+  \tikzset{
+    house/.style= {
+      single arrow,
+      shape border rotate=90,
+      single arrow head extend=0pt,
+      single arrow tip angle=120,
+      minimum height=0.75cm,
+      minimum width=0.75cm
+    }
+  }
+
+  \node[draw, squarx] (A) {A};
+  \node (X) [right=of A] {};
+  \node[draw, squarx] (B) [right=of X] {B};
+  \node[draw, house] (H) [below=of X] {};
+
+  \draw[->] (A) edge[bend right] (H);
+  \draw[->, dashed] (H) edge[bend right] (A); 
+\end{tikzpicture}
+```
+
+Different patterns of movement possible between hospitals and home.
+(left) direct transfer from one hospital to another, with no discharge home (centre) indirect transfer from one hospital to another, where the patient is discharged to home for some duration (right) delayed readmission, a special case of an idrect transfer, where the patient readmits at the same hospital they were discharged from.
+Dashed arrows represent movements where the time at the origin may be different to the destiantion.
 :::
@@
   - Some improvement can be made by using networks and models that incorporate more temporal fidelity, but they do not capture the magnitude of change in the temporal network model
-
-```{python}
-#| echo: false
-import netsim_summariser as summ
-import polars as pl
-import seaborn as sns
-import numpy as np
-from matplotlib import pyplot as plt
-from scipy import stats
-from pathlib import Path
-
-sns.set()
-sns.set_palette('colorblind')
-sns.set_context('paper')
-
-root = Path("../")
-```
 
@@
 ana_E_hitT_naive = np.sort(analytical_expected_hitting_naive.select('hitting_time').to_series().to_numpy().flatten())
+```
 
+```{python}
+#| label: fig-ecdf
+#| fig-cap: Empirical cumulative distributions of the observed hitting times to hospitals, where a single hospital is seeded with 30 tracked individuals.
+A plot of CDF of the analytical expected hitting times is also shown for the naïve static network model.
 plt.figure()
@@
 plt.legend(loc='lower right')
+
+sns.despine()
 ```
+
+# Supplementary Material
+
+## Multichain Hitting
+
+We compute the analytical solution for $N$ independent Markov chains charaterised by the transition matrix $Q$ as follows.
+
+Denote $U_{\mathcal{I}}(t)$ the probability that no chain has hit the subset of states $\mathcal{I}$ by time $t$, and $v_{\mathcal{I}}(t)$ the same probability, but for a single chain.
+
+Then the expectation of the time for any of the $N$ chains to hit a state in $\mathcal{I}$ is
+
+$$
+\begin{aligned}
+\mathbb{E}[T_\mathcal{I}] &=\int_0^\infty U_{\mathcal{I}}(t) dt\\
+&=\int_0^\infty [v_\mathcal{I}(t)]^N dt
+\end{aligned}
+$$
+
+By noting that $v_{\mathcal{I}}(t)$ must obey the backward equations:
+
+$$ 
+\frac{\mathrm{d}v(t)}{\mathrm{d}t} = Q v(t)
+$$
+
+then we can construct a numerical scheme to evaluate $\mathbb{E}[T_\mathcal{I}]$.
+
+Because the backward equations are linear, we can expect $v_\mathcal{I}(t)$ to be a linear combination of exponentials, and we also know that the real parts of the exponents must be negative, since the survival probability must tend towards zero over time.
+We could write that 
+$$v_\mathcal{I}(t) = \sum_{n=1}^{M} a_n \exp(c_n t)$$
+
+Thus, the integral for $\mathbb{E}[T_\mathcal{I}]$ can be evaluated with an Gauss-Laguerre quadrature scheme: 
+$$\int_0^\infty f(t) \exp(-t) dt \approx \sum_{k=1}{K} w_k f(t_k)$$
+
+We can predetermine the locations $t_k$ to evaluate $v_\mathcal{I}(t)$ at (which is in turn solved using a numerical ODE solver), and then compute the integral using a weighted sum.
+
+We observed that the G-L scheme could only converge when the number of locations, $K$, was large.
+This was it needed to capture "enough" of the decay behaviour of components of $v(t)$ that have exponents where $|Re(c_i)|$ is small(est).
+
+We can make a scaling adjustment to the time variable $t$ where we have instead $t = \sigma s$.
+Then $$\int_0^\infty g(t) dt = \sigma \int_0^\infty g(s) ds$$.
+
+## Comparing Static and Snapshot Models
+
+As described in the main text, the static and snapshot network models implement the readmission mechanic differently.
+This leads to a non-intuitive ordering of their hitting time distributions at early time.
+
+```{python}
+#| label: fig-early-time-ecdf
+#| fig-cap: The ECDF at early time
+plt.figure()
+for model, dist in hitting_time_dists.items():
+    plt.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model)
+
+plt.plot(ana_E_hitT_naive, np.linspace(0, 1, len(ana_E_hitT_naive)), 
+         label='naive\nanalytical', linestyle='dashed')
+
+plt.xlim(-5, 370)
+plt.ylim(-0.01, 0.15)
+
+plt.xlabel('hitting time')
+plt.ylabel('ECDF')
+
+plt.legend(loc='lower right')
+
+sns.despine()
+```
+
+1. The snapshot curve initially matches the temporal curve, and begins slower/lower than the static curve.
+2. The snapshot and static curve overlap at around $t=170$ days, whereafter the snapshot curve is faster/higher than the static curve.
+3. The snapshot curve is "bumpy" in comparison to the other curves.
+
+To address the last point, this is due to th exhaustion of tracekd individuals within the duration of a snapshot.
+Because tracked individuals are held at home until the end of the snapshot, the number of tracke individuals that can move to new hospitals decreases over the duration of a given snapshot.
+Relatedly, the snapshot and temporal curve agree for the duration of the first snapshot, since for that particular time discretisation window, there are only movements on the direct transfer layer (no indirect transfer would occur until the start of the next snapshot/time discretisation).
+
+The relative positions/speeds of the snapshot and static curves are not necessarily so simple to explain.
+
+First we will examine why the static network yields faster spread than the series of static snapshots.
+We propose that this is caused by the delay in connectedness of the snapshot network model.
+```{python}
+NO_CACHE = False 
+connectedness_file = root / "outputs/snapshot_movement_analysis/overlay_connectedness.parquet"
+
+if NO_CACHE:
+    Gs = [
+        ig.Graph.Read(root / f"data/conc_tempo_14_detailed/{i:04d}.graphml")
+        for i in range(0, 3626+14, 14)
+    ]
+
+    # remove problem vertex (isolated hospital)
+    problem_vertex = 4454
+    for graph in Gs:
+        graph.delete_vertices(graph.vs.select(name_eq=problem_vertex).indices)
+
+    connectedness = []
+    component_sizes = []
+    for i in range(len(Gs)):
+        U = ig.union(Gs[:i+1])
+        connectedness.append(U.is_connected())
+        component_sizes.append(Counter(map(len, U.connected_components())))
+
+    df = pl.from_dicts([
+        {'snapshot': i, 'fully_connected': full_conn, 'component_size': k, 'count': v} 
+        for i, (full_conn, comp_counter) in enumerate(zip(connectedness, component_sizes)) 
+        for k,v in comp_counter.items()
+    ])
+
+    df.write_parquet(connectedness_file)
+else:
+    df = pl.read_parquet(connectedness_file)
+```
+
+```{python}
+#| label: fig-connected-components
+#| fig-cap: Plot of (strongly) connected components of the aggregated graph of the network snapshots.
+(Above) the size of the dominant connected component over time (Below) number of (strongly) connected components.
+fig, axs = plt.subplot_mosaic([['c'], ['c'], ['b']], sharex=True, layout='constrained')
+
+component_stats = df.group_by('snapshot').agg(
+    pl.col('component_size').max().alias('dominant component size'),
+    pl.col('count').sum().alias('number of components')
+).with_columns(time = pl.col('snapshot') * 14).sort('snapshot')
+
+sns.lineplot(component_stats, x='time', y='dominant component size', ax=axs['c'])
+sns.lineplot(component_stats, x='time', y='number of components', ax=axs['b'])
+```
+
+We see in @fig-connected-components that
+
+```{python}
+
+S = ig.Graph.Read(root / "data/concordant_networks/trimmed_static_base_1_threshold_4.graphml")
+
+static_df = pl.from_dict(
+    {
+        'indirect': S.es['indirect_weight'],
+        'direct': S.es['direct_weight'],
+        'source': S.es['source_loc'],
+        'target': S.es['dest_loc'],
+    }
+).cast(pl.Int64)
+
+size_df = pl.read_csv(root/ "data/concordant_networks/size_14.csv")
+
+static_df = (
+  static_df
+  .join(
+    size_df, 
+    left_on='source', 
+    right_on='hospital'
+  )
+  .with_columns(
+    dep_rate = (
+      pl.col('indirect').cast(pl.Float64) 
+      / pl.col('estimated_size') 
+      / S['time_span']
+    ),
+  )
+)
+```
+
+```{python}
+NOCACHE = False
+
+snapshot_table_file = root / "outputs/snapshot_movement_analysis/combined_outwards.parquet"
+
+if NO_CACHE:
+    snapshot_graph_tables = []
+    for i in range(0, 3626+14, 14):
+        graph = ig.Graph.Read(root / f"data/conc_tempo_14_detailed/{i:04d}.graphml")
+        problem_vertex = 4454
+        graph.delete_vertices(graph.vs.select(name_eq=problem_vertex).indices)
+        snapshot_graph_tables.append(
+            (pl.from_dict({
+                'source': graph.es['source'],
+                'target': graph.es['target'],
+                'snapshot_direct': graph.es['weight'],
+                'snapshot_indirect': graph.es['departures'],
+                'time': pl.lit(i)
+            })
+            .cast({
+                'source': pl.Int64,
+                'target': pl.Int64,
+                'snapshot_direct': pl.Int64,
+                'snapshot_indirect': pl.Int64,
+                'time': pl.Int64(),
+            })
+            .with_columns(time=pl.lit(i))
+            )
+        )
+    snapshot_table = pl.concat(snapshot_graph_tables)
+    snapshot_table.write_parquet(snapshot_table_file)
+else:
+    snapshot_table = pl.read_parquet(snapshot_table_file)
+
+joined_table = snapshot_table.join(static_df, on=['source', 'target'])
+
+SS_WIDTH = 14
+
+joined_table = (
+    joined_table
+    .with_columns(
+        ss_direct_rate = pl.col('snapshot_direct').cast(pl.Float64) / pl.col('estimated_size') / SS_WIDTH,
+        ss_indirect_rate = pl.col('snapshot_indirect').cast(pl.Float64) / pl.col('estimated_size') / SS_WIDTH,
+        st_direct_rate = pl.col('direct').cast(pl.Float64) / pl.col('estimated_size') / S['time_span'],
+    )
+)
+```
+
+```{python}
+deviations = (
+  joined_table
+  .filter(pl.col('dep_rate') > 0.0)
+  .with_columns(
+    deviation = pl.col('ss_indirect_rate') - pl.col('dep_rate')
+  )
+)
+
+plt.hist(deviations.select('deviation').to_series().to_numpy(), bins=101);
+plt.xlabel("Departure Rate Deviation (Snapshot Edge Weight - Static Edge Weight)")
+plt.yscale('log')
+```

### 2025-04-01 reports/model.qmd
Update model

---
title: "Section: Model of Stochastic Processes and Network Models"
filters:
  - tikz 
format: html
execute:
  echo: false
---

Notations

- $u,v$ for locations; discrete, usually subscript
- $s,t$ for times; continuous, usually function inputs/variables
- $T, \omega, \Omega$ for characteristic times
- $N$ as a size variable
- $x, y, z$ are state variables; often functions of location and time

:::{.hidden}
$$
\newcommand{\dd}[2][t]{\frac{\mathrm{d}#2}{\mathrm{d}#1}}
\newcommand{\ceil}[1]{\lceil#1\rceil}
$$
:::

## General Sketch

:::{#fig-general-diagram}
```{.tikz}
%%| format: svg
%%| filename: movement_general_case
%%| img-width: 800
%%| fig-cap: general movement processes
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}

\begin{tikzpicture}[
  node distance=2cm and 4cm,
  >=latex
]
  \tikzset{
    squarx/.style= {
      regular polygon,
      regular polygon sides=4
    }
  }
  \tikzset{
    house/.style= {
      single arrow,
      shape border rotate=90,
      single arrow head extend=0pt,
      single arrow tip angle=120,
      minimum height=0.75cm,
      minimum width=0.75cm
    }
  }

  \node[draw, squarx] (A) {$u$};
  \node (X) [right=of A] {};
  \node[draw, squarx] (B) [right=of X] {$v$};
  \node[draw, house] (H) [below=of X] {};
  \node[draw, circle] (D) [above=of X] {$\emptyset$};

  \draw[->] (A) edge node[above]{$d_{uv}(s, t)$}  (B);
  \draw[->] (A) edge node[above left]{$\zeta_u(s)$} (D);
  \draw[->] (A) edge node[below left]{$\eta_{uv}(s,t)$} (H);
  \draw[->] (H) edge node[below right]{$\rho_{uv}(s,t)$} (B);
\end{tikzpicture}
```

Sketch of processes modelled
:::

Here, $u$ and $v$ are locations (potentially identical locations), $s$ and $t$ are times. We take that $t \geq s$.
$\emptyset$ represents patients that are not observed to return to the hospital system after their discharge from location $u$. We assume the associated hazard for this process ($\zeta_u(t)$) to be known for all locations $u$ at all times $t$: the modelling of this process is the same for all network models, and will be neglected in the following.

Further, we define the function $w_{uv}(s, t)$ to be the number of individuals that are observed to discahrge from $u$ at time $s$ and readmit at time $t$. We allow a maximum of one of the pairs $u,v$ and $s,t$ to be reflexive, i.e. $u=v$ or $s=t$, but not both.

We will also assume that we will have quantities $N_u$ that represent the "size" of a location $u$. We assume that all locations $u$ will have $N_u$ individuals in them at all times, and that the observed mvoements $w_{uv}(s, t)$ can be thought of as being scaled by $N_u$.

The governing equations can be roughly expressed as a function of the hazards of each process.

$$\begin{aligned}
\lambda(u(s) \to \emptyset) &= \zeta_u(s)\\
\lambda(u(s) \to v(t)) &= d_{uv}(s, t)\\
\lambda(u(s) \to z_{uv}(t)) &= \eta{uv}(s, t)\\
\lambda(z_{uv}(t) \to v(t)) &= \rho_{uv}(s, t)
\end{aligned}$$

Note that

$$ \lambda(U \to V) = \lambda $$

represents that in some time step $\Delta t$, the number of individuals that move from state $U$ to state $V$ is distributed as

$$ \mathrm{Poisson}(\lambda x_U \Delta t) $$.

### Naïve Static Model

For both the naïve and improved static network models, we have that the observations of $w$ are taken over a time period of $T_\Sigma$.

In the naïve static model , we have that:

$$d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = 0$$
$$\rho_{uv}(s,t) = \rho_{uv} = 0$$

That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period.

### Improved Static Model

For the improved static model, we choose some threshold value $\omega$ that delineates direct and indirect transfers.

We will have:

$$d_{uv}(s,t) = d_{uv} =  \frac{\sum_{s,t: (t-s) < \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = \frac{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv} = \left[ \frac{\sum_{s,t: (t-s) \geq \omega} (t-s) w_{uv}(s,t)}{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}  \right]^{-1}$$

Here, we model the rate of indirect transfers $\eta$ similarly to the rate of direct transfers  $d$, by counting the average rate of observed movements. For the rate of readmission, $\rho$, we use the inverse of the mean readmission duration/delay.

### Snapshot Model

For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.

This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
This alters the indirect transfer processes to

$$\begin{aligned}
\lambda(u(s) \to {z'}_{uv}(t)) &= \eta_{uv}(s, t)\\
\lambda({z'}_{uv}(t) \to z_{uv}(t)) &= \delta(t \mathrm{ mod } \omega)\\
\end{aligned}$$

where $\delta(\cdot)$ is the Dirac delta function.

Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$. We denote this quantity $H_{uv}(t)$

We will have that:

$$d_{uv}(s,t) = d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s,t)}{\omega N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv}(s) = \frac{\sum_{t: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv}(t) = \frac{\sum_{s: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega H_{uv}}$$

### Temporal Model

For the temporal network, we make a choice of the time discretisation $\omega$. That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$. Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$. We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$

For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly.

This yields:

$$d_{uv}(s, t) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\eta_{uv}(s, t) = \frac{\sum_{t: (t-s) \geq \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\rho_{uv}(s,t) = \frac{1}{\ceil{t}_\omega - t}$$

The readmission process thus has a uniform hazard.

### 2025-04-02 reports/report.qmd
Update report
Some small updates

@@
 
-Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon.
+Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
 The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
@@
 ## Existing Network Modelling {#sec-network-modelling}
+
+Existing studies treating hospital systems as networks
 
@@
 :::
-
-::: {#new-eq}
 

### 2025-04-02 reports/sketch.qmd
Some small updates

@@
 ---
-title: "DRAFT: Towards a paper on patient movement in the Victorian hospital system"
-subtitle: "an exercise in trying to avoid reality"
-author: "X-- xXx THE AUTHORS xXx --X"
+title: 'DRAFT: Towards a paper on patient movement in the Victorian hospital system'
+subtitle: an exercise in trying to avoid reality
+author: X-- xXx THE AUTHORS xXx --X
 filters:
@@
   echo: false
+jupyter:
+  jupytext:
+    text_representation:
+      extension: .qmd
+      format_name: quarto
+      format_version: '1.0'
+      jupytext_version: 1.16.4
+  kernelspec:
+    display_name: Python 3 (ipykernel)
+    language: python
+    name: python3
 ---
@@
 #| label: fig-ecdf
-#| fig-cap: Empirical cumulative distributions of the observed hitting times to hospitals, where a single hospital is seeded with 30 tracked individuals.
-A plot of CDF of the analytical expected hitting times is also shown for the naïve static network model.
+#| fig-cap: 'Empirical cumulative distributions of the observed hitting times to hospitals, where a single hospital is seeded with 30 tracked individuals.
+A plot of CDF of the analytical expected hitting times is also shown for the naïve static network model.'
 plt.figure()
@@
 sns.despine()
+
 ```
@@
 sns.despine()
+
+plt.savefig("early_time_ecdf.png", dpi=360)
 ```
@@
 We propose that this is caused by the delay in connectedness of the snapshot network model.
+
 ```{python}
@@
 ```{python}
-
 S = ig.Graph.Read(root / "data/concordant_networks/trimmed_static_base_1_threshold_4.graphml")
@@
 ```{python}
+#| label: fig-deviations-snapshot-vs-static
+#| fig-cap: Deviation between indirect transfer rate of the static (mean) vs snapshot (time-varying) networks.
 deviations = (

### 2025-04-02 reports/report.qmd
small push on bibs refs

@@
 filters:
-    - diagram
+    - tikz 
 jupyter: python3
@@
     error: false
+    eval: false
+callout-appearance: simple
 ---
@@
 
-Existing studies treating hospital systems as networks
+There has been some existing work that analyse hospital systems using the framework of network analysis. 
+They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
+Using these networks, hospitals and links that are important to movement around the hospital system by analysing measures of centrality and clustering.
+<!-- Thus the network can encode a representation of the movement between hospitals over a given time period. -->
 
-A network or graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ is made up of nodes or vertices $\mathcal{V} = \{v_1, v_2, \dots\}$ and edges $\mathcal{E} = \{e_{ij} \coloneqq (v_i, v_j)\}$ that join those nodes.
-Each edge $e_{ij}$ can be assigned attributes, including an edge weight $w_{ij}$.
-For our application, this edge weight will typically represent the number of observed movements in a given time period.
-Thus the network can encode a representation of the movement between hospitals over a given time period.
+A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand
 
-We can use the network to construct a Markov chain, where the expected rate of movement is governed by the observed movement in the data.
 For some edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, the rate of movement along that edge $\lambda_{ij} = w_{ij} / \tau$, such that the time between movement events along that edge are exponentially distributed $\text{Exp}(\lambda_{ij})$.
 
+:::{.callout-note} 
 [Stuff about previous implementations of this idea] 
+:::
 
@@
 We can then say that
-:::
 

### 2025-04-07 reports/report.qmd
small poke

@@
 
-Understanding of the spread of AMR through hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
-The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
-However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
-It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
-In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
-In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granularity of the temporality]]]]].
+Understanding of the spread of AMR tharity of the temporality]]]]].
 
@@
 They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
-Using these networks, hospitals and links that are important to movement around the hospital system by analysing measures of centrality and clustering.
+Using these networks, hospitals and links that are important to mrough hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
+The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
+However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
+It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
+In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
+In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granulovement around the hospital system by analysing measures of centrality and clustering.
 <!-- Thus the network can encode a representation of the movement between hospitals over a given time period. -->
@@
 
-For some edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, the rate of movement along that edge $\lambda_{ij} = w_{ij} / \tau$, such that the time between movement events along that edge are exponentially distributed $\text{Exp}(\lambda_{ij})$.
+For each edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, 
+we can construct a reaction that has a hazard function $\lambda_{ij} = w_{ij} / \tau$.
+This implicitly constructs a process that has inter-event times (i.e.
+times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.
+
+This type of construction is
 

### 2025-04-07 reports/_model.qmd
restruct model part

--- 2025-04-01 reports/model.qmd
@@
----
-title: "Section: Model of Stochastic Processes and Network Models"
-filters:
-  - tikz 
-format: html
-execute:
-  echo: false
----
-
-Notations
+<!-- Notations
 
@@
 - $N$ as a size variable
-- $x, y, z$ are state variables; often functions of location and time
+- $x, y, z$ are state variables; often functions of location and time -->
 
@@
 :::
-
-## General Sketch
 
@@
   \node[draw, squarx] (B) [right=of X] {$v$};
-  \node[draw, house] (H) [below=of X] {};
+  \node[draw, house] (H) [below=of X] {$h_{uv}$};
   \node[draw, circle] (D) [above=of X] {$\emptyset$};
@@
 
-Sketch of processes modelled
+Sketch of the generalised model with processes of 1. permanent discharge ($u \to \emptyset$), 2. direct transfer ($u \to v$), 3. return home ($u \to h_{uv}$, 4. readmission $h_{uv} \to v$).
 :::
@@
 
-Note that
-
-$$ \lambda(U \to V) = \lambda $$
-
-represents that in some time step $\Delta t$, the number of individuals that move from state $U$ to state $V$ is distributed as
-
-$$ \mathrm{Poisson}(\lambda x_U \Delta t) $$.
+We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the hodling time an individual patient spends in each state.
 
@@
 
-For both the naïve and improved static network models, we have that the observations of $w$ are taken over a time period of $T_\Sigma$.
+We can begin with a simple naive static model.
+This model will have the edge weights between healthcare facilities that represent the number of transfers from the source node to the target node over a certain period of observation time ($T_\Sigma$), regardless of how long the patient is absent from the healthcare facilities between initial discharge and subsequent readmission.
 
-In the naïve static model , we have that:
-
+With reference to the general model above, the hazards of the  naïve static model are:
 $$d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u}$$
@@
 That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period.
+This is an oversimplification of the work done in @donker, but represents a "worst" case model of temporal fidelity that can be benchmarked against.
 
@@
 
-For the improved static model, we choose some threshold value $\omega$ that delineates direct and indirect transfers.
+We can introduce the concept of "indirectness" by allowing individuals to return home, in an "improved" static model.
+In this model, we choose some threshold value $\omega$ that delineates direct transfers that occur "instantaneously" and indirect transfers that require a patient to first move to an intermediary "home" state before readmitting.
 
-We will have:
+In comparison to the naive model, we will have non-zero hazards $\eta$ and $\rho$:
 
@@
 
+Of course, as the hospital system changes over time, the rate of patient transfers, and thus the hazards represented in our model should also change.
+We see evidence of this in the overall rates of movement in the network over time
+
+.... need a plot here of total direct/direct over time
+
+We approach this using static network snapshots of the hospital system over time.
+
 For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.
-
 This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
@@
 where $\delta(\cdot)$ is the Dirac delta function.
-
 Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$.
 We denote this quantity $H_{uv}(t)$
-
 We will have that:
@@
 We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$
-
 For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly. 
-
 This yields:
@@
 $$\rho_{uv}(s,t) = \frac{1}{\ceil{t}_\omega - t}$$
-
-The readmission process thus has a uniform hazard.

### 2025-04-07 reports/report.qmd
restruct model part

@@
 :::
+
+## Network Models for Temporal Fidelity
+
+Static networks provide a very interpretable model of influence of nodes in the system: network measures allow for some semblance of understanding of aggregate aspects of the network and its components.
+However, dynamic models derived from those static networks miss important dynamics and temporal qualities of the system they are modelling.
+It is well-known that temporal networks exhibit different behaviour to analogous static networks: they have lower connectivity ....
+
+In the context of the spread of AMR between healthcare facilities, these differences can lead to misleading results about the dynamics of the [disease], and related control mechanisms and policies.
+Using the detailed linkage data from the VAED, we can constrct a relatively detailed temporal network, which can then be projected in various network representations with different temporal fidelity.
+We roughly consider three aspects of the dynamics:
+
+1. the representation of the time patients spend outside the healthcare system (i.e. at home)
+1. the time-varying number of transfers (direct and indirect) between facililties
+1. the distribution of the inter-admission times for patients that are discharged and subsequently readmit to the system
+
+{{< include _model.qmd >}}
 
@@
 ## Movement Simulation Studies
+We simulate the movement of individuals without infection dynamics (for example, in the case of a noncommunicable disease) on each of the different types of network.
+$X_0 = 5$ individuals are seeded in a single healthcare facility at the start of each simulation, and move through the network for $T = 8$years.
 

### 2025-04-08 reports/report.qmd
small edit: small analytical example

@@
 
-# Methods
+# Models and Methods
 
@@
 
-A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand
+A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.
 
@@
 {{< include _model.qmd >}}
+
+These models have distinct advantages and disadvantages: the "simpler" models are more amenable to analysis, but the more "complex" models represent more fidelity.
+For example, the snapshot network model allows us to analyse each static snapshot as its own graph, and then subsequently aggregate those results as a time series; this is a much smaller algorithmic complexity compared to any analogous analysis that could be done on the full temporal network.
+
+In the extreme case, we can analyse a toy model in order to demonstrate the difference between the naïve and improved static network models.
+
+Consider a two-node network, where patients transfer from hospital A to hospital B. Suppose over some time period $T$ there are $n_d$ direct transfers and $n_i$ indirect transfers.
+In the naïve model, the expected hitting time is equal to $\frac{T}{n_d+ n_i}$.
+In the improved model, we suppose we also have the mean inter-admission time of $\tau$.
+Then the expected hitting times $k$ can be solved by a linear system [@Norris_Markov_Chains_1997]:
+
+$$
+\begin{bmatrix}
+\frac{T}{n_d+n_i} & -\frac{T}{n_d} & -\frac{T}{n_i} \\
+0 & 1 & 0 \\
+0 & -\frac{1}{\tau} & \frac{1}{\tau}
+\end{bmatrix} \begin{bmatrix} k_A \\ k_B \\ k_H \end{bmatrix}
+= \begin{bmatrix} 1 \\ 0 \\ 1\end{bmatrix}
+$$
+
+which solves to give $k_A = \frac{T}{n_d+n_i}\left(1 + \frac{n_i}{T} \tau \right)$.
+Somewhat expectedly, the expectation of the time taken to reach hospital B from hospital A is longer by a factor related to the number of indirect movements and their average inter-admission period.
+
+# Simulation Experiments
+
+Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.
+
+## Experimental Setup
+
+<hr />
+
+# APDX
 

### 2025-04-09 notebooks/rare_path.qmd
Update rare path for example graph diagrmans and analysis

--- 2025-03-25 rare_path.qmd
@@
 import numpy as np
-import network_simulation as netsim
+from echidna import network_simulation as netsim
 
@@
     edges=[
-        (0, 1),
-        (1, 2),
-        (2, 0),
+        # (0, 1),
+        # (1, 2),
+        # (2, 0),
         (0, 3),
@@
     edge_attrs={'weight':[
-        0,
-        0,
-        0,
+        # 0,
+        # 0,
+        # 0,
         100,
@@
 )
+```
+
+```{python}
+#| label: fig-graphs
+fig, axs = plt.subplots(ncols=2)
+axz = axs.flatten()
+
+ig.plot(
+    G,
+    target=axz[0],
+    edge_label=G.es['weight'],
+    layout='circle',
+)
+
+ig.plot(
+    R,
+    target=axz[1],
+    edge_label=R.es['weight'],
+    layout='circle',
+)
+
+axz[0].set_xlabel('regular')
+axz[0].spines[:].set_visible(True)
+axz[1].set_xlabel('rare')
+axz[1].spines[:].set_visible(True)
 ```
@@
 
+run and analyse the snapshot stuff
 ```{python}
@@
 
-histories = []
+snapshot_histories = []
 for it_ in range(N_SIMS):
@@
     snapshot_sim.simulate(30, nostop=True)
-    histories.append(np.array(snapshot_sim.history))
+    snapshot_histories.append(snapshot_sim.history)
+```
+
+```{python}
+timez = [snapshot_sim.ts[x] 
+         for hist in snapshot_histories 
+         for x in np.flatnonzero(hist[-1,:])[:1]
+        ]
+
+plt.hist(timez)
+```
+
+run and analyse static
+```{python}
+static_histories = []
+for it_ in range(N_SIMS):
+    static_sim.reset(soft=True)
+    static_sim.simulate(30, nostop=True)
+    static_histories.append(static_sim.history)
+```
+
+```{python}
+static_timez = [static_sim.ts[x] 
+         for hist in static_histories 
+         for x in np.flatnonzero(hist[-1,:])[:1]
+        ]
+```
+```{python}
+#| label: fig-cumhist
+plt.xlabel('hitting time')
+# plt.ylabel('count')
+# plt.hist(static_timez, histtype='step', bins=31, label='static')
+# plt.hist(timez, histtype='step', bins=4, label='snapshot')
+plt.ylabel('cumulative count')
+plt.hist(static_timez, cumulative=True, histtype='step', bins=31, label='static')
+plt.hist(timez, cumulative=True, histtype='step', bins=4, label='snapshot')
+plt.legend()
+# plt.yscale('log')
 ```

### 2025-04-09 reports/report.qmd
small content adds

@@
 The modelling of patient transfers between hospitals has historically been done by using static networks.
+We show here that this can lead to misleading estimates of the extent of the spread of patients over time.
+This has implications for surveillance and control of infections that are transferred between healthcare facilities by patients, such as CPE.
 
@@
 We can instead look at the series of static network induced by projection onto the locations at each time point. This allows
+
+## Estimation of Hospital Capacity
+
+Central to our model of the movement dynamics is the quantity $N_u$, the "size" of each hospital.
+We assume that this is constant over time, which is rather dubious, since hospitals can be reconfigured, expanded, or shrunk to meet administrative and operational needs.
+
+We make a few assumptions to reach the hospital size estimates:
+
+1. Healthcare facilities, on average, churn a number patients the equivalent of their size within a day.
+2. Some proportion of these churned individuals are readmissions to the same facility.
+3. There is a minimum size of 5.
+
+Then the size can be approximated as the number of transfers (that are not self-transfers) ...

### 2025-04-09 reports/report.qmd
proto exp setup

@@
 ## Experimental Setup
+
+We initialise the state of a simulation by seeding 30 individuals in a single facility.
+We then simulate the system forwards in time until a fixed end time.
+
+Each starting facility is looped over to be the starting facility, and realisations are acquired 20 times for each starting facility.
+This gives pairs of hitting times between the (known) starting facility and target facility (which we get for all non-starting facilities by the end of the simulation realisation)
 

### 2025-04-14 reports/old_sections.qmd
split old (pre aggd hiting time) and new

--- 2025-04-09 reports/report.qmd
@@
----
-title: "Temporality is Important for Networks of Hospital Patient Transfers"
-authors:
-    - name: David Wu
-    - name: Michael Lydeamore
-    - name: Andrew Stewardson
-    - name: Tjibbe Donker
-    - name: Nic Geard
-    - name: "******"
-date: last-modified
-format:
-    html: default
-    pdf:
-        include-in-header:
-            - text: |
-                \usepackage{mathtools}
-        papersize: a4
-bibliography: refs.bib
-csl: plos.csl
-filters:
-    - tikz 
-jupyter: python3
-execute:
-    echo: false
-    warning: false
-    error: false
-    eval: false
-callout-appearance: simple
----
-
-```{python}
-import numpy as np
-import polars as pl
-from matplotlib import pyplot as plt, colors
-import seaborn as sns
-import h5py
-import itertools
-import re
-
-from pathlib import Path
-
-# setup printing formats for tables
-pl.Config.set_tbl_hide_column_data_types(True)
-pl.Config.set_tbl_hide_dataframe_shape(True)
-
-sns.set_theme(style='ticks')
-sns.set_palette('colorblind')
-sns.set_context('paper')
-```
-
-# Key Points
-
-- There is a significant amount of patient movement in the Victorian hospital system where the patient returns home for a substantive amount of time between discharge from one hospital and subsequent admission at another hospital.
-- This "indirect" movement has implications on the rate of spread of a hypothetical contagion through the hospital system. It causes dispersal of patients to be slower than expected.
-- We can recover some of this behaviour by introducing an additional layer in the network to represent indirect movement.
-
-<!-- # Glossary
-
-| symbol | name | description |
-|--------|------|-------------|
-| v | vertex | |
-| e | edge | |
-| i | index of vertex | |
-| j | index of other vertex | |
-| t | time | |
-| x | location | hospital |
-| W | bin width | discretisation of the temporal network |
-| $\gamma$ | discharge | rate of discharge |
-| N | num | size of something |
-| k | hitting time | (expected) |
-| $\lambda$ | rate | of an event / process |
-| T | time period | of the _entire_ simulation |
-
-define a vertex $v_i = (x_i, t_i)$ made up of a location and time resp. -->
-
-# Introduction
-
-Antimicrobial resistance poses a great threat to human health and development. [@murray_global_2022]
-Globally, it poses a large burden, operationally and economically, on hospital systems.
-In Australia, we see [significant] economic burden from AMR.
-this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR. 
-One X of interest is carbapenemase-producing enterobacteriacea (CPE). 
-Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level. 
-This already causes some level of burden, and introduced strains would further multiply this.
-
-Understanding of the spread of AMR tharity of the temporality]]]]].
-
-# Models and Methods
-
-## Setting and data sources
-
-Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.
-
-Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
-In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study.
-All Victorian hospital admissions during this study period are included.
-The total number of admission records is 26876787.
-This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.
-
-Some exploratory analysis of this data was done in by [@lydeamore_carbapenemase-producing_2024].
-
-## Existing Network Modelling {#sec-network-modelling}
-
-There has been some existing work that analyse hospital systems using the framework of network analysis. 
-They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
-Using these networks, hospitals and links that are important to mrough hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
-The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
-However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
-It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
-In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
-In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granulovement around the hospital system by analysing measures of centrality and clustering.
-<!-- Thus the network can encode a representation of the movement between hospitals over a given time period. -->
-
-A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.
-
-For each edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, 
-we can construct a reaction that has a hazard function $\lambda_{ij} = w_{ij} / \tau$.
-This implicitly constructs a process that has inter-event times (i.e.
-times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.
-
-This type of construction is
-
-:::{.callout-note} 
-[Stuff about previous implementations of this idea] 
-:::
-
-## Network Models for Temporal Fidelity
-
-Static networks provide a very interpretable model of influence of nodes in the system: network measures allow for some semblance of understanding of aggregate aspects of the network and its components.
-However, dynamic models derived from those static networks miss important dynamics and temporal qualities of the system they are modelling.
-It is well-known that temporal networks exhibit different behaviour to analogous static networks: they have lower connectivity ....
-
-In the context of the spread of AMR between healthcare facilities, these differences can lead to misleading results about the dynamics of the [disease], and related control mechanisms and policies.
-Using the detailed linkage data from the VAED, we can constrct a relatively detailed temporal network, which can then be projected in various network representations with different temporal fidelity.
-We roughly consider three aspects of the dynamics:
-
-1. the representation of the time patients spend outside the healthcare system (i.e. at home)
-1. the time-varying number of transfers (direct and indirect) between facililties
-1. the distribution of the inter-admission times for patients that are discharged and subsequently readmit to the system
-
-{{< include _model.qmd >}}
-
-These models have distinct advantages and disadvantages: the "simpler" models are more amenable to analysis, but the more "complex" models represent more fidelity.
-For example, the snapshot network model allows us to analyse each static snapshot as its own graph, and then subsequently aggregate those results as a time series; this is a much smaller algorithmic complexity compared to any analogous analysis that could be done on the full temporal network.
-
-In the extreme case, we can analyse a toy model in order to demonstrate the difference between the naïve and improved static network models.
-
-Consider a two-node network, where patients transfer from hospital A to hospital B. Suppose over some time period $T$ there are $n_d$ direct transfers and $n_i$ indirect transfers.
-In the naïve model, the expected hitting time is equal to $\frac{T}{n_d+ n_i}$.
-In the improved model, we suppose we also have the mean inter-admission time of $\tau$.
-Then the expected hitting times $k$ can be solved by a linear system [@Norris_Markov_Chains_1997]:
-
-$$
-\begin{bmatrix}
-\frac{T}{n_d+n_i} & -\frac{T}{n_d} & -\frac{T}{n_i} \\
-0 & 1 & 0 \\
-0 & -\frac{1}{\tau} & \frac{1}{\tau}
-\end{bmatrix} \begin{bmatrix} k_A \\ k_B \\ k_H \end{bmatrix}
-= \begin{bmatrix} 1 \\ 0 \\ 1\end{bmatrix}
-$$
-
-which solves to give $k_A = \frac{T}{n_d+n_i}\left(1 + \frac{n_i}{T} \tau \right)$.
-Somewhat expectedly, the expectation of the time taken to reach hospital B from hospital A is longer by a factor related to the number of indirect movements and their average inter-admission period.
-
-# Simulation Experiments
-
-Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.
-
-## Experimental Setup
-
-<hr />
-
 # APDX

### 2025-04-14 reports/report.qmd
split old (pre aggd hiting time) and new

--- 2025-04-08 reports/report.qmd
@@
 
+We initialise the state of a simulation by seeding 30 individuals in a single facility.
+We then simulate the system forwards in time until a fixed end time.
+
+Each starting facility is looped over to be the starting facility, and realisations are acquired 20 times for each starting facility.
+This gives pairs of hitting times between the (known) starting facility and target facility (which we get for all non-starting facilities by the end of the simulation realisation)
+
 <hr />
-
-# APDX
-
-## Explicitly Temporal Models
-Previous analyses of movement and spread of contagions through these networks have made assumptions that result in the weight $w_{ij}$ of any given edge being constant over time, i.e.
-the network is static.
-[.....Typically, this is a result of performing an aggregation, and approximating the patient movement between hospitals as being Poisson, or equivalently that the residence time  hospitals being exponentially distributed......]
-As shown in @fig-readmission-window, these approximations lead to misleading dynamics on the hospital system of Victoria: the empirical distribution of inter-transfer times [....].
-A naïve static approximation that matches the mean readmission window will be inaccurate; in the following, we will attempt to resolve this problem in three increasingly granular ways:
-
-1. using a simple piecewise exponential approximation, by choosing a single threshold
-2. using a series of static networks with ...
-3. using a temporal network that exactly captures the readmission window distribution
-
-<!-- One advantage that the VAED provide us is the minute-scale granularity of patient admissions, which allows us to construct temporal networks and models of patient movement. -->
-
-```{python}
-#| label: fig-readmission-window
-#| fig-cap: Distribution of readmission time (time between discharge and successive admission of a single patient)
-
-# sourced from readmission_time.py
-readmit_window = pl.read_parquet("../concordant_networks/readmission_time.parquet")
-
-# naive static distribution
-mean_window = (
-    readmit_window
-    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
-    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
-    .item()
-)
-num_readmits = readmit_window.select(pl.col('count').sum()).item()
-
-xs = np.arange(0, 366, dtype=np.float64)
-naive_static_dist = num_readmits / mean_window * np.exp(-xs / mean_window) # scaling to N
-
-# piecewise distribution (improved static)
-threshold = 4
-shorts = readmit_window.filter(pl.col('readmission_time') <= threshold)
-longs = readmit_window.filter(pl.col('readmission_time') > threshold)
-
-short_window = (
-    shorts
-    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
-    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
-    .item()
-)
-n_shorts = shorts.select(pl.col('count').sum()).item()
-
-long_window = (
-    longs
-    .with_columns((pl.col('readmission_time') * pl.col('count')).alias('weighted_sum'))
-    .select(pl.col('weighted_sum').sum() / pl.col('count').sum())
-    .item()
-)
-n_longs = longs.select(pl.col('count').sum()).item()
-
-sxs = np.arange(0, threshold+1, dtype=np.float64)
-short_dist = n_shorts / short_window * np.exp(-sxs / short_window)
-
-lxs = np.arange(threshold, 366, dtype=np.float64)
-long_dist = n_longs / long_window * np.exp(-lxs / long_window)
-
-fig, ax = plt.subplots()
-ax.scatter(*(readmit_window.to_numpy().T), marker='.', color='k')
-h_0 = ax.plot(xs, naive_static_dist, color='C0', label='Naïve Static')
-h_1 = ax.plot(sxs, short_dist, color='C1', label='Improved Static')
-ax.plot(lxs, long_dist, color='C1')
-ax.axvline(threshold, color='C1', linestyle='dashed')
-ax.set_xscale('asinh')
-ax.set_yscale('log')
-ax.set_xlabel('Readmission Window (days)')
-ax.set_ylabel('Count')
-ax.set_aspect(0.5)
-ax.set_xticks([0, 1, 1e1, 1e2])
-ax.set_xticks([2, 5, 2e1, 5e1, 2e2, 5e2], minor=True)
-ax.set_xlim(-1e-1, 1e3) # !! this is manual
-ax.set_ylim(5e3, 1e8) # !! this is manual
-ax.legend(handles=[*h_0, *h_1])
-sns.despine()
-```
-
-{skip}
-
-Patient transfers can be considered as edges from some source hospital $x_i$ at some time $t_i$ to a target hospital $x_j$ at some time $t_j$. 
-Because we cannot export exact times, we instead discretise time.
-This is done by treating movements involving some arbitrary hospital $x_i$ between times $t_i$ and $t_i + W$ as instead involving a single node at $v_i \coloneqq (x_i, t_i)$.
-
-<!-- ::: {#fig-}
-```{.tikz}
-\begin{tikzpicture}
-\tikzstyle{every node}=[font=\LARGE]
-\draw [short] (2.75,14.75) -- (16.5,14.75);
-\draw [short] (2.75,11) -- (16.5,11);
-\draw [dashed] (3.5,16.25) -- (3.5,9.25);
-\draw [dashed] (8.25,16.5) -- (8.25,9.25);
-\draw [dashed] (12.25,16.5) -- (12.25,9.25);
-\draw [dashed] (16,16.5) -- (16,9.25);
-\draw [->, >=Stealth] (7.25,14.75) -- (9.25,11);
-\draw [->, >=Stealth] (8.75,14.75) -- (13.25,11);
-\draw [->, >=Stealth] (9.5,11) -- (10.75,14.75);
-\draw [->, >=Stealth] (4.25,11) -- (5,14.75);
-\draw [->, >=Stealth] (5.5,14.75) -- (5.5,11);
-\draw [->, >=Stealth] (6.25,11) -- (7,14.75);
-\draw [->, >=Stealth] (10,14.75) .. controls (11.75,15.75) and (12.5,15.75) .. (14.25,14.75) ;
-\draw [->, >=Stealth] (15,11) -- (15,14.75);
-\end{tikzpicture}
-``` -->
-
-<!-- ::: {#fig-time-discretisation}
-
-```{.tikz}
-%%| caption="This is an image, created by **TikZ i.e. LaTeX**."
-%%| additionalPackages="\usepackage{adjustbox}"
-\usetikzlibrary{arrows}
-
-\resizebox{16cm}{!}{%
-\begin{tikzpicture}[auto]
-    \tikzstyle{every node}=[font=\LARGE]
-    \draw [short] (0.275,1.475) -- (1.65,1.475);
-    \draw [short] (0.275,1.1) -- (1.65,1.1);
-    \draw [dashed] (0.35,1.65) -- (0.35,0.925);
-    \draw [dashed] (0.825,1.65) -- (0.825,0.925);
-    \draw [dashed] (1.225,1.65) -- (1.225,0.925);
-    \draw [dashed] (1.6,1.65) -- (1.6,0.925);
-    \draw [O-Latex] (0.725,1.475) -- (0.925,1.1);
-    \draw [->, >=Stealth] (0.875,1.475) -- (1.325,1.1);
-    \draw [->, >=Stealth] (0.95,1.1) -- (1.075,1.475);
-    \draw [->, >=Stealth] (0.425,1.1) -- (0.5,1.475);
-    \draw [->, >=Stealth] (0.55,1.475) -- (0.55,1.1);
-    \draw [->, >=Stealth] (0.625,1.1) -- (0.7,1.475);
-    \draw [->, >=Stealth] (1.0,1.475) .. controls (1.175,1.575) and (1.25,1.575) .. (1.425,1.475) ;
-    \draw [->, >=Stealth] (1.5,1.1) -- (1.5,1.475);
-\end{tikzpicture}
-}
-```
-
-Example schematic for time discretisation
-::: -->
-
-Then we can have a temporal network that can be treated as a static network as in @sec-network-modelling, i.e.
-we have edge weights $w_{ij}$ along the edge $e_{ij} \coloneqq (v_i, v_j)$, but with some explicit temporal representation.
-
-The network analysis of the temporal network as described above is computationally expensive since it is relatively large ($N_x = 388$ hospitals and $N_t = 259$ fortnightly bins, for $\approx 10000$ nodes).
-We can instead analyse slices of the temporal network, where a slice at $t_n$ consists of nodes at that time discretisation bin, and the related nodes.
-
-## Dynamics
-
-For the temporal network, we can model the number of patients discharged from some given location ($x_i$) at a given time ($t_n$) as a Poisson random variable with rate $\gamma(x_i, t_n)$.
-Some proportion $q(x_i, \emptyset)$ of these individuals will not re-admit at any other hospital, and can be removed from the system; the remaining discharged patients will have some probability $p(x_i, t_n, x_j, t_m)$ to re-admit at location $x_j$ at time $t_m$.
-These probabilites $p$ wil depend on the edges...
-Patients that have $t_i \neq t_j$ will be moved into an auxiliary state $z_{i,j}$...
-Movemnts $z_{ij}$ are distriuted as Uniform(0, W)
-
-<!-- 
-We have line-listed patient movement data between {start_date} and {end_date}. This allows us to generate exact patient movement networks up to minute precision.
-By arranging patient journeys within this dataset, we can construct a network of patient transfers when two consecutive (but not necessarily temporally adjacent) admissions of the same patient occur at different healthcare facilities.
-We can model this data as a temporal network where a (single) patient being discharged from hospital $A$ at time $t_1$ and admitted to hospital $B$ at a later time $t_2$, given the patient is not admitted to any other hospital in the interim, contributes to a weight of 1 on the edge $(A, t_1) \to (B, t_2)$.
-To reduce the computational expense of such a network, and to protect patient anonymity, we aggregate these patient transfer events to a discrete set of times.
-By choosing some window size $W$, we treat events that occur in the time interval $[t_i, t_i+W)$ as all occurring at $t_i$.
-This aggregation classifies patient transfers into two classes. 
-Transfers can be "direct", where the discharge and subsequent admission both occur within the same time window $[t_i, t_i+W)$.
-Alternatively they can be "indirect", where the patient is not present in the hospital system for some amount of time between discharge and admission, i.e.
-when $t_2 \neq t_1$.
-From inspecting the dataset, we see that there is a significant number of transfers that are indirect, and the time between consecutive admissions is distributed sub-exponentially, with a heavy tail.
-These types of transfers imply causal edges between hospitals that are neither synchronous (occur at the same time), nor persistent (where the effect of a cause at time $t_1$ is not felt over some interval $(t_1, t_2)$), which make it challenging to fit a temporal network generated by this dataset into classical temporal network frameworks of representation.
--->
-
-<!-- 
-We use a discrete time Markov chain model to simulate the spread of infection.
-Simulations step forward in timesteps of $\delta t$, and events within the window $\delta t$ are assumed to be independent.
-For simplicity, we model the infection as SI or SIS, depending on the choice of removal rate; we assume that all facilities are at capacity all the time, which allows us to remove the explciit tracking of susceptibles.
-When individuals are removed, they are randomly chosen to either never return or eventually return (to the hospital system); if they do, they are randomly placed into a transfer bucket which has two properties:
-
-- where they will readmit
-- when they will readmit
-
-For a "full" temporal simulation, both where and when are specified; for a "reduced" simulation, only "where" is specified, and "when" falls into one of two buckets: immediately, or not immediately.
-We use a "full" temporal simulation with the unprojected temporal network, and a "reduced" simulation for any projection of that temporal network that has lost temporal information.
-
-<Specification of the model here>
-
-| Parameter | Description |
-|-----------|---------------------------------------|
-| $\beta$   | Rate of infection per unit infected per day |
-| $\eta$    | Rate of removal/recovery |
-| $\gamma$  | Rate of discharge from hospital |
-| $M$       | Transition matrix inferred from network | -->
-
-### A Reduced Model
-
-For the naïve and improved static network models, we can compute the hitting times directly and compare them.
-We can model the dyanmics on these networks as a (continuous time) Markov Chain and extract the mean hitting time to any target hospital from all other hospitals.
-
-A simplistic model of the effect of indirect transfer can be modelled with the following figures:
-
-:::{#fig-model-going-home layout-ncol="2"}
-
-```{dot}
-//| fig-width: 30%
-digraph G {
-    A -> B [ label="λ₀" ]
-}
-```
-
-```{dot}
-//| fig-width: 50%
-digraph G {
-    A -> B [ label="λ†p" ]
-    A -> H [ label="λ†(1-p)" ]
-    H -> B [ label="λₕ" ]
-}
-```
-
-Toy model of the effect of indirect transfer on hitting time
-:::
-
-Let us denote the (empirical) distribution of return times as $f(T)$, where $T$ is the return time.
-Then for a naive model, we have that the hazard of moving between two hospitals is approximately
-$\lambda_0 = (\mathbb{E}(A \to B))^{-1} = (\int_0^{\infty} f(T) T dT)^{-1}$
-
-For the refined model, we have instead two processes
-
-$A \to B$ and $A \to H \to B$,
-
-where
-
-$\mathbb{E}(T_{A\to B}) = \int_0^{\Omega} f(T) T dT$
-
-and
-
-$\mathbb{E}(T_{A\to H\to B}) = \int_\Omega^{\infty} f(T) T dT$
-
-We could simulate with some rate $\lambda_\dagger$ so that
-
-$\lambda_d = \mathbb{E}(T_{A\to B})^{-1} = \lambda_\dagger \times p$
-
-and
-
-$\lambda_x = \lambda_\dagger \times (1-p)$ (indirect)
-
-AND
-
-$p = \int_0^{\Omega} f(T) dT$ is the proportion of individuals that take the direct path.
-
-Thus we can recover that $\lambda_\dagger = \left(\int_0^{\Omega} f(T) T dT\right)^{-1} \int_0^{\infty} f(T) dT$.
-
-And if we require that
-
-$(\lambda_x)^{-1} + (\lambda_h)^{-1} = \int_\Omega^{\infty} f(T) T dT$
-
-and so
-
-$$
-\begin{aligned}
-(\lambda_h)^{-1} &= \left[\int_\Omega^{\infty} f(T) T dT \right] - (\lambda_\dagger (1-p))^{-1}\\
-&= \left[\int_\Omega^{\infty} f(T) T dT \right] - \left[\left[\int_0^{\Omega} f(T) T dT \right]^{-1} {\int_\Omega^{\infty} f(T) dT} \right]^{-1}\\
-&= \left[\int_\Omega^{\infty} f(T) T dT \right] - \left[\int_0^{\Omega} f(T) T dT \frac{1}{\int_\Omega^\infty f(T) dT}\right]
-\end{aligned}
-$$
-
-We also need that $(\lambda_\dagger (1-p))^{-1} < \int_\Omega^{\infty} f(T) T dT$ (is this always possible?)
-
-If $T \sim Exp(\lambda)$, then $f(T) = \lambda \exp(-\lambda T)$ and $F(T) = 1 - \exp(-\lambda T)$.
-We can also find that $\int_a^b f(T) T dT = \left[ -(T + \lambda^{-1}) \exp(-\lambda T) \right]^b_a$, so that
-
-$\int_0^\Omega f(T) T dT = -(\Omega + \lambda^{-1}) \exp(-\lambda \Omega) + (0 + \lambda^{-1}) \exp(-\lambda 0) = \lambda^{-1} (1- \exp(-\lambda \Omega)) - \Omega \exp(-\lambda \Omega)$
-
-and
-
-$\int_\Omega^\infty f(T) T dT = (\Omega \lambda^{-1}) \exp(-\lambda \Omega)$
-
-so that
-
-$(\lambda_\dagger (1-p))^{-1} = [\lambda^{-1} (1- \exp(-\lambda \Omega)) - \Omega \exp(-\lambda \Omega)]  \frac{1}{\exp(-\lambda \Omega)}$
-
-We can then say that
-
-<!-- In the model without indirect transfer, we can compute the expected hitting time $k_{1}$ from $S_0$ to $S_1$ as
-
-$$k_{1} = {\lambda_0}^{-1}$$
-
-In a model with indirect transfer, we introduce an additional state $S_2$ which represents the indirect pathway.
-If the probability of direct transfer is $p$ given that an individual will move out of $S_0$, and the average indirect transfer period is $T = \lambda_2^{-1}$, the we can compute the hitting time $k_{1^*}$ of $S_1$ given we start at $S_0$:
-
-$$k_{1^*} = {\lambda_0}^{-1} + \frac{1-p}{\lambda_2}$$
-
-i.e. the hitting time is increased by $\frac{1-p}{\lambda_2}$. -->
-
-For a larger network, the number of possible paths causes complications for explicit computation of the effect, but we can always expect it to follow a similar principle, where splitting transfers into a direct path and a slower indirect path will add to the expected hitting time.
-These can be obtained by solving a linear system [@Norris_Markov_Chains_1997].
-
-![An example of a small 2-layer static network](diagrams/precompiled/BiLayer_ManimCE_v0.18.1.png){#fig-bilayer}
-
-# Results
-
-## Temporal Network Analysis
-
-## Effect of Home-return Mechanic
-
-For the naïve static network, we can model the rate of movement from hospital $i$ to hospital $j$ as the mean number of observed movements over the observation period (XXX days).
-For the improved static network, we expand the state by including an additional state for each pair of hospitals, so that the residence time at home (between hospital admissions) for indirect transfers can be modelled.
-Then as with the naïve model, the movement from $i$ to $j$ directly is the mean observed number of transfers over the observation epriod.
-The indirect movements from $i$ to the intermediate $i,j$ state are derived from the observed number of movements, and the rate from $i,j$ to $j$ is derived from the mean readmission period (as seed in @fig-readmission-window).
-
-Comparing the (uniform) relative difference in the mean hitting times ($= k_{1^*} / (k_1 + 1)$) gives us @fig-hitting-time-difference.
-We see that there is significant banding in the structure of the matrix.
-
-```{python}
-#| label: fig-hitting-time-difference
-#| fig-cap: Relative difference between the expected hitting times for the naïve static and improved static models
-
-with h5py.File('../hitting_time_analysis/hitting_time_arrays.h5', 'r') as fp:
-    naive_static_times = fp['naive_static'][:]
-    delayed_static_times = fp['delayed_static'][:]
-
-relative_change = delayed_static_times / (naive_static_times + 1)
-plt.matshow(relative_change, cmap='viridis', fignum=99)
-plt.colorbar(label='Relative difference in hitting time')
-plt.xlabel('Hospital to hit')
-plt.ylabel('Starting hospital');
-```
-
-## Movement Simulation Studies
-We simulate the movement of individuals without infection dynamics (for example, in the case of a noncommunicable disease) on each of the different types of network.
-$X_0 = 5$ individuals are seeded in a single healthcare facility at the start of each simulation, and move through the network for $T = 8$years.
-
-<!-- 
-To isolate the effect of the projections on the movement of patients, we can simulate the model with parmeters $\beta = 0, \eta=0$. -->
-
-```{python}
-# simulation archives from run_zero*.py
-simulation_archives = {
-    'temporal': '../zero_sims/temporal/sim_all_30s.h5',
-    'static': '../zero_sims/static/sim_all_30s_fixed_axis.h5',
-    'snapshot': '../zero_sims/snapshot/sims_with_30seeds_pseudocap_fixed_axis.h5',
-    'naive_static': '../zero_sims/naive_static/sim_all_30s.h5',
-}
-
-metrics_archives = {
-    'temporal': '../zero_sims/temporal/metrics_30s.parquet',
-    'static': '../zero_sims/static/metrics_30s_fa.parquet',
-    'snapshot': '../zero_sims/snapshot/metrics_30s_fa.parquet',
-    'naive_static': '../zero_sims/naive_static/metrics_30s.parquet'
-}
-
-processed_results = {}
-
-# metrics precomputed in compute_summary_stats.ipynb
-for label, archive in metrics_archives.items():
-        processed_results[label] = pl.read_parquet(archive)
-```
-
-We define the _extent_ of a simulation by the number of facilities that some patient has moved to or through within a given number of days.
-We see in @fig-extent-30 that we get starkly different distributions of extent for the naïve static model, compared to the other models.
-
-```{python}
-#| label: fig-extent-30
-#| fig-cap: "Distribution of 30-day extent in different simulation models on different network types"
-
-combined_all_df = pl.concat(
-    [(df.select('seed', 'extent')
-        .with_columns(
-            pl.lit(label).alias('simulation'), 
-            pl.col('extent').cast(pl.Int64)
-        )
-     ) for label, df in processed_results.items()
-    ]
-)
-
-sim_type_order = ['temporal', 'snapshot', 'static', 'naive_static']
-extent_bins = np.arange(0, combined_all_df.select('extent').max().item()+1)
-
-sns.histplot(
-    combined_all_df, 
-    x = 'extent', 
-    hue = 'simulation', 
-    multiple = 'dodge', 
-    bins = extent_bins, 
-    hue_order = sim_type_order,
-    discrete = True,
-)
-sns.despine()
-```
-
-```{python}
-# comparisons from compute_summary_stats.ipynb
-comparative_results = dict()
-name_pattern = re.compile(r"(\w+)_x_([\w_]+)\.30s")
-for comp_file in Path("../zero_sims/comparisons/").glob("*.30s.parquet"):
-    pair = name_pattern.match(comp_file.stem).groups()
-    if pair is None:
-        raise RuntimeError(f"Unable to parse file name: {comp_file.name}")
-    paired_comparison = pl.read_parquet(comp_file)
-    comparative_results[pair] = paired_comparison
-    comparative_results[tuple(reversed(pair))] = paired_comparison.with_columns(-pl.col('statistic'))
-```
-
-We can frame the empirical cumulative distribution (ECDF) of hitting times from a given seed hospital to a target hospital in a survival analysis view, and compare different experimental setups by comparing the ECDFs of the combinations of hitting times with log-rank tests.
-Each pair of hospitals yields a separate log-rank test statistic.
-Using a crude Bonferroni correction to counteract the multiple comparisons (which occur due to the ${N_x}^2=338^2$ combinations of seed and target hospitals), we find that there are significant differences between each type of model.
-We note spikes at very low p-values when comparing the naïve static model to other models.
-
-```{python}
-#| label: fig-logrank-comparison
-#| fig-cap: Log-rank test comparisons between experiments
-#| cache: true
-
-comparable_pairs = map(lambda x: sorted(x, key=sim_type_order.index), 
-                       itertools.combinations(sim_type_order, 2))
-
-NN = len(sim_type_order) - 1
-
-fig, axs = plt.subplots(
-    nrows=NN, ncols=NN, 
-    squeeze=True, 
-    sharex=True, sharey=True, 
-    layout="constrained"
-)
-
-confidence_threshold = 0.01 / (338**2)
-
-for pair in comparable_pairs:
-    index = (sim_type_order.index(pair[1]) - 1, sim_type_order.index(pair[0]))
-    comparison = comparative_results[tuple(pair)]
-
-    sns.histplot(comparison, x='pvalue', log_scale=True, ax=axs[index], kde=True)
-
-    axs[index].axvline(confidence_threshold, color='r')
-    axs[index].set_yscale('asinh')
-    sns.despine(ax=axs[index])
-
-for label, ax in zip(sim_type_order[1:], axs[:,0]):
-    ax.set_ylabel(label)
-
-for label, ax in zip(sim_type_order, axs[-1,:]):
-    ax.set_xlabel(label)
-
-fig.supxlabel('pvalue')
-fig.supylabel('Count')
-
-for index in zip(*np.triu_indices_from(axs, k=1)):
-    fig.delaxes(axs[index])
-```
-
-# Discussion
-
-The modelling of patient transfers between hospitals has historically been done by using static networks.
-
-{{< pagebreak >}}
-# Appendix
-
-## Projecting the temporal network
-We can project these causal edges by making their effects persistent in a hospital. This reduces the amount of information stored, so in theory should improve performance of simulations.
-
-With a naive approach, an edge $(A, t_1) \to (B, t_2)$ can be decomposed into:
-
-1. $(A, t_1) \to (B, t_1)$ and $(B, t_1) \to \dots \to (B, t_2)$
-2. $(A, t_1) \to \dots \to (A, t_2)$ and $(A, t_2) \to (B, t_2)$
-
-This causes an implied increased infection pressure at certain hospital nodes.
-This sort of projection also loses specificity of temporal causality.
-By inspecting the temporal network that results from this projection, you cannot tell if a patient is moving from $t_1$ to $t_2$ or to $t_3$.
-
-There is also a mechanistically distinct projection, where we introduce "home" nodes for each facility. For some facility $A$, the home node is denoted $A'$.
-Thus we could also project the edge as
-
-- $(A, t_1) \to (A', t_1)$ and $(A', t_1) \to \dots \to (A', t_2)$ and $(A', t_2) \to (B, t_2)$.
-
-This is an ideal projection, but is difficult to achieve. Specifically, it is difficult to control the duration that an individual spends in the home node.
-However, this projection still loses temporal causal specificty, as above.
-
-We project the temporal network into a series of multilayer networks.
-Instantaneous movements (consisting of direct hospital-hospital transfers) make up one layer; movements from hospital to home make up another layer; and movements from home to hospital make up the third layer.
-By doing this, we can more accurately capture the time that a patient spends at home between hospital visits, but without explicitly tracking individuals (just the number of patients along each path).
-
-<!-- ## Estimation of parameters for the projected network simulations
-
-The rate / transition matrix that defines the rate of return from home to hospital is characterised by the matrix $M_{in}$, such that
-
-$$M_{in}(i, j, t) = \frac{n_r(i, j, t)}{n_h(i, j, t)},$$
-
-where $i$ is the index for the hospital an individual was at prior to going home, $j$ the index for the hospital an individual is going to from home, $t$ the current time, $n_r$ the number of individuals that are returning from home, and $n_h$ the number of individuals that are currently at home (and will return to a hospital at some point in the future).
-
-We observe $n_r$ in the data, but we do not have any direct observations of $n_h$.
-We could compute this by iterating through the data, counting the number of individuals going home and returning to hospital, but this is expensive, and not necessarily available in all cases.
-Instead, we utilise the available values of $n_g$, the number of individuals that go home from hospital.
-We have a estimated return time distribution for individuals characterised by the pdf $r(\tau)$, so we apply this by estimating $n_h$ as
-
-$$n_h(i, j, t) \approx \int_{t-T}^t n_g(i, j, \tau) r(t - \tau) d\tau.$$
-
-We truncate this approximation at some bandwith time $T$.
-
-The retrun time distribution is presented in @figure-??. We fit a power law model $r(\tau) \sim \tau ^k$ with coefficient $k \approx -0.6$. -->
-
-## Simulation Details and Model
-
-We keep records of two sets of states:
-
-1. $n(t, x)$, the current number of infected at location $x$. This represents individuals in hospital.
-2. $u(t, x, y)$, the current number of individuals that came from location $x$ and will depart for location $y$ at some point in the future. This represents the number of people at home.
-
-Each snapshot can be considered a graph composed of vertices $V$ that correspond to locations, and have edges $E$.
-
-Edges of the temporal network, $e(x, s, y, t)$ from location $x$ at time $s$ to location $y$ at time $t$ with some weight $w$ can be decomposed:
-
-1. if $s = t$, construct an edge $e(x, y)$ with weight $w$
-2. if $s \neq t$ construct two edges:
-   1. $e_o(x, y, s)$ if it doesn't exist, and add weight $w$
-   2. $e_i(x, y, t)$ if it doesn't exist, and add weight $w$
-
-We can alternatively conceptualise this as decomposing the temporal edges into movements to a "home" node for each pair of hospitals.
-The edges $e_o(x, y, s)$ can be conceptualised as an edge between $x$ and a new "home" node denoted by the ordered pair $(x, y)$ at time $s$.
-
-<!-- ```{python}
-#| echo: false
-
-# we're going to use networkx and its layout engine to draw the diagram cause the other engines are dogwater
-import networkx as nx
-from matplotlib import pyplot as plt
-
-fig = plt.figure()
-grid = fig.add_gridspec(2, 1, hspace=0)
-ax1, ax2 = grid.subplots(sharex=True, squeeze=True)
-for ax in (ax1, ax2):
-    ax.set_box_aspect(1)
-
-ABC = nx.DiGraph(['AB', 'BC', 'BA', 'CB', 'AC', 'CA'])
-nx.draw_networkx(ABC, ax=ax1, with_labels=True)
-
-``` -->
-
-## Searching snapshots for information on unlikely paths
-
-We observed some empirical simulated hitting time survival curves that suggested that paths that did exist and were likely on the temporal network did not exist or were very unlikely on the equivalent snapshots.
-This seems confusing, since the snapshot representation _should_ represent a superset of possible movements.
-
-## Approximations made with the Static Network
-
-A standard static network consists of a single layer of nodes and edges.
-In order to capture patients leaving and then subsequently returning to hospital, we model the network with two layers: a "standard" layer represeting 'direct' transfers, and a secondary layer representing indirect transfers.
-This secondary layer is bipartite in nature, where each (potential) edge in the primary layer is represented by replacing it with a "home" node, and adding two additional edges to it: one from the original source node to the home node, and another from the home node to the original target node.
-This is implicitly captured in the network data structure by setting two additional edge attributes that represent the amount of flow on each of the two new secondary layer edges respectively.
-
-We remove self-loops from the static network; we attempt one model without self-loops in only the direct layer (since this does not make physical sense), and another without self-loops in both the direct layer, and the indirect layer (we uspected that these self-loops slowed the movement of individuals down, but the resulting effect is the explosion of movement in the first simulation step)
-
-## Snapshots
-
-Network analysis of the explicitly-temporal network is difficult, since the number of nodes is relatively large ($N \sim N_x \times N_t$), and inversely scales with the temporal discretisation chosen ($N_t \sim 1/W$).
-
-We can instead look at the series of static network induced by projection onto the locations at each time point. This allows

### 2025-05-06 reports/report.qmd
pre-qin edits

--- 2025-04-09 reports/report.qmd
@@
 
-Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.
+Beyond simple models, there is vanishingly small possibility of finding computable analytical expressions for the hitting times.
+Instead we examine the empirical distributions from simulating the dynamics.
+To do this, we iterate over all facilities, seeding 30 individuals in that single facility, and simulating until a fixed end time, using a discrete-time Markov Chain formulation.
+We note the number of individuals in each facility at each time step.
+The central metric we will examine is the hitting time.
+In these experiments, to e
 
-## Experimental Setup
+## Results
+
+```{python}
+
+```
+
+Plotting the aggregated hitting times as an empirical cumulative distribution function (ECDF) in @fig-ecdf, we can see that
 

### 2025-05-06 reports/sketch.qmd
small edit: small analytical example
small content adds
pre-qin edits

@@
 
+plt.savefig("early_time_ecdf.png", dpi=360)
 ```
@@
 
-plt.plot(ana_E_hitT_naive, np.linspace(0, 1, len(ana_E_hitT_naive)), 
-         label='naive\nanalytical', linestyle='dashed')
+#plt.plot(ana_E_hitT_naive, np.linspace(0, 1, len(ana_E_hitT_naive)), 
+#         label='naive\nanalytical', linestyle='dashed')
 
-plt.xlim(-5, 370)
-plt.ylim(-0.01, 0.15)
+plt.xlim(-5, 500)
+plt.ylim(-0.01, 0.2)
 
@@
 
-plt.savefig("early_time_ecdf.png", dpi=360)
+# plt.savefig("early_time_ecdf.png", dpi=360)
 ```
@@
 ```
+
+Examining the deviations greater than 1.0, we see they are all edges that are self-loops of a particular location.
+This occurs due to the rate of indirect departures being greater than 1, which is a byproduct of our location size estiamtion procedure.
+```{python}
+from echidna import graph_importer as gim
+
+# investigating the weirdness in the RHS
+investigation_location = 8386
+
+tempo = gim.make_graph(str(root / "data/concordant_networks/temponet_14_365.lgl"))
+investi_locs = tempo.vs.select(loc_eq=investigation_location)
+self_loops = tempo.es.select(_from_in=investi_locs, _to_in=investi_locs)
+sloop_source_times = [edge.source_vertex['time'] for edge in self_loops]
+sloop_target_times = [edge.target_vertex['time'] for edge in self_loops]
+self_loop_df = pl.DataFrame({
+  'source_time': sloop_source_times, 
+  'target_time': sloop_target_times,
+  'weight': self_loops['weight'],
+})
+
+```

### 2025-05-06 reports/report.qmd

--- 2025-04-14 reports/report.qmd
@@
 
-Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.
+Beyond simple models, there is vanishingly small possibility of finding computable analytical expressions for the hitting times.
+Instead we examine the empirical distributions from simulating the dynamics.
+To do this, we iterate over all facilities, seeding 30 individuals in that single facility, and simulating until a fixed end time, using a discrete-time Markov Chain formulation.
+We note the number of individuals in each facility at each time step.
+The central metric we will examine is the hitting time.
+In these experiments, to e
 
-## Experimental Setup
+## Results
 
-We initialise the state of a simulation by seeding 30 individuals in a single facility.
-We then simulate the system forwards in time until a fixed end time.
+```{python}
 
-Each starting facility is looped over to be the starting facility, and realisations are acquired 20 times for each starting facility.
-This gives pairs of hitting times between the (known) starting facility and target facility (which we get for all non-starting facilities by the end of the simulation realisation)
+```
+
+Plotting the aggregated hitting times as an empirical cumulative distribution function (ECDF) in @fig-ecdf, we can see that
 

### 2025-05-06 echidna/examine_transfers_for_sizes.py
revisit size estimation

@@ -1,6 +1,7 @@
 import igraph as ig
 from collections import defaultdict
 import numpy as np
+from more_itertools import minmax
 
 
 def determine_self_loops(G: ig.Graph):
@@ -20,6 +21,13 @@ def determine_out_edges(G: ig.Graph):
     return loc_weights
 
 
+def determine_existence_duration(G: ig.Graph):
+    loc_span = {
+        loc: minmax(G.vs.select(loc_eq=loc)["time"]) for loc in set(G.vs["loc"])
+    }
+    return {loc: span[1] - span[0] for loc, span in loc_span.items()}
+
+
 def determine_self_prop(G: ig.Graph):
     self_loops = determine_self_loops(G)
     out_edges = determine_out_edges(G)
@@ -32,10 +40,25 @@ def rough_hospital_size(G: ig.Graph, round_precision=16, clip_under=5):
     TIME_SCALE = 3652  # days
     return {
         loc: np.round(
-            np.clip(sz / PROP_READMIT / TIME_SCALE / CHURN_RATE, clip_under, None),
+            np.clip(moves / PROP_READMIT / TIME_SCALE / CHURN_RATE, clip_under, None),
             round_precision,
         )
-        for loc, sz in determine_out_edges(G).items()
+        for loc, moves in determine_out_edges(G).items()
+    }
+
+
+def nuanced_hospital_size(G: ig.Graph, round_precision=16, clip_under=5):
+    PROP_READMIT = 0.42
+    CHURN_RATE = 1
+    existence = determine_existence_duration(G)
+    return {
+        loc: np.round(
+            np.clip(
+                moves / PROP_READMIT / existence[loc] / CHURN_RATE, clip_under, None
+            ),
+            round_precision,
+        )
+        for loc, moves in determine_out_edges(G).items()
     }
 
 

### 2025-05-07 echidna/network_simulation.py
Changes to np.array typing

@@ -1,6 +1,6 @@
 """this is a base refactor of on_network_simulation"""
 
-from typing import Sequence, Mapping
+from typing import Iterable, Sequence, Mapping, SupportsFloat
 import numpy as np
 from .util import BlackHole
 from .numba_sample import multinomial_sparse_full, truncated_poisson, multinomial_sample_sparse_collapsed
@@ -8,7 +8,7 @@ import h5py
 
 
 class Simulation:
-    def __init__(self, full_size: Sequence[int], parameters: Mapping, dt=1.0):
+    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, dt=1.0):
         self.N = np.asanyarray(full_size).reshape((-1, 1))
         self.NHOSP, _ = self.N.shape
         self.state = np.zeros_like(self.N)
@@ -146,7 +146,7 @@ class SimulationWithMovers(Simulation):
 class SnapshotWithHome(SimulationWithMovers):
     def __init__(
         self,
-        full_size: Sequence[int],
+        full_size: Iterable[SupportsFloat], 
         parameters: Mapping,
         timings: Sequence,
         track_movement=False,
@@ -225,7 +225,7 @@ class SnapshotWithHome(SimulationWithMovers):
 class StaticWithHome(SimulationWithMovers):
     def __init__(
         self,
-        full_size: Sequence[int],
+        full_size: Iterable[SupportsFloat], 
         parameters: Mapping,
         dt=1.0,
         track_movement=False,
@@ -293,7 +293,7 @@ class TemporalSim(SimulationWithMovers):
     
     def __init__(
             self, 
-            full_size: Sequence[int],
+            full_size: Iterable[SupportsFloat],
             parameters: Mapping, 
             num_times: int,
             discretisation_size: float,
@@ -390,7 +390,7 @@ class TemporalSim(SimulationWithMovers):
         self.time_travellers = np.zeros_like(self.time_travellers, dtype=np.int64)
 
 class SnapshotNaive(Simulation):
-    def __init__(self, full_size: Sequence[int], parameters: Mapping, timings: Sequence, dt=1.0):
+    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, timings: Sequence, dt=1.0):
         super().__init__(
             full_size=full_size,
             parameters=parameters,

### 2025-05-07 scripts/run_zero_naive_static.py
Update sim running scripts (for new size files)

@@ -1,15 +1,18 @@
 # run_zero_naive_static.py
 
+from pathlib import Path
 import numpy as np
 import datetime
-import network_conversion as conv
-import network_simulation as simu
+from echidna import network_conversion as conv
+from echidna import network_simulation as simu
+
+root = Path(__file__).resolve().parent.parent.resolve()
 
 def preload():
 
-    ordering = conv.Ordering.from_file("./concordant_networks/size_14.csv")
+    ordering = conv.Ordering.from_file(root / "data/concordant_networks/size_14_nu.csv")
     converter = conv.NaiveStaticConverter.from_file(
-        "./concordant_networks/shuf_static_network.graphml",
+        root / "data/concordant_networks/shuf_static_network.graphml",
         ordering=ordering,
         time_span=3626,
     )
@@ -17,7 +20,7 @@ def preload():
     return ordering, converter
 
 def generate_params(converter):
-    prob_final = conv.ColumnDict.from_prob_final_file("./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
+    prob_final = conv.ColumnDict.from_prob_final_file(root / "data/concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
     prob_final_arr = prob_final.organise_by(converter.ordering)
     prob_final_zero =  np.zeros_like(prob_final_arr)
 
@@ -77,7 +80,7 @@ def run_sims(n_reps=20):
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/naive_static/sim_all_30s_fixTbug.h5"
+            outname = root / "simulations/zero_sims_resized/naive_static/sim_all_30s_fixTbug.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, simdate=simdate, simtime=simtime, seed=i)
 
 if __name__ == "__main__":

### 2025-05-07 scripts/run_zero_snapshot_sim.py

@@ -1,7 +1,7 @@
 """new_newtork_sim"""
 import numpy as np
-import network_conversion as netcon
-import network_simulation as netsim
+from echidna import network_conversion as netcon
+from echidna import network_simulation as netsim
 
 def create_snaphome_parameter_converter(ordering_file, snapshots_directory):
 
@@ -70,14 +70,17 @@ def simulate_sim_and_record(simulation: netsim.SnapshotWithHome, simid, until=10
 
 if __name__ == "__main__":
     import datetime
+    from pathlib import Path
+
+    root = Path(__file__).resolve().parent.parent.resolve()
 
     conv = create_snaphome_parameter_converter(
-        ordering_file="./concordant_networks/size_14.csv",
-        snapshots_directory="./conc_tempo_14_detailed/",
+        ordering_file= root / "data/concordant_networks/size_14_nu.csv",
+        snapshots_directory= root / "data/conc_tempo_14_detailed/",
     )
 
     params = init_zero_params(
-        conv, "./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv"
+        conv, root / "data/concordant_networks/probability_of_final_stay_by_shuffled_campus.csv"
     )
 
     sim = create_sim(
@@ -98,5 +101,5 @@ if __name__ == "__main__":
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/snapshot/sims_30s_pc_fa_r2.h5"
+            outname = root / "simulations/zero_sims_resized/snapshot/sims_30s_pc_fa_r2.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)

### 2025-05-07 scripts/run_zero_static_sim.py

@@ -1,13 +1,16 @@
+from pathlib import Path
 import numpy as np
 import datetime
-import network_conversion as conv
-import network_simulation as simu
+from echidna import network_conversion as conv
+from echidna import network_simulation as simu
+
+root = Path(__file__).resolve().parent.parent.resolve() 
 
 def preload():
 
-    ordering = conv.Ordering.from_file("./concordant_networks/size_14.csv")
+    ordering = conv.Ordering.from_file( root / "data/concordant_networks/size_14_nu.csv")
     converter = conv.StaticConverter.from_file(
-        "./concordant_networks/static_base_1_threshold_4.graphml", 
+        root / "data/concordant_networks/static_base_1_threshold_4.graphml", 
         ordering,
         purge_selfloops=True,
     )
@@ -17,7 +20,7 @@ def preload():
 
 def gen_params(converter):
 
-    prob_final = conv.ColumnDict.from_prob_final_file("./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
+    prob_final = conv.ColumnDict.from_prob_final_file(root / "data/concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
     prob_final_arr = prob_final.organise_by(converter.ordering)
     prob_final_zero =  np.zeros_like(prob_final_arr)
 
@@ -81,7 +84,7 @@ def run_sims(n_reps=20):
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = int(simdate + simtime)
-            outname = f"zero_sims/static/sim_all_30s_fixed_axis.h5"
+            outname = root / "simulations/zero_sims_resized/static/sim_all_30s_fixed_axis.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)
 
 if __name__ == "__main__":

### 2025-05-07 scripts/run_zero_temporal_sim.py

@@ -1,13 +1,16 @@
+from pathlib import Path
 import numpy as np
 import datetime
-import network_conversion as conv
-import network_simulation as simu
+from echidna import network_conversion as conv
+from echidna import network_simulation as simu
+
+root = Path(__file__).resolve().parent.parent.resolve()
 
 def preload():
 
-    ordering = conv.Ordering.from_file("./concordant_networks/size_14.csv")
+    ordering = conv.Ordering.from_file(root / "data/concordant_networks/size_14_nu.csv")
     converter = conv.TemporalNetworkConverter.from_file(
-        "concordant_networks/temponet_14_365.lgl", 
+        str(root / "data/concordant_networks/temponet_14_365.lgl"), 
         ordering=ordering,
         weight='weight',
     )
@@ -16,7 +19,7 @@ def preload():
 
 def gen_params(converter):
 
-    prob_final = conv.ColumnDict.from_prob_final_file("./concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
+    prob_final = conv.ColumnDict.from_prob_final_file(root/ "data/concordant_networks/probability_of_final_stay_by_shuffled_campus.csv")
     prob_final_arr = prob_final.organise_by(converter.ordering)
     prob_final_zero =  np.zeros_like(prob_final_arr)
 
@@ -85,7 +88,7 @@ def run_sims(n_reps=20):
             simdate = now.strftime('%y%m%d')
             simtime = now.strftime('%H%M%S')
             simid = (i, _n, simdate, simtime)
-            outname = f"zero_sims/temporal/sim_all_30s.h5"
+            outname = root / "simulations/zero_sims_resized/temporal/sim_all_30s.h5"
             simulate_sim_and_record(sim, simid=simid, until=8*365, nostop=True, outfile=outname, with_movers=True, simdate=simdate, simtime=simtime, seed=i)
 
 if __name__ == "__main__":

### 2025-05-07 scripts/run_zero_analysis.py
Extracct some of the analysis from notebooks to scripts
Probably next steps are to consolidate the running and analysis in a particular folder for paper export

@@ -1,0 +1,43 @@
+# assumes that metrics are computed
+import numpy as np
+from pathlib import Path
+import polars as pl
+from matplotlib import pyplot as plt
+import seaborn as sns
+
+root = Path(__file__).resolve().parent.parent.resolve()
+
+metrics_files = {
+    folder.stem: folder / "metrics_30s.parquet"
+    for folder in (root / "simulations/zero_sims_resized").iterdir()
+}
+
+metrics = {
+    label: pl.read_parquet(file)
+    for label , file in metrics_files.items()
+}
+
+_hitting_time_columns = pl.selectors.starts_with("hitting_time_")
+agg_metrics = dict()
+hitting_time_dists = dict()
+
+for model, df in metrics.items():
+    agg_metric = (df
+                  .unpivot(
+                      on=_hitting_time_columns,
+                      index='seed',
+                      variable_name='target',
+                      value_name='hitting_time',
+                  )
+                  .with_columns(
+                      target_seed = (
+                          pl.col('target')
+                          .str.strip_prefix('hitting_time_')
+                          .str.to_integer()
+                      )
+                  )
+                  .drop('target')
+                 ) 
+    agg_metrics[model] = agg_metric
+    hitting_time_dists[model] = np.sort(agg_metric.select("hitting_time").to_series().to_numpy())
+

### 2025-05-07 scripts/run_zero_metric_computation.py

@@ -1,0 +1,20 @@
+from echidna import netsim_summariser as summ
+from pathlib import Path
+import numpy as np
+import polars as pl
+
+root = Path(__file__).resolve().parent.parent.resolve()
+simulation_archives = {
+    folder.stem: folder / "sim_all_30s.h5"
+    for folder in (root / "simulations/zero_sims_resized").iterdir()
+}
+
+processed_results = dict()
+
+for label, archive in simulation_archives.items():
+    print(f"Analysing: {label}")
+    summariser = summ.Summariser(archive)
+    result_df = summariser.metrics(ncpus=6, no_move=True, drop=[]).add_extent(30)
+    result_df.write_parquet(archive.with_name("metrics_30s.parquet"))
+    processed_results[label] = result_df
+

### 2025-05-07 echidna/netsim_summariser.py
Allow for multiple extent captures on diff extent names

@@ -183,13 +183,13 @@ class Summariser():
 
 class Metrics(pl.DataFrame):
 
-    def add_extent(self, extent_time=30):
+    def add_extent(self, extent_time=30, extent_name="extent"):
         return self.with_columns(
             pl.concat_list(pl.selectors.starts_with('hitting_time'))
             .list.eval(pl.element().lt(extent_time))
             .list.sum().sub(1)
-            .alias('extent'),
-            pl.lit(extent_time).alias('extent_time')
+            .alias(extent_name),
+            pl.lit(extent_time).alias(f'{extent_name}_time')
         )
 
 

### 2025-05-07 echidna/network_simulation.py
Ignore extraneous sim init kwargs

@@ -8,7 +8,7 @@ import h5py
 
 
 class Simulation:
-    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, dt=1.0):
+    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, dt=1.0, **kwargs):
         self.N = np.asanyarray(full_size).reshape((-1, 1))
         self.NHOSP, _ = self.N.shape
         self.state = np.zeros_like(self.N)
@@ -21,6 +21,9 @@ class Simulation:
         ]  # softref here should update as initial self.state updates in seed
 
         self.rng = np.random.default_rng(None)
+
+        # remove additional init kwargs
+        _ = kwargs
 
     @property
     def history(self):
@@ -151,12 +154,14 @@ class SnapshotWithHome(SimulationWithMovers):
         timings: Sequence,
         track_movement=False,
         dt=1.0,
+        **kwargs,
     ):
         super().__init__(
             full_size=full_size,
             parameters=parameters,
             track_movement=track_movement,
             dt=dt,
+            **kwargs,
         )
 
         self.current_index = 0
@@ -229,12 +234,14 @@ class StaticWithHome(SimulationWithMovers):
         parameters: Mapping,
         dt=1.0,
         track_movement=False,
+        **kwargs,
     ):
         super().__init__(
             full_size=full_size,
             parameters=parameters,
             dt=dt,
             track_movement=track_movement,
+            **kwargs,
         )
 
         self.shadow_state = np.zeros((self.NHOSP, self.NHOSP), dtype=np.int64)
@@ -299,12 +306,14 @@ class TemporalSim(SimulationWithMovers):
             discretisation_size: float,
             track_movement=False,
             dt=1.0,
+            **kwargs,
     ):
         super().__init__(
             full_size=full_size,
             parameters=parameters,
             track_movement=track_movement,
             dt=dt,
+            **kwargs,
         )
 
         self.NT = num_times
@@ -390,11 +399,12 @@ class TemporalSim(SimulationWithMovers):
         self.time_travellers = np.zeros_like(self.time_travellers, dtype=np.int64)
 
 class SnapshotNaive(Simulation):
-    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, timings: Sequence, dt=1.0):
+    def __init__(self, full_size: Iterable[SupportsFloat], parameters: Mapping, timings: Sequence, dt=1.0, **kwargs):
         super().__init__(
[…]

### 2025-05-07 echidna/network_conversion.py
Add flexi name for reading from files
Make consistent so that all converters have a from_file

@@ -283,6 +283,11 @@ class SnapshotWithHomeConverter(Converter):
     def from_directory(cls, directory: str | Path, *args, **kwargs):
         snapshots = cls.load_snapshots(directory)
         return cls(snapshots, *args, **kwargs)
+    
+    @classmethod
+    def from_file(cls, file: str | Path, *args, **kwargs):
+        # thin wrapper for from_directory
+        return cls.from_directory(file, *args, **kwargs)
 
     @staticmethod
     def load_snapshots(rootpath: str | Path) -> Mapping[int, ig.Graph]:

### 2025-05-07 echidna/simulation_runner.py
Start prototype common runner

@@ -1,0 +1,129 @@
+import numpy as np
+
+from . import network_conversion as conv
+from . import network_simulation as simm
+
+from os import PathLike
+from typing import Mapping, Self, SupportsFloat, SupportsInt
+
+class Runner():
+
+    CLASSES = {
+        "naive_static": {
+            'converter': conv.NaiveStaticConverter,
+            'simulator': simm.Simulation,
+        },
+        "static": {
+            'converter': conv.StaticConverter,
+            'simulator': simm.StaticWithHome,
+        },
+        "snapshot": {
+            'converter': conv.SnapshotWithHomeConverter,
+            'simulator': simm.SnapshotWithHome,
+        },
+        "temporal": {
+            'converter': conv.TemporalNetworkConverter,
+            'simulator': simm.TemporalSim,
+        },
+    }
+
+    def __init__(
+        self,
+        simulation_class: str,
+        **kwargs,
+    ):
+        raise NotImplemented
+
+    def __full_init__(
+        self,
+        simulation_class: str,
+        ordering_file: PathLike,
+        graph_file: PathLike,
+        prob_final_file: PathLike,
+        beta: SupportsFloat,
+        output_file: PathLike,
+        zero_sim = False,
+        sim_n_seed_events: SupportsInt = 1,
+        sim_n_seed_infections: SupportsInt = 1,
+        sim_dt: SupportsFloat = 1.0,
+        sim_track_movement = False,
+        pseudo_capacity: SupportsFloat | None = None,
+        converter_args: Mapping | None = None,
+        simulation_args: Mapping | None = None,
+    ):
+        
+        self.simulation_class = simulation_class
+        self.base_classes = self.CLASSES[simulation_class]
+        
+        self.ordering_file = ordering_file
+        self.graph_file = graph_file
+        self.prob_final_file = prob_final_file
+        self.output_file = output_file
+        self.beta = beta
+
+        if converter_args is None: converter_args = dict()
+        self._create_parameter_converter(**converter_args)
+
+        if zero_sim:
+            self._create_zero_parameters()
+        else:
+            self._create_parameters()
+
+        if simulation_args is None: simulation_args = dict()
+        self._create_simulator(**simulation_args)
+
+    def _create_parameter_converter(self, **kwargs):
+
+        self.ordering = conv.Ordering.from_file(self.ordering_file)
[…]

### 2025-05-20 reports/_model.qmd
small content adds
Add some work

@@
 :::{#fig-general-diagram}
-```{.tikz}
-%%| format: svg
-%%| filename: movement_general_case
-%%| img-width: 800
-%%| fig-cap: general movement processes
-\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}
 
-\begin{tikzpicture}[
-  node distance=2cm and 4cm,
-  >=latex
-]
-  \tikzset{
-    squarx/.style= {
-      regular polygon,
-      regular polygon sides=4
-    }
-  }
-  \tikzset{
-    house/.style= {
-      single arrow,
-      shape border rotate=90,
-      single arrow head extend=0pt,
-      single arrow tip angle=120,
-      minimum height=0.75cm,
-      minimum width=0.75cm
-    }
-  }
-
-  \node[draw, squarx] (A) {$u$};
-  \node (X) [right=of A] {};
-  \node[draw, squarx] (B) [right=of X] {$v$};
-  \node[draw, house] (H) [below=of X] {$h_{uv}$};
-  \node[draw, circle] (D) [above=of X] {$\emptyset$};
-
-  \draw[->] (A) edge node[above]{$d_{uv}(s, t)$}  (B);
-  \draw[->] (A) edge node[above left]{$\zeta_u(s)$} (D);
-  \draw[->] (A) edge node[below left]{$\eta_{uv}(s,t)$} (H);
-  \draw[->] (H) edge node[below right]{$\rho_{uv}(s,t)$} (B);
-\end{tikzpicture}
-```
+![](diagrams/movement_general_case.pdf)
 
@@
 \lambda(u(s) \to v(t)) &= d_{uv}(s, t)\\
-\lambda(u(s) \to z_{uv}(t)) &= \eta{uv}(s, t)\\
-\lambda(z_{uv}(t) \to v(t)) &= \rho_{uv}(s, t)
+\lambda(u(s) \to z_{uv}(s, t)) &= \eta_{uv}(s, t)\\
+\lambda(z_{uv}(s, t) \to v(t)) &= \rho_{uv}(s, t)
 \end{aligned}$$
@@
 $$\begin{aligned}
-\lambda(u(s) \to {z'}_{uv}(t)) &= \eta_{uv}(s, t)\\
-\lambda({z'}_{uv}(t) \to z_{uv}(t)) &= \delta(t \mathrm{ mod } \omega)\\
+\lambda(u(s) \to {z'}_{uv}(s)) &= \sum_t \eta_{uv}(s, t)\\
+\lambda({z'}_{uv}(t) \to z_{uv}(t)) &= \delta(t \mod\omega)\\
 \end{aligned}$$
@@
 Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$.
-We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$
+We denote the smallest value in that set larger than a given time $t$ to be $\lceil{}t\rceil{}_\omega$
 For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly. 
@@
 $$\eta_{uv}(s, t) = \frac{\sum_{t: (t-s) \geq \omega} w_{uv}(s, t)}{\omega N_u}$$
-$$\rho_{uv}(s,t) = \frac{1}{\ceil{t}_\omega - t}$$
+$$\rho_{uv}(s,t) = \frac{1}{\lceil{}t\rceil{}_\omega - t}$$

### 2025-05-20 reports/paper.qmd
Add some work

---
title: "Temporality of the Victorian Hospital Patient Transfer Network"
format:
  plos-pdf:
    number-sections: false
    journal:
      id: ploscompbio
    include-in-header:
      text: |
        \usepackage{setspace}
        \doublespacing
    keep-tex: true
author:
  - name: Michael J Lydeamore
    affiliations:
      - ref: ebs
      - ref: vicdh
    email: michael.lydeamore@monash.edu
    corresponding: true
  - name: David Wu
    affiliations:
      - ref: ebs
  - name: Tjibbe Donker
    affiliations:
      - ref: freiburg
  - name: Ben Cooper
    affiliations:
      - ref: oxford
  - name: Marion Easton
    affiliations:
      - ref: vicdh
  - name: Nicholas Geard
    affiliations:
      - ref: uom
  - name: Claire Gorrie
    affiliations:
      - ref: doherty
      - ref: mdu
  - name: Daneeta Hennessy
    affiliations:
      - ref: vicdh
  - name: Benjamin Howden
    affiliations:
      - ref: mdu
  - name: Anton Y Peleg
    affiliations:
      - ref: id
  - name: Charlie Higgs
    affiliations:
      - ref: mdu
  - name: Andrew Wilson
    affiliations:
      - ref: vicdh
  - name: Andrew J Stewardson
    affiliations:
      - ref: id

affiliations:
  - id: ebs
    name: Department of Econometrics and Business Statistics, Monash University
    city: Melbourne
    state: Victoria
    country: Australia
  - id: vicdh
    name: Victorian Department of Health, Government of Victoria
    city: Melbourne
    state: Victoria
    country: Australia
  - id: freiburg
    name: Institute for Infection Prevention and Hospital Epidemiology, University Medical Center
    state: Freiburg
    country: Germany
  - id: uom
    name: School of Computing and Information systems, University of Melbourne
    city: Melbourne
    state: Victoria
    country: Australia
  - id: mdu
    name: Microbiological Diagnostic Unit Public Health Laboratory, Department of Microbiology & Immunology, University of Melbourne, at the Peter Doherty Institute for Infection and Immunity
    city: Melbourne
    state: Victoria
    country: Australia
  - id: doherty
    name: Department of Microbiology & Immunology, University of Melbourne, at the Peter Doherty Institute for Infection and Immunity
    city: Melbourne
    state: Victoria
    country: Australia
  - id: id
    name: Department of Infectious Diseases, The Alfred and Central Clinical School, Monash University
    city: Melbourne
    state: Victoria
    country: Australia
  - id: oxford
    name: Centre for Tropical Medicine and Global Health, Nuffield Department of Medicine, University of Oxford
    state: Oxford
    country: United Kingdom

bibliography: refs.bib
crossref:
  fig-prefix: Fig
execute:
  warning: false
  error: false
---

```{python}
#| label: compute-imports
import numpy as np
from pathlib import Path
import polars as pl
from matplotlib import pyplot as plt, dates as mdates
import seaborn as sns
import calendar
from itertools import accumulate
import datetime

root = Path("../").resolve()

sns.set_theme(
  context='paper',
  style='white',
  palette='colorblind',
  font='cmr10',
)
```

# Introduction

Antimicrobial resistance poses a great threat to human health and development. [@murray_global_2022]
Globally, it poses a large burden, operationally and economically, on hospital systems.
In Australia, we see [significant] economic burden from AMR.
this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR. 
One X of interest is carbapenemase-producing enterobacteriacea (CPE). 
Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level. 
This already causes some level of burden, and introduced strains would further multiply this.

## Setting

Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.

Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admission records is 26876787.
This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.

Some exploratory analysis of this data was done in by [@lydeamore_carbapenemase-producing_2024].

# Methods

## Existing modelling {#sec-network-modelling}

There has been some existing work that analyse hospital systems using the framework of network analysis. 
They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
Using these networks, hospitals and links that are important to mrough hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granulovement around the hospital system by analysing measures of centrality and clustering.
A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.
For each edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, 
we can construct a reaction that has a hazard function $\lambda_{ij} = w_{ij} / \tau$. This implicitly constructs a process that has inter-event times (i.e. times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.

Temporal networks are also not new phenomena in modelling infectious diseases processes. For example, 
[ref] [does thing]. However, edges of temporal networks typically represent

## Network models of patient transfers for temporal fidelity

:::{.hidden}
$$
\newcommand{\dd}[2][t]{\frac{\mathrm{d}#2}{\mathrm{d}#1}}
\newcommand{\ceil}[1]{\lceil#1\rceil}
$$
:::

![Sketch of the generalised model with processes of (1) permanent discharge ($u \to \emptyset$), (2) direct transfer ($u \to v$), (3) return home ($u \to h_{uv}$), (4) readmission $h_{uv} \to v$).](diagrams/movement_general_case.pdf){#fig-general-diagram}

Here, $u$ and $v$ are locations (potentially identical locations), $s$ and $t$ are times. We take that $t \geq s$.
$\emptyset$ represents patients that are not observed to return to the hospital system after their discharge from location $u$. We assume the associated hazard for this process ($\zeta_u(t)$) to be known for all locations $u$ at all times $t$: the modelling of this process is the same for all network models, and will be neglected in the following.

Further, we define the function $w_{uv}(s, t)$ to be the number of individuals that are observed to discharge from $u$ at time $s$ and readmit at time $t$. We allow a maximum of one of the pairs $u,v$ and $s,t$ to be reflexive, i.e. $u=v$ or $s=t$, but not both.

We will also assume that we will have quantities $N_u$ that represent the "size" of a location $u$. We assume that all locations $u$ will have $N_u$ individuals in them at all times, and that the observed mvoements $w_{uv}(s, t)$ can be thought of as being scaled by $N_u$.

The governing equations can be roughly expressed as a function of the hazards of each process.

$$
\begin{aligned}
\lambda(u(s) \to \emptyset) &= \zeta_u(s)\\
\lambda(u(s) \to v(t)) &= d_{uv}(s, t)\\
\lambda(u(s) \to z_{uv}(s, t)) &= \eta_{uv}(s, t)\\
\lambda(z_{uv}(s, t) \to v(t)) &= \rho_{uv}(s, t)
\end{aligned}
$$

We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the hodling time an individual patient spends in each state.

### Naïve Static Model

We can begin with a simple naive static model. This model will have the edge weights between healthcare facilities that represent the number of transfers from the source node to the target node over a certain period of observation time ($T_\Sigma$), regardless of how long the patient is absent from the healthcare facilities between initial discharge and subsequent readmission.

With reference to the general model above, the hazards of the  naïve static model are:

$$ d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u} $$
$$ \eta_{uv}(s,t) = \eta_{uv} = 0 $$
$$ \rho_{uv}(s,t) = \rho_{uv} = 0 $$

That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period. This is an oversimplification of the work done in @donker, but represents a "worst" case model of temporal fidelity that can be benchmarked against.

### Improved Static Model

We can introduce the concept of "indirectness" by allowing individuals to return home, in an "improved" static model.
In this model, we choose some threshold value $\omega$ that delineates direct transfers that occur "instantaneously" and indirect transfers that require a patient to first move to an intermediary "home" state before readmitting.

In comparison to the naive model, we will have non-zero hazards $\eta$ and $\rho$:

$$d_{uv}(s,t) = d_{uv} =  \frac{\sum_{s,t: (t-s) < \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = \frac{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv} = \left[ \frac{\sum_{s,t: (t-s) \geq \omega} (t-s) w_{uv}(s,t)}{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}  \right]^{-1}$$

Here, we model the rate of indirect transfers $\eta$ similarly to the rate of direct transfers  $d$, by counting the average rate of observed movements. For the rate of readmission, $\rho$, we use the inverse of the mean readmission duration/delay.

### Snapshot Model

Of course, as the hospital system changes over time, the rate of patient transfers, and thus the hazards represented in our model should also change.
We see evidence of this in @fig-movement-over-time where we plot the overall rates of movement in the network over time.

```{python}
#| label: compute-movement-over-time
base_date = datetime.date(2011, 1, 1)
aggregated_movement = pl.read_csv(root / "data/concordant_networks/analyses/movement_aggregated.csv")
aggregated_movement = (
  aggregated_movement
  .group_by('t')
  .agg(pl.col('w').sum())
  .sort('t')
  .with_columns(
    date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
    my=pl.col('w').rolling_mean(window_size=365, center=True),
    mw=pl.col('w').rolling_mean(window_size=7, center=True),
    mm=pl.col('w').rolling_mean(window_size=28, center=True),
  )
)
```

```{python}
#| label: fig-movement-over-time
#| fig-cap: Number of patient transfers throughout the network increases over time. Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period. We also observe seasonal behaviour (most visually obvious with periods of a year and a week), and consistent decreases of transfers during public holiday periods, especially at the end of the year.
fig = plt.figure()
ax = fig.add_subplot()
sns.despine(ax=ax)

nth_year = 4
x1, x2 = datetime.date(2011+nth_year-1, 12, 15), datetime.date(2011+nth_year+1, 1, 15)
y1, y2 = 12_000, 25_000
inset = ax.inset_axes(
  [0.15, 0.1, 0.7, 0.35], 
  xlim=(x1, x2), 
  ylim=(y1, y2),
  xlabel='',
  ylabel='',
  yticks=[],
  yticklabels=[],
)
for axes in (ax, inset):
  sns.lineplot(aggregated_movement, x='date', y='w', color='C0', ax=axes)

sns.lineplot(aggregated_movement, x='date', y='my', color='red', ax=ax)
sns.lineplot(aggregated_movement, x='date', y='mw', color='red', ax=inset)

ax.indicate_inset_zoom(inset, edgecolor='black')

inset.set_xlabel('')
inset.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
_ = ax.set_xlabel('Date')
_ = ax.set_ylabel('Number of patient transfers')
```

We approach this using static network snapshots of the hospital system over time.

For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.
This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
This alters the indirect transfer processes to

$$
\begin{aligned}
\lambda(u(s) \to {z'}_{uv}(s)) &= \sum_t \eta_{uv}(s, t)\\
\lambda({z'}_{uv}(t) \to z_{uv}(t)) &= \delta(t \mod\omega)\\
\end{aligned}
$$

where $\delta(\cdot)$ is the Dirac delta function.
Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$. We denote this quantity $H_{uv}(t)$
We will have that:

$$d_{uv}(s,t) = d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s,t)}{\omega N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv}(s) = \frac{\sum_{t: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv}(t) = \frac{\sum_{s: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega H_{uv}}$$

### Temporal Model

For the temporal network, we make a choice of the time discretisation $\omega$. That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$. Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$. We denote the smallest value in that set larger than a given time $t$ to be $\lceil{}t\rceil{}_\omega$
For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly. 
This yields:

$$d_{uv}(s, t) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\eta_{uv}(s, t) = \frac{\sum_{t: (t-s) \geq \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\rho_{uv}(s,t) = \frac{1}{\lceil{}t\rceil{}_\omega - t}$$

# Results

## Simulation Study

Since analytical results are intractable for the more complex network models, we resort to simulation as a statistical proxy.
Movements of marked individuals are tracked for each network type, when they are seeded at each facility in the network at $t=0$. The hitting time $k_j$ to each other hospital is recorded.
We mark $N_0 = 30$ individuals at the start of each realisation; we generate $N_r=20$ realisations for each seed facility.

We plot the (empirical) cumulative distribution function of the hitting times for all pairs of seed facility and target facility in @fig-ecdf

```{python}
#| label: compute-aggregating-sim-metrics
metrics_files = {
  folder.stem: folder / "metrics_30s.parquet"
    for folder in (root / "simulations/zero_sims_resized").iterdir()
}

metrics = {
    label: pl.read_parquet(file)
    for label , file in metrics_files.items()
}

_hitting_time_columns = pl.selectors.starts_with("hitting_time_")
agg_metrics = dict()
hitting_time_dists = dict()
extent_dists = dict()

for model, df in metrics.items():
    agg_metric = (df
                .unpivot(
                    on=_hitting_time_columns,
                    index=['seed', 'extent'],
                    variable_name='target',
                    value_name='hitting_time',
                )
                .with_columns(
                    target_seed = (
                        pl.col('target')
                        .str.strip_prefix('hitting_time_')
                        .str.to_integer()
                    )
                )
                .drop('target')
                ) 
    agg_metrics[model] = agg_metric
    hitting_time_dists[model] = np.sort(agg_metric.select("hitting_time").to_series().to_numpy())
    extent_dists[model] = np.sort(agg_metric.select('extent').to_series().to_numpy())
```
```{python}
#| label: fig-ecdf
#| fig-cap: Empirical cumulative distributions of the hitting times between any pair of facilities in the network, for different network models. $N_0 =30$ individuals are seeded initially, all at one facility, and then allowed to propagate through the network.

fig = plt.figure()
ax = fig.add_subplot()
for model, dist in hitting_time_dists.items():
    ax.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model)

ax.set_xlabel('hitting time')
ax.set_ylabel('ECDF')

ax.legend(loc='lower right')
sns.despine(ax=ax)
```

```{python}
#| label: fig-extent
#| fig-cap: Number of facilities that have had a marked individual admitted after $t=30$ days. The naïve static model overestimates the range of the spread.

max_extent = max(max(dist) for dist in extent_dists.values())
extent_bins = np.arange(max_extent+2)

# fig = plt.figure()
# ax = fig.add_subplot()
fig, axz = plt.subplots(nrows=4, sharey=True, sharex=True)
axs = axz.flatten()

model_order = sorted(extent_dists.keys())
for ax, model in zip(axs, model_order):
  ax.hist(
    extent_dists[model],
    bins=extent_bins,
    density=True,
  )
  ax.set_ylabel(model)

axs[-1].set_xlabel('Extent')
```

# Discussion

We have investigated the effects of modelling the network representations of patient transfers between healthcare facilities in Victoria with differing levels of temporal fidelity. We see that there are qualitative behavioural differences between the movement patterns induced by these different network representations.
One unique temporal property of the networks here is that edges can connect nodes not just in space, but also in time. This leads to a reservoir of individuals of interest that exist outside of the healthcare system that will return at a later time, which leads to rich temporal dynamics. These behaviours are poorly captured by a naïve static network; including this mechanism on top of a static network representation can help significantly in moving the qualitative behaviour fo the

# Supplementary

## Verification of simulation results

For any continuous-time Markov chain, we can derive the Q-matrix, which represents how ...
Importantly, we can use the forward equation to generate a viable representation of the evolution of the probability density function as a system of ODEs that scales with the number of states.

Typically, the structure of these models represents the state of the _system_; here we use the model as a representation of teh state of an _individual_. Thus, we can envision a simulation with $N_0$ initial seed individuals as being equivalent to $N_0$ independent, but identically distributed, Markov chains.
Here, then, a state maps onto a facility; for more complex models, a state may map onto either a facility, or a "home" state that could be uniquely identified by a pair of facilities.

This make computing the probability density function very computationally expensive when the number of Markov states becomes large. This is due to the fact that the ODE system will tend towards very dense matrices $P$ if the network is connected, even if $Q$ itself is sparse. This makes some representations, in particular the temporal network, completely intractable, since we would have to allocate a ${(N_f)}^2 \times {(N_T)}^2$ floating point matrix for each time step. For $N_f \sim O(10^2), N_T \sim O(10^3)$, this yields a roughly $O(10^{(10\sim 11)}$ byte (10\~100 GB) float32 matrix every time step (in memory).

Thus, we only solve this problem for the naïve static network case.
Below, we plot the analytical solution against the empirical mean hitting time of pairs that have all realisations hitting.

::: {#fig-naive-static-verification}

```{python}
#| label: compute-naive-static-verify
model = 'naive_static'
dist = hitting_time_dists['naive_static']
n_null_accept = 1 
agg_stats = (
  agg_metrics[model]
  .group_by('seed', 'target_seed')
  .agg(
    pl.col('hitting_time').mean().alias('hitting_time_mean'),
    # pl.col('extent').mean().alias('extent_mean'),
    pl.col('hitting_time').median().alias('hitting_time_median'),
    # pl.col('extent').median().alias('extent_median'),
    pl.col('hitting_time').null_count().alias('hitting_time_nulls'),
    # pl.col('extent').null_count().alias('extent_nulls'),
  )
  .with_columns(
    pl.when(pl.col('hitting_time_nulls') > n_null_accept).then(pl.lit(None)).otherwise(pl.col('hitting_time_mean')).alias('hitting_time_mean_nld'),
    pl.when(pl.col('hitting_time_nulls') > n_null_accept).then(pl.lit(None)).otherwise(pl.col('hitting_time_median')).alias('hitting_time_median_nld'),
  )
)

hit_mean = agg_stats.select('hitting_time_mean_nld').to_series().sort(nulls_last=True).to_numpy()
hit_median = agg_stats.select('hitting_time_median_nld').to_series().sort(nulls_last=True).to_numpy()

# import analytical
analytical_naive_static_soln = pl.read_csv(root / "outputs/hitting_time_analysis/numer_hitting_30s_nu_size.csv")
analytical_dist = analytical_naive_static_soln.select('hitting_time').to_series().sort(nulls_last=True).to_numpy()

N_pairs = agg_stats.shape[0]
dist_len = len(dist)
N_anal = len(analytical_dist)

fig, ax = plt.subplots()
# ax.plot(
  # dist, np.linspace(0, 1, dist_len, endpoint=False), 
  # label='empirical distribution'
# )
ax.plot(
  hit_mean, np.linspace(0, 1, N_pairs, endpoint=False),
  label='empirical mean'
)
# ax.plot(hit_median, np.linspace(0, 1, N_pairs, endpoint=False), label='median')
xlim = ax.get_xlim()
ylim = ax.get_ylim()
ax.plot(
  analytical_dist, np.linspace(0, 1, N_anal, endpoint=False),
  linestyle='dashed',
  label='analytical mean'
)
ax.set_xlim(xlim[0], max(dist) * 1.15)
# ax.set_ylim(ylim)
ax.legend()

ax.set_xlabel('time')
ax.set_ylabel('ECDF')
sns.despine(ax=ax)
```

Verification of the empirical simulation results against analytical results. We see that at early time the mean empirical and analytical expected hitting times agree, with a large right-hand censoring effect. When computing the empirical means, we allow `{python} n_null_accept` nulls across all realisations.
:::

### 2025-05-20 echidna/hitting_time_multichain.py

@@ -47,6 +47,7 @@ def solve_survival_ode(Q, index=-1, locations=None, endpoint=200):
 
 
 def compute_multichain_hitting_time(Nchains, Q, index=-1, precision=30, scaling=1.0):
+    print(index, end='\t', flush=True)
     roots, weights = determine_laguerre_roots_and_weights(precision)
     gl_weights = weights * np.exp(roots)
 

### 2025-05-20 scripts/hitting_time_analysis.py

@@ -1,21 +1,27 @@
+from os import PathLike
 from typing import Mapping
+from pathlib import Path
 import numpy as np
 import igraph as ig
+import polars as pl
 from matplotlib import pyplot as plt, colors
 from scipy import sparse
 
-import network_conversion as ntc
-import hitting_markov as hit
-from realisation_summary_stats import hitting_time
+from echidna import network_conversion as ntc
+from echidna import hitting_markov as hit
+from echidna import examine_transfers_for_sizes as esz
+
+root = Path(__file__).resolve().parent.parent.resolve()
 
 def graph_read(filename: str):
     return ig.Graph.Read(filename)
 
 ## NAIVE STATIC MODEL
 def get_naive_static_Q(
-    graph: str | ig.Graph = "./concordant_networks/shuf_static_network.graphml",
-    size_file="./concordant_networks/size_14.csv",
-    time_span = (10*365),
+    graph: str | PathLike | ig.Graph = root / "data/concordant_networks/shuf_static_network.graphml",
+    size_file: str | PathLike = root / "data/concordant_networks/size_14_nu.csv",
+    duration_file: str | PathLike = root / "data/concordant_networks/durations_1.csv",
+    default_time_span = (10*365),
     ordering_key = 'name',
     value_key = 'id',
 ):
@@ -33,14 +39,19 @@ def get_naive_static_Q(
     size_order = ntc.Ordering.from_file(size_file)
     sizes = np.array([size_order(i) for i in graph_ordering.order])
 
+    durations = esz.quick_read(duration_file)
+    duration_values = np.array([durations.get(i, default_time_span) 
+                                for i in graph_ordering.order])
+
+    denominator = sizes * duration_values
+
     # we use the naive construction, since the null set (never return) version creates 
     # results that misrepresent the amount of time it takes to move through the system
     # We also use the internal function in order to directly pass the transition matrix 
     # constructed, as opposed to the non-scaled weighted adjacency matrix
     T = ntc.transition_matrix_from_graph(G, 
                                         ordering=graph_ordering,
-                                        scaling_per_node=sizes,
-                                        global_scaling=time_span,
+                                        scaling_per_node=denominator,
                                         ordering_key=ordering_key, 
                                         adjacency_attribute='weight',
                                         )

### 2025-05-20 scripts/run_zero_analysis.py

@@ -125,11 +125,25 @@ def main_plot():
     ax.set_xlim(xlim)
     ax.set_ylim(ylim)
 
-    ax.figure.savefig("outputs/zero_hitting_time_dur_upd.png")
+    ax.set_xlim(-5, 250)
+    ax.set_ylim(-1e-3, 0.1)
+
+
+    sns.despine(ax=ax)
+
+    ax.figure.savefig("outputs/zero_hitting_time_dur_upd_zoomed_in.pdf")
     return ax
 
 if __name__ == "__main__":
+    import argparse
 
-    analytical_file = root / "outputs/hitting_time_analysis/numer_hitting_30s_nu_size.csv"
-    compute_analytical_naive_static_hitting(analytical_file)
+    parser = argparse.ArgumentParser()
+    parser.add_argument("--force", '-f', action='store_true', 
+                        help='force recomputation of analytical solution')
+    args = parser.parse_args()
+
+    if args.force:
+        analytical_file = root / "outputs/hitting_time_analysis/numer_hitting_30s_nu_size.csv"
+        compute_analytical_naive_static_hitting(analytical_file)
+    
     ax = main_plot()

### 2025-05-20 scripts/run_zero_home_prop.py

@@ -1,0 +1,70 @@
+import functools
+import numpy as np
+from scipy import linalg, stats
+from matplotlib import pyplot as plt
+import seaborn as sns
+import polars as pl
+import h5py
+from pathlib import Path
+
+# arrays are NxM, (observations are rows, columns are time)
+
+
+root = Path(__file__).resolve().parent.parent.resolve()
+
+def gather(write=True):
+    histories = dict()
+    for dir in (root / "simulations/zero_sims_resized/" ).iterdir():
+        model = dir.stem
+        fl = dir / "sim_all_30s.h5"
+        with h5py.File(fl) as fp:
+            grps = list(fp)
+            presence = [fp[grp]['history'][:].sum(axis=0) for grp in grps]
+        histories[model] = np.vstack(presence)
+        if write:
+            write_to_file = dir / "hosp_presence.csv"
+            np.savetxt(write_to_file, histories[model], delimiter=',')
+            print(f"written to {write_to_file}")
+    return histories
+
+def load():
+    histories = {
+        dir.stem: np.loadtxt(dir/"hosp_presence.csv", delimiter=',')
+        for dir in (root / "simulations/zero_sims_resized/" ).iterdir()
+    }
+
+    return histories
+
+def heatmap(hist, nmax=30):
+    # hist is 2D
+    bins = np.arange(nmax+2)
+    hist_heat = np.array([
+        np.histogram(hist[:, i], bins=bins)[0]
+        for i in range(hist.shape[1])
+    ]).T
+
+    return bins, hist_heat
+
+@functools.wraps(heatmap)
+def plot_heatmap(*args, **kwargs):
+    bins, hist_heat = heatmap(*args, **kwargs)
+
+    ax = sns.heatmap(hist_heat)
+
+    return ax
+
+def mega_heatmap(hists, vmax=1500):
+    hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
+    fig, axs = plt.subplots(2, 2, figsize=[16, 9])
+    for ax, (k, hmv) in zip(axs.flatten(), hms.items()):
+        sns.heatmap(data=hmv[::-1,:], vmax=vmax, ax=ax, cbar=False)
+        ax.set_title(k)
+    fig.colorbar(ax.collections[0], ax=axs, label='Count', extend='max')
+    fig.suptitle('Number at home')
+
+    return fig
+
+
+
+if __name__ == "__main__":
+    gather()

### 2025-05-27 reports/paper.qmd
Update the home prop fig

@@
 from itertools import accumulate
+from collections import defaultdict
 import datetime
@@
 
-axs[-1].set_xlabel('Extent')
+_ = axs[-1].set_xlabel('Extent')
+```
+
+One striking feature
+
+```{python}
+histories = {
+  dir.stem: np.loadtxt(dir/"hosp_presence.csv", delimiter=',')
+  for dir in (root / "simulations/zero_sims_resized/" ).iterdir()
+}
+
+_BASE_DATE = np.datetime64('2011-01', 'M')
+_MAX_DATE = np.datetime64('2018-12', 'M')
+def set_xdatelabels(hist, ax, spacing=12, dt=1.0):
+    # here, we want to get succinct date formats
+    # we map the times onto the timedeltas of the simulations
+    N = hist.shape[1]
+    emp_max_date = (np.timedelta64(int(N*dt), 'D') + _BASE_DATE).astype('datetime64[M]')
+    ax_max_date = np.min([emp_max_date, _MAX_DATE])
+
+    dates = np.arange(_BASE_DATE, ax_max_date, np.timedelta64(spacing, 'M'))
+
+    label_pos = (dates.astype('datetime64[D]').astype('int64') - _BASE_DATE.astype('datetime64[D]').astype('int64')) / dt
+
+    ax.set_xticks(label_pos, dates.astype('datetime64[Y]'), rotation=45, rotation_mode='anchor')
+
+    return ax.get_xticks()
+
+def heatmap(hist, nmax=30):
+    # hist is 2D
+    bins = np.arange(nmax+2)
+    hist_heat = np.array([
+        np.histogram(hist[:, i], bins=bins)[0]
+        for i in range(hist.shape[1])
+    ]).T
+
+    return bins, hist_heat
+```
+
+```{python}
+def mega_heatmap(hists, vmax=1500):
+  hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
+  fig, axs = plt.subplots(nrows=4, figsize=[16, 9], sharex=True)
+  for ax, (k, hmv) in zip(axs.flatten(), hms.items()):
+      if k == 'temporal':
+        hmv = hmv[:,::2]
+      sns.heatmap(data=hmv[::-1,:], vmax=vmax, ax=ax, cmap='inferno', cbar=False)
+      ax.set_title(k)
+      ax.set_ylim([-0.5, 30.5])
+      ax.set_yticks([0, 10, 20, 30])
+  set_xdatelabels(hmv, ax)
+  fig.colorbar(ax.collections[0], ax=axs, label='Count', extend='max')
+
+  return fig
+
+_ = mega_heatmap(histories)
 ```

### 2025-05-27 scripts/run_zero_home_prop.py

@@ -1,7 +1,10 @@
+from collections import defaultdict
 import functools
+from typing import Mapping
 import numpy as np
 from scipy import linalg, stats
 from matplotlib import pyplot as plt
+from matplotlib import dates as mpldates
 import seaborn as sns
 import polars as pl
 import h5py
@@ -35,6 +38,24 @@ def load():
 
     return histories
 
+_BASE_DATE = np.datetime64('2011-01', 'M')
+_MAX_DATE = np.datetime64('2018-12', 'M')
+def set_xdatelabels(hist, ax: plt.Axes, spacing=12, dt=1.0):
+    # here, we want to get succinct date formats
+    # we map the times onto the timedeltas of the simulations
+    N = hist.shape[1]
+    emp_max_date = (np.timedelta64(int(N*dt), 'D') + _BASE_DATE).astype('datetime64[M]')
+    
+    ax_max_date = np.min([emp_max_date, _MAX_DATE])
+
+    dates = np.arange(_BASE_DATE, ax_max_date, np.timedelta64(spacing, 'M'))
+
+    label_pos = (dates.astype('datetime64[D]').astype('int64') - _BASE_DATE.astype('datetime64[D]').astype('int64')) / dt
+
+    ax.set_xticks(label_pos, dates.astype('datetime64[Y]'), rotation=45)
+
+    return ax.get_xticks()
+
 def heatmap(hist, nmax=30):
     # hist is 2D
     bins = np.arange(nmax+2)
@@ -55,16 +76,17 @@ def plot_heatmap(*args, **kwargs):
 
 def mega_heatmap(hists, vmax=1500):
     hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
-    fig, axs = plt.subplots(2, 2, figsize=[16, 9])
+    fig, axs = plt.subplots(nrows=4, figsize=[16, 9], sharex=True)
     for ax, (k, hmv) in zip(axs.flatten(), hms.items()):
-        sns.heatmap(data=hmv[::-1,:], vmax=vmax, ax=ax, cbar=False)
+        if k == "temporal": hmv=hmv[:,::2]
+        sns.heatmap(data=hmv[::-1,:], vmax=vmax, ax=ax, cmap='inferno', cbar=False)
         ax.set_title(k)
+        ax.set_ylim([-0.5, 30.5])
+    set_xdatelabels(hmv, ax)
+    fig.subplots_adjust(hspace=0.4)
[…]

### 2025-05-27 echidna/simulation_runner.py
add hack for sim runner dep on converter

@@ -30,15 +30,8 @@ class Runner():
     def __init__(
         self,
         simulation_class: str,
-        **kwargs,
-    ):
-        raise NotImplemented
-
-    def __full_init__(
-        self,
-        simulation_class: str,
-        ordering_file: PathLike,
         graph_file: PathLike,
+        ordering_file: PathLike,
         prob_final_file: PathLike,
         beta: SupportsFloat,
         output_file: PathLike,
@@ -47,7 +40,6 @@ class Runner():
         sim_n_seed_infections: SupportsInt = 1,
         sim_dt: SupportsFloat = 1.0,
         sim_track_movement = False,
-        pseudo_capacity: SupportsFloat | None = None,
         converter_args: Mapping | None = None,
         simulation_args: Mapping | None = None,
     ):
@@ -69,8 +61,20 @@ class Runner():
         else:
             self._create_parameters()
 
+        self.base_sim_args = dict(
+            n_seed_events=sim_n_seed_events,
+            n_seed_infections=sim_n_seed_infections,
+            dt=sim_dt,
+            track_movement=sim_track_movement,
+        )
         if simulation_args is None: simulation_args = dict()
-        self._create_simulator(**simulation_args)
+        self.simulation_args = dict(
+            **self.base_sim_args
+        )
+        self.simulation_args.update(simulation_args)
+        self._create_simulator(
+            **self.simulation_args,
+        )
 
     def _create_parameter_converter(self, **kwargs):
 
@@ -105,6 +109,11 @@ class Runner():
         else:
             full_sizes = self.ordering.sizes
 
+        # here we map converter-based args
+        for arg, val in simulator_args.items():
+            if callable(val):
+                simulator_args[arg] = val(self.converter)
+
         self.sim = self.base_classes['simulator'](
             full_size=full_sizes,
             parameters=self.parameters,
@@ -123,7 +132,21 @@ class Runner():
         self.sim.simulate(until=until, nostop=nostop)
 
 
-    def export(self, to, identity, with_movers=False, **kwargs):
+    def export(self, to=None, identity=None, with_movers=False, **kwargs):
+
+        if to is None:
+            to = self.output_file
+        
+        if identity is None:
+            identity = self._export_identity()
 
         self.sim.export_history(to, identity=identity, with_movers=with_movers, **kwargs)
 
+
+    def _export_identity(self) -> str:
+        
+        iden = ""
+        if hasattr(self.sim, 'rng'):
+            iden = self.sim.rng.state['state']['state']
+        
+        return iden

### 2025-06-03 reports/paper.qmd
Upd rep

--- 2025-05-20 reports/paper.qmd
@@
     keep-tex: true
+    fig-format: png
+    fig-dpi: 300
 author:
@@
 import polars as pl
-from matplotlib import pyplot as plt, dates as mdates
+from matplotlib import pyplot as plt, dates as mdates, colors as mcolors
 import seaborn as sns
@@
   context='paper',
-  style='white',
+  style='ticks',
   palette='colorblind',
-  font='cmr10',
+  font='serif',
 )
@@
 For the temporal network, we make a choice of the time discretisation $\omega$.
-That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$.
+Events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$.
 Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$.
@@
 ax = fig.add_subplot()
-for model, dist in hitting_time_dists.items():
-    ax.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model)
+model_order = ["naive_static", "static", "snapshot", "temporal"] 
+for model in model_order:
+    dist = hitting_time_dists[model]
+    ax.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model.replace('_', ' '))
 
@@
 
-model_order = sorted(extent_dists.keys())
+model_order = ['temporal', 'snapshot', 'static', 'naive_static'] 
 for ax, model in zip(axs, model_order):
@@
   )
-  ax.set_ylabel(model)
+  ax.set_ylabel(model.replace('_', '\n'))
+  sns.despine(ax=ax)
 
-axs[-1].set_xlabel('Extent')
+_ = axs[-1].set_xlabel('Extent')
+```
+
+One striking feature
+
+```{python}
+histories = {
+  dir.stem: np.loadtxt(dir/"hosp_presence.csv", delimiter=',')
+  for dir in (root / "simulations/zero_sims_resized/" ).iterdir()
+}
+```
+```{python}
+#| label: fig-homeprop-heatmap
+#| fig-cap: Distribution of the number of indivdiuals at home (not in the hospital system) over the duration of the simulation for each model type.
+
+_BASE_DATE = np.datetime64('2011-01', 'M')
+_MAX_DATE = np.datetime64('2018-12', 'M')
+def set_xdatelabels(hist, ax, spacing=12, dt=1.0):
+    # here, we want to get succinct date formats
+    # we map the times onto the timedeltas of the simulations
+    N = hist.shape[1]
+    emp_max_date = (np.timedelta64(int(N*dt), 'D') + _BASE_DATE).astype('datetime64[M]')
+    ax_max_date = np.min([emp_max_date, _MAX_DATE])
+
+    dates = np.arange(_BASE_DATE, ax_max_date, np.timedelta64(spacing, 'M'))
+
+    label_pos = (dates.astype('datetime64[D]').astype('int64') - _BASE_DATE.astype('datetime64[D]').astype('int64')) / dt
+
+    ax.set_xticks(label_pos, dates.astype('datetime64[Y]'), rotation=45, rotation_mode='anchor')
+
+    return ax.get_xticks()
+
+_DTS = defaultdict(lambda: 1.0)
+_DTS['temporal'] = 0.5
+def single_heatmap(model, hist, bins=(540, 31), norm=None, ax=None, cbar=False):
+    if ax is None:
+        fig, ax = plt.subplots()
+
+    if norm is None:
+        norm = mcolors.Normalize()
+
+    dt = _DTS[model]
+    ts = (np.arange(hist.shape[1]) * dt).reshape((1, -1))
+    ts_arr = np.repeat(ts, hist.shape[0], axis=0)
+
+    *_, coll = ax.hist2d(
+        ts_arr.flatten(), 
+        30 - hist.flatten(),
+        bins=bins,
+        density=True,
+        norm=norm,
+    )
+
+    if cbar:
+        ax.figure.colorbar(coll, ax=ax)
+
+    return coll, ax
+
+_ax_order = {
+    'temporal': 0,
+    'snapshot': 1,
+    'static': 2,
+    'naive_static': 3,
+}
+def final_heatmap(hists, norm=mcolors.PowerNorm(0.5)):
+
+    fig, axs = plt.subplots(nrows=4, figsize=[8.25, 5.25], sharex=True)
+    axs_flat = axs.flatten()
+    for k, h in hists.items():
+        ax = axs_flat[_ax_order[k]]
+        coll, _ = single_heatmap(k, h, norm=norm, ax=ax)
+        norm = coll.norm
+        ax.set_ylabel(k.replace('_', '\n'))
+    set_xdatelabels(h, axs_flat[-1])
+    fig.subplots_adjust(hspace=0.3)
+    fig.colorbar(coll, ax=axs, label='Density', extend='max')
+    fig.supylabel("Number at home ($N_0=30$)")
+
+    return fig, axs
+
+_ = final_heatmap(histories)
 ```
@@
 These behaviours are poorly captured by a naïve static network; including this mechanism on top of a static network representation can help significantly in moving the qualitative behaviour fo the
+
+# References
+::: {#refs}
+:::
+
+::: {.content-visisble when-format="pdf"}
+\newpage
+:::
 
@@
 :::
+
+## Uniform dissapation of at-home individuals
+
+We know that for a random variable $t$ uniformly distributed between $[\tau, \tau+\omega]$, the pdf and survival functions are, respectively:
+
+$$
+\begin{gathered}
+f(t) = \frac{1}{\omega}, \\
+S(t) = \frac{(\tau+\omega) - t}{\omega},
+\end{gathered}
+$$
+
+which results in a hazard function
+
+$$
+\lambda(t) = \frac{1}{(\tau + \omega) - t}.
+$$
+
+This hazard function can be simulated approximately in our DTMC framework by taking small time steps, and modelling the number of individuals $dN$ that move in a time step $dt$, where $n(t)$ is the number of inidividuals still present at time $t$ as
+
+$$dN(t) \sim \text{Binomial}(n(t), dt\lambda(t)).$$

### 2025-06-03 reports/diagrams/network_diags/temporal.tex
new diagrams

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}[scale = 0.20]

\tikzset{
    wavy/.style={
        decorate,
        decoration={
            snake,
            pre length=2mm,
            post length=2mm
        }
    }
}

\node at (19.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A0){A};
\node at (19.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B0){B};
\node at (61.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A1){A};
\node at (61.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B1){B};

% direct transfers
\draw [-stealth,solid](A0.247)|-++(0,-1)-|(B0.113);
\draw [-stealth,solid](B0.67)|-++(0,1)-|(A0.293);
\draw [-stealth,solid](A1.247)|-++(0,-1)-|(B1.113);
\draw [-stealth,solid](B1.67)|-++(0,1)-|(A1.293);

% time consistency
\draw [-stealth,wavy](A0.east)--(A1.west);
\draw [-stealth,wavy](B0.east)--(B1.west);

% indirect
\draw [out=45,in=135,-stealth,dashed,bend left=30](A0.45) to (A1.135);
\draw [out=315,in=225,-stealth,dashed, bend right=30](B0.315) to (B1.225);
\draw [out=315,in=135,-stealth,dashed](A0.315) to (B1.135);
\draw [out=45, in=225,-stealth,dashed](B0.45) to (A1.225);

\end{tikzpicture}

\end{document}

### 2025-06-03 scripts/run_zero_home_prop.py
concord with paper

@@ -1,12 +1,8 @@
 from collections import defaultdict
 import functools
-from typing import Mapping
 import numpy as np
-from scipy import linalg, stats
-from matplotlib import pyplot as plt
-from matplotlib import dates as mpldates
+from matplotlib import pyplot as plt, colors as mcolors
 import seaborn as sns
-import polars as pl
 import h5py
 from pathlib import Path
 
@@ -27,16 +23,36 @@ def gather(write=True):
         if write:
             write_to_file = dir / "hosp_presence.csv"
             np.savetxt(write_to_file, histories[model], delimiter=',')
-            print(f"written to {write_to_file}")
+            print(f"presence written to {write_to_file}")
     return histories
 
-def load():
+def load(fname="hosp_presence.csv"):
     histories = {
-        dir.stem: np.loadtxt(dir/"hosp_presence.csv", delimiter=',')
+        dir.stem: np.loadtxt(dir/fname, delimiter=',')
         for dir in (root / "simulations/zero_sims_resized/" ).iterdir()
     }
 
     return histories
+
+
+def compute_heatmaps(hists, write=True):
+    hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
+    # adjust temporal for dt
+    hms['temporal'] = np.hstack([
+        (hms['temporal'][:,:-1:2] + hms['temporal'][:, 1::2]) / 2,
+        hms['temporal'][:,-1:],
+    ])
+
+    if write:
+        for model, hmarr in hms.items():
+            write_to_file = root / "simulations/zero_sims_resized" / model / "hosp_presence_heat.csv"
+            np.savetxt(write_to_file, hmarr, delimiter=',')
+            print(f"heatmap written to {write_to_file}")
+
+    return hms
+
+def load_heatmaps():
+    return load("hosp_presence_heat.csv")
 
 _BASE_DATE = np.datetime64('2011-01', 'M')
 _MAX_DATE = np.datetime64('2018-12', 'M')
@@ -74,11 +90,57 @@ def plot_heatmap(*args, **kwargs):
 
     return ax
 
-def mega_heatmap(hists, vmax=1500):
-    hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
[…]

### 2025-06-04 reports/paper.qmd
smol upd

@@
 this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR. 
-One X of interest is carbapenemase-producing enterobacteriacea (CPE). 
+One X of interest, for example, is carbapenemase-producing enterobacteriacea (CPE). 
 Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level. 
@@
 
+The spread of AMR in these systems has been studied by modelling the hospital system as a network or interconnected healthcare facilities.
+For example, [@donker_patient_2010] examines hospitals in the Netherlands to study the spread of MRSA, followed by a similar analysis of the UK hospital system in [@donker_hospital_2012].
+Ultimately, the aim of studying these systems as networks is leveraging the network framework when designing surveillance and control protocols for emerging and recurring AMR outbreaks.
+This has been analysed to some extent in [@ciccolini_efficient_2014], for example; but such analyses use
+
+There are existing results that show that accounting for temporality in network models results in lower control costs [@li_fundamental_2017].
+In the public health space, a lot of attention has been spent on patient-staff and staff-staff contacts within a hospital, for example in [@barnes_dynamic_2010; @martinet_link_2018].
+
+# Methods
+
 ## Setting
 
-Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.
-
-Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
-In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study.
+Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ healthcare facilities that vary in size, case mix and services. 
+Admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
+In particular, data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study.
 All Victorian hospital admissions during this study period are included.
@@
 
-Some exploratory analysis of this data was done in by [@lydeamore_carbapenemase-producing_2024].
+Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.
 
-# Methods
+## Network models of patient transfers for temporal fidelity
 
-## Existing modelling {#sec-network-modelling}
-
-There has been some existing work that analyse hospital systems using the framework of network analysis. 
-They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
-Using these networks, hospitals and links that are important to mrough hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
-The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
-However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
-It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
 In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
@@
 times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.
-
-Temporal networks are also not new phenomena in modelling infectious diseases processes. For example, 
-[ref] [does thing]. However, edges of temporal networks typically represent
-
-## Network models of patient transfers for temporal fidelity
 
@@
 For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.
-This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
+This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit at their next healthcare facility.
 This alters the indirect transfer processes to
@@
 Since analytical results are intractable for the more complex network models, we resort to simulation as a statistical proxy.
-Movements of marked individuals are tracked for each network type, when they are seeded at each facility in the network at $t=0$. The hitting time $k_j$ to each other hospital is recorded.
+Movements of marked individuals are tracked for each network type, when they are seeded at each facility in the network at $t=0$. The hitting time $k_j$ to each other healthcare facililty is recorded.
 We mark $N_0 = 30$ individuals at the start of each realisation; we generate $N_r=20$ realisations for each seed facility.
@@
 We see that at early time the mean empirical and analytical expected hitting times agree, with a large right-hand censoring effect.
-When computing the empirical means, we allow `{python} n_null_accept` nulls across all realisations.
+When computing the empirical means, we allow `{python} n_null_accept` nulls (where the target facility is not hit) across all realisations.
 :::

### 2025-06-09 reports/diagrams/network_diags/temporal.tex
meaningless small updates

@@
 
-\node at (19.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A0){A};
-\node at (19.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B0){B};
-\node at (61.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A1){A};
-\node at (61.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B1){B};
+\node at (19.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A0){$u_s$};
+\node at (19.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B0){$v_s$};
+\node at (61.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A1){$u_t$};
+\node at (61.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B1){$v_t$};
 

### 2025-06-09 reports/paper.qmd
meaningless small updates

@@
 
-:::{.hidden}
-$$
-\newcommand{\dd}[2][t]{\frac{\mathrm{d}#2}{\mathrm{d}#1}}
-\newcommand{\ceil}[1]{\lceil#1\rceil}
-$$
-:::
-
 ![Sketch of the generalised model with processes of (1) permanent discharge ($u \to \emptyset$), (2) direct transfer ($u \to v$), (3) return home ($u \to h_{uv}$), (4) readmission $h_{uv} \to v$).](diagrams/movement_general_case.pdf){#fig-general-diagram}
@@
 We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the hodling time an individual patient spends in each state.
+
+:::{#fig-specific-model-diagrams layout-ncol=2 layout-nrow=2 layout-valign="bottom"}
+
+![Naïve static](diagrams/network_diags/naive_static.pdf)
+
+![Static](diagrams/network_diags/static.pdf)
+
+![Snapshot](diagrams/network_diags/snapshot.pdf)
+
+![Temporal](diagrams/network_diags/temporal.pdf)
+
+Schematic diagrams of each type of network representation for a two-facility network.
+Solid arrows represent direct transfers, dashed arrows represent indirect transfers, and wavy lines represent temporal consistency.
+:::
 
@@
     date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
-    my=pl.col('w').rolling_mean(window_size=365, center=True),
-    mw=pl.col('w').rolling_mean(window_size=7, center=True),
-    mm=pl.col('w').rolling_mean(window_size=28, center=True),
+    my=pl.col('w').rolling_mean(window_size=365, center=True), # yearly ma
+    mw=pl.col('w').rolling_mean(window_size=7, center=True), # weekly ma
+    mm=pl.col('w').rolling_mean(window_size=28, center=True), # monthly ma
   )
@@
 #| fig-cap: Number of patient transfers throughout the network increases over time.
+Moving averages plotted in red with periods of 1 year (main plot) and 1 week (inset).
 Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period.
@@
 ```{python}
+#| label: compute-load-hosp-presence
 histories = {

### 2025-06-23 reports/diagrams/new-networks/naive_static.tex
Last week's stuff

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}

    \node at (0.0, 0.0) (NORTHWEST){};
    \node at (10.0, 10.0) (SOUTHEAST){};

    \node at (5.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
    \node at (5.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};

    \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
    \draw [out=45,in=315,-stealth,solid](B.45)to(A.315);

\end{tikzpicture}

\end{document}

### 2025-06-23 reports/diagrams/new-networks/snapshot.tex
Last week's stuff

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}

    \node at (0.0, 0.0) (NORTHWEST){};
    \node at (16.0, 10.0) (SOUTHEAST){};

    \node at (2.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
    \node at (2.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
    \node at (6.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H){$h_{uv}$};

    \node at (10.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A2){$u$};
    \node at (10.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B2){$v$};
    \node at (14.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H2){$h_{uv}$};

    \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
    \draw [out=0,in=270,-stealth,densely dashed](B.0)to(H.270);

    \draw [out=90,in=0,-stealth,densely dashed](H2.90)to(A2.0);

    \draw [loosely dashed, very thick](8.0, 0.0)--(8.0, 10.0);

\end{tikzpicture}

\end{document}

### 2025-06-23 reports/diagrams/new-networks/static.tex
Last week's stuff

--- 2025-06-23 reports/diagrams/new-networks/snapshot.tex
@@
     \node at (0.0, 0.0) (NORTHWEST){};
-    \node at (16.0, 10.0) (SOUTHEAST){};
+    \node at (10.0, 10.0) (SOUTHEAST){};
 
-    \node at (2.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
-    \node at (2.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
-    \node at (6.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H){$h_{uv}$};
-
-    \node at (10.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A2){$u$};
-    \node at (10.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B2){$v$};
-    \node at (14.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H2){$h_{uv}$};
+    \node at (3.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
+    \node at (3.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
+    \node at (7.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H){$h_{uv}$};
 
     \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
+    %\draw [out=45,in=315,-stealth,solid](B.45)to(A.315);
     \draw [out=0,in=270,-stealth,densely dashed](B.0)to(H.270);
-
-    \draw [out=90,in=0,-stealth,densely dashed](H2.90)to(A2.0);
-
-    \draw [loosely dashed, very thick](8.0, 0.0)--(8.0, 10.0);
+    \draw [out=90,in=0,-stealth,densely dashed](H.90)to(A.0);
 

### 2025-06-23 reports/diagrams/new-networks/temporal.tex
Last week's stuff

--- 2025-06-23 reports/diagrams/new-networks/snapshot.tex
@@
 
-    \node at (2.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
-    \node at (2.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
-    \node at (6.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H){$h_{uv}$};
+    \node at (4.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
+    \node at (4.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
 
-    \node at (10.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A2){$u$};
-    \node at (10.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B2){$v$};
-    \node at (14.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H2){$h_{uv}$};
+    \node at (12.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A2){$u$};
+    \node at (12.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B2){$v$};
 
     \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
-    \draw [out=0,in=270,-stealth,densely dashed](B.0)to(H.270);
+    \draw [out=0,in=180,-stealth,densely dashed](B.0)to(A2.180);
 
-    \draw [out=90,in=0,-stealth,densely dashed](H2.90)to(A2.0);
-
-    \draw [loosely dashed, very thick](8.0, 0.0)--(8.0, 10.0);
+    \draw [loosely dashed,very thick](8.0, 0.0)--(8.0, 10.0);
 

### 2025-06-23 reports/paper.qmd
Last week's stuff

@@
 import polars as pl
-from matplotlib import pyplot as plt, dates as mdates, colors as mcolors
+from matplotlib import pyplot as plt, dates as mdates, colors as mcolors, cm
 import seaborn as sns
@@
 import datetime
+import copy
 
@@
 
+Perhaps most pertinently, [@belik_leveraging_2017] explores patient referral networks using a temporal network methodology, and also consider the effect of community stays by augmenting their temporal network with additional edges corresponding to readmissions that occur within a period of $k$ days (representative of the infectious period).
+Here, we will show that
+
 # Methods
@@
 In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison.
-In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granulovement around the hospital system by analysing measures of centrality and clustering.
+Qualitatively, there are mechanisms implicit in the data that are not previously well-modelled, and may be of importance to constructing robust control measures.
+Critically, we notice that the distribution of readmission times (the time from a discharge of a given patient from a facility to next admission of that patient at any facility), as plotted in @fig-readmission-time, are not well modelled by an exponential distribution, i.e.
+poorly approximated by Markov processes.
+
+```{python}
+#| label: fig-readmission-time
+#| fig-cap: Distribution of empirical readmission times (time from discharge to next admission), truncated at 365 days; and best-fit exponential (via MLE).
+
+readmission_df = pl.read_parquet(root / "data/concordant_networks/readmission_time.parquet")
+readmission_df = (readmission_df
+  .with_columns(
+    Density = pl.col('count') / pl.col('count').sum()
+  )
+  .rename(
+    {
+      'readmission_time': 'Readmission Time [days]', 
+      'count': 'Count',
+    }
+  )
+)
+
+fig, ax = plt.subplots()
+
+sns.scatterplot(readmission_df, x='Readmission Time [days]', y='Density', ax=ax, label='data')
+sns.despine(ax=ax)
+
+ax.set_xscale('asinh')
+_xlims = ax.set_xlim(-0.1, None)
+ax.set_yscale('log')
+_ylims = ax.get_ylim()
+
+# get best fit exponential, we know that gamma = 1/mean
+readmission_mean = readmission_df.select(
+  pl.col('Readmission Time [days]') * pl.col('Density')
+).sum().item()
+
+gamma = 1. / readmission_mean
+_xgrid = np.arange(0, 366, 1)
+ax.plot(_xgrid, gamma * np.exp(-(_xgrid * gamma)), color='orange', linestyle='dashed', label='best-fit exponential')
+
+ax.set_ylim(*_ylims)
+ax.set_xlim(*_xlims)
+ax.set_xticks([0, 1, 10, 100])
+
+ax.legend()
+```
+
 A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.
@@
 
-![Naïve static](diagrams/network_diags/naive_static.pdf)
+![Naïve static](diagrams/new-networks/naive_static.pdf)
 
-![Static](diagrams/network_diags/static.pdf)
+![Static](diagrams/new-networks/static.pdf)
 
-![Snapshot](diagrams/network_diags/snapshot.pdf)
+![Snapshot](diagrams/new-networks/snapshot.pdf)
 
-![Temporal](diagrams/network_diags/temporal.pdf)
+![Temporal](diagrams/new-networks/temporal.pdf)
 
-Schematic diagrams of each type of network representation for a two-facility network.
-Solid arrows represent direct transfers, dashed arrows represent indirect transfers, and wavy lines represent temporal consistency.
+Schematic diagrams of each type of network representation for a two-facility network. There exists a direct transfer from $u$ to $v$ and an indirect transfer from $v$ to $u$. 
 :::
@@
 
-We plot the (empirical) cumulative distribution function of the hitting times for all pairs of seed facility and target facility in @fig-ecdf
+We plot the (empirical) cumulative distribution function of the hitting times for all pairs of seed facility and target facility in @fig-ecdf, with a zoom of the early-time distribution in @fig-ecdf-close.
+We see that the naive static network model produces a hitting time distribution that has a large amount of movement of the marked individuals in a short amount of time, and that the full temporal network model produces much less movement.
+This is consistent with what is expected.
+We see that the
 
@@
 
-ax.set_xlabel('hitting time')
+ax.set_xlabel('Hitting Time (days)')
+ax.set_ylabel('ECDF')
+
+ax.legend(loc='lower right')
+sns.despine(ax=ax)
+```
+
+```{python}
+#| label: fig-ecdf-close
+#| fig-cap: Empirical cumulative distributions of the hitting times between pairs of facilities over approxiumately the first year, for different network models.
+
+fig, ax = plt.subplots()
+for model in model_order:
+    dist = hitting_time_dists[model]
+    ax.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model.replace('_', ' '))
+
+ax.set_xlim(-15, 350)
+ax.set_ylim(-0.025, 0.15)
+
+ax.set_xlabel('Hitting Time (days)')
 ax.set_ylabel('ECDF')
@@
 
-One striking feature
+When comparing the aggregate number of individuals present or not present in the hospital system between models, we see some striking differences.
+The naïve static model has no capability for discharging individuals out of the system, so is significantly different to the other network models.
+For the other models, after some burn-in period, they reach a (pseudo-)steady state. 
+Notable features:
+- the band at N=0 for the snapshot and temporal network models, where some individuals do not discharge - this is usually due to their seed hospital not existing until sometime much later in the simulation.
+This cannot be modelled with static networks, so instead in the static network, there are bands of lagging that correspond to lower global rates of discharge.
+- the differing values of steady state for each model. Temporal model has the most at home, followed by the static model, and then the snapshot model.
+- it seems the mechanisms for readmission in the snapshot model lead to periodic bursty readmissions, which results in a lower number of individuals at home.
 
@@
 #| label: fig-homeprop-heatmap
-#| fig-cap: Distribution of the number of indivdiuals at home (not in the hospital system) over the duration of the simulation for each model type.
+#| fig-cap: Distribution of the number of individuals at home (not in the hospital system) over the first 3 years of the simulation for each model type.
 
@@
     emp_max_date = (np.timedelta64(int(N*dt), 'D') + _BASE_DATE).astype('datetime64[M]')
-    ax_max_date = np.min([emp_max_date, _MAX_DATE])
+    ax_lim_date = (np.timedelta64(int(np.max(ax.get_xlim())), 'D') + _BASE_DATE).astype('datetime64[M]')
+    ax_max_date = np.min([emp_max_date, _MAX_DATE, ax_lim_date])
 
@@
 _DTS['temporal'] = 0.5
-def single_heatmap(model, hist, bins=(540, 31), norm=None, ax=None, cbar=False):
+def single_heatmap(model, hist, bins=(540, 31), cutoff=None, norm=None, ax=None, cbar=False):
     if ax is None:
@@
     ts = (np.arange(hist.shape[1]) * dt).reshape((1, -1))
-    ts_arr = np.repeat(ts, hist.shape[0], axis=0)
+    ts_arr = np.repeat(ts, hist.shape[0], axis=0).flatten()
+
+    if cutoff is not None:
+      ts_mask = ts_arr < cutoff
+      ts_arr = ts_arr[ts_mask]
+      hist_arr = hist.flatten()[ts_mask]
+    else:
+      hist_arr = hist.flatten()
 
     *_, coll = ax.hist2d(
-        ts_arr.flatten(), 
-        30 - hist.flatten(),
+        ts_arr, 
+        30 - hist_arr,
         bins=bins,
@@
 }
-def final_heatmap(hists, norm=mcolors.PowerNorm(0.5)):
+def final_heatmap(hists, norm=mcolors.PowerNorm(0.5), bins=(540, 31), cutoff=None):
 
@@
         ax = axs_flat[_ax_order[k]]
-        coll, _ = single_heatmap(k, h, norm=norm, ax=ax)
+        coll, _ = single_heatmap(k, h, bins=bins, cutoff=cutoff, norm=norm, ax=ax)
         norm = coll.norm
@@
     fig.subplots_adjust(hspace=0.3)
-    fig.colorbar(coll, ax=axs, label='Density', extend='max')
+    coll_norm = copy.deepcopy(coll.norm) 
+    coll_norm.autoscale([0.0, bins[0]* coll.norm.vmax])
+    scaled_scalar_map = cm.ScalarMappable(norm=coll_norm, cmap=coll.get_cmap())
+    fig.colorbar(scaled_scalar_map, ax=axs, label='Density', extend='max')
     fig.supylabel("Number at home ($N_0=30$)")
@@
 
-_ = final_heatmap(histories)
+_ = final_heatmap(histories, norm=mcolors.PowerNorm(0.4), bins=(313,31), cutoff=3*365)
 ```

### 2025-06-23 reports/paper.qmd
smol  mon ni

@@
 Ultimately, the aim of studying these systems as networks is leveraging the network framework when designing surveillance and control protocols for emerging and recurring AMR outbreaks.
-This has been analysed to some extent in [@ciccolini_efficient_2014], for example; but such analyses use
+This has been analysed to some extent in [@ciccolini_efficient_2014, @++needmore], but these studies neglect the temporal structure of the healthcare system, through the use of assumptions such as patients will be disease-free when they are discharged.
 
-There are existing results that show that accounting for temporality in network models results in lower control costs [@li_fundamental_2017].
+The temporal structure of the healthcare systems can be considered by using temporal network models.
+There are existing results that show that control costs are lower in temporal networks[@li_fundamental_2017].
 In the public health space, a lot of attention has been spent on patient-staff and staff-staff contacts within a hospital, for example in [@barnes_dynamic_2010; @martinet_link_2018].
-
 Perhaps most pertinently, [@belik_leveraging_2017] explores patient referral networks using a temporal network methodology, and also consider the effect of community stays by augmenting their temporal network with additional edges corresponding to readmissions that occur within a period of $k$ days (representative of the infectious period).
-Here, we will show that
 
@@
 
-Since analytical results are intractable for the more complex network models, we resort to simulation as a statistical proxy.
+Since analytical results are intractable for our more complex network models, we resort to simulation as a statistical proxy.
 Movements of marked individuals are tracked for each network type, when they are seeded at each facility in the network at $t=0$. The hitting time $k_j$ to each other healthcare facililty is recorded.

### 2025-06-30 reports/paper.qmd
yep week

@@
         \usepackage{setspace}
-        \doublespacing
+        %\doublespacing
     keep-tex: true
@@
 
-# Methods
+# Materials and Methods
 
-## Setting
-
-Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ healthcare facilities that vary in size, case mix and services. 
-Admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
-In particular, data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study.
-All Victorian hospital admissions during this study period are included.
-The total number of admission records is 26876787.
-This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.
-
+We use admissions data sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
+This contains line-listed admissions for patients in Victoria, Australia, covering 338 healthcare facilities that vary in size, case mix and services.
+the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
+Because there is no unique healthcare identifier in Victoria, Australia, linkage is performed on the VAED dataset by the Centre for Victorian Data Linkage, in order to build information on succesive admissions of patients.
+As a result of this linkage and hospital administrative procedures, there can be overlapping admissions, where a patient admits at a facility before being discharged at a prior facility.
+These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
 Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.
 
-## Network models of patient transfers for temporal fidelity
+## Data Characterisation
+
+To motivate the following methods, we first characterise temporal trends in the data.
+
+Firstly, we examine the duration of stays
+
+```{python}
+#| label: fig-stay-duration
+#| fig-cap: Distribution of duration of stays
+
+stay_dur_df = pl.read_csv(root / "data/export/stay_duration_hist.csv")
+
+fig, ax = plt.subplots()
+sns.scatterplot(stay_dur_df, x='duration', y='count')
+```
 

### 2025-06-30 reports/paper.qmd
add monday night

@@
 These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
-Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.
+Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework, and it characterises the structure of the healthcare system.
 
@@
 
-To motivate the following methods, we first characterise temporal trends in the data.
+The temporal characterisation of the Victorian healthcare system is less well-explored.
+For example, the amount of patient movement in the system fluctuates temporally, as shown in @fig-movement-over-time. 
+There is an increasing amount of transfers over time, and there are dips in transfers related to public holidays in the state of Victoria.
+This implies that we should have at least an inhomogenous Poisson process governing the rate of transfers between facilites over time.
 
-Firstly, we examine the duration of stays
+```{python}
+#| label: compute-movement-over-time
+base_date = datetime.date(2011, 1, 1)
+aggregated_movement = pl.read_csv(root / "data/concordant_networks/analyses/movement_aggregated.csv")
+aggregated_movement = (
+  aggregated_movement
+  .group_by('t')
+  .agg(pl.col('w').sum())
+  .sort('t')
+  .with_columns(
+    date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
+    my=pl.col('w').rolling_mean(window_size=365, center=True), # yearly ma
+    mw=pl.col('w').rolling_mean(window_size=7, center=True), # weekly ma
+    mm=pl.col('w').rolling_mean(window_size=28, center=True), # monthly ma
+  )
+)
+```
+
+```{python}
+#| label: fig-movement-over-time
+#| fig-cap: Number of patient transfers throughout the network increases over time.
+Moving averages plotted in red with periods of 1 year (main plot) and 1 week (inset).
+Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period.
+We also observe seasonal behaviour (most visually obvious with periods of a year and a week), and consistent decreases of transfers during public holiday periods, especially at the end of the year.
+fig = plt.figure()
+ax = fig.add_subplot()
+sns.despine(ax=ax)
+
+nth_year = 4
+x1, x2 = datetime.date(2011+nth_year-1, 12, 15), datetime.date(2011+nth_year+1, 1, 15)
+y1, y2 = 12_000, 25_000
+inset = ax.inset_axes(
+  [0.15, 0.1, 0.7, 0.35], 
+  xlim=(x1, x2), 
+  ylim=(y1, y2),
+  xlabel='',
+  ylabel='',
+  yticks=[],
+  yticklabels=[],
+)
+for axes in (ax, inset):
+  sns.lineplot(aggregated_movement, x='date', y='w', color='C0', ax=axes)
+
+sns.lineplot(aggregated_movement, x='date', y='my', color='red', ax=ax)
+sns.lineplot(aggregated_movement, x='date', y='mw', color='red', ax=inset)
+
+ax.indicate_inset_zoom(inset, edgecolor='black')
+
+inset.set_xlabel('')
+inset.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
+_ = ax.set_xlabel('Date')
+_ = ax.set_ylabel('Number of patient transfers')
+```
+
+The underlying admission duration is ___ distributed (see @fig-stay-duration), and the time between admissions (i.e.
+the transfer period) is power-law distributed (see @fig-readmission-time).
+Thus, a fully Markovian POisson process on a static network is unlikely to be a good representation of the movement in the network.
 
@@
 ```
-
-In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison.
-Qualitatively, there are mechanisms implicit in the data that are not previously well-modelled, and may be of importance to constructing robust control measures.
-Critically, we notice that the distribution of readmission times (the time from a discharge of a given patient from a facility to next admission of that patient at any facility), as plotted in @fig-readmission-time, are not well modelled by an exponential distribution, i.e.
-poorly approximated by Markov processes.
 
@@
 ```
+
+## Network Models
 
@@
 We see evidence of this in @fig-movement-over-time where we plot the overall rates of movement in the network over time.
-
-```{python}
-#| label: compute-movement-over-time
-base_date = datetime.date(2011, 1, 1)
-aggregated_movement = pl.read_csv(root / "data/concordant_networks/analyses/movement_aggregated.csv")
-aggregated_movement = (
-  aggregated_movement
-  .group_by('t')
-  .agg(pl.col('w').sum())
-  .sort('t')
-  .with_columns(
-    date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
-    my=pl.col('w').rolling_mean(window_size=365, center=True), # yearly ma
-    mw=pl.col('w').rolling_mean(window_size=7, center=True), # weekly ma
-    mm=pl.col('w').rolling_mean(window_size=28, center=True), # monthly ma
-  )
-)
-```
-
-```{python}
-#| label: fig-movement-over-time
-#| fig-cap: Number of patient transfers throughout the network increases over time.
-Moving averages plotted in red with periods of 1 year (main plot) and 1 week (inset).
-Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period.
-We also observe seasonal behaviour (most visually obvious with periods of a year and a week), and consistent decreases of transfers during public holiday periods, especially at the end of the year.
-fig = plt.figure()
-ax = fig.add_subplot()
-sns.despine(ax=ax)
-
-nth_year = 4
-x1, x2 = datetime.date(2011+nth_year-1, 12, 15), datetime.date(2011+nth_year+1, 1, 15)
-y1, y2 = 12_000, 25_000
-inset = ax.inset_axes(
-  [0.15, 0.1, 0.7, 0.35], 
-  xlim=(x1, x2), 
-  ylim=(y1, y2),
-  xlabel='',
-  ylabel='',
-  yticks=[],
-  yticklabels=[],
-)
-for axes in (ax, inset):
-  sns.lineplot(aggregated_movement, x='date', y='w', color='C0', ax=axes)
-
-sns.lineplot(aggregated_movement, x='date', y='my', color='red', ax=ax)
-sns.lineplot(aggregated_movement, x='date', y='mw', color='red', ax=inset)
-
-ax.indicate_inset_zoom(inset, edgecolor='black')
-
-inset.set_xlabel('')
-inset.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
-_ = ax.set_xlabel('Date')
-_ = ax.set_ylabel('Number of patient transfers')
-```
 

### 2025-07-18 reports/paper.qmd
stuff

@@
 import polars as pl
+from scipy import stats
 from matplotlib import pyplot as plt, dates as mdates, colors as mcolors, cm
@@
 
-## Data Characterisation
-
-The temporal characterisation of the Victorian healthcare system is less well-explored.
-For example, the amount of patient movement in the system fluctuates temporally, as shown in @fig-movement-over-time. 
-There is an increasing amount of transfers over time, and there are dips in transfers related to public holidays in the state of Victoria.
-This implies that we should have at least an inhomogenous Poisson process governing the rate of transfers between facilites over time.
-
-```{python}
-#| label: compute-movement-over-time
-base_date = datetime.date(2011, 1, 1)
-aggregated_movement = pl.read_csv(root / "data/concordant_networks/analyses/movement_aggregated.csv")
-aggregated_movement = (
-  aggregated_movement
-  .group_by('t')
-  .agg(pl.col('w').sum())
-  .sort('t')
-  .with_columns(
-    date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
-    my=pl.col('w').rolling_mean(window_size=365, center=True), # yearly ma
-    mw=pl.col('w').rolling_mean(window_size=7, center=True), # weekly ma
-    mm=pl.col('w').rolling_mean(window_size=28, center=True), # monthly ma
-  )
-)
-```
-
-```{python}
-#| label: fig-movement-over-time
-#| fig-cap: Number of patient transfers throughout the network increases over time.
-Moving averages plotted in red with periods of 1 year (main plot) and 1 week (inset).
-Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period.
-We also observe seasonal behaviour (most visually obvious with periods of a year and a week), and consistent decreases of transfers during public holiday periods, especially at the end of the year.
-fig = plt.figure()
-ax = fig.add_subplot()
-sns.despine(ax=ax)
-
-nth_year = 4
-x1, x2 = datetime.date(2011+nth_year-1, 12, 15), datetime.date(2011+nth_year+1, 1, 15)
-y1, y2 = 12_000, 25_000
-inset = ax.inset_axes(
-  [0.15, 0.1, 0.7, 0.35], 
-  xlim=(x1, x2), 
-  ylim=(y1, y2),
-  xlabel='',
-  ylabel='',
-  yticks=[],
-  yticklabels=[],
-)
-for axes in (ax, inset):
-  sns.lineplot(aggregated_movement, x='date', y='w', color='C0', ax=axes)
-
-sns.lineplot(aggregated_movement, x='date', y='my', color='red', ax=ax)
-sns.lineplot(aggregated_movement, x='date', y='mw', color='red', ax=inset)
-
-ax.indicate_inset_zoom(inset, edgecolor='black')
-
-inset.set_xlabel('')
-inset.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
-_ = ax.set_xlabel('Date')
-_ = ax.set_ylabel('Number of patient transfers')
-```
-
-The underlying admission duration is ___ distributed (see @fig-stay-duration), and the time between admissions (i.e.
-the transfer period) is power-law distributed (see @fig-readmission-time).
-Thus, a fully Markovian POisson process on a static network is unlikely to be a good representation of the movement in the network.
-
-```{python}
-#| label: fig-stay-duration
-#| fig-cap: Distribution of duration of stays
-
-stay_dur_df = pl.read_csv(root / "data/export/stay_duration_hist.csv")
-
-fig, ax = plt.subplots()
-sns.scatterplot(stay_dur_df, x='duration', y='count')
-```
-
-```{python}
-#| label: fig-readmission-time
-#| fig-cap: Distribution of empirical readmission times (time from discharge to next admission), truncated at 365 days; and best-fit exponential (via MLE).
-
-readmission_df = pl.read_parquet(root / "data/concordant_networks/readmission_time.parquet")
-readmission_df = (readmission_df
-  .with_columns(
-    Density = pl.col('count') / pl.col('count').sum()
-  )
-  .rename(
-    {
-      'readmission_time': 'Readmission Time [days]', 
-      'count': 'Count',
-    }
-  )
-)
-
-fig, ax = plt.subplots()
-
-sns.scatterplot(readmission_df, x='Readmission Time [days]', y='Density', ax=ax, label='data')
-sns.despine(ax=ax)
-
-ax.set_xscale('asinh')
-_xlims = ax.set_xlim(-0.1, None)
-ax.set_yscale('log')
-_ylims = ax.get_ylim()
-
-# get best fit exponential, we know that gamma = 1/mean
-readmission_mean = readmission_df.select(
-  pl.col('Readmission Time [days]') * pl.col('Density')
-).sum().item()
-
-gamma = 1. / readmission_mean
-_xgrid = np.arange(0, 366, 1)
-ax.plot(_xgrid, gamma * np.exp(-(_xgrid * gamma)), color='orange', linestyle='dashed', label='best-fit exponential')
-
-ax.set_ylim(*_ylims)
-ax.set_xlim(*_xlims)
-ax.set_xticks([0, 1, 10, 100])
-
-ax.legend()
-```
-
 ## Network Models
 
-A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.
-For each edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, 
-we can construct a reaction that has a hazard function $\lambda_{ij} = w_{ij} / \tau$.
-This implicitly constructs a process that has inter-event times (i.e.
-times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.
+This data can be used to inform the construction of a wide variety of (weighted, directed) networks, which can, in turn, be used to construct a stochastic process model.
+In this study, we will use discrete time Markov chain models, where a state vector $x$ that represents the number of marked individuals at facilities $u$.
+Weights $w_{ij}$ of the edges $e_{ij}$ between nodes $u_i, u_j$ of the network inform the hazard function of the movement process of marked individuals between those nodes.
+By default, the movement process is a Markovian process, i.e.
+is a Poisson random variable with hazard $\lambda_{ij}(x_i, t) = c_iw_{ij}x_i\Delta t$, where $c_i$ is a normalising constant, and $\Delta t$ is the time step.
+
+Applying this directly to a static network of transfers between healthcare facilities, with appropriate infection dynamics, is what is done in @donker_hospital_2012, for example, and represents a model where ....
+However, since individuals tend to leave the hospital system for non-negligible amoutns of time, it may be appropriate to include time spent in the community (e.g.
+at home) as additional nodes of the network.
+In @belik_leveraging_2017, this approach is taken with a temporal network approach in snapshot representation, where the system over a time period is represented as a series of static networks that "snapshot" the system at discrete points in time.
+Through examination of the behaviour of transfers, especially those that are indirect and cross-facilitly, we want to expand on this temporal representation, and include edges that explicityl ...
+
+We can capture the This motivates the general representation in @fig-general-diagram, where edges from a location $u$ can go to null (discharged and never seen again), directly to some location $v$ (that could be $u$), or indirectly to $v$ via a community node $h_{uv}$.
 
 ![Sketch of the generalised model with processes of (1) permanent discharge ($u \to \emptyset$), (2) direct transfer ($u \to v$), (3) return home ($u \to h_{uv}$), (4) readmission $h_{uv} \to v$).](diagrams/movement_general_case.pdf){#fig-general-diagram}
-
-Here, $u$ and $v$ are locations (potentially identical locations), $s$ and $t$ are times. We take that $t \geq s$.
-$\emptyset$ represents patients that are not observed to return to the hospital system after their discharge from location $u$.
-We assume the associated hazard for this process ($\zeta_u(t)$) to be known for all locations $u$ at all times $t$: the modelling of this process is the same for all network models, and will be neglected in the following.
-
-Further, we define the function $w_{uv}(s, t)$ to be the number of individuals that are observed to discharge from $u$ at time $s$ and readmit at time $t$.
-We allow a maximum of one of the pairs $u,v$ and $s,t$ to be reflexive, i.e.
-$u=v$ or $s=t$, but not both.
-
-We will also assume that we will have quantities $N_u$ that represent the "size" of a location $u$.
-We assume that all locations $u$ will have $N_u$ individuals in them at all times, and that the observed mvoements $w_{uv}(s, t)$ can be thought of as being scaled by $N_u$.
 
@@
 
-We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the hodling time an individual patient spends in each state.
+We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the holding time an individual patient spends in each state.
+
+Under this general structure, we can describe four network models that capture different levels of temporal fidelity.
 
@@
 # Results
+
+## Data Characterisation
+
+The amount of patient movement in the system fluctuates temporally, as shown in @fig-movement-over-time. 
+There is an increasing amount of transfers over time, and there are dips in transfers related to public holidays in the state of Victoria.
+This implies that we should have at least an inhomogenous Poisson process governing the rate of transfers between facilites over time.
+
+```{python}
+#| label: compute-movement-over-time
+base_date = datetime.date(2011, 1, 1)
+aggregated_movement = pl.read_csv(root / "data/concordant_networks/analyses/movement_aggregated.csv")
+aggregated_movement = (
+  aggregated_movement
+  .group_by('t')
+  .agg(pl.col('w').sum())
+  .sort('t')
+  .with_columns(
+    date=pl.lit(base_date) + pl.duration(days=pl.col('t')),
+    my=pl.col('w').rolling_mean(window_size=365, center=True), # yearly ma
+    mw=pl.col('w').rolling_mean(window_size=7, center=True), # weekly ma
+    mm=pl.col('w').rolling_mean(window_size=28, center=True), # monthly ma
+  )
+)
+```
+
+```{python}
+#| label: fig-movement-over-time
+#| fig-cap: Number of patient transfers throughout the network increases over time.
+Moving averages plotted in red with periods of 1 year (main plot) and 1 week (inset).
+Right-hand end of data exhibits censoring behaviour, since we do not see long-terms readmissions beyond the end of the data observation period.
+We also observe seasonal behaviour (most visually obvious with periods of a year and a week), and consistent decreases of transfers during public holiday periods, especially at the end of the year.
+fig = plt.figure()
+ax = fig.add_subplot()
+sns.despine(ax=ax)
+
+nth_year = 4
+x1, x2 = datetime.date(2011+nth_year-1, 12, 15), datetime.date(2011+nth_year+1, 1, 15)
+y1, y2 = 12_000, 25_000
+inset = ax.inset_axes(
+  [0.15, 0.1, 0.7, 0.35], 
+  xlim=(x1, x2), 
+  ylim=(y1, y2),
+  xlabel='',
+  ylabel='',
+  yticks=[],
+  yticklabels=[],
+)
+for axes in (ax, inset):
+  sns.lineplot(aggregated_movement, x='date', y='w', color='C0', ax=axes)
+
+sns.lineplot(aggregated_movement, x='date', y='my', color='red', ax=ax)
+sns.lineplot(aggregated_movement, x='date', y='mw', color='red', ax=inset)
+
+ax.indicate_inset_zoom(inset, edgecolor='black')
+
+inset.set_xlabel('')
+inset.xaxis.set_major_formatter(mdates.DateFormatter('%b'))
+_ = ax.set_xlabel('Date')
+_ = ax.set_ylabel('Number of patient transfers')
+```
+
+Binning the length of stay and the time between sequential admissions of the same patient into 1-day bins in @fig-stay-duration and @fig-readmission-time, we can see that both quantities are power-law distributed when aggregated over the entire dataset.
+Thus, a fully Markovian Poisson process on a static network is unlikely to be a good representation of the movement in the network.
+
+```{python}
+def gof_exponential(bin_edges, counts):
+  bin_midpoints = (bin_edges[1:] + bin_edges[:-1]) / 2
+  sample_mean = np.sum(bin_midpoints * counts) / np.sum(counts)
+  lambd = 1./ sample_mean
+
+  cumul_f = np.exp(-lambd * bin_edges)
+  prop_f = -np.diff(cumul_f) * np.sum(counts)
+
+  test_result = stats.chisquare(
+    f_obs=counts,
+    f_exp=prop_f,
+    ddof=len(counts) - 2,
+  )
+
+  return test_result, lambd
+
+```
+
+```{python}
+#| label: fig-stay-duration
+#| fig-cap: Distribution of length of stays, rounded up to the nearest day.
+
+stay_dur_df = pl.read_csv(root / "data/export/length_of_stay_binned.csv")
+
+fig, ax = plt.subplots()
+sns.scatterplot(stay_dur_df, x='bin_edge_right', y='count', ax=ax)
+ax.set_xlabel('Length of Stay [days]')
+ax.set_ylabel('Count')
+ax.set_xscale('log')
+ax.set_yscale('log')
+
+sns.despine(ax=ax)
+```
+
+```{python}
+#| label: fig-readmission-time
+#| fig-cap: Distribution of empirical readmission times (time from discharge to next admission), truncated at 365 days; and best-fit exponential (via MLE).
+
+readmission_df = pl.read_parquet(root / "data/concordant_networks/readmission_time.parquet")
+readmission_df = (readmission_df
+  .with_columns(
+    Density = pl.col('count') / pl.col('count').sum()
+  )
+  .rename(
+    {
+      'readmission_time': 'Readmission Time [days]', 
+      'count': 'Count',
+    }
+  )
+)
+
+fig, ax = plt.subplots()
+
+sns.scatterplot(readmission_df, x='Readmission Time [days]', y='Density', ax=ax, label='data')
+sns.despine(ax=ax)
+
+ax.set_xscale('asinh')
+_xlims = ax.set_xlim(-0.1, None)
+ax.set_yscale('log')
+_ylims = ax.get_ylim()
+
+# get best fit exponential, we know that gamma = 1/mean
+readmission_mean = readmission_df.select(
+  pl.col('Readmission Time [days]') * pl.col('Density')
+).sum().item()
+
+gamma = 1. / readmission_mean
+_xgrid = np.arange(0, 366, 1)
+ax.plot(_xgrid, gamma * np.exp(-(_xgrid * gamma)), color='orange', linestyle='dashed', label='best-fit exponential')
+
+ax.set_ylim(*_ylims)
+ax.set_xlim(*_xlims)
+ax.set_xticks([0, 1, 10, 100])
+
+ax.legend()
+```
 
@@
 
+The only process that does not have exponential waiting times is the dissapation of indivdiuals at home in the fulyl tremporal network model waiting to release into the appropriate location during their specified time bin.
+We need a waiting time distribution that has finite support, since all indivdiuals will return to hospital within the givne timeframe; an exponential waiting time does not have this property.
+We choose the uniform distribution as the most parisomonious distriubtion given the circumstance.
+
 We know that for a random variable $t$ uniformly distributed between $[\tau, \tau+\omega]$, the pdf and survival functions are, respectively:

### 2025-07-18 echidna/simulation_runner.py
add timestamp cap to runner

@@ -1,10 +1,33 @@
+from collections import namedtuple
 import numpy as np
+import datetime
 
 from . import network_conversion as conv
 from . import network_simulation as simm
 
 from os import PathLike
 from typing import Mapping, Self, SupportsFloat, SupportsInt
+
+# namedtuples can be parsed by hdf5 into an array
+Timestamp = namedtuple("Timestamp", ['year', 'month', 'day', 'hour', 'minute', 'second', 'microsecond'])
+
+def get_timestamp() -> Timestamp:
+
+    tnow = datetime.datetime.now()
+    return Timestamp(
+        tnow.year, tnow.month, tnow.day, tnow.hour, tnow.minute, tnow.second, tnow.microsecond
+    )
+
+def coerce_datetime(timestamp: Timestamp) -> datetime.datetime:
+    return datetime.datetime(
+        year=timestamp.year,
+        month=timestamp.month,
+        day=timestamp.day,
+        hour=timestamp.hour,
+        minute=timestamp.minute,
+        second=timestamp.second,
+        microsecond=timestamp.microsecond,
+    )
 
 class Runner():
 
@@ -52,6 +75,9 @@ class Runner():
         self.prob_final_file = prob_final_file
         self.output_file = output_file
         self.beta = beta
+
+        self.converter: conv.Converter 
+        self.sim: simm.Simulation 
 
         if converter_args is None: converter_args = dict()
         self._create_parameter_converter(**converter_args)
@@ -132,13 +158,16 @@ class Runner():
         self.sim.simulate(until=until, nostop=nostop)
 
 
-    def export(self, to=None, identity=None, with_movers=False, **kwargs):
+    def export(self, to=None, identity=None, with_movers=False, with_timestamp=True, **kwargs):
 
         if to is None:
             to = self.output_file
         
         if identity is None:
             identity = self._export_identity()
+
+        if with_timestamp:
+            kwargs['timestamp'] = get_timestamp()
 
         self.sim.export_history(to, identity=identity, with_movers=with_movers, **kwargs)
 
@@ -147,6 +176,6 @@ class Runner():
         
         iden = ""
         if hasattr(self.sim, 'rng'):
-            iden = self.sim.rng.state['state']['state']
+            iden = self.sim.rng.bit_generator.state['state']['state']
         
         return iden

### 2025-08-05 reports/paper.qmd
some progress, mostly power law iftting

@@
 import polars as pl
-from scipy import stats
+from scipy import stats, special, signal, optimize
 from matplotlib import pyplot as plt, dates as mdates, colors as mcolors, cm
@@
 ```{python}
-def gof_exponential(bin_edges, counts):
-  bin_midpoints = (bin_edges[1:] + bin_edges[:-1]) / 2
-  sample_mean = np.sum(bin_midpoints * counts) / np.sum(counts)
-  lambd = 1./ sample_mean
+# power law fitting utils
+def pull_xfx(df, make_midpoint=True, left="bin_edge_right", right="bin_edge_right", quantity='bin_midpoint', count="count"):
+  if make_midpoint:
+    xfx_df = df.with_columns(bin_midpoint=(pl.col(left)+pl.col(right))/2)
+  else:
+    xfx_df = df
+  x_arr = xfx_df.select(quantity).to_series().to_numpy()
+  y_arr = xfx_df.select(count).to_series().to_numpy()
+  return x_arr, y_arr
 
-  cumul_f = np.exp(-lambd * bin_edges)
-  prop_f = -np.diff(cumul_f) * np.sum(counts)
+def estimate_alpha(x, fx, xmin, discr_corr=0.5):
+  mask = x > xmin
+  relative_x = np.log(x[mask] / (xmin-discr_corr))
+  sum_relx = fx[mask] * relative_x
+  return 1 + np.sum(fx[mask]) / np.sum(sum_relx)
 
-  test_result = stats.chisquare(
-    f_obs=counts,
-    f_exp=prop_f,
-    ddof=len(counts) - 2,
-  )
+def scaled_N(x, fx, xmin):
+  mask = x > xmin
+  return np.sum(fx[mask])
 
-  return test_result, lambd
+def power_law_density_fn(xs, alpha, xmin, N=1):
+  return N * (xs ** (-alpha) / special.zeta(alpha, xmin))
 
+# expon fitting utils
+def estimate_gamma(x, fx):
+  mean_val = np.sum(x*fx) / np.sum(fx)
+  return 1./ mean_val
+
+def expon_density_fn(xs, gamma, N=1):
+  return N * gamma * np.exp(-gamma * xs)
+
+# power law graphical method
+def power_law_fit_graphical(x, fx, xmin):
+  mask = (fx > 0) & (x > xmin)
+  coefs = stats.linregress(np.log(x[mask]), np.log(fx[mask]))
+  return coefs.slope, np.exp(coefs.intercept)
 ```
@@
 
+stay_dur_xfx = pull_xfx(stay_dur_df)
+stay_dur_expon_gamma = estimate_gamma(*stay_dur_xfx)
+stay_dur_pow_alpha = estimate_alpha(*stay_dur_xfx, xmin=40)
+
 fig, ax = plt.subplots()
-sns.scatterplot(stay_dur_df, x='bin_edge_right', y='count', ax=ax)
+sns.scatterplot(stay_dur_df, x='bin_edge_right', y='count', linewidth=0, ax=ax)
 ax.set_xlabel('Length of Stay [days]')
@@
 $$dN(t) \sim \text{Binomial}(n(t), dt\lambda(t)).$$
+
+## Fitting power-law distributions for binned data
+
+Due to the underlying data being identiiable health information, we extract binned data for analysing the distributions of patient admission durations and inter-admission duration.
+We also right truncate the inter-admission duration [check]
+
+For the distribution of patient stay durations, we use the method from @clauset_power-law_2009, where the probability distribution function is given by
+
+$$
+f(x) = \frac{\alpha - 1}{x_{\textrm{min}}} \left(\frac{x}{x_{\textrm{min}}\right)^\alpha,
+$$
+
+where $\alpha and x_{\textrm{min}}$ are distributional parameters. For a given $x_\textrm{min}$, the MLE for $\alpha$ is recovered as
+
+$$
+\hat{\alpha}(x_\text{min}) = 1 + n \left[\sum_{i=1}^{n} \log\left(\frac{x_i}{x_\textrm{min}}\right) \right].
+$$
+
+$x_\text{min}$ is estimated by a grid search, and looking the region where the corresponding $\hat{\alpha}\text{s}$ are stable.
+
+```{python}
+#| label: fig-stay-dur-xmin-gridsearch
+#| fig-cap: Stay duration grid search 
+stay_dur_df = pl.read_csv(root / "data/export/length_of_stay_binned.csv")
+
+stay_dur_xfx = pull_xfx(stay_dur_df)
+
+xmin_arr = np.arange(1, 400)
+alpha_hats = [estimate_alpha(*stay_dur_xfx, xmin=xmin) for xmin in xmin_arr]
+Ncounts = [scaled_N(*stay_dur_xfx, xmin=xmin) for xmin in xmin_arr]
+
+fig, axs = plt.subplots(ncols=2, figsize=(10, 4))
+axz = axs.flatten()
+
+axz[1].loglog(xmin_arr, alpha_hats)
+axz[1].set_xlabel(r'$x_\text{min}$')
+axz[1].set_ylabel(r'$\hat\alpha$')
+
+axz[0].scatter(*stay_dur_xfx, alpha=0.3, color='k', zorder=999)
+axz[0].set_xscale('log')
+axz[0].set_yscale('log')
+axz[0].set_xlabel('Stay Duration [Days]')
+axz[0].set_ylabel('Count')
+ylim_ = axz[0].get_ylim()
+
+xgrid = np.linspace(0.5, 2500, 10_000)
+xgrid_norm = mcolors.Normalize(vmin=min(xmin_arr), vmax=max(xmin_arr))
+xgrid_clrs = cm.ScalarMappable(cmap=cm.YlOrRd, norm=xgrid_norm)
+for xmin, alpha, NN in zip(xmin_arr, alpha_hats, Ncounts):
+  axz[0].plot(
+    xgrid, 
+    power_law_density_fn(xgrid, alpha, xmin, N=NN),
+    lw=0.75,
+    color=xgrid_clrs.to_rgba(xmin),
+  )
+axz[0].set_ylim(ylim_)
+fig.colorbar(xgrid_clrs, ax=axz[0])
+```
+
+For readmission, preliminary linear regression on log-log shows that we have $\alpha < 1$, so we cannot use the result from @clauset_power-law_2009. 
+```{python}
+#| label: fig-readmit-gridsearch 
+#| fig-cap: Readmission time grid search 
+readmit_df = pl.read_parquet(root / "data/concordant_networks/readmission_time.parquet")
+
+readmit_xfx = pull_xfx(readmit_df, make_midpoint=False, quantity='readmission_time')
+
+r_xmin_arr = np.arange(1, 200)
+power_law_fits = [power_law_fit_graphical(*readmit_xfx, xmin=xm) for xm in r_xmin_arr]
+pl_alphas, pl_icpts = zip(*power_law_fits)
+
+fig, axs = plt.subplot_mosaic(
+  [
+    ['datafit', 'alpha'], 
+    ['datafit', 'coeficpt'],
+  ], 
+  figsize=[10, 4],
+)
+
+axs['alpha'].plot(r_xmin_arr, pl_alphas)
+# axs['alpha'].set_xlabel(r'$x_\text{min}$')
+axs['alpha'].set_ylabel(r'$\hat\alpha$')
+
+axs['coeficpt'].plot(r_xmin_arr, pl_icpts)
+axs['coeficpt'].set_xlabel(r'$x_\text{min}$')
+axs['coeficpt'].set_ylabel(r'$\hat{c}$')
+
+axs['datafit'].scatter(*readmit_xfx, alpha=0.3, color='k', zorder=997)
+axs['datafit'].set_xscale('log')
+axs['datafit'].set_yscale('log')
+axs['datafit'].set_xlabel('Readmission Duration [Days]')
+axs['datafit'].set_ylabel('Count')
+ylim_ = axs['datafit'].get_ylim()
+
+r_xgrid = np.linspace(0.5, 370, 10_000)
+r_xmin_norm = mcolors.Normalize(vmin=min(r_xmin_arr), vmax=max(r_xmin_arr))
+r_xmin_clrs = cm.ScalarMappable(cmap=cm.YlOrRd, norm=r_xmin_norm)
+for xmin, (alpha, icpt) in zip(r_xmin_arr, power_law_fits):
+  axs['datafit'].plot(
+    r_xgrid, 
+    icpt * r_xgrid ** alpha,
+    lw=0.75,
+    color=r_xmin_clrs.to_rgba(xmin),
+  )
+axs['datafit'].set_ylim(ylim_)
+fig.colorbar(r_xmin_clrs, ax=axs['datafit'])
+```

### 2025-08-19 reports/paper.qmd
updates

@@
     fig-dpi: 300
+    cite-method: citeproc
 author:
@@
 
-Antimicrobial resistance poses a great threat to human health and development. [@murray_global_2022]
+Antimicrobial resistance (AMR) poses a great threat to human health and development. [@murray_global_2022]
 Globally, it poses a large burden, operationally and economically, on hospital systems.
-In Australia, we see [significant] economic burden from AMR.
-this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR. 
-One X of interest, for example, is carbapenemase-producing enterobacteriacea (CPE). 
-Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level. 
-This already causes some level of burden, and introduced strains would further multiply this.
+In Australia, the impact of AMR, including its economic and healthcare burden on the population, has only recently been analysed in detail [@wozniak_health_2019, @wozniak_disease_2022].
+This highlights a gap in the understanding of how AMR may proliferate through the healthcare system in Australia.
+One pathogen of interest, especially in the state of Victoria, is carbapenemase-producing enterobacteriacea (CPE), which is detected throughout the state, and reported through various channels to the Department of Health surveillance report.
+[@needref] Genomic analysis of the data indicates that suspected local acquisition is dominated by a particular gene (?), and it is of interest that its spread through the healthcare system, both actual and potential, be understood beyond reactive monitoring.
+[@vicorian_guideline_on_cpo_for_health_services]
 
-The spread of AMR in these systems has been studied by modelling the hospital system as a network or interconnected healthcare facilities.
+The spread of AMR in healthcare systems has been studied by modelling the hospital system as a network or interconnected healthcare facilities.
 For example, [@donker_patient_2010] examines hospitals in the Netherlands to study the spread of MRSA, followed by a similar analysis of the UK hospital system in [@donker_hospital_2012].
-Ultimately, the aim of studying these systems as networks is leveraging the network framework when designing surveillance and control protocols for emerging and recurring AMR outbreaks.
-This has been analysed to some extent in [@ciccolini_efficient_2014, @++needmore], but these studies neglect the temporal structure of the healthcare system, through the use of assumptions such as patients will be disease-free when they are discharged.
+Ultimately, the analysis of these systems as networks is leveraged when designing surveillance and control protocols for emerging and recurring AMR outbreaks.
+This has been done to some extent in [@ciccolini_efficient_2014, @++needmore], but these studies neglect the temporal structure of the healthcare system, through the use of assumptions such as patients will be disease-free when they are discharged.
 
@@
 
+In this study, we will use the Victorian healthcare system as a case study to further investigate the effects of different levels of temporal fidelity on potential models of AMR spread.
+To remove confounding factors, we will primarily focus on movement of individuals as carriers of AMR, and neglect the disease transmission component of the analysis.
+
 # Materials and Methods
@@
 the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
-Because there is no unique healthcare identifier in Victoria, Australia, linkage is performed on the VAED dataset by the Centre for Victorian Data Linkage, in order to build information on succesive admissions of patients.
-As a result of this linkage and hospital administrative procedures, there can be overlapping admissions, where a patient admits at a facility before being discharged at a prior facility.
-These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
+Overlapping admissions for a patient are resolved by deferring to the patient's most recent admission event, and the method is described in detail in the Appendix, and is used in [@donker_patient_2010,@nekkab_assessing_2020].
+
 Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework, and it characterises the structure of the healthcare system.
 
-## Network Models
+<!--# Model description starts here -->
 
-This data can be used to inform the construction of a wide variety of (weighted, directed) networks, which can, in turn, be used to construct a stochastic process model.
-In this study, we will use discrete time Markov chain models, where a state vector $x$ that represents the number of marked individuals at facilities $u$.
+Admissions data can be used to inform the construction of a wide variety of (weighted, directed) networks, which can, in turn, be used to construct a stochastic process model.
+In this study, we use discrete time Markov chain models, where a state vector $x$ that represents the number of marked individuals at facilities $u$.
 Weights $w_{ij}$ of the edges $e_{ij}$ between nodes $u_i, u_j$ of the network inform the hazard function of the movement process of marked individuals between those nodes.
@@
 
-The governing equations can be roughly expressed as a function of the hazards of each process.
+The governing equations for such models can be captured by sets of transitions modelled as a Markov chain. The hazards of the transition events is given in @eq-general-model-transitions.
 
@@
 \end{aligned}
-$$
+$$ {#eq-general-model-transitions}
 
-We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the holding time an individual patient spends in each state.
+In order in increase the flexibility of the models, we will use a _discrete time_ Markov chain, where we have a Poisson-distributed number of transition events occur within some given timestep.
+For some models, we will also allow for the number of transitions for a particular class of event to be non-Poisson.
+
+<!--# Model abstraction ends here, model details start here -->
 
@@
 That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period.
-This is an oversimplification of the work done in @donker, but represents a "worst" case model of temporal fidelity that can be benchmarked against.
+This is an oversimplification of the work done in @donker_hospital_2012, but represents a "worst" case model of temporal fidelity that can be benchmarked against.
 
@@
 
-Binning the length of stay and the time between sequential admissions of the same patient into 1-day bins in @fig-stay-duration and @fig-readmission-time, we can see that both quantities are power-law distributed when aggregated over the entire dataset.
+Binning the length of stay and the time between sequential admissions of the same patient into 1-day bins in @fig-stay-duration and @fig-readmission-time, we can see that both quantities are power-law distributed, or at least heavy-tailed, when aggregated over the entire dataset.
 Thus, a fully Markovian Poisson process on a static network is unlikely to be a good representation of the movement in the network.
@@
 ```{python}
-# power law fitting utils
+#| label: compute-fitting-util-fns
+
 def pull_xfx(df, make_midpoint=True, left="bin_edge_right", right="bin_edge_right", quantity='bin_midpoint', count="count"):
@@
 
+def scaled_N(x, fx, xmin):
+  mask = x > xmin
+  return np.sum(fx[mask])
+
+# power law via pareto MLE
 def estimate_alpha(x, fx, xmin, discr_corr=0.5):
@@
 
-def scaled_N(x, fx, xmin):
-  mask = x > xmin
-  return np.sum(fx[mask])
-
 def power_law_density_fn(xs, alpha, xmin, N=1):
   return N * (xs ** (-alpha) / special.zeta(alpha, xmin))
+
+# power law graphical method
+def power_law_fit_graphical(x, fx, xmin):
+  mask = (fx > 0) & (x > xmin)
+  coefs = stats.linregress(np.log(x[mask]), np.log(fx[mask]))
+  return coefs.slope, np.exp(coefs.intercept)
 
@@
   return N * gamma * np.exp(-gamma * xs)
-
-# power law graphical method
-def power_law_fit_graphical(x, fx, xmin):
-  mask = (fx > 0) & (x > xmin)
-  coefs = stats.linregress(np.log(x[mask]), np.log(fx[mask]))
-  return coefs.slope, np.exp(coefs.intercept)
 ```
@@
 
+stay_dur_xmin = 28
 stay_dur_xfx = pull_xfx(stay_dur_df)
 stay_dur_expon_gamma = estimate_gamma(*stay_dur_xfx)
-stay_dur_pow_alpha = estimate_alpha(*stay_dur_xfx, xmin=40)
+stay_dur_pow_alpha = estimate_alpha(*stay_dur_xfx, xmin=stay_dur_xmin)
+
+stay_dur_xgrid = np.linspace(1, 3000, 3000)
 
@@
 sns.scatterplot(stay_dur_df, x='bin_edge_right', y='count', linewidth=0, ax=ax)
+ax.set_xscale('log')
+ax.set_yscale('log')
+ylim_ = ax.get_ylim()
+ax.plot(
+  stay_dur_xgrid, 
+  expon_density_fn(
+    stay_dur_xgrid, 
+    stay_dur_expon_gamma, 
+    N=stay_dur_xfx[1].sum()
+  ),
+  color='C1', 
+  label='best-fit exponential',
+)
+ax.plot(
+  stay_dur_xgrid,
+  power_law_density_fn(
+    stay_dur_xgrid, 
+    stay_dur_pow_alpha,
+    xmin = stay_dur_xmin,
+    N=scaled_N(*stay_dur_xfx, xmin=stay_dur_xmin),
+  ),
+  color='C2',
+  label='power law fit'
+)
 ax.set_xlabel('Length of Stay [days]')
 ax.set_ylabel('Count')
-ax.set_xscale('log')
-ax.set_yscale('log')
+ax.set_ylim(ylim_)
+ax.legend()
 
@@
 gamma = 1. / readmission_mean
-_xgrid = np.arange(0, 366, 1)
+_xgrid = np.linspace(0.0, 366, 10_000)
 ax.plot(_xgrid, gamma * np.exp(-(_xgrid * gamma)), color='orange', linestyle='dashed', label='best-fit exponential')
+
+readmits_xfx = pull_xfx(readmission_df, make_midpoint=False, quantity='Readmission Time [days]', count='Density')
+alpha, cees = power_law_fit_graphical(*readmits_xfx, xmin=30)
+ax.plot(_xgrid, cees * _xgrid ** alpha, color='green', label='power law fit')
 
@@
 This leads to a reservoir of individuals of interest that exist outside of the healthcare system that will return at a later time, which leads to rich temporal dynamics.
-These behaviours are poorly captured by a naïve static network; including this mechanism on top of a static network representation can help significantly in moving the qualitative behaviour fo the
+At a high-level, a model with better temporal fidelity suggests a much less severe outbreak -- the number of facilities that could be affected by colonised in a given time period after initial detection is lower than suggested by a more static model.
+However, this slowdown is caused by a large "reservoir" that exists "outside" of the healthcare system, and is thus much more difficult to detect. 
+These behaviours are poorly captured by a naïve static network; including this mechanism on top of a static network representation can help significantly in moving the qualitative behaviour of the hitting time distribution towards the one produced by the temporal network representation.
+
+In the current study, we have used a discrete time stepping scheme in order to allow for some non-Markovian behaviour; it is likely that an agent-based model would be a more "accurate" representation.
+We have also neglected "higher-order" pathing. [@myall_network_2021]
 
@@
 # Supplementary
+
+## Removal of admission overlaps
+
+Due to a variety of reasons ranging from the lack of a unique patient identifier and administrative data entry errors, there are instances of overlapping admissions for a single patient.
+In these cases, we resolve these overlaps by deferring to the admission event that had
 
@@
 This makes some representations, in particular the temporal network, completely intractable, since we would have to allocate a ${(N_f)}^2 \times {(N_T)}^2$ floating point matrix for each time step.
-For $N_f \sim O(10^2), N_T \sim O(10^3)$, this yields a roughly $O(10^{(10\sim 11)}$ byte (10\~100 GB) float32 matrix every time step (in memory).
+For $N_f \sim O(10^2), N_T \sim O(10^3)$, this yields a roughly $O(10^{(10\sim 11)})$ byte (10\~100 GB) float32 matrix every time step (in memory).
 
@@
 $$
-f(x) = \frac{\alpha - 1}{x_{\textrm{min}}} \left(\frac{x}{x_{\textrm{min}}\right)^\alpha,
+f(x) = \frac{\alpha - 1}{x_{\textrm{min}}} \left(\frac{x}{x_{\textrm{min}}}\right)^\alpha,
 $$
@@
 $x_\text{min}$ is estimated by a grid search, and looking the region where the corresponding $\hat{\alpha}\text{s}$ are stable.
+We note that it is probably more appropriate to seek a piecewise power law fit, but we do not use these power law fits directly in our simulation models.
 
@@
 
-axz[1].loglog(xmin_arr, alpha_hats)
+axz[1].plot(xmin_arr, alpha_hats)
 axz[1].set_xlabel(r'$x_\text{min}$')
@@
 
-axz[0].scatter(*stay_dur_xfx, alpha=0.3, color='k', zorder=999)
+axz[0].scatter(*stay_dur_xfx, alpha=0.4, color='C0', zorder=999, linewidth=0)
 axz[0].set_xscale('log')
@@
 xgrid = np.linspace(0.5, 2500, 10_000)
-xgrid_norm = mcolors.Normalize(vmin=min(xmin_arr), vmax=max(xmin_arr))
+xgrid_norm = mcolors.Normalize(vmin=-100, vmax=max(xmin_arr))
 xgrid_clrs = cm.ScalarMappable(cmap=cm.YlOrRd, norm=xgrid_norm)
@@
 axz[0].set_ylim(ylim_)
-fig.colorbar(xgrid_clrs, ax=axz[0])
+clbr = fig.colorbar(xgrid_clrs, ax=axz[0])
+clbr.ax.set_ylim((0, None))
 ```
@@
 For readmission, preliminary linear regression on log-log shows that we have $\alpha < 1$, so we cannot use the result from @clauset_power-law_2009.
+Instead, we inspect the behaviour of a the linear regressop on log-log as a function of a lower threshold $x_\text{min}$.
+The probability density function will follow:
+$$ 
+f(x) = c x^\alpha, \qquad x_\text{min} \leq x \leq x_\text{max}
+$$
+
 ```{python}
@@
 
-axs['datafit'].scatter(*readmit_xfx, alpha=0.3, color='k', zorder=997)
+axs['datafit'].scatter(*readmit_xfx, alpha=0.4, color='C0', zorder=997, linewidth=0)
 axs['datafit'].set_xscale('log')
@@
 r_xgrid = np.linspace(0.5, 370, 10_000)
-r_xmin_norm = mcolors.Normalize(vmin=min(r_xmin_arr), vmax=max(r_xmin_arr))
+r_xmin_norm = mcolors.Normalize(vmin=-50, vmax=max(r_xmin_arr))
 r_xmin_clrs = cm.ScalarMappable(cmap=cm.YlOrRd, norm=r_xmin_norm)
@@
 axs['datafit'].set_ylim(ylim_)
-fig.colorbar(r_xmin_clrs, ax=axs['datafit'])
+clbr = fig.colorbar(r_xmin_clrs, ax=axs['datafit'])
+clbr.ax.set_ylim((0, None))
 ```

### 2025-09-01 reports/paper.qmd
? model rewrite

@@
 the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
+Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework, and it characterises the structure of the healthcare system.
 Overlapping admissions for a patient are resolved by deferring to the patient's most recent admission event, and the method is described in detail in the Appendix, and is used in [@donker_patient_2010,@nekkab_assessing_2020].
 
-Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework, and it characterises the structure of the healthcare system.
-
-<!--# Model description starts here -->
-
-Admissions data can be used to inform the construction of a wide variety of (weighted, directed) networks, which can, in turn, be used to construct a stochastic process model.
-In this study, we use discrete time Markov chain models, where a state vector $x$ that represents the number of marked individuals at facilities $u$.
-Weights $w_{ij}$ of the edges $e_{ij}$ between nodes $u_i, u_j$ of the network inform the hazard function of the movement process of marked individuals between those nodes.
-By default, the movement process is a Markovian process, i.e.
-is a Poisson random variable with hazard $\lambda_{ij}(x_i, t) = c_iw_{ij}x_i\Delta t$, where $c_i$ is a normalising constant, and $\Delta t$ is the time step.
-
-Applying this directly to a static network of transfers between healthcare facilities, with appropriate infection dynamics, is what is done in @donker_hospital_2012, for example, and represents a model where ....
-However, since individuals tend to leave the hospital system for non-negligible amoutns of time, it may be appropriate to include time spent in the community (e.g.
-at home) as additional nodes of the network.
-In @belik_leveraging_2017, this approach is taken with a temporal network approach in snapshot representation, where the system over a time period is represented as a series of static networks that "snapshot" the system at discrete points in time.
-Through examination of the behaviour of transfers, especially those that are indirect and cross-facilitly, we want to expand on this temporal representation, and include edges that explicityl ...
-
-We can capture the This motivates the general representation in @fig-general-diagram, where edges from a location $u$ can go to null (discharged and never seen again), directly to some location $v$ (that could be $u$), or indirectly to $v$ via a community node $h_{uv}$.
-
-![Sketch of the generalised model with processes of (1) permanent discharge ($u \to \emptyset$), (2) direct transfer ($u \to v$), (3) return home ($u \to h_{uv}$), (4) readmission $h_{uv} \to v$).](diagrams/movement_general_case.pdf){#fig-general-diagram}
-
-The governing equations for such models can be captured by sets of transitions modelled as a Markov chain. The hazards of the transition events is given in @eq-general-model-transitions.
+As stated above, existing literature does examine patient transfer networks using network analysis methods. These encompass a large range of dynamics and temporal fidelity of movement.
+Here, we will consider a number of different models of movement at various levels of temporal fidelity.
+In general, we will decompose the model of patient movement as in @fig-general-diagram, where a patient is discharged at location $u$ at time $s$, to either never readmit (move to $\emptyset$), directly transfer to location $v$ at time $s$, or return home $h_{uv}$ and admit at location $v$ at some later time $t$.
+This dynamic is replicated at each pair of locations $(u, v)$, and potentially over all pairs of times $(s, t)$.
+The hazards for these processes are used to define discrete-time stochastic processes that can be analysed, and will be simulated, in order to understand the dynamics of the movement models.
+These can be captured in @eq-general-model-transitions:
 
@@
 \lambda(u(s) \to \emptyset) &= \zeta_u(s)\\
-\lambda(u(s) \to v(t)) &= d_{uv}(s, t)\\
-\lambda(u(s) \to z_{uv}(s, t)) &= \eta_{uv}(s, t)\\
-\lambda(z_{uv}(s, t) \to v(t)) &= \rho_{uv}(s, t)
+\lambda(u(s) \to v(s)) &= d_{uv}(s)\\
+\lambda(u(s) \to h_{uv}(s, t)) &= \eta_{uv}(s, t)\\
+\lambda(h_{uv}(s, t) \to v(t)) &= \rho_{uv}(s, t)
 \end{aligned}
 $$ {#eq-general-model-transitions}
-
-In order in increase the flexibility of the models, we will use a _discrete time_ Markov chain, where we have a Poisson-distributed number of transition events occur within some given timestep.
-For some models, we will also allow for the number of transitions for a particular class of event to be non-Poisson.
-
-<!--# Model abstraction ends here, model details start here -->
 
@@
 
-$$ d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u} $$
+$$ d_{uv}(s) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u} $$
 $$ \eta_{uv}(s,t) = \eta_{uv} = 0 $$
@@
 
-$$d_{uv}(s,t) = d_{uv} =  \frac{\sum_{s,t: (t-s) < \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
+$$d_{uv}(s) = d_{uv} =  \frac{\sum_{s,t: (t-s) < \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
 $$\eta_{uv}(s,t) = \eta_{uv} = \frac{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
@@
 
-$$d_{uv}(s,t) = d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s,t)}{\omega N_u}$$
+$$d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s,t)}{\omega N_u}$$
 $$\eta_{uv}(s,t) = \eta_{uv}(s) = \frac{\sum_{t: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega N_u}$$
@@
 
-$$d_{uv}(s, t) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s, t)}{\omega N_u}$$
+$$d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s, t)}{\omega N_u}$$
 $$\eta_{uv}(s, t) = \frac{\sum_{t: (t-s) \geq \omega} w_{uv}(s, t)}{\omega N_u}$$
@@
 
+There are some clear avenues of further investigation from the current study.
+One of these is the addition of disease transmission dynamics.
+We chose to neglect these mechanisms due to the complexity of defining and characterising the manner of transmission within a facility -- and this can have confounding effects on the interpretation of observed results.
+It is believed that CPE, for example, is transmitted through the environment, with
+
 In the current study, we have used a discrete time stepping scheme in order to allow for some non-Markovian behaviour; it is likely that an agent-based model would be a more "accurate" representation.
-We have also neglected "higher-order" pathing. [@myall_network_2021]
+This would allow for "higher-order" behaviours, such as those observed in [@myall_network_2021], where patient pathways through wards in that case, and hospitals in ours, are dependent on their admission history.
+However, agent-based models are an order of magnitude more expensive to simulate, due to increased computational and memory requirements. The number of
 

### 2025-09-02 reports/paper.qmd
disc

@@
 We chose to neglect these mechanisms due to the complexity of defining and characterising the manner of transmission within a facility -- and this can have confounding effects on the interpretation of observed results.
-It is believed that CPE, for example, is transmitted through the environment, with
+It is believed that CPE, for example, is transmitted through the environment, but commentary on reservoirs and control strategies in the clinical literature [@kotsanas_down_2013,@de_geyter_sink_2017,@meschiari_five-component_2021] suggests that there is no well agreed-on conceptual model of the mechanisms of transmission.
 
-In the current study, we have used a discrete time stepping scheme in order to allow for some non-Markovian behaviour; it is likely that an agent-based model would be a more "accurate" representation.
-This would allow for "higher-order" behaviours, such as those observed in [@myall_network_2021], where patient pathways through wards in that case, and hospitals in ours, are dependent on their admission history.
-However, agent-based models are an order of magnitude more expensive to simulate, due to increased computational and memory requirements. The number of
+Another research avenue is comparing the results against an agent-based simulation.
+In the current study, we have used a discrete time stepping scheme in order to allow for some non-Markovian behaviour while still retaining a reasonable simulation time.
+This results in the loss of the abililty to "track" individual patients through the system.
+In order to recover this ability, and impose more complex rules about individual patient pathways, an agent-based simulation would be required.
+These higher-order pathways have been shown to cause significant difference in ward-to-ward movement within hospital facilities in [@myall_network_2021]; similar results would be expected for transfers between facilities.
+Of course, allowing this would require an order of magnitude more computational power, memory, and time.
 
